### 2019-09-11 ResearchProposal.tex
Additional changes to structure, additions to background, editing of introduction

\documentclass[12pt]{article}

\usepackage[round]{natbib}
\usepackage{booktabs}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,       % color of links to bibliography
    filecolor=magenta,   % color of file links
    urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\begin{document}

\title{Assessing the Impact of MDE and Code Generation on the Sustainability of
  Scientific Computing Software: A Research Proposal} 
\author{Spencer Smith}
\date{\today}

\maketitle

\begin{abstract}
  ...
\end{abstract}

\tableofcontents

\newpage

\section{Introduction}

Without dramatic intervention, our collective confidence in Scientific Computing
Software (SCS) is due for a catastrophic collapse.  We are increasingly trusting
ever more complex and ambitious computations, but our software foundations are
built on sand. Areas of concern include nuclear safety analysis and
computational medicine. Successfully building such software requires
communication between software developers and experts from multiple
domains. Collaboration is difficult at the best of times, and is made worse
because developers favour handcrafted solutions over adapting software
engineering processes, methods and tools \citep{FaulkEtAl2009}. Handcrafted
solutions do not account for the inevitable changes in requirements, design and
implementation. The twin challenges of changing requirements and inadequate
documentation conspire to make computational results notoriously difficult to
reproduce, especially in the important endeavour of one researcher independently
replicating the results of another \citep{Smith2018}.  Inadequate traceability
within documentation means re­certification of previously certified engineering
software is as expensive and time consuming as the original certification
exercise, because of a lack of explicit, effort reducing, traceability
information.

Documentation can address the confidence problem mentioned above.  For software
in general, documentation, written before and during development, provides many
benefits~\cite{Parnas2010}: easier reuse of old designs, better communication
about requirements, more useful design reviews, easier integration of separately
written modules, more effective code inspection, more effective testing, and
more efficient corrections and improvements.  For SCS in particular, developers
do not always value documentation, but documentation has been shown to help
improve SCS.  For instance, \citet{SmithEtAl2015-SS-TR} shows that the quality
of statistical software for psychology is generally improved when developed
using the structured CRAN (Comprehensive R Archive Network) process and tools,
versus an ad hoc process.  \citet{SmithAndKoothoor2016} highlights the value of
proper documentation by redeveloping nuclear safety analysis software.  Twenty
seven (27) worrisome documentation problems were found, including
incompleteness, ambiguity, inconsistency, verifiability, modifiability,
traceability and a lack of abstraction.  A redevelopment experiment with five
existing projects \citep{SmithJegatheesanAndKelly2016} enabled the code owners
(as ascertained through interviews) to clearly see the value of documentation.
However, mirroring other studies \citep{CarverEtAl2007}, the code owners felt
documentation takes too much time.

One promising approach to gain the benefits of documentation, without the often
infeasible time commitment, is to use a Model Driven Engineering (MDE), where
MDE is an engineering approach that exploits structured representations
(metamodels for abstract syntax, and constraints to capture static semantics) to
automate repetitive, tedious and error prone tasks.  In this research proposal,
we will use MDE as a short-hand to include model transformation, Domain Specific
Languages (DSLs) and code generation.  Several current papers point to DSLs and
code/document generation as a transformative technology for documentation,
design and verification of SCS \citep{JohansonAndHasselbring2018, Smith2018}.
The benefit of an MDE approach is to dramatically reduce the cost of developing
reliable, reproducible and re-certifiable software by: i) creating
infrastructure for a knowledge base of scientific and computing models; and, ii)
writing explicit ``recipes'' that weave together this knowledge to generate
theoretical models, design documents, code, test cases and build scripts. The
generator can render to multiple languages, such as html, LaTeX, RTF for
documentation and Python, Java, C, and Matlab, for code. One source of
knowledge, with rules for transformation, means completeness, consistency and
traceability can be achieved by construction. Moreover, these qualities can be
maintained as requirements are modified, design decisions are changed,
documentation standards are varied, and software is re­certified.  The knowledge
will be built up incrementally.  New projects will reuse existing knowledge and
expand as necessary.

Scientists are awakening to the critical importance of SCS sustainability, as
evidenced by the success of Software Carpentry and the Software
Sustainability Institute, but the current solutions are not bold enough.
Although valuable, we need more than studies of developers, training sessions
and project support tools.  \emph{We need transformative change (enabled by
  creating new tools for knowledge capture and document/code generation) where
  documentation, design and verification are expected for all SCS.}

- background and context

- problem statement and research question
- research question: What ideas (if any) from Drasil and Drasil procedures are perceived
by graduate students and end user developers with improving software quality and
productivity - for long lived (safety related?) projects

roadmap of document

\section{Background and Literature Review}

roadmap of this section.  Some overview of the topics themselves, but an
emphasis on how they appear in the SCS context.

\subsection{Software Qualities of Interest}

For instance,
a rational design process can provide quality improvements, and Quality
Assurance (QA), as shown in Table~\ref{Tbl_Qualities}.  Moreover, documentation,
written before and during development, can provide many
benefits~\cite{Parnas2010}: easier reuse of old designs, better communication
about requirements, more useful design reviews, easier integration of separately
written modules, more effective code inspection, more effective testing, and
more efficient corrections and improvements.

- qualities of interest, add sustainability

\begin{description}%[leftmargin=\the\parindent]

\item [\textbf{Verifiability}] involves ``solving the equations
  right''~\cite[p.~23]{Roache1998}; it benefits from rational documentation
  that systematically shows, with explicit traceability, how the governing
  equations are transformed into code.

\item [\textbf{Validatability}] means ``solving the right
  equations''~\cite[p.~23]{Roache1998}.  Validatability is improved by a
  rational process via clear documentation of the theory and assumptions, along
  with an explicit statement of the systematic steps required for experimental
  validation.

% \item [\textbf{Reliability}] is a critical quality for scientific software,
%   since the results of computations are meaningless, if they are not dependable.
%   Reliability is closely tied to verifiability, since the key quality to verify
%   is reliability, while the act of verification itself improves reliability.

% \item [\textbf{Performance}] considerations can make certification challenging, since QA
%   becomes more difficult for more complex code.  However,
%   as Roache~\cite[p.~355]{Roache1998} points out, using simpler algorithms and
%   reducing the number of options in general purpose code, is not always a
%   practical option.

\item [\textbf{Usability}] can be a problem.  Different users, solving the same
  physical problem, using the same software, can come up with different answers,
  due to differences in parameter selection~\cite[p.~370]{Roache1998}.  To
  reduce misuse, a rational process must state expected user characteristics,
  modelling assumptions, definitions and the range of applicability of the code.

\item [\textbf{Maintainability}] is necessary in scientific software, since change,
  through iteration, experimentation and exploration, is inevitable.  Models of
  physical phenomena and numerical techniques necessarily evolve over
  time~\cite{CarverEtAl2007, SegalAndMorris2008}.  Proper documentation,
  designed with change in mind, can greatly assist with change management.%   QA
  % activities need to take the need for creativity into account, while not
  % smothering it~\cite[p.~352]{Roache1998}.

\item [\textbf{Reusability}] provides support for the quality of reliability,
  since reliability is improved by reusing trusted components~\cite{Dubois2005}.
  (Care must still be taken with reusing trusted components, since blind reuse
  in a new context can lead to errors, as dramatically shown in the Ariane 5
  disaster~\cite[p.~37--38]{OliveiraAndStewart2006}.)  The odds of reuse are
  improved when it is considered right from the start.

\item [\textbf{Understandability}] is necessary, since reviewers can only certify
  something they understand.  Scientific software developers have the
  view ``that the science a developer embeds in the code must be apparent to
  another scientist, even ten years later''~\cite{Kelly2013}.
  Understandability applies to the documentation and code, while usability
  refers to the executable software.  Documentation that follows a rational
  process is the easiest to follow.

\item [\textbf{Reproducibility}] is a required component of the scientific
  method~\cite{Davison2012}.  Although QA has, ``a bad name among
  creative scientists and engineers''~\cite[p.~352]{Roache1998}, the community
  need to recognize that participating in QA management also improves
  reproducibility.  Reproducibility, like QA, benefits from a consistent and
  repeatable computing environment, version control and separating code from
  configuration/parameters~\cite{Davison2012}.
\end{description}

\subsection{Desirable Qualities of Documentation}

To address the challenges for adequate documentation for QA and certification
efforts for SCS, the qualities described below need to be a priority.  All but
the final quality listed (abstraction), are adapted from the IEEE recommended
practise for producing good software requirements~\cite{IEEE1998}.

\begin{itemize}

\item \emph{Completeness}: Documentation is said to be complete when all the
  requirements of the software are detailed. That is, each goal, functionality,
  attribute, design constraint, value, data, model, symbol, term (with its unit
  of measurement if applicable), abbreviation, acronym, assumption and
  performance requirement of the software is defined.  The software's response
  to all classes of inputs, both valid and invalid and for both desired and
  undesired events, also needs to be specified.

\item \emph{Consistency}: Documentation is said to be consistent when no subset
  of individual statements are in conflict with each other. That is, a
  specification of an item made at one place in the document should not
  contradict the specification of the same item at another location.

\item \emph{Modifiability}: The documentation should be developed in such a way
  that it is easily modifiable so that likely future changes do not destroy the
  structure of the document. Also it should be easy to reflect the change,
  wherever needed in the document to maintain consistency, traceability and
  completeness. For documentation to be modifiable, its format must be
  structured in a way that repetition is avoided and cross-referencing is
  employed.

\item \emph{Traceability}: Documentation should be traceable, as this
  facilitates maintenance and review. If a change is made to the design or code
  of the software, then all the documentation relating to those segments have to
  be modified.  This property is also important for recertification.

\item \emph{Unambiguity}: Documentation is said to be unambiguous only when
  every requirement's specification has a unique interpretation.  The
  documentation should be unambiguous to all audiences, including developers,
  users and reviewers.

\item \emph{Correctness}: There is no direct tool or method for measuring
  correctness. One way of building confidence in correctness is by reviewing to
  ensure that each requirement stated is one that the stakeholders and experts
  desire.  By maintaining traceability, consistency and unambiguity, we can
  reduce the occurrence of errors and make the goal of reviewing for correctness
  easier.

\item \emph{Verifiability}: Every requirement in the documentation must be the
  one fulfilled by the implemented software. Therefore all the requirements
  should be clear, unambiguous and testable, so that a person or a machine can
  verify whether the software product meets the requirements.

\item \emph{Abstract}: Documented requirements are said to be abstract if they
  state what the software must do and the properties it must possess, but do not
  speak about how these are to be achieved. For example, a requirement can
  specify that an Ordinary Differential Equation (ODE) must be solved, but it
  should not mention that Euler's method should be used to solve the ODE. How to
  accomplish the requirement is a design decision, which is documented during
  the design phase.

\end{itemize}

\subsection{Current State of Practice for SCS Development}

The SCS community is finally realizing that current practices are not
sustainable.  \citet{FaulkEtAl2009} observe, ``growing concern about the
reliability of scientific results based on ... software.''  Embarrassing
failures have occurred, like a retraction of derived molecular protein
structures \citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  A
recent report on directions for SCS research and education states: ``While the
volume and complexity of [SCS] have grown substantially in recent decades, [SCS]
traditionally has not received the focused attention it so desperately needs
... to fulfill this key role as a cornerstone of long-term collaboration and
scientific progress'' \citep{RudeEtAl2018}.  Estimates suggest that the number
of released faults per thousand executable lines of code during a given
program’s life cycle is at best 0.1, and more likely 10 to 100 times worse
\citep{Hatton2007}.

Although reproducibility is the cornerstone of the scientific method, until
recently it has not been treated seriously in software
\citep{BenureauAndRougier2017}.  Fortunately, in recent years multiple
conferences, workshops and individuals are calling for dramatic change
\citep{BaileyEtAl2016}.  The need for action is highlighted by a study of 402
computer systems papers - only 48.3\% of the code was both available and
compilable \citep{CollbergEtAl2015}.  (Drasil addresses this problem because as
programming languages evolve the code renderers in Drasil can be updated.)
Reproducibility problems are even more extreme when the goal is replicability.
A third party should be able to repeat a study using only the description of the
methodology from a published article %, with no access to the original code or
%computing environment 
\citep{BenureauAndRougier2017}.  However, replicability is rarely achieved, as
shown for microarray gene expression \citep{IoannidisEtAl2009} and for economics
modelling \citep{IonescuAndJansson2013}.  Drasil addresses completeness and
ambiguity problems, since it emphasizes capturing and documenting all of the
required knowledge, including derivation of equations and
rationales. \citet{CrickAndHall2014} point out potential roadblocks for
reproducibility, including page length constraints and differing detail needs
depending on the audience.  Again Drasil addresses these concerns because the
recipes used to generate the documentation can be tailored to the level of
detail required.

\emph{Although scientists recognize the seriousness of their SCS problems, their
  corrective steps are too incremental.}  For instance, a recent proposal for
the future of High Energy Physics (HEP) software \cite{StewartEtAl2017} uses
words like sustainability, maintainability and reproducibility, but is almost
completely silent on how these qualities are to be achieved.  The proposal
mentions developing new and improved algorithms (including parallel computing
and machine learning), programming tools, recruitment and training, but there is
little on documentation, design or verification techniques (other than unit
testing).  Similarly, a recent proposal on future directions for SCS research
and education \citep{RudeEtAl2018} recognizes the desperate need for change, but
then only suggests training on project management tools, open sharing and
ethics.  \emph{Incremental change is not adequate; we need transformative
  change.}

Thankfully SCS leaders recognize that an interdisciplinary approach provides the
path forward.  They believe that the solution to SCS quality problems is
applying, adapting and developing SE methods, tools and techniques.  However,
typical software processes are a barrier to progress.  ``To break the gridlock,
we must establish a degree of cooperation and collaboration with the [SE]
community that does not yet exist'' \citep{FaulkEtAl2009}.  ``There is a need to
improve the transfer of existing practices and tools from ... [SE] to scientific
programming. In addition, ... there is a need for research to specifically
develop methods and tools that are tailored to the domain'' \citep{Storer2017}.
\emph{This tailored research will require individuals, like myself, that have a
  multidisciplinary background.}

\subsection{Documentation for SCS}

Table~\ref{TblDocuments} shows the recommended documentation for a scientific
software project.  The documents are typical of what is suggested for scientific
software certification, where certification consists of official recognition by
an authority, or regulatory body, that the software is fit for its intended use.
For instance, the Canadian Standards Association (CSA) requires a similar set of
documents for quality assurance of scientific programs for nuclear power
plants~\cite{CSA1999}.

\begin{table}[!h]
  \caption{Recommended Documentation} \label{TblDocuments}

  \begin{tabular}{l p{8cm}}
    %\hline
    %\toprule
    \textit{Problem Statement} & Description of problem to be solved\\
    \textit{Development Plan} & Overview of development process/infrastructure\\
    \textit{Requirements} & Desired functions and qualities of the software\\
    \textit{V\&V Plan} & Verification that all documentation artifacts,
                           including the code, are internally correct.
                           Validation, from an external viewpoint, that the right
                           problem, or model, is being solved.\\
    \textit{Design Specification} & Documentation of how the requirements are to
                                    be realized, through both a software
                                    architecture and detailed design of modules
                                    and their interfaces\\
    \textit{Code} & Implementation of the design in code\\
    \textit{V\&V Report} & Summary of the V\&V efforts,
                             including testing\\
    \textit{User Manual} & Instructions on installation, usage; worked examples\\
    %\hline
    %\bottomrule
  \end{tabular}
\end{table}

To achieve the qualities listed in Table~\ref{Tbl_Qualities}, the documentation
in Table~\ref{TblDocuments} should have the following qualities:
\emph{complete}, \emph{correct}, \emph{consistent}, \emph{modifiable},
\emph{traceable}, \emph{unambiguous}, and \emph{verifiable}.  All of these
qualities are listed in the IEEE recommended practice for software
requirements~\cite{IEEE1998}.  The IEEE guidelines are for requirements, but
most qualities are relevant for all documentation artifacts.  Another relevant
quality, which is not on the IEEE list, is \emph{abstract}.  Requirements should
state what is to be achieved, but be silent on how it is to be achieved.
Abstraction is an important software development principle for dealing with
complexity~\cite[p.~40]{GhezziEtAl2003}.  Smith and Koothoor present further
details on the qualities of documentation for scientific
software~\cite{SmithAndKoothoor2016}.

\subsection{MDE and Code Generation}

DSLs and code/document generation provide a transformative technology for
documentation, design and verification \citep{JohansonAndHasselbring2018,
  Smith2018}.  DSLs allow scientists their preferred approach of focusing on
science not software \citep{Kelly2007}.  A generative approach removes the
maintenance nightmare of documentation duplicates and near duplicates
\citep{LucivEtAl2018}, since knowledge is only captured once and automatically
transformed as needed.  Code generation has previously been applied to improve
SCS \citep{WhaleyEtAl2001, Veldhuizen1998, Pueschel2001}.
%   For instance,
% ATLAS (Automatically Tuned Linear Algebra Software) \citep{WhaleyEtAl2001} 
%and
%Blitz++ \citep{Veldhuizen1998} 
% produces efficient and portable linear algebra
% software.  
% Spiral \citep{Pueschel2001} uses software/hardware generation for
% digital signal processing.  
\citet{Carette2008} shows how to generate a family
of efficient, type-safe Gaussian elimination algorithms.  FEniCS (Finite Element
and Computational Software) \citep{LoggEtAl2012} uses code generation when
solving differential equations.  Unlike previous work on SCS code generation,
Drasil will focus on generating all software artifacts (requirements, design
etc.), not just code.

\citet{SzymczakEtAl2016} removes excuses for avoiding documentation by providing
transformative SCS development technology.  Drasil provides an infrastructure
for knowledge capture and document/code generation.  Jacques Carette and I,
along with our students, have developed a prototype (available on GitHub),
implemented via Domain Specific Languages (DSLs) embedded in Haskell.  Drasil
already generates requirements documentation and code for several case studies,
including simulating the temperature of a solar water heating system, glass
breakage and two-dimensional game physics.  The full documentation
(requirements, design etc), code and test cases have been created manually for
each case study.  These manual case studies (many from \citet{SmithJegatheesanAndKelly2016})
provide the `gold standard' against which Drasil is tested.  All of these
artifacts are available publicly.  The generation techniques for Drasil began to
take shape during my work on generating geometric data structures and functions
\citep{CaretteEtAl2011}.

\subsection{Empirical Methods for Software Engineering}

empirical software engineering paper, case study paper

\section{Research Design and Methodologies}

roadmap - fit the different pieces together

\subsection{State of the Practice}

A survey of existing SCS software will identify patterns, find opportunities for
improvement and identify candidate program families for future implementation.
The survey will require developing terminology and a classification system to
answer such questions as: What are the current best practises?  What software
packages are reused and why these packages?  What do practitioners look for in
the tools that support their work?

To achieve objective S2, a measurement template, possibly using pairwise
comparisons of quality metrics between software packages, will be developed for
assessing the quality of the documentation, code, test cases and development
processes for existing SCS program families.  The quality measurement template
will be used to assess the quality of approximately 30 different SCS families.
About half of the families will be model oriented and the other half tool
oriented.  Additional details on best practices will be collected by students
interviewing code owners.  Based on preliminary discussions, Dr.\ Jeffrey Carver
from the University of Alabama and some of his colleagues may also participate
in collecting interview data. Once all the surveys and interviews are
completed, a meta-analysis will draw conclusions for each domain and between
domains.  Knowing the state of practice for family development will highlight
which domains are using best practices that should be emulated in LSS.

The plan here is to return to the task of measuring/assessing the state of
software development practice in several scientific computing domains.  We will
update the work that was done previously for domains such as Geographic
Information Systems~\cite{SmithEtAl2018_arXivGIS}, Mesh
Generators~\cite{SmithEtAl2016}, Seismology software~\cite{SmithEtAl2018}, and
Statistical software for psychology~\cite{SmithEtAl2018_StatSoft}.  We could
return to these domains and/or introduce new domains.  Potential domains/collaborators
are listed in the subsections.

MEng students will be asked to assess the state of practice in each domain.
Last time each student measured two domains, so we would start off with this
model again.  If the scope becomes large enough, we might switch to one domain
per student.

In the previous project, we measured 30 software projects for each domain.  With
the increased scrutiny required in this re-boot, we won't likely be able to
measure the details on this high a number of software projects.  We may still
start with 30 software examples in each domain, but then use some criteria to
create a shorter sub-list for detailed measurement.  The details of how to
proceed here would be determined and documented in the measurement protocol.

In the previous state of practice assessing exercise, a series of questions and
simple metrics was devised to measure the quality of the documentation and
adherence to best software development practices.  The new project will
critically assess the previous set of questions and revise as necessary.  In
addition, the following data will be collected/developed for each domain:

\begin{itemize}
\item Characterization of the functionality provided by the software in the
  domain.  What services does each member of the domain provide?  Ideally this
  information would be summarized in a commonality analysis document.  The
  document will summarize the domain software via its commonalities,
  variabilities and parameters of variation.
\item Usability testing of each software in the domain.  Specific metrics for
  assessing usability still need to be researched, but measures that involve the
  time to complete a task and the users overall impression would be considered.
\item Collect empirical software engineering related data, such as the number of
  files, number of lines of code, cyclomatic complexity, number of open issues,
  etc.  As a starting point for tool support, HubListener
  (\url{https://github.com/pjmc-oliveira/HubListener}) could be used.
\item Collect at least one empirical measure of the quality of the documentation
  - the number of lines of documentation.  [Has anyone previously looked at this
  metric?]
\end{itemize}

The above data will be combined to rank the software in the domain.  The
specifics are yet to be determined, but in the past the Analytic Hierarchy
Process proved helpful in this.

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we didn't involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

\begin{itemize}
\item Review the protocol for assessing/measuring quality.  This same protocol
  will be used in each domain.
\item Provide their expertise on potential publication.  Specifically, they will
  recommend a suitable journal and act as the corresponding author for any paper
  submissions. 
\item Provide an authoritative list of domain software, possibly augmented by
  existing on-line lists.
\item Provide some assistance from their own team of supervised students to
  facilitate software testing and domain characterization.  Assessing the
  functionality and measuring usability will require an individual with domain
  knowledge.  Multiple measurements will be necessary to have confidence.  The
  measurements will likely take a few hours over a few days of time from each
  student volunteer.
\end{itemize}

There is no budget for this project, but the student volunteers will be
considered co-authors in the resulting paper.  Having them as co-authors also
means that ethics approval should not be necessary.

GitHub will be used to coordinate the work of the large team of people that will
be involved in this project.  In addition to the project record left on GitHub,
the final data will be exported to Mendeley.

Determine criteria for what makes a domain a good fit.

Potential collaborators/domains are listed in the following subsections.

\subsubsection{Medical Imaging}

Medical imaging and analysis software (Dr.\ Michael Noseworthy)?

\subsubsection{Climate Modelling Software}

Climate modelling software (Dr.\ Zoe Li or Dr.\ Xander Wang)?

\subsubsection{Computational Medicine}

Extraction of 3D geometries from medical images (Dr.\ Zahra Motamed)?  Lattice
Boltzmann Solver?

\subsubsection{Finite Element Analysis}

Finite element solvers and/or mesh generators (Dr.\ Dieter Stolle)?  Potential
software includes FeNiCS, MOOSE, FreeFEM++, Dolphin, etc.  Install the packages
and use them to do something non-trivial.

\subsubsection{Psychology Software}

Statistics software for psychology (Dr.\ Karin Humphreys)?

\subsubsection{Chemical Engineering Software}

Li Xi (\url{https://www.eng.mcmaster.ca/chemeng/people/faculty/li-xi})?

\subsubsection{Geographic Information Systems}

Geographic Information Systems (Jason Brodeur and/or Patrick DeLuca)?

\subsubsection{Physical Metallurgy Software}

Solidification and casting software (Dr.\ Andr\'e Phillion)?

\subsection{Quantum Chemistry Software}

Quantum chemistry software (Dr.\ Paul Ayers)?

\subsection{Impact of MDE on SCS}

roadmap

\subsection{Fitting Drasil into the Scientific Software Development Process}

Scientists generally use an approximation of agile methods for their software
development.  In addition, documentation is generally not emphasized.  How
should Drasil be fit into the scientific software development process?  The
Drasil process should encourage an early intensity of thought and effort, but at
the same time, it shouldn't require waterfall development.  Drasil supports
frequent chance and incremental development, but where should developers start
and how should the work proceed?  How does one create an almost empty project in
Drasil and then borrow from existing knowledge and create new knowledge?  This
project could start with something like the task described in~\ref{SecPDF}.

Drasil process should move importance of requirements and domain analysis
earlier in the process.  The quality of the resulting software depends on the
quality of the requirements.  As is known from historical data, the greatest
return on investment is at the requirements state (Boehm papers).
Unfortunately, developers don't like requirements.  Quite possibly they don't
like the intensity of thought that is necessary to get the requirements right.
Inspection techniques, such as task based inspection (Kelly et al) hold promise
for getting the scientist to think deeply at an earlier stage in the development
process.

\subsection{etc.}

Experimentally Determine Appropriate Documentation Template: The best approach
for documenting requirements and design is an open question.  Drasil could be
used to attempt to address this question.  Different recipes for documentation
could be developed and compared.  Some comparison could be automated between the
documents, such as empirical measures of their quality
(Section~\ref{SecEmpiricalMeasureDoc}).  Other measures would require some kind
of usability experiments.

Empirical Measurement of Documentation: Considerable effort has been invested
into developing metrics, like cyclomatic complexity, to attempt to quantify code
quality, but no measures (as far as I know) exist for quantifying the quality of
documentation.  (Khedri does have some work with Bhahati (sp?) Sanga that
quantifies the connectivity of the documentation via the traceability graph.)
Likely the reason that effort has not been invested to quantifying the quality
of documentation is that is not represented formally.  With Drasil much of the
documentation is formal.  This should facilitate automatic measures of
documentation quality, or at least metrics that might be correlated with
quality.

Measuring the Impact of Drasil on Developing Scientific Software: This project
is challenging to design, but would be HUGE if we were successful.  We assume
that Drasil will improve the quality of scientific software and the efficiency
of its production, but data to back this up would go a long way to convincing
the scientific community.  Experiments where a given project is developed in
parallel by two independent teams come to mind, but this would be expensive, and
difficult to control.  Some additional thought, and resources, are necessary
here.  If we could do a project where the scientist, or a student under their
supervision, is a true partner in the development, we might get more convincing
anecdotal evidence.

To quantify the quality of the generated documentation, measures of information
compression will be used.  The measure will be based on the length of the
(family of) generated documentation, code and test cases, compared to the length
of the generator itself.  Our measure will be based on the Minimum Description
Length (MDL) Principle, which states that ``the more we are able to compress the
data, the more we have learned about the data'' \citep{Grunwald2004}.  Another
important measure will be whether the generated documentation achieves
reproducibility. The proposed measure is reproducibility depth
\citep{Soergel2014}.  A depth of 1 means that an independent researcher can
start from the generated code and calculate the same results.  Achieving deeper
reproducibility will mean the same results can be calculated for an independent
researcher starting with the design documentation.  Each higher measure gets
further from the source code, until the highest depth (replicability), which
starts from the requirements and leads to identical (within acceptable error)
results.

\subsection{Impact of Assurance Cases for Building Confidence}

Some SCS needs to be verifiably safe.  We will achieve this by building a safety
assurance case. An assurance case is an explicit structured argument pertaining
to specific properties, such as trustworthiness

Assurance cases have potential to be used in SC to improve our collective
confidence in the developed software~\cite{SmithEtAl2018_ICSEPoster}.  Further
work is necessary to demonstrate the value of assurance cases, but they
certainly hold promise.  Drasil could go a long way to supporting development of
assurance cases.  Some potential Drasil related support ideas include the
following:

\begin{itemize}
\item Automatic generation of visual depiction of the assurance case from the
  goals, claims, evidence etc.
\item Automated reporting of missing evidence, incomplete arguments, etc.
\item Coordination of requirements, design decisions, etc.  Assurance cases use
  the same information as recommended in ``document driven'' design, but the
  traceability and connection between the data is even more important.  A tool
  like Drasil could go a long way in facilitating the presentation and
  navigation of an assurance case.
\end{itemize}

Assurance cases in Drasil could potentially go beyond the scientific context.
For instance, they could also incorporate safety cases for
control systems.  This could potentially be of interest to McSCert.

\subsection{Impact of Computational Variability Testing}

This activity supports objective S3.  The applicant's students will apply LSS
and code generation to Computational Variability Testing (CVT), using the
example of a family of finite element programs.  As proposed by the applicant,
CVT uses code generation to build confidence in the generated code in an
analogous way to the use of grid refinement studies.  Grid refinement looks at
how the solution changes by varying the run-time parameter of grid density and
comparing the results to theoretical expectations.  CVT, on the other hand, can
generate code to ``refine'' build time parameters, such as order of
interpolation, or degree of implicitness.  These parameters can be
systematically varied and the results compared against the expected trend.
Varying computational variabilities is not a new idea, but it is only recently
that the option has become available of coupling this with code generation, so
that the validity of an entire program family can be assessed at the same time.
The experimentation with CVT will initially use the FEniCS
package~\citep{LoggEtAl2012} for elliptical partial differential equations to
compare elements with different orders of interpolation to verify that the
convergence to the solution agrees with the theoretical expectations.

{Design and Documentation For Family of Data Fitting Algorithms}: The fitting
used in GlassBR and in SFS (Software for Solidification) could be made much more
generic.  We could have a family of fitting algorithms that could be used in any
situation where fitting is required.  A proper commonality analysis of this
domain could potentially show the potential design decisions that bridge between
the requirements and the design.  In the SFS example many different fitting
routines were tried.  If the experiments could have been done easily via a
declarative specification, considerable time would have been saved.  If the
experiments are combined with automated testing and ``properties of a correct
solution'' the human involvement could be reduced, so that we have partially
automated algorithm selection.

\section{Implications and Contributions to Knowledge}

We are currently putting too much trust in SCS.  Examples such as nuclear safety
analysis and computational medicine show the significance of SCS for health and
safety.  Although SCS developers do excellent work, we do not currently have
enough checks and balances in place for full confidence.  More should be
expected for documentation, design and testing.  Fortunately, SE offers
techniques, like DSLs and code generation, that can be employed to not only
improve SCS quality, but also reduce development costs.  Now is the time for
change.  SCS practitioners have recognized problems with the status quo and
multidisciplinary researchers are needed to bridge the gap between SE and SCS.

Codifying medical, biomechanical and computational knowledge is challenging, but
success will completely transform the development of safety-related software in
medicine, science and engineering.  MDE will remove the drudgery of generating
and maintaining documentation, code, test cases and build environments, enabling
scientists to focus on science, engineers to focus on engineering and medical
professionals to focus on medicine.  MDE will detect inconsistencies between
models via inter-model consistency constraints.  As a consequence, mistakes like
inconsistent units or assumptions cannot occur.  MDE will raise the bar so that
we can expect an explicit argument for safety, not just code that we are
expected to blindly trust.  With the right up-front investment of knowledge
capture, we can have software that is long lived because the stable knowledge is
separated from the rapidly changing modelling assumptions and design decisions.

\newpage

\section{Research Schedule}

\bibliographystyle {plainnat}
\bibliography {ResearchProposal}

\end{document}