### 2019-09-23 ResearchProposal.tex
Updates to research proposal

\documentclass[12pt]{article}

\usepackage[round]{natbib}
\usepackage{booktabs}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,       % color of links to bibliography
    filecolor=magenta,   % color of file links
    urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\begin{document}

\title{Assessing the Impact of MDE and Code Generation on the Sustainability of
  Scientific Computing Software: A Research Proposal} 
\author{Spencer Smith}
\date{\today}

\maketitle

\begin{abstract}
  ...
\end{abstract}

\tableofcontents

\newpage

\section{Introduction} \label{SecIntroduction}

Without dramatic intervention, our collective confidence in Scientific Computing
Software (SCS) is due for a catastrophic collapse.  We are increasingly trusting
ever more complex and ambitious computations, but our software foundations are
built on sand. Areas of concern include nuclear safety analysis and
computational medicine. Successfully building such software requires
communication between software developers and experts from multiple
domains. Collaboration is difficult at the best of times, and is made worse
because developers favour handcrafted solutions over adapting software
engineering processes, methods and tools \citep{FaulkEtAl2009}. Handcrafted
solutions do not account for the inevitable changes in requirements, design and
implementation. The twin challenges of changing requirements and inadequate
documentation conspire to make computational results notoriously difficult to
reproduce, especially in the important endeavour of one researcher independently
replicating the results of another \citep{Smith2018}.  Inadequate traceability
within documentation means re­certification of previously certified engineering
software is as expensive and time consuming as the original certification
exercise, because of a lack of explicit, effort reducing, traceability
information.  The existing challenges for modification, maintenance and
extension point to problems with the sustainability of SCS.

Documentation can improve software sustainability.  For software in general,
documentation, written before and during development, provides many
benefits~\cite{Parnas2010}: easier reuse of old designs, better communication
about requirements, more useful design reviews, easier integration of separately
written modules, more effective code inspection, more effective testing, and
more efficient corrections and improvements. For SCS in particular,
documentation provides significant benefits.  For instance,
\citet{SmithEtAl2015-SS-TR} shows that the quality of statistical software for
psychology is generally improved when developed using the structured CRAN
(Comprehensive R Archive Network) process and tools, versus an ad hoc approach.
\citet{SmithAndKoothoor2016} highlights the value of proper documentation by
redeveloping nuclear safety analysis software.  Twenty seven (27) worrisome
documentation problems were found, including problems with incompleteness,
ambiguity, inconsistency, verifiability, modifiability, traceability and a lack
of abstraction.  A redevelopment experiment with five existing projects
\citep{SmithJegatheesanAndKelly2016} enabled the code owners (as ascertained
through interviews) to clearly see the value of documentation.  However,
mirroring other studies \citep{CarverEtAl2007}, the code owners felt
documentation takes too much time.  Due to the time commitment and resource
limitation, documentation is rarely emphasized in SCS.

One promising approach to gain the sustainability benefits of documentation, but
significantly reduce the time commitment, is to use a Model Driven Engineering
(MDE) process.  MDE is an engineering approach that exploits structured
representations (metamodels for abstract syntax, and constraints to capture
static semantics) to automate repetitive, tedious and error prone tasks.  In
this research proposal, we will use MDE as a short-hand to include knowledge
models, model transformation, Domain Specific Languages (DSLs) and code
generation.  Several current papers point to DSLs and code/document generation
as a transformative technology for documentation, design and verification of SCS
\citep{JohansonAndHasselbring2018, Smith2018}.  MDE has the potential to
dramatically reduce the cost of developing reliable, reproducible and
re-certifiable software.

One example of applying MDE to SCS is Drasil, which is introduced in
\citet{SzymczakEtAl2016} and available at
\href{https://github.com/JacquesCarette/Drasil}
{https://github.com/JacquesCarette/Drasil}.  Drasil is implemented as an open
source set of Domain Specific Languages (DSLs).  The Drasil approach involves:
i) creating infrastructure for a knowledge base of scientific and computing
models; and, ii) writing explicit ``recipes'' that weave together this knowledge
to generate theoretical models, design documents, code, test cases and build
scripts. The generator can render to multiple languages, such as html, and LaTeX
for documentation and Python, Java, C++ and Lua, for code. One source of
knowledge, with rules for transformation, means completeness, consistency and
traceability can be achieved by construction. Moreover, these qualities can be
maintained as requirements are modified, design decisions are changed,
documentation standards are varied, and software is re­certified.  Although
creating the knowledge base is time consuming, the knowledge can be built up
incrementally.  New projects will reuse existing knowledge and expand as
necessary.

As described above, sustainability is a concern for SCS; MDE holds promise
for addressing this concern.  However, a transformative change in the
development of SCS requires more than promise.  Empirical evidence is necessary
to demonstrate that an MDE process can improve the sustainability of SCS.  This
leads to the long-term objective for this research proposal: \emph{assess and
  measure the impact of MDE on the sustainability of SCS.}

The scope of SCS software for this proposal includes long-lived software, as
implied by the sustainability objective.  To gain the full benefits of
generating documentation, the scope also includes software that interests
multiple stakeholders, with different interests and backgrounds.  Finally, the
scope of this proposal emphasizes safety related software, such as software for
nuclear safety analysis, medical imaging and computational medicine.  Safety
related software is prioritized because documentation is often expected as part
of a certification exercise.  Although this study is not specifically restricted
to open-source software, for practical reasons (especially the lack of access to
the code), commercial software will not be strongly emphasized.

Given the large number of potential MDE processes, techniques and technologies,
together with the wide variety in purpose, scope and context for SCS, the
long-term objective of measuring the impact of MDE is decomposed into four
short-term objectives.  For each objective, a reference is given to the
corresponding section where details are provided.

\begin{enumerate}
\item Assess the current state of the software development practice for SCS.
  How is the software developed in different sub-domains of SCS?  Are there
  existing development methods that lead to higher software sustainability?
  (Section~\ref{SecStateOfThePractice})
\item Assess the impact of an MDE process on end user developed SCS with respect
  to developer productivity and software sustainability for long-lived lived software.  
  (Section~\ref{SecImpactMDE})
\item Assess the impact on safety when the MDE generated documentation targets a
  safety assurance case.  An assurance case is an explicit structured argument
  pertaining to specific properties, such as trustworthiness.  The evidence for
  the argument will come from the traceability between artifacts generated by
  model transformations, expert reviews, test cases, etc.
  (Section~\ref{SecAssuranceCase})
\item Assess the impact on software quality when using MDE to facilitate
  Computational Variability Testing (CVT).  CVT uses code generation to build
  confidence in the generated code in an analogous way to the use of grid
  refinement studies.  Grid refinement looks at how the solution changes by
  varying the run-time parameter of grid density and comparing the results to
  theoretical expectations.  CVT, on the other hand, can generate code to
  ``refine'' build time parameters, such as order of interpolation, or degree of
  implicitness.  These parameters can be systematically varied and the results
  compared against the expected trend.  (Section~\ref{SecCVT})
\end{enumerate}

Before presenting the details of the short-term objectives
(Section~\ref{SecMethodologies}), the necessary background information is
provided, along with a literature review (Section~\ref{SecBackgroundLitReview}).
The literature review covers software quality (Sections~\ref{SecSoftwareQuality}
and~\ref{SecDesirableQs}), the literature on the current state of the practice
for SCS development (Section~\ref{SecCurrStateOfThePractice}), potential target
documentation artifacts for SCS (Section~\ref{SecDocumentation}), MDE and code
generation (Section~\ref{SecMDE}) and empirical methods for software engineering
(Section~\ref{SecEmpirical}).  This document ends with a summary of the
potential contributions to knowledge from this study
(Section~\ref{SecContribKnowledge}) and a research schedule
(Section~\ref{SecSchedule}).

\section{Background and Literature Review} \label{SecBackgroundLitReview}

To assess the impact of MDE on SCS quality, we need a clear definition of what
we mean by quality.  Section~\ref{SecSoftwareQuality} shows how the concept of
quality is decomposed into a set of separate qualities.  This set of qualities
can be applied to the software artifacts (documentation, test cases, etc) and to
the software development process itself.  Several of the qualities from the list
in Section~\ref{SecSoftwareQuality} cannot be measured directly, such as
maintainability.  Therefore, Section~\ref{SecDesirableQs} introduces measurable
documentation qualities that are believed to contribute to the overall software
qualities.  In some cases it may be necessary to use the qualities in
Section~\ref{SecDesirableQs} to indirectly measure qualities listed in
Section~\ref{SecSoftwareQuality}.

The other subsection in this background section provide an overview of the
literature on the current approaches to developing SCS software
\ref{SecCurrStateOfThePractice}, recommended document artifacts for SCS
\ref{SecDocumentation}, an overview of MDE and Code Generation in an SCS context
\ref{SecMDE}, and a summary of existing literature on empirical methods for
software engineering \ref{SecEmpirical}.

\subsection{Software Qualities of Interest} \label{SecSoftwareQuality}

Our analysis is centred around a set of software qualities.  Quality is not
considered as a single measure, but a collection of different qualities, often
called ``ilities.''  These qualities highlight the desirable nonfunctional
properties for software artifacts, which include both documentation and
code. Some qualities, such as visibility and productivity, apply to the process
used for developing the software. The following list of qualities is based on
\cite{GhezziEtAl2003}. To the list from \cite{GhezziEtAl2003}, we have added
three qualities important for SC: installability, reproducibility and
sustainability.

\begin{description}%[leftmargin=\the\parindent]

\item [\textbf{Installability}] A measure of the ease of installation.

\item [\textbf{Correctness}] Software is correct if it matches its specification.

\item [\textbf{Verifiability}] involves ``solving the equations
  right''~\cite[p.~23]{Roache1998}; it benefits from rational documentation
  that systematically shows, with explicit traceability, how the governing
  equations are transformed into code.

\item [\textbf{Validatability}] means ``solving the right
  equations''~\cite[p.~23]{Roache1998}.  Validatability is improved by a
  rational process via clear documentation of the theory and assumptions, along
  with an explicit statement of the systematic steps required for experimental
  validation.

\item [\textbf{Reliability}] is a critical quality for scientific software,
  since the results of computations are meaningless, if they are not dependable.
  Reliability is closely tied to verifiability, since the key quality to verify
  is reliability, while the act of verification itself improves reliability.

\item [\textbf{Performance}] considerations can make certification challenging,
  since QA becomes more difficult for more complex code.  However, as
  Roache~\cite[p.~355]{Roache1998} points out, using simpler algorithms and
  reducing the number of options in general purpose code, is not always a
  practical option.

\item [\textbf{Usability}] can be a problem.  Different users, solving the same
  physical problem, using the same software, can come up with different answers,
  due to differences in parameter selection~\cite[p.~370]{Roache1998}.  To
  reduce misuse, a rational process must state expected user characteristics,
  modelling assumptions, definitions and the range of applicability of the code.

\item [\textbf{Maintainability}] is necessary in scientific software, since change,
  through iteration, experimentation and exploration, is inevitable.  Models of
  physical phenomena and numerical techniques necessarily evolve over
  time~\cite{CarverEtAl2007, SegalAndMorris2008}.  Proper documentation,
  designed with change in mind, can greatly assist with change management.%   QA
  % activities need to take the need for creativity into account, while not
  % smothering it~\cite[p.~352]{Roache1998}.

\item [\textbf{Reusability}] provides support for the quality of reliability,
  since reliability is improved by reusing trusted components~\cite{Dubois2005}.
  (Care must still be taken with reusing trusted components, since blind reuse
  in a new context can lead to errors, as dramatically shown in the Ariane 5
  disaster~\cite[p.~37--38]{OliveiraAndStewart2006}.)  The odds of reuse are
  improved when it is considered right from the start.

\item [\textbf{Understandability}] is necessary, since reviewers can only certify
  something they understand.  Scientific software developers have the
  view ``that the science a developer embeds in the code must be apparent to
  another scientist, even ten years later''~\cite{Kelly2013}.
  Understandability applies to the documentation and code, while usability
  refers to the executable software.  Documentation that follows a rational
  process is the easiest to follow.

\item [\textbf{Reproducibility}] is a required component of the scientific
  method~\cite{Davison2012}.  Although QA has, ``a bad name among
  creative scientists and engineers''~\cite[p.~352]{Roache1998}, the community
  need to recognize that participating in QA management also improves
  reproducibility.  Reproducibility, like QA, benefits from a consistent and
  repeatable computing environment, version control and separating code from
  configuration/parameters~\cite{Davison2012}.

\item [\textbf{Productivity}] [Fill in]

\item [\textbf{Sustainability}] A combination of other qualities, likely
  maintainability and productivity.  Reference authorless paper found on-line.

\end{description}

\subsection{Desirable Qualities of Documentation} \label{SecDesirableQs}

To achieve the qualities listed in Section~\ref{SecSoftwareQuality}, the
documentation should achieve the qualities listed in this section.  All but the
final quality listed (abstraction), are adapted from the IEEE recommended
practise for producing good software requirements~\cite{IEEE1998}.  Abstraction
means only revealing relevant details, which in a requirements document means
stating what is to be achieved, but remaining silent on how it is to be
achieved.  Abstraction is an important software development principle for
dealing with complexity~\cite[p.~40]{GhezziEtAl2003}.
\citet{SmithAndKoothoor2016} present further details on the qualities of
documentation for SCS.

\begin{description}

\item [\textbf{Completeness}] Documentation is said to be complete when all the
  requirements of the software are detailed. That is, each goal, functionality,
  attribute, design constraint, value, data, model, symbol, term (with its unit
  of measurement if applicable), abbreviation, acronym, assumption and
  performance requirement of the software is defined.  The software's response
  to all classes of inputs, both valid and invalid and for both desired and
  undesired events, also needs to be specified.

\item [\textbf{Consistency}] Documentation is said to be consistent when no subset
  of individual statements are in conflict with each other. That is, a
  specification of an item made at one place in the document should not
  contradict the specification of the same item at another location.

\item [\textbf{Modifiability}] The documentation should be developed in such a way
  that it is easily modifiable so that likely future changes do not destroy the
  structure of the document. Also it should be easy to reflect the change,
  wherever needed in the document to maintain consistency, traceability and
  completeness. For documentation to be modifiable, its format must be
  structured in a way that repetition is avoided and cross-referencing is
  employed.

\item [\textbf{Traceability}] Documentation should be traceable, as this
  facilitates maintenance and review. If a change is made to the design or code
  of the software, then all the documentation relating to those segments have to
  be modified.  This property is also important for recertification.

\item [\textbf{Unambiguity}] Documentation is said to be unambiguous only when
  every requirement's specification has a unique interpretation.  The
  documentation should be unambiguous to all audiences, including developers,
  users and reviewers.

\item [\textbf{Correctness}] There is no direct tool or method for measuring
  correctness. One way of building confidence in correctness is by reviewing to
  ensure that each requirement stated is one that the stakeholders and experts
  desire.  By maintaining traceability, consistency and unambiguity, we can
  reduce the occurrence of errors and make the goal of reviewing for correctness
  easier.

\item [\textbf{Verifiability}] Every requirement in the documentation must be the
  one fulfilled by the implemented software. Therefore all the requirements
  should be clear, unambiguous and testable, so that a person or a machine can
  verify whether the software product meets the requirements.

\item [\textbf{Abstract}] Documented requirements are said to be abstract if they
  state what the software must do and the properties it must possess, but do not
  speak about how these are to be achieved. For example, a requirement can
  specify that an Ordinary Differential Equation (ODE) must be solved, but it
  should not mention that Euler's method should be used to solve the ODE. How to
  accomplish the requirement is a design decision, which is documented during
  the design phase.

\end{description}

\subsection{Literature Review on Current State of the Practice for SCS
  Development} \label{SecCurrStateOfThePractice}

The SCS community is finally realizing that current practices are not
sustainable.  \citet{FaulkEtAl2009} observe, ``growing concern about the
reliability of scientific results based on ... software.''  Embarrassing
failures have occurred, like a retraction of derived molecular protein
structures \citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  A
recent report on directions for SCS research and education states: ``While the
volume and complexity of [SCS] have grown substantially in recent decades, [SCS]
traditionally has not received the focused attention it so desperately needs
... to fulfill this key role as a cornerstone of long-term collaboration and
scientific progress'' \citep{RudeEtAl2018}.  Estimates suggest that the number
of released faults per thousand executable lines of code during a given
program’s life cycle is at best 0.1, and more likely 10 to 100 times worse
\citep{Hatton2007}.

Although reproducibility is the cornerstone of the scientific method, until
recently it has not been treated seriously in software
\citep{BenureauAndRougier2017}.  Fortunately, in recent years multiple
conferences, workshops and individuals are calling for dramatic change
\citep{BaileyEtAl2016}.  The need for action is highlighted by a study of 402
computer systems papers - only 48.3\% of the code was both available and
compilable \citep{CollbergEtAl2015}.  (Drasil addresses this problem because as
programming languages evolve the code renderers in Drasil can be updated.)
Reproducibility problems are even more extreme when the goal is replicability.
A third party should be able to repeat a study using only the description of the
methodology from a published article, with no access to the original code or
computing environment \citep{BenureauAndRougier2017}.  However, replicability is
rarely achieved, as shown for microarray gene expression
\citep{IoannidisEtAl2009} and for economics modelling
\citep{IonescuAndJansson2013}.  Drasil addresses completeness and ambiguity
problems, since it emphasizes capturing and documenting all of the required
knowledge, including derivation of equations and
rationales. \citet{CrickAndHall2014} point out potential roadblocks for
reproducibility, including page length constraints and differing detail needs
depending on the audience.  Again Drasil addresses these concerns because the
recipes used to generate the documentation can be tailored to the level of
detail required.

\emph{Although scientists recognize the seriousness of their SCS problems, their
  corrective steps are too incremental.}  For instance, a recent proposal for
the future of High Energy Physics (HEP) software \cite{StewartEtAl2017} uses
words like sustainability, maintainability and reproducibility, but is almost
completely silent on how these qualities are to be achieved.  The proposal
mentions developing new and improved algorithms (including parallel computing
and machine learning), programming tools, recruitment and training, but there is
little on documentation, design or verification techniques (other than unit
testing).  Similarly, a recent proposal on future directions for SCS research
and education \citep{RudeEtAl2018} recognizes the desperate need for change, but
then only suggests training on project management tools, open sharing and
ethics.  \emph{Incremental change is not adequate; we need transformative
  change.}

Thankfully SCS leaders recognize that an interdisciplinary approach provides the
path forward.  They believe that the solution to SCS quality problems is
applying, adapting and developing SE methods, tools and techniques.  However,
typical software processes are a barrier to progress.  ``To break the gridlock,
we must establish a degree of cooperation and collaboration with the [SE]
community that does not yet exist'' \citep{FaulkEtAl2009}.  ``There is a need to
improve the transfer of existing practices and tools from ... [SE] to scientific
programming. In addition, ... there is a need for research to specifically
develop methods and tools that are tailored to the domain'' \citep{Storer2017}.
This tailored research will require teams with a multidisciplinary background.

\subsection{Documentation for SCS} \label{SecDocumentation}

Table~\ref{TblDocuments} shows the recommended documentation for a scientific
software project.  The documents are typical of what is suggested for scientific
software certification, where certification consists of official recognition by
an authority, or regulatory body, that the software is fit for its intended use.
For instance, the Canadian Standards Association (CSA) requires a similar set of
documents for quality assurance of scientific programs for nuclear power
plants~\cite{CSA1999}.

\begin{table}[!h]
  \caption{Recommended Documentation} \label{TblDocuments}

  \begin{tabular}{l p{8cm}}
    %\hline
    %\toprule
    \textit{Problem Statement} & Description of problem to be solved\\
    \textit{Development Plan} & Overview of development process/infrastructure\\
    \textit{Requirements} & Desired functions and qualities of the software\\
    \textit{V\&V Plan} & Verification that all documentation artifacts,
                           including the code, are internally correct.
                           Validation, from an external viewpoint, that the right
                           problem, or model, is being solved.\\
    \textit{Design Specification} & Documentation of how the requirements are to
                                    be realized, through both a software
                                    architecture and detailed design of modules
                                    and their interfaces\\
    \textit{Code} & Implementation of the design in code\\
    \textit{V\&V Report} & Summary of the V\&V efforts,
                             including testing\\
    \textit{User Manual} & Instructions on installation, usage; worked examples\\
    %\hline
    %\bottomrule
  \end{tabular}
\end{table}

\wss{Add figure showing the V-model of the documentation.}

\subsection{MDE and Code Generation} \label{SecMDE}

DSLs and code/document generation provide a transformative technology for
documentation, design and verification \citep{JohansonAndHasselbring2018,
  Smith2018}.  Scientists prefer focusing on science not software
\citep{Kelly2007}.  DSLs allow them to do this.  A generative approach removes
the maintenance nightmare of documentation duplicates and near duplicates
\citep{LucivEtAl2018}, since knowledge is only captured once and automatically
transformed as needed.  Code generation has previously been applied to improve
SCS \citep{WhaleyEtAl2001, Veldhuizen1998, Pueschel2001}.  For instance, ATLAS
(Automatically Tuned Linear Algebra Software) \citep{WhaleyEtAl2001} and Blitz++
\citep{Veldhuizen1998} produces efficient and portable linear algebra software.
Spiral \citep{Pueschel2001} uses software/hardware generation for digital signal
processing.  \citet{Carette2008} shows how to generate a family of efficient,
type-safe Gaussian elimination algorithms.  FEniCS (Finite Element and
Computational Software) \citep{LoggEtAl2012} uses code generation when solving
differential equations.  Unlike previous work on SCS code generation, Drasil
(\href{https://github.com/JacquesCarette/Drasil}
{https://github.com/JacquesCarette/Drasil}) focus on generating all software
artifacts (requirements, design etc.), not just code.

Drasil, as presented in \citet{SzymczakEtAl2016} removes excuses for avoiding
documentation by providing transformative SCS development technology.  Drasil
provides an infrastructure for knowledge capture and document/code generation.
Drasil is implemented via Domain Specific Languages (DSLs) embedded in Haskell.
Drasil generates requirements documentation and code for several case studies,
including simulating the temperature of a solar water heating system, glass
breakage and two-dimensional game physics.  The full documentation
(requirements, design etc), code and test cases have been created manually for
each case study.  These manual case studies (many from
\citet{SmithJegatheesanAndKelly2016}) provide the `gold standard' against which
Drasil is tested.  All of these artifacts are available publicly at
\href{https://jacquescarette.github.io/Drasil/}
{https://jacquescarette.github.io/Drasil/}.  The generation techniques for
Drasil began to take shape during work on generating geometric data structures
and functions \citep{CaretteEtAl2011}.

\subsection{Empirical Methods for Software Engineering} \label{SecEmpirical}

The empirical methods in this study will follow the guidelines given in
\citet{KitchenhamEtAl2002}.  Since case studies will be part of the methodology,
the guidelines for conducting and reporting case study research in software
engineering \citep{RunesonAndHost2009} will be followed.

\section{Research Design and Methodologies} \label{SecMethodologies}

This section is decomposed into 4 subsections -- one for each of the different
approaches for assessing the impact of MDE on the sustainability of SCS, as
outlined in the Introduction (Section~\ref{SecIntroduction}).  The goal of the
first subsection is to establish a ``baseline'' on the current state of the
practice for SCS.  This is done by identifying domains within SCS and then
systematically assessing the quality of the work in those domains.  The second
subsection looks specifically at the impact of MDE on SCS through a series of
case studies.  An ``ideal'' MDE process is proposed and then followed with
different stakeholders and qualitative data is collected.  The last two
sub-sections discuss assessing the impact on SCS of assurance cases and
computational variability testing, respectively.

\subsection{State of the Practice} \label{SecStateOfThePractice}

To understand the current quality of SCS and to assist in developing methods for
measuring qualities, we will identify several domains within SCS and endeavour
to summarize the current state of the practice within those domains.  This
``state of the practice'' exercise will build off of prior work on
measuring/assessing the state of software development practice in several SCS
domains.  We will update the work that was done previously for domains such as
Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators
\citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  We may
return to these domains and/or introduce new domains.  Potential
domains/collaborators are listed in the subsections below.

MEng students will be asked to assess the state of practice in each domain.
During the last state of the practice exercise each student measured two
domains, but with the higher demands from the new measurement protocol, we will
likely switch to one domain per student.

In the previous project, we measured 30 software projects for each domain.  With
the increased scrutiny required in this re-boot, we will not likely be able to
measure the details on this high a number of software projects.  We may still
start with 30 software examples in each domain, but then use some criteria to
create a shorter sub-list for detailed measurement.  The details of how to
proceed here would be determined and documented in the measurement protocol.

In the previous state of the practice assessing exercise
\citep{SmithEtAl2018_arXivGIS}, a series of questions and simple metrics was
devised to measure the quality of the documentation and adherence to best
software development practices.  The previous set of questions is provided
\href{https://github.com/adamlazz/DomainX/blob/master/TemplateToGradeSCSoft.pdf}
{here}. The new project will critically assess the previous set of questions and
revise as necessary.  In addition, the following data will also be considered
for collection for each domain:

\begin{itemize}
\item Characterization of the functionality provided by the software in the
  domain.  What services does each member of the domain provide?  Ideally this
  information would be summarized in a commonality analysis document.  The
  document will summarize the domain software via its commonalities,
  variabilities and parameters of variation.
\item Usability testing of each software in the domain.  Specific metrics for
  assessing usability still need to be researched, but measures that involve the
  time to complete a task and the users overall impression will be considered.
\item Collect empirical software engineering related data, such as the number of
  files, number of lines of code, cyclomatic complexity, number of open issues,
  etc.  As a starting point for tool support, HubListener
  (\url{https://github.com/pjmc-oliveira/HubListener}) could be used.
\item Collect at least one empirical measure of the quality of the documentation
  - the number of lines of documentation.  We should investigate to see if this
  is a metric that has previously been collected.  So many of the existing
  metrics are completely focused on the code, or possibly on issue creation and
  management; it seems that little work has been done to assess the
  documentation quality.
\item One of the guidelines for successfully on-boarding new developers is to
  ``Keep knowledge up to date and findable'' \citep{ShollerEtAl2019}.  The
  measurements selected should be able to assist with judging whether a project
  succeeds on this front.
\end{itemize}

The above data will be combined to rank the software in the domain.  The
specifics are yet to be determined, but in the past the Analytic Hierarchy
Process (AHP) \citep{Saaty1980} proved helpful in this.  In the past the AHP
results were calculated by a grade out of 10 that each student evaluator gave to
each quality for each measured product.  A better approach may be to have a
group of people familiar with the software do a pairwise comparison between each
product to come up with a number.  The number would be based on the data
collected, but not in a numerical way.  The data collected would sever the same
role as a decision support system.  It would provided insight, but human beings
would do the final pairwise comparisons.

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we did not involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

\begin{itemize}
\item Review the protocol for assessing/measuring quality.  This same protocol
  will be used in each domain.
\item Provide their expertise on potential publication.  Specifically, they will
  recommend a suitable journal and act as the corresponding author for any paper
  submissions. 
\item Provide an authoritative list of domain software, possibly augmented by
  existing on-line lists.
\item Provide some assistance from their own team of supervised students to
  facilitate software testing and domain characterization.  Assessing the
  functionality and measuring usability will require an individual with domain
  knowledge.  Multiple measurements will be necessary to have confidence.  The
  measurements will likely take a few hours over a few days of time from each
  student volunteer.
\end{itemize}

There is no budget for this project, but the student volunteers will be
considered co-authors in the resulting paper.  Having them as co-authors also
means that ethics approval should not be necessary.

GitHub will be used to coordinate the work of the large team of people that will
be involved in this project.  In addition to the project record left on GitHub,
the final data will be exported to Mendeley (as was done for the previous
quality measurement exercise, for instance the grades for GIS software are
summarized at \href{https://data.mendeley.com/datasets/6kprpvv7r7/1}
{https://data.mendeley.com/datasets/6kprpvv7r7/1}.)

As a next step for this work, the procedure for measuring software qualities in
SCS will be brainstormed, discussed and written up.  The procedure will be
applied to each domain with SCS that is assessed.  The write-up on the procedure
will be published.  Ideally at a relevant conference, but if that is not a
feasible option, it will be hosted on the arxiv.

Potential domains for future measurement are listed below.  Determination of
these domains are good candidates will require evaluation of the following
criteria:

\begin{itemize}
\item Availability of a domain expert that is willing to assist, and who has at
  least a few graduate students that can assist with the measurement tasks.
\item A domain where end-user developed software is common
\item A domain where open-source software is available.  Commercial products are
  not as effective for measurement, since we cannot generally gain access to the
  source code, developer documentation, or issue tracking system.
\item A domain where scientific computing is used.  The selected software should
  fit in the larger domain of SCS.
\end{itemize}

The data for each of the different subdomains within SCS will be standardized to
facilitate a meta-analysis that combines all of the data.

\subsubsection{Medical Imaging}

Medical imaging and analysis software.  This work could build off of the
previous project on this topic completed by Vasudha Kapil.  (Vasudha's report is
located in the \texttt{mmsc} svn repo) .  % (Dr.\ Michael Noseworthy)?

\subsubsection{Climate Modelling Software}

Climate modelling software.  This could be software used for multi-model
ensemble aggregating \citep{SamoulyEtAl2018} or possibly with software that uses
physical models to predict climate change.
% (Dr.\ Zoe Li or Dr.\ Xander Wang)?

\subsubsection{Computational Medicine}

Potential sub-domains with computational medicine could include: extraction of
3D geometries from medical images, or the use of Lattice Boltzmann
Solvers.  % (Dr.\ Zahra Motamed)?

\subsubsection{Finite Element Analysis}

Finite element solvers and/or mesh generators.  Potential
software includes FeNiCS, MOOSE, FreeFEM++, Dolphin, etc.  Install the packages
and use them to do something non-trivial.  %(Dr.\ Dieter Stolle)?

\subsubsection{Psychology Software}

Statistics software for psychology.  A project in this domain would allow us to
update \citet{SmithEtAl2018_StatSoft}. %(Dr.\ Karin Humphreys?)

\subsubsection{Chemical Engineering Software}

Software is used for many purposes in chemical engineering.  We would want to
find an application where non-commercial software is available.
% Li Xi? (\url{https://www.eng.mcmaster.ca/chemeng/people/faculty/li-xi})?

\subsubsection{Geographic Information Systems (GIS)}

We studied GIS software in the past \citep{SmithEtAl2018_arXivGIS}, so this
would be an interesting area to update.  The key for the updated effort will be
to identify a domain expert available for collaboration. %(Jason Brodeur and/or Patrick DeLuca)?

\subsubsection{Physical Metallurgy Software}

Solidification and casting software and other software used for physical
metallurgy.  An example of software written for analyzing the solidification of
a molten alloy is given in \citet{SmithEtAl2019_arXiv}.  % (Dr.\ Andr\'e Phillion)?

\subsubsection{Quantum Chemistry Software}

Quantum chemistry software.  One example is software for searching for
conformers, like
\href{https://github.com/PeaWagon/Kaplan}{Kaplan}.  %(Dr.\ Paul Ayers)?

\subsection{Impact of MDE on the Sustainability of SCS} \label{SecImpactMDE}

Once we have details on the current state of the practice in SCS, we can begin
to apply MDE to SCS projects and assess the ramifications of this change.  This
section starts by laying out the details on how MDE can be applied to SCS
(Section~\ref{SecMDEProcess}).  We plan to start with an ideal process and then
approximate it through different experiments to measure and assess the impact of
MDE.  The second subsection (Section~\ref{SecQOfMDEArtifacts}) outlines how the
quality of MDE artifacts can be assessed.  These measurements will be used in
the various experiments to assess the impact of MDE.  The third and fourth
subsections describe two experiments for collecting data: i) conducting case
studies with real SCS projects (Section~\ref{SecCaseStudy}); and, ii) performing
an experiment where multiple developers create a small SCS project in one day
(Section~\ref{SecSmallPhysExample}).

\subsubsection{Fitting MDE into the Scientific Software Development
  Process} \label{SecMDEProcess}

Many researchers have stated that a document driven process is not used by, nor
suitable for, scientific software.  These researchers argue that scientific
developers naturally use an agile philosophy~\cite{AckroydEtAl2008,
  CarverEtAl2007, EasterbrookAndJohns2009, Segal2005}, or an amethododical
process~\cite{Kelly2013}, or a knowledge acquisition driven
process~\cite{Kelly2015}.  However, as discussed in the introduction to this
proposal (Section~\ref{SecIntroduction}, documentation provides many benefits.
An MDE approach holds promise to provide these benefits, while removing the
drudgery from the process of documentation and facilitating inevitable changes
in the documentation and the software.

The general idea that MDE will be helpful is appealing, but to study the impact
of MDE, we need to know how it will actually be used.  How exactly would MDE be
fit into the scientific software development process?  The new process should
encourage an early intensity of thought and effort, but at the same time, it
should not impose waterfall development, since waterfall is an unrealistic
expectation for a problem that is not yet well understood.  MDE supports
frequent change and incremental development, but where should developers start
and how should the work proceed?  How does one create an almost empty project in
MDE and then borrow from existing knowledge and create new knowledge?  To create
a specific MDE process for SCS, the Drasil tools will be utilized.  A good
starting project could be
\href{https://jacquescarette.github.io/Drasil/examples/Projectile/srs/Projectile_SRS.html}
{Projectile}, which models simple projectile motion.

Compared to existing SCS development techniques, the new Drasil process should
emphasize requirements and domain analysis earlier in the process.  The quality
of the resulting software depends on the quality of the requirements.  As
Boehm's data shows~\cite{Boehm1981}, early identification of errors implies a
significant return on investment, since the cost of fixing errors increases
dramatically for later development stages.  An SRS is invaluable for
verification, since the quality of software cannot be assessed without a
standard against which to judge it. Better design decisions are facilitated by
the information captured in an SRS.  Moreover, the SRS aids the software
lifecycle by facilitating incremental development.  This last advantage is
important for SCS software, since changes frequently occur as developers explore
the problem domain.  Unfortunately, developers do not generally write
requirements.  Quite possibly they do not like the intensity of thought that is
necessary to get the requirements right.  Inspection techniques, such as task
based inspection \citep{KellyAndShepard2000, KellyAndShepard2004} hold promise
for getting the scientist to think deeply at an earlier stage in the development
process.

Since Drasil is not the only tool that can be used for MDE, we will also
consider what an MDE approach would look like using other tools.  As a starting
point, we will look at \href{https://www.eclipse.org/epsilon/}{Epsilon}.
Epsilon is a family of languages and tools for code generation, model-to-model
transformation, model validation, comparison, migration and refactoring that
work out of the box with EMF, UML, Simulink, XML and other types of models.

\subsubsection{Measuring the Qualities of MDE Developed
  Artifacts} \label{SecQOfMDEArtifacts}

Some of the quality measures developed in Section~\ref{SecStateOfThePractice}
for the ``state of the practice'' exercise will also be used to measure the
quality of the generated documentation and code.
Section~\ref{SecStateOfThePractice} mentions using code quality metrics, like
cyclomatic complexity, to attempt to quantify code quality.  Given the
importance of documentation, it seems reasonable to also assess its quality.
Measuring documentation quality is not something that is often promoted.
(\citet{Sanga2003} does discuss quantifying the connectivity within an SRS
document by analyzing the traceability graph.)  Part of the reason for the lack
of effort on quantifying the quality of documentation is that is not represented
formally.  With Drasil much of the documentation is formal.  This should
facilitate automatic measures of documentation quality, or at least metrics that
might be correlated with quality.

To quantify the quality of the generated documentation, measures of information
compression will be used.  The measure will be based on the length of the
(family of) generated documentation, code and test cases, compared to the length
of the generator itself.  Our measure will be based on the Minimum Description
Length (MDL) Principle, which states that ``the more we are able to compress the
data, the more we have learned about the data'' \citep{Grunwald2004}.  Another
important measure will be whether the generated documentation achieves
reproducibility. The proposed measure is reproducibility depth
\citep{Soergel2014}.  A depth of 1 means that an independent researcher can
start from the generated code and calculate the same results.  Achieving deeper
reproducibility will mean the same results can be calculated for an independent
researcher starting with the design documentation.  Each higher measure gets
further from the source code, until the highest depth (replicability), which
starts from the requirements and leads to identical (within acceptable error)
results.

\subsubsection{Case Study} \label{SecCaseStudy}

For assessing the impact of MDE, experiments where a given project is developed
in parallel by two independent teams are infeasible due to expense and
challenges for having a truly controlled experiment.  Instead, we will focus on
the more practical, but still very useful, case study approach.  Guidelines for
conducting and reporting case study research in software engineering are given
in \citet{RunesonAndHost2009}.  As explained there, a case study should be built
around a research question.  The research question should be specific and
measurable.  For the current case study exercise, we propose the question:
``What tools and methods from MDE are perceived by end user developers to
improve their productivity and the sustainability of their software in the long
term?''

We will identify multiple case studies, with potential projects coming from the
domains listed for the ``state of the practice'' exercise
(Section~\ref{SecStateOfThePractice}), like computational medicine, medical
imaging or quantum chemistry software.  We will first identify the partners and
look for feasible projects that can be developed using an MDE approach.  We will
use the ideal process identified in Section~\ref{SecMDEProcess}.  Compare to Some additional thought, and resources, are necessary
here.  If we could do a project where the scientist, or a student under their
supervision, is a true partner in the development, we might get more convincing
anecdotal evidence.  Pre and post interviews.  Measurement of quality following the
techniques developed in the state of the practice exercise.

List of projects above could also be used for case studies.

Observe - how answer questions when problems come up.  Should have so
pre and post interviews.

\subsubsection{Development of Small Physics
  Examples} \label{SecSmallPhysExample}

describe a different experiment.  conference multiple develop physics code.
Like projectile.  Survey before and after.  Quantify success of efforts and time required.

\subsection{Impact of Assurance Cases for Building
  Confidence} \label{SecAssuranceCase}

Sketching out an Assurance Case in advance of developing the software could be
just what is needed to manufacture the up-front intensity that is needed for
successful projects.

Experimentally Determine Appropriate Documentation Template: The best approach
for documenting requirements and design is an open question.  Drasil could be
used to attempt to address this question.  Different recipes for documentation
could be developed and compared.  Some comparison could be automated between the
documents, such as empirical measures of their quality.  Other measures would
require some kind of usability experiments.

Some SCS needs to be verifiably safe.  We will achieve this by building a safety
assurance case. An assurance case is an explicit structured argument pertaining
to specific properties, such as trustworthiness

Assurance cases have potential to be used in SC to improve our collective
confidence in the developed software~\cite{SmithEtAl2018_ICSEPoster}.  Further
work is necessary to demonstrate the value of assurance cases, but they
certainly hold promise.  Drasil could go a long way to supporting development of
assurance cases.  Some potential Drasil related support ideas include the
following:

\begin{itemize}
\item Automatic generation of visual depiction of the assurance case from the
  goals, claims, evidence etc.
\item Automated reporting of missing evidence, incomplete arguments, etc.
\item Coordination of requirements, design decisions, etc.  Assurance cases use
  the same information as recommended in ``document driven'' design, but the
  traceability and connection between the data is even more important.  A tool
  like Drasil could go a long way in facilitating the presentation and
  navigation of an assurance case.
\end{itemize}

Assurance cases in Drasil could potentially go beyond the scientific context.
For instance, they could also incorporate safety cases for
control systems.  This could potentially be of interest to McSCert.

\subsection{Impact of Computational Variability Testing} \label{SecCVT}

This activity supports objective S3.  The applicant's students will apply LSS
and code generation to Computational Variability Testing (CVT), using the
example of a family of finite element programs.  As proposed by the applicant,
CVT uses code generation to build confidence in the generated code in an
analogous way to the use of grid refinement studies.  Grid refinement looks at
how the solution changes by varying the run-time parameter of grid density and
comparing the results to theoretical expectations.  CVT, on the other hand, can
generate code to ``refine'' build time parameters, such as order of
interpolation, or degree of implicitness.  These parameters can be
systematically varied and the results compared against the expected trend.
Varying computational variabilities is not a new idea, but it is only recently
that the option has become available of coupling this with code generation, so
that the validity of an entire program family can be assessed at the same time.
The experimentation with CVT will initially use the FEniCS
package~\citep{LoggEtAl2012} for elliptical partial differential equations to
compare elements with different orders of interpolation to verify that the
convergence to the solution agrees with the theoretical expectations.

{Design and Documentation For Family of Data Fitting Algorithms}: The fitting
used in GlassBR and in SFS (Software for Solidification) could be made much more
generic.  We could have a family of fitting algorithms that could be used in any
situation where fitting is required.  A proper commonality analysis of this
domain could potentially show the potential design decisions that bridge between
the requirements and the design.  In the SFS example many different fitting
routines were tried.  If the experiments could have been done easily via a
declarative specification, considerable time would have been saved.  If the
experiments are combined with automated testing and ``properties of a correct
solution'' the human involvement could be reduced, so that we have partially
automated algorithm selection.

\section{Implications and Contributions to Knowledge} \label{SecContribKnowledge}

We are currently putting too much trust in SCS.  Examples such as nuclear safety
analysis and computational medicine show the significance of SCS for health and
safety.  Although SCS developers do excellent work, we do not currently have
enough checks and balances in place for full confidence.  More should be
expected for documentation, design and testing.  Fortunately, SE offers
techniques, like DSLs and code generation, that can be employed to not only
improve SCS quality, but also reduce development costs.  Now is the time for
change.  SCS practitioners have recognized problems with the status quo and
multidisciplinary researchers are needed to bridge the gap between SE and SCS.

Codifying medical, biomechanical and computational knowledge is challenging, but
success will completely transform the development of safety-related software in
medicine, science and engineering.  MDE will remove the drudgery of generating
and maintaining documentation, code, test cases and build environments, enabling
scientists to focus on science, engineers to focus on engineering and medical
professionals to focus on medicine.  MDE will detect inconsistencies between
models via inter-model consistency constraints.  As a consequence, mistakes like
inconsistent units or assumptions cannot occur.  MDE will raise the bar so that
we can expect an explicit argument for safety, not just code that we are
expected to blindly trust.  With the right up-front investment of knowledge
capture, we can have software that is long lived because the stable knowledge is
separated from the rapidly changing modelling assumptions and design decisions.

\section{Research Schedule} \label{SecSchedule}

\newpage

\bibliographystyle {plainnat}
\bibliography {ResearchProposal}

\end{document}