### 2020-01-16 StateOfPractice/Peter-Notes.tex
added notes for LBM solutions and metrics/metric readings

\documentclass{article}

\usepackage{tabularx}
\usepackage{booktabs}

\title{Project Notes}

\author{Peter Michalski}

\date{}

\begin{document}

\maketitle

~\newpage

\section{List of LBM solutions}
\begin{enumerate}
	\item aromanro/LatticeBoltzmann

	\item Atruszkowska/LBM\_MATLAB

	\item ch4-project

	\item CPE490\_Lattice\_Boltzmann\_Project

	\item CUDA-LBM-simulator

	\item CudneLB (TCLB)

	\item DL\_Meso

	\item ESPResSo

	\item ESPResSo++

	\item firesim

	\item fvLBM

	\item HemeLB

	\item JavaCFD

	\item JFlowSim

	\item laboetie

	\item LatBo.jl

	\item lettuce

	\item LB2D\_Prime

	\item LB3D

	\item LBDEMcoupling-public

	\item LBSim

	\item Limbes

	\item listLBM

	\item LUMA

	\item MP-LABS

	\item MUPHY

	\item Openlb

	\item Palabos

	\item PyLBM

	\item Sailfish

	\item siramirsaman/LBM

	\item SunlightLB

	\item Taxila LBM

	\item Tensorflow

	\item turbulent\_lbm\_multigpu

	\item waLBerla

	\item wlb

	\item Zmhhaha/LBM-Cplusplus-A.A.Mohamad

\end{enumerate}

~\newpage
\section{Quality Measurement}
\subsection{Robustness}

Does not crash and provides (descriptive) error upon an invalid input.\\

Is software that crashes but provides an error more robust than a program that does not crash but does not provide an error?\\

Mean time to recovery from failures?\\

DOI: 10.1007/978-3-642-29032-9\_16:\\

The goal of robustness testing is to activate those faults (typically design or programming faults) or vulnerabilities in the system that result in incorrect operation, i.e., robustness failure, affecting the resilience of the system. Robustness testing mostly concentrates on the internal design faults activated through the system interface. The robustness failures are typically classified according to the CRASH criteria [540]: Catastrophic (the whole system crashes or reboots), Restart (the application has to be restarted), Abort (the application terminates abnormally), Silent (invalid operation is performed without error signal), and Hindering (incorrect error code is returned–note that returning a proper error code is considered as robust operation).\\

The measure of robustness can be given as the ratio of test cases that exposes robustness faults, or, from the system’s point of view, as the number of robustness faults exposed by a given test suite.\\

1) Injecting Random input 
2) Using invalid input 
3) Invalid inputs for each function in interface (valid and invalid inputs are type specific) 
4) applying mutation techniques 
\subsection{Performance}
How many concurrent loads (this will also require hardware parallelism)\\

How many application instances can be run?\\

CPU usage (monitor this)\\

Average Response Time (compared to other solutions)\\

Error rate (exceptions)\\

MTBF 
\subsection{Maintainability}
How much time it takes to add a new function (requirement)\\

Metrics for Assessing a Software System's Maintainability (Oman, Hagemeister)\\

Software Metrics for Predicting Maintainability (Frappier, Matwin, Mili) 
\subsection{Reusability}
Measuring Software Reusability (Poulin)\\

Chidamber and Kemerer object-oriented metrics:\\

https://www.aivosto.com/project/help/pm-oo-ck.html:\\

eg weighted methods per class, number of children, coupling per class\\

Software reusability metrics estimation: Algorithms, models and optimization techniques (Padhy)\\

Cyclomatic complexity: independent paths through source code\\

Software Reuse and Reusability Metrics and Models (Frakes, Terry)\\

A new reusability metric for object-oriented software (Barnard)\\

A metrics suite for measuring reusability of software components (Washizaki)\\

Reusability Index: A Measure for Assessing Software Assets Reusability (Ampatzoglou) 
\subsection{Portability}
compare effort to port to effort to redevelop\\

determine if the system can be ported: ram, processor, resolution, OS, browser\\

Issues in the Specification and Measurement of Software Portability (Mooney)\\

Designing a Measurement Method for the Portability Non-functional Requirement (Talib)\\

ISO 9126 breaks down portability testing: installability, compatability, adapatability, and replaceability.\\

https://www.softwaretestinghelp.com/what-is-portability-testing/ 
\end{document}