### 2020-02-09 StateOfPractice/Ao-Notes.tex
Update notes

\documentclass{article}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{longtable}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

\title{Ao's Notes on Medical Imaging Software}

\author{Ao Dong}

\date{\today}

\begin{document}

\maketitle

\begin{table}[hp]
\caption{Revision History} \label{TblRevisionHistory}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Date} & \textbf{Author(s)} & \textbf{Change}\\
\midrule
Jan/30/2020 & Ao Dong & Initial draft\\
Feb/02/2020 & Ao Dong & Update\\
Feb/08/2020 & Ao Dong & Changed table to a spreadsheet link\\
\bottomrule
\end{tabularx}
\end{table}

\section{Software List}

Link to the software list (Google Sheet): \href{https://docs.google.com/spreadsheets/d/122wU0v3ZtvDcqy8C4zKJ89kU-8fXAbo3Mzn6vcVXOi0/edit?usp=sharing}{LINK}

\section{Quality Measurements}

\subsection{Interoperability}
There are not many measuring methods in papers, and most of them are very
complicated.
\begin{itemize}
\item Can the workstation software (for visualization, analysis, etc.) connect
with the PACS (server)?
    \item Does the software use output from or provide input to other software?
    \item Can the software work with customized plug-ins? 
\end{itemize}

\noindent Measuring aspects from \citep{SmithEtAl2018}:
\begin{itemize}
    \item Does the software interoperate with external systems? ({yes*, no})
    \item Is there a workflow that uses other softwares? ({yes*, no})
\item If there are external interactions, is the API clearly defined? ({yes*,
no, n/a})
\end{itemize}

\subsection{Visibility/Transparency}
\begin{itemize}
\item Does the software use any version and issue tracking system, such as
Github, Gitlab for development?
\item Does the software have documents recording the development process and
status?
\item Does the software have clear release log with essential information, such
as release date, bug fixed and new features?
\end{itemize}

\noindent Measuring aspects from \citep{SmithEtAl2018}:
\begin{itemize}
\item Is the development process defined? If yes, what process is used. ({yes*,
no, n/a})
\item Ease of external examination relative to other products considered? ({1 ..
10})
\end{itemize}

\subsection{Productivity}
Can be measured by the summation of all output (such as the number of lines of
new code, the number of pages of new documents and the number of new test cases)
produced per person-day.

However, it is hard for a third party to carry out the above method without
knowing the exact number of developers or how much time they spent on the
project.

\subsection{Completeness}

\subsection{Consistency}

\newpage

\bibliographystyle {plainnat}
\bibliography {../CommonFiles/ResearchProposal}

\end{document}