### 2020-02-14 StateOfPractice/Peter-Notes-QualityMeasurement.tex
peter's notes - further cleaned up maintainability, robustness, performance

\documentclass{article}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{hyperref}
\hypersetup{%
	colorlinks = true,
	linkcolor  = black
}

\title{Project Notes - Quality Measurement}

\author{Peter Michalski}

\date{}

\begin{document}

\maketitle

\newpage

\tableofcontents
\addtocontents{toc}{\protect\thispagestyle{empty}}

~\newpage
\section{Variabilities}

Input Variabilities
\begin{enumerate}
	\item boundary parameters
	\item dimension
	\item number of velocity directions
\end{enumerate}

\noindent Calculation Variabilities
\begin{enumerate}
	\item computational model
	\item decomposition technique
	\item coefficient weights
	\item input check
	\item encoding of output
	\item exception check
\end{enumerate}

\noindent Other Variabilities
\begin{enumerate}
	\item language
	\item license
\end{enumerate}

~\newpage
\section{Quality Measurement}
\subsection{Robustness}
\subsubsection{Definition}
Software possesses the characteristic of robustness if it behaves ``reasonably'' in two situations: i) when it encounters circumstances not anticipated in the requirements specification; and ii) when the assumptions in its requirements specification are violated.

\subsubsection{Metrics}
Using CRASH criteria: (DOI: 10.1007/978-3-642-29032-9\_16)
\begin{enumerate}
	\item How does the system react when random but valid input is provided?\\

	Inject random input at interface (automated testing (selenium?) - generate list of valid input and observe if CRASH)
	\item How does the system react when invalid input is provided?\\

	Use invalid input at interface (testing - generate list of invalid input and observe if CRASH)
	\item How does the system react when required libraries are not provided?\\

	Not provide necessary libraries (find if external libraries are needed, not provide, observe if CRASH)
	\item How does the system react when directories and/or files are mutated?\\

	Removing directories (mutation) (for example removal of output directory - find directories for mutation, remove, observe if CRASH)

	\item How does the system react when run on OS and/or hardware that is outside of the requirement specification?\\

	Using OS and hardware outside of those in the reqs: (find req OS and hardware, make list outside of reqs, attempt to run the software and observe if CRASH)
\end{enumerate}
\subsubsection{Notes}
DOI: 10.1007/978-3-642-29032-9\_16:\\ 
The robustness failures are typically classified according to the CRASH criteria [540]: Catastrophic (the whole system crashes or reboots), Restart (the application has to be restarted), Abort (the application terminates abnormally), Silent (invalid operation is performed without error signal), and Hindering (incorrect error code is returned–note that returning a proper error code is considered as robust operation). The measure of robustness can be given as the ratio of test cases that exposes robustness faults, or, from the system’s point of view, as the number of robustness faults exposed by a given test suite.\\

~\newpage

\subsection{Performance}
\subsubsection{Definition}
The degree to which a system or component accomplishes its designated functions within given constraints, such as speed (database response times, for instance), throughput (transactions per second), capacity (concurrent usage loads), and timing (hard real-time demands).

\subsubsection{Metrics}

\begin{enumerate}
	\item How does the application compare in CPU utilization compared to other solutions running on the same hardware?\\

	Measure average CPU utilization of each application using the same hardware.
	\item What is the average memory usage of the application?\\

	Measure the average memory usage of the application using the same hardware.
	\item How many concurrent usage loads can the system handle?\\

	Run multiple instances of the application. 
	\item What is the average amount of time that the system needs to provide a solution?\\

	Measure the average amount of time that each application takes to provide a solution.
	\item How many errors does the system experience in X hours of run-time?\\

	Run automated tests and count the number of run-time errors.
	\item How accurate, on average, are the results of the solution?\\

	Measure the average error of the solution when compared to a pseudo oracle.
\end{enumerate}

\subsubsection{Notes}

~\newpage

\subsection{Maintainability}
\subsubsection{Definition}
The effort with which a software system or component can be modified to:

\begin{enumerate}
	\item correct faults
	\item improve performance or other
	attributes
	\item satisfy new requirements
\end{enumerate}

\subsubsection{Metrics}
\begin{enumerate}
	\item Open and closed issues on Git\\

	Count the ratio of open to closed issues on Git
	\item Average time to close issues\\

	Calculate the average time taken to close an issue on Git
	\item Update frequency\\

	Count average number of days between major update releases?
	\item Number of maintainers
\\

	Count how many maintainers have worked on this project.\\
\end{enumerate}

Metrics for Assessing a Software System's Maintainability (Oman, Hagemeister):
\begin{enumerate}
\item Age\\

Count number of months since release.
\item Size\\

Count thousands of non commented source statements (TNCSS)
\item Stability

Calculate: Stability = 1 - Change Factor when CF is ( e$\land$number of months * average percentage change of lines of code in number of months)
\item Defect Intensity\\

Calculate: Defect intensity =  e$\land$number of months * average percentage of defective lines of code per month
\item Subjective Product Appraisals\\

Measure: Subjective Product Appraisal = 5 point (very low to very high) for language complexity, application complexity, requirements volatility, product dependencies, complexity of build, installation complexity, intensity of product use, efficiency of the software system
\item Modularity\\

Measure: number of modules, average module size
\item Consistency\\

Measure: std deviation of module size (TNCSS)
\item Global Data Types\\

Measure: the number of global data types divided by the total number of defined data types
\item Global Data Structures\\

Measure: the number of global data structures divided by the total number of defined data structures
\item Data Type Consistency\\

Calculate: 1 - percentage of data structures that undergo type conversion during assignment operations
\item I/O Complexity\\

Measure: the number of lines of code devoted to I/O divided by TNCSS
\item Local Data Types\\

Calculate: the number of local data types divided by the total number of defined data types averaged over all modules
\item Local Data Types\\

Calculate: the number of local data types divided by the total number of defined data types averaged over all modules
\item Local Data Structures\\

Calculate: the number of local data structures divided by the total number of defined data structures averaged over all modules
\item Initialization Integrity\\

Calculate: percentage of variables initialized prior to use averaged over all modules
\item Overall Formatting\\

Calculate: percentage of blank lines in the whole program, percentage of modules with blank lines
\item Commenting\\

Calculate: percentage of comment lines in program, percentage of modules with header comments
\item Statement Formatting\\

Calculate: percentage of uncrowded statements (no more than one statement per line) per module averaged over all modules
\item Intramodule Commenting\\

Calculate: percentage lines of comments in module, averaged over all modules
\item Subjective Evaluation of Document Descriptiveness\\

Categorize based on accuracy, consistency, unambiguous
\item Subjective Evaluation of Document Completeness\\

Categorize based on extent of document set, contents
\item Subjective Evaluation of Document Correctness\\

Categorize based on traceability, verifiability
\item Subjective Evaluation of Document Readability\\

Categorize based on organization, accessibility via indices and table of contents, consistency of the writing style, typography, and comprehensibility
\item Subjective Evaluation of Document Modifiability\\

Categorize based on document set redundancies
\end{enumerate}
\subsubsection{Notes}

Metrics for Assessing a Software System's Maintainability (Oman, Hagemeister):\\
Software system metrics divided into 3 categories:\\

1. server maturity attributes (age since release, size, stability, maintenance intensity, defect intensity, reliability, reuse, subjective product appraisals)\\

2. source code (control structure, information structure, typography and naming and commenting) - each of these is broken into system and component subcategories\\

system control structure: modularity, complexity, consistency, nesting, control coupling, encapsulation, module reuse, control flow consistency\\

component control structure: complexity, use of structured constructs, use of unconditional branching, nesting, span of control structures, cohesion\\

system information structure: global data types, global data structures, system coupling, data flow consistency, data type consistency, nesting, I/O complexity, I/O integrity\\

component information structure: local data types, local data structures, data coupling, initialization integrity, span of data \\

system typography, naming and commenting: overall program formatting, overall program commenting, module separation, naming, symbols and case\\

component typography, naming and commenting: statement formatting, vertical spacing, horizontal spacing, intramodule commenting\\

3. supporting documentation (abstraction, physical attributes)\\

supporting documentation abstraction: descriptiveness appraisals, completeness appraisals, correctness appraisals\\

supporting documentation physical attributes: readability appraisals, modifiability appraisals\\

~\newpage

\subsection{Reusability}
\subsubsection{Definition}
The extent to which a software component can be used with or without adaptation in a problem solution other than the one for which it was originally developed.

\subsubsection{Metrics}
\subsubsection{Notes}
Measuring Software Reusability (Poulin)\\

Taxonomy of reusability mctrics
1. Empirical methods
-
-
Module oriented
- Complexity based
- Size based
- Reliability based
Component oriented

 2.Qualitative methods
-
-Module oriented
- Style guidelines
Component oriented
- Certification guidelines
- Quality guidelines\\

3.1 Prieto-Diaz and Freeman:
Their process model encourages
white-box reuse and consists of finding candidate
reusable modules, evaluating each, deciding which
module the programmer can modify the easiest, then
adapting the module. In this model they identify
four module-oricntcd metrics and a fifth metric used
to modify the fust four.

Program size. Reuse depends on a small module
size, as indicated by lines of source code.
Program structure. Reuse depends o n a simple
program structure as indicated by fewer links to
other modules (low coupling) and low
cyclomatic complexity.
Program documentation.
Reuse depends on
excellent documentation as indicated by a sub-
jective overall rating o n a scale of 1 to 10.
Programming language. Reuse depends on pro-
gamming language to the extent that it helps to
reuse a module written in the same programming
language. If a reusable module in the same lan-
guage does not exist, the degree of similarity
between the target language and the one used in
the module affects the difficulty of modifying the
module to meet the new requirement.
Reuse experience. The experience of the reuser
in the programming language and in the appliea-
tion domain affects the previous metrics because
every programmer views a module from their
own perspective. For example, programmers
wtll have diffcrent views of what makes a
“small” module, depending on their background.
?’lis fifth metric serves to modify the values of
the other mctrics.

-----

Chidamber and Kemerer object-oriented metrics:\\

https://www.aivosto.com/project/help/pm-oo-ck.html:\\

eg weighted methods per class, number of children, coupling per class\\

Software reusability metrics estimation: Algorithms, models and optimization techniques (Padhy)\\

Cyclomatic complexity: independent paths through source code\\

Software Reuse and Reusability Metrics and Models (Frakes, Terry)\\

A new reusability metric for object-oriented software (Barnard)\\

A metrics suite for measuring reusability of software components (Washizaki)\\

Reusability Index: A Measure for Assessing Software Assets Reusability (Ampatzoglou)

~\newpage
\subsection{Portability}
\subsubsection{Definition}
Effort required to transfer a program between system environments (including hardware and software).\\
\subsubsection{Metrics}
\subsubsection{Notes}
compare effort to port to effort to redevelop\\

determine if the system can be ported: ram, processor, resolution, OS, browser\\

Issues in the Specification and Measurement of Software Portability (Mooney):\\
The term portability
 refers to the ability of a software
 unit to be ported (to a given environment).
 A program is
 portable if and to the degree that the cost of porting is less than
 the cost of redevelopment. A software unit would be perfectly
 portable if it could be ported at zero cost, but this is never
 possible in practice.
 Instead, a software unit may be
 characterized by its degree ofportability,
 which is a function of
 the porting and development costs, with respect to a specific
 target environment.\\

 Some of these issues have been examined by the author
 [Mooney 931. This work has proposed as a metric the degree of
 portability
 of a software unit with respect to a target
 environment, defined as
 DPfsu) = 1 - (Cport(su,q) I Crd$\&$req.e2)).\\
 This metric relates portability to a ratio between the cost
 of porting (which depends on the properties of the existing
 software unit and on the target environment), and the cost of
 redevelopment (which depends on the requirements and the
 target environment). A series of experiments is underway to
 refine and validate this metric and to determine how to measure
 or estimate Cpoyf and Crdev.
 One study by Sheets [94]
 suggests that the metric can be badly skewed by secondary
 elements such as inadequate documentation for the existing
 software.\\

Designing a Measurement Method for the Portability Non-functional Requirement (Talib):\\
 EffortNew is the total effort needed in working hours units
 to develop the software functionalities on a new
 environment\\

-----------\\
https://www.softwaretestinghelp.com/what-is-portability-testing/:\\

ISO 9126 breaks down portability testing: installability, compatability, adapatability, and replaceability.\\

check Installability, Adaptability, Replaceability, Compatibility or Coexistence\\

Installability: validate OS reqs, memory and RAM reqs, clear installation and uninstallation procedures, additional prerequisites\\

Adapatability: hardware and software dependency, language dependency and communication system\\

\end{document}