### 2020-04-15 StateOfPractice/Methodology/GradingTemplate.tex
Hard wrap with to 80

The table below lists the full set of measures that are assessed for each
software product. The measures are grouped under headings for each quality,
an done for summary information. Following each measure, the type for a valid
result is given in brackets.  Many of the types are given as enumerated sets.
For instance, the response on many of the questions is one of ``yes,'' ``no,''
or ``unclear.'' The type ``number'' means natural number, a positive integer.
The
types for date and url are not explicitly defined, but they are what one would
expect from their names. In some cases the response for a given question is
not necessarily limited to one answer, such as the question on what platforms are
supported by the software product.  Case like this are indicated by ``set of''
preceding the type of an individual answer.  The type in these cases are then
the power set of the individual response type. In some cases a superscript
$^*$is used to indicate that a response of this type should be accompanied by
explanatory text.  For instance, if problems were caused by uninstall, the
reviewer should note what problems were caused.

\begin{longtable}{p{16cm}}
  \caption{Grading Template}   \label{table:TemplateFull}\\
  \toprule
  \textbf{Summary Information}\\
  \midrule
  Software name? (string)\\
  URL? (url)\\
  Educational institution (string)\\
  Software purpose (string)\\
Number of developers (number)\ad{It might be hard to decide. For example 3D
Slicer has 171 people with commits to its repo, but on its website, 18
contributors(including people who are not developers) are acknowledged.}\\
  How is the project funded (string)\\
  Number of downloads for current version (number)\\
  Release date (date)\\
  Last updated (date)\\
  Status (\{alive, dead, unclear\})\\
  License (\{GNU GPL, BSD, MIT, terms of use, trial, none, unclear\})\\
  Platforms (set of \{Windows, Linux, OS X, Android, Other OS\})\\
  Category (\{concept, public, private\})\\
  Development model (\{open source, freeware, commercial\})\\
  Publications using the software (set of url)\\
  Publications about the software (set of url) \\
Is source code available? (\{yes, no\})\ad{It might be better to directly ask
for a set of url, and the answer can be n/a if it's not open source, or unclear
if it can't be determined. An example is as follows.}\\
  \ad{Example}Source code URL (\{set of url, n/a, unclear\})\\
  Programming language(s) (set of \{FORTRAN, Matlab, C, \CC, Java, R, Ruby,
  Python, Cython, BASIC, Pascal, IDL, unclear\})\\

  \midrule
  \textbf{Installability} (Measured via installation on a virtual machine.)\\
  \midrule

  Are there installation instructions? (\{yes, no\})\\
  Are the installation instructions linear? (\{yes, no, n/a\})\\
  Is there something in place to automate the installation? (\{yes$^*$, no\})\\
Is there a specified way to validate the installation, such as a test suite?
\ad{What's the difference between this one and the getting started tutorial?}
(\{yes$^*$, no\})\\
How many steps were involved in the installation? (number) \ad{what if the
installation has different steps on different OS?}\\
How many software packages need to be installed before or during
installation?(number) \ad{should it be 0 if no packages needed? I saw some
answers with "no"
or "none"}\\
Run uninstall, if available. Were any obvious problems caused? (\{unavail,
yes$^*$, no\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Correctness and Verifiability}\\
  \midrule

  Are external libraries used? (\{yes$^*$, no, unclear\})\\
  Does the community have confidence in this library? (\{yes, no, unclear\})\\
  Any reference to the requirements specifications of the program?
(\{yes$^*$, no, unclear\})\ad{Seems like a good question to put in the
questionnaire}\\
What tools or techniques are used to build confidence of correctness?
(string)\ad{Seems like a good question to put in the questionnaire}\\
If there is a getting started tutorial, is the output as expected? (\{yes,
no$^*$, n/a\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Reliability}\\
  \midrule

  Did the software ``break'' during installation? (\{yes$^*$, no\})\\
Did the software ``break'' during the initial tutorial testing? (\{yes$^*$, no,
n/a\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Robustness}\\
  \midrule

Does the software handle garbage input reasonably? (\{yes, no$^*$\}) \ad{How do
we prepare garbage input?}\\
  For any plain text input files, if all new lines are replaced with new lines
  and carriage returns, will the software handle this gracefully? (\{yes,
  no$^*$, n/a\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Performance}\\
  \midrule

  Is there evidence that performance was considered? (\{yes$^*$, no\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Usability}\\
  \midrule

  Is there a getting started tutorial? (\{yes, no\})\\
  Is there a standard example that is explained? (\{yes, no\})\\
  Is there a user manual? (\{yes, no\})\\
  Does the application have the usual ``look and feel'' for the platform it is
  on? (\{yes, no$^*$\})\\
  Are there any features that show a lack of visibility? (\{yes, no$^*$\})\\
  Are expected user characteristics documented? (\{yes, no\})\\
What is the user support model? (string)\ad{Can GitHub be one? Many users
directly ask questions there by open new issues.}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Maintainability}\\
  \midrule

Is there a history of multiple versions of the software? (\{yes, no,
unclear\})\\
  Is there any information on how code is reviewed, or how to contribute?
  (\{yes$^*$, no\})\\
  Is there a changelog?  (\{yes, no\})\\
What is the maintenance type? (set of \{corrective, adaptive, perfective,
unclear\})\\
  What issue tracking tool is employed? (set of \{Trac, JIRA, Redmine, e-mail,
discussion board, sourceforge, google code, git, none, unclear\}) \ad{Does git
mean GitHub and GitLab? Should we add BitBucket here?}\\
Are the majority of identified bugs fixed? (\{yes, no$^*$, unclear\})\ad{It
might be better to record a number here. An example is as follows.}\\
\ad{Example}What is the percentage of identified issues that are closed?
(percentage)\\
Which version control system is in use? (\{svn, cvs, git, github, unclear\})\\
Is there evidence that maintainability was considered in the design?
  (\{yes$^*$, no\})\\
Are there code clones? (\{yes$^*$, no, unclear\})\ad{Might need to test some
static analyzer.}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Reusability}\\
  \midrule

Are any portions of the software used by another package? (\{yes$^*$,
no\})\ad{It seems hard to find an answer for some software.}\\
  Is there evidence that reusability was considered in the design? (API
documented, web service, command line tools, ...) (\{yes$^*$, no, unclear\})\\
Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Portability}\\
  \midrule

  What platforms is the software advertised to work on?
  (set of \{Windows, Linux, OS X, Android, Other OS\})\\
  Are special steps taken in the source code to handle portability? (\{yes$^*$,
  no, n/a\}) \\
  Is portability explicitly identified as NOT being important? (\{yes, no\})\\
Convincing evidence that portability has been achieved? (\{yes$^*$,
no\})\ad{Seems like a good question to put in the questionnaire}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Understandability} (Based on 10 random source files)\\
  \midrule

  Consistent indentation and formatting style? (\{yes, no, n/a\})\\
  Explicit identification of a coding standard? (\{yes$^*$, no, n/a\})\\
  Are the code identifiers consistent, distinctive, and
  meaningful? (\{yes, no$^*$, n/a\})\\
Are constants (other than 0 and 1) hard coded into the program? (\{yes, no$^*$,
n/a\})\\
Comments are clear, indicate what is being done, not how? (\{yes, no$^*$,
n/a\})\\
  Is the name/URL of any algorithms used mentioned?
  (\{yes, no$^*$, n/a\})\\
  Parameters are in the same order for all functions? (\{yes, no$^*$, n/a\})\\
  Is code modularized? (\{yes, no$^*$, n/a\})\\
  Descriptive names of source code files? (\{yes, no$^*$, n/a\})\\
Is a design document provided? (\{yes$^*$, no, n/a\})\ad{What is a design
document? There can be some examples.}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Interoperability}\\
  \midrule

  Does the software interoperate with external systems? (\{yes$^*$, no\})\\
  Is there a workflow that uses other softwares? (\{yes$^*$, no\})\\
If there are external interactions, is the API clearly defined? (\{yes$^*$, no,
n/a\})\\
  \ad{Some more suggested questions are as follows:}\\
  \ad{1.}Can the output be used by other software as input?(\{yes$^*$, no\})\\
  \ad{2.}Can it take output from other software as input?(\{yes$^*$, no\})\\
\ad{3.}Does the software generate output files with uncommon formats or
extensions? Such as the ones that can only be used by itself.(\{yes$^*$, no,
n/a\})\\
  \ad{4.}Can the software work with customized plug-ins? (\{yes, no\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Visibility/Transparency}\\
  \midrule

Is the development process defined? If yes, what process is used. (\{yes$^*$,
no, n/a\})\ad{Seems like a good question to put in the questionnaire}\\
  Ease of external examination relative to other products
  considered?  (\{1 .. 10\})\\
  \ad{Some more suggested questions are as follows:}\\
\ad{1.}Are there any documents recording the development process and status?
(\{yes$^*$, no,\})\\
  \ad{2.}Are there release notes? (\{yes, no\})\\
\ad{3.}In the issue tracking system, are there at least some issues assigned to
specific developers? (\{yes, no\})\ad{It would be better to get a percentage,
but the empirical tool can't achieve this. It might be worth to manually count
the most recent 50 issues.}\\
\ad{4.}In the issue tracking system, are there at least some issues with tags of 
types and priorities? (\{yes, no\})\ad{In GitHub, tags are optional. In
BitBucket, it seems to be mandatory. In SourceForge, it seems to be not
supported.}\\
\ad{5.}Are there any project management tools used during the
development?\ad{Seems like a good question to put in the questionnaire}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Reproducibility}\\
  \midrule

  Is there a record of the environment used for their development and testing?
  (\{yes$^*$, no\})\\
  Is test data available for verification?  (\{yes, no\})\\
Are automated tools used to capture experimental context? (\{yes$^*$, no\})\\
Overall impression? (\{1 .. 10\})\\

  \bottomrule

\end{longtable}