### 2020-04-15 StateOfPractice/Peter-Notes/GradingTemplate.tex
comments on old template

The table below lists the full set of measures that are assessed for each
software product. The measures are grouped under headings for each quality,
an done for summary information. Following each measure, the type for a valid
result is given in brackets.  Many of the types are given as enumerated sets.
For instance, the response on many of the questions is one of ``yes,'' ``no,''
or ``unclear.'' The type ``number'' means natural number, a positive integer.
The
types for date and url are not explicitly defined, but they are what one would
expect from their names. In some cases the response for a given question is
not necessarily limited to one answer, such as the question on what platforms are
supported by the software product.  Case like this are indicated by ``set of''
preceding the type of an individual answer.  The type in these cases are then
the power set of the individual response type. In some cases a superscript
$^*$is used to indicate that a response of this type should be accompanied by
explanatory text.  For instance, if problems were caused by uninstall, the
reviewer should note what problems were caused.

\begin{longtable}{p{16cm}}
  \caption{Grading Template}   \label{table:TemplateFull}\\
  \toprule
  \textbf{Summary Information}\\
  \midrule
  Software name? (string)\\
  URL? (url)\\
  Educational institution (string)\\
  Software purpose (string)\\
Number of developers (number)\\
  How is the project funded (string)\\
  Number of downloads for current version (number)\\
  Release date (date)\\
  Last updated (date)\\
  Status (\{alive, dead, unclear\})\\
  License (\{GNU GPL, BSD, MIT, terms of use, trial, none, unclear\})\\
  Platforms (set of \{Windows, Linux, OS X, Android, Other OS\})\\
  Category (\{concept, public, private\})\\
  Development model (\{open source, freeware, commercial\})\\
  Publications using the software (set of url)\\
  Publications about the software (set of url) \\
  Is source code available? (\{yes, no\})\\
  Source code URL (\{set of url, n/a, unclear\})\\
  Programming language(s) (set of \{FORTRAN, Matlab, C, \CC, Java, R, Ruby,
  Python, Cython, BASIC, Pascal, IDL, unclear\})\\

  \midrule
  \textbf{Installability} (Measured via installation on a virtual machine.)\\
  \midrule
 \pmi{Are required (known to work) operating system versions listed?}\\
  Are there installation instructions? (\{yes, no\})\\
  Are the installation instructions linear? (\{yes, no, n/a\})\\
  Is there something in place to automate the installation? (\{yes$^*$, no\})\\
Is there a specified way to validate the installation, such as a test suite?
(\{yes$^*$, no\})\\
How many steps were involved in the installation? (number) \\
How many software packages need to be installed before or during
installation?(number) \\
\pmi{are specific versions of required packages listed?}\\
\pmi{are there links to or instructions for the installation of required packages?}\\
Run uninstall, if available. Were any obvious problems caused? (\{unavail,
yes$^*$, no\})\\
\pmi{In your opinion could a non-developer of the software effectively install this software by verbatim following the instructions}\\
\pmi{Was troubleshooting the instructions necessary to install the system?}
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Correctness and Verifiability}\\
  \midrule

  Are external libraries used? (\{yes$^*$, no, unclear\})\\
  \pmi{are versions of external libraries specified?}\\
  \pmi{are instructions for finding/installing the external libraries specified?}\\
  Does the community have confidence in this library? (\{yes, no, unclear\})\\
  Any reference to the requirements specifications of the program?
(\{yes$^*$, no, unclear\})\\
What tools or techniques are used to build confidence of correctness?
(string)\\
If there is a getting started tutorial, is the output as expected? (\{yes,
no$^*$, n/a\})\\
\pmi{Does the getting started tutorial provide an answer to verify against?}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Reliability}\\
  \midrule

  Did the software ``break'' during installation? (\{yes$^*$, no\})\\
  \pmi{was a helpful error message provided?}
    \pmi{was the installation recoverable?}
Did the software ``break'' during the initial tutorial testing? (\{yes$^*$, no,
n/a\})\\
  \pmi{was a helpful error message provided?}\\
  \pmi{was the testing recoverable?}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Robustness}\\
  \midrule

Does the software handle garbage input reasonably? (\{yes, no$^*$\}) \\
  For any plain text input files, if all new lines are replaced with new lines
  and carriage returns, will the software handle this gracefully? (\{yes,
  no$^*$, n/a\})\\
  \pmi{Does the system react gracefully when required libraries or input files are not provided? Descriptive message?}
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Performance}\\
  \midrule

  Is there evidence that performance was considered? (\{yes$^*$, no\})\\
  \pmi{Does the software (optionally) use GPU for processing?}\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Usability}\\
  \midrule

  Is there a getting started tutorial? (\{yes, no\})\\
  Is there a standard example that is explained? (\{yes, no\})\\
  \pmi{Is a solution provided?}\\
  Is there a user manual? (\{yes, no\})\\
  Does the application have the usual ``look and feel'' for the platform it is
  on? (\{yes, no$^*$\})\\
  \pmi{this question is vague to me}\\
  Are there any features that show a lack of visibility? (\{yes, no$^*$\})\\
  Are expected user characteristics documented? (\{yes, no\})\\
What is the user support model? (string)\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Maintainability}\\
  \midrule

Is there a history of multiple versions of the software? (\{yes, no,
unclear\})\\
  Is there any information on how code is reviewed, or how to contribute?
  (\{yes$^*$, no\})\\
  Is there a changelog?  (\{yes, no\})\\
What is the maintenance type? (set of \{corrective, adaptive, perfective,
unclear\})\\
  What issue tracking tool is employed? (set of \{Trac, JIRA, Redmine, e-mail,
discussion board, sourceforge, google code, git, none, unclear\})\\
Are the majority of identified bugs fixed? (\{yes, no$^*$, unclear\})\\
What is the percentage of identified issues that are closed?
(percentage)\\
\pmi{is there a way to measure average time to close issues?}\\
\pmi{are files well commented? - percentage?}\\
\pmi{comment on module / file size}\\
Which version control system is in use? (\{svn, cvs, git, github, unclear\})\\
Is there evidence that maintainability was considered in the design?
  (\{yes$^*$, no\})\\
Are there code clones? (\{yes$^*$, no, unclear\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Reusability}\\
  \midrule

Are any portions of the software used by another package? (\{yes$^*$,
no\})\\
  Is there evidence that reusability was considered in the design? (API
documented, web service, command line tools, ...) (\{yes$^*$, no, unclear\})\\
\pmi{Are module interfaces clearly defined?}\\
\pmi{Are modules described in the user guide?}\\
\pmi{Is the system modularized?}\\
Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Portability}\\
  \midrule

  What platforms is the software advertised to work on?
  (set of \{Windows, Linux, OS X, Android, Other OS\})\\
  \pmi{What versions of these platforms is the software advertised to work on?}\\
  \pmi{How many external libraries are needed?}\\
  Are special steps taken in the source code to handle portability? (\{yes$^*$,
  no, n/a\}) \\
  Is portability explicitly identified as NOT being important? (\{yes, no\})\\
Convincing evidence that portability has been achieved? (\{yes$^*$,
no\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Surface Understandability} (Based on 10 random source files)\\
  \midrule

  Consistent indentation and formatting style? (\{yes, no, n/a\})\\
  Explicit identification of a coding standard? (\{yes$^*$, no, n/a\})\\
  Are the code identifiers consistent, distinctive, and
  meaningful? (\{yes, no$^*$, n/a\})\\
Are constants (other than 0 and 1) hard coded into the program? (\{yes, no$^*$,
n/a\})\\
Comments are clear, indicate what is being done, not how? (\{yes, no$^*$,
n/a\})\\
  Is the name/URL of any algorithms used mentioned?
  (\{yes, no$^*$, n/a\})\\
  Parameters are in the same order for all functions? (\{yes, no$^*$, n/a\})\\
  Is code modularized? (\{yes, no$^*$, n/a\})\\
  Descriptive names of source code files? (\{yes, no$^*$, n/a\})\\
Is a design document provided? (\{yes$^*$, no, n/a\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Interoperability}\\
  \midrule

  Does the software interoperate with external systems? (\{yes$^*$, no\})\\
  Is there a workflow that uses other softwares? (\{yes$^*$, no\})\\
If there are external interactions, is the API clearly defined? (\{yes$^*$, no,
n/a\})\\

  \midrule
  \textbf{Visibility/Transparency}\\
  \midrule

Is the development process defined? If yes, what process is used. (\{yes$^*$,
no, n/a\})\\
  Ease of external examination relative to other products
  considered?  (\{1 .. 10\})\\
  Overall impression? (\{1 .. 10\})\\

  \midrule
  \textbf{Reproducibility}\\
  \midrule

  Is there a record of the environment used for their development and testing?
  (\{yes$^*$, no\})\\
  Is test data available for verification?  (\{yes, no\})\\
Are automated tools used to capture experimental context? (\{yes$^*$, no\})\\
Overall impression? (\{1 .. 10\})\\

  \bottomrule

\end{longtable}