### 2020-05-13 StateOfPractice/Ao-Notes/Questions to Domain Experts.tex
Created a separate file for the questions to experts, and modified these questions

\documentclass[12pt]{article}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xspace}
\usepackage{fullpage}
\usepackage{hyperref}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\title{Questions to Domain Experts} 
\author{Ao Dong}
\date{\today}

\begin{document}

\maketitle

\section{Paper}
\begin{itemize}
\item Where should we publish this paper?
\item Who are our targeted readers?
\end{itemize}

\section{Schedule \& Team}
\begin{itemize}
\item We propose to start this project from ... and finish it before ... . What
do you think?
\item Our meeting schedule is ... does it fit your plan?
\item We need ... extra members from your team, each working ... hours per week
for ... weeks. What do you think?
\item Is there any other experts you think we should keep them involved?
\end{itemize}

\section{Software List}
\begin{itemize}
\item The top 10 software of your choice?
\item What do you think of our process of selecting software candidates?
\item What do you think of the list of the selected software?
\item Is there any software you think should be removed from / added to the
list?
\end{itemize}

\section{Qualities \& Measurements}
\begin{itemize}
\item Are there any pain points when you develop/use software?
\item In order to test the usability, what is a good small task for the test?
\item Are there any suggestions on surface measure to qualities?
\item Are there any suggestions on deep measure to qualities?
\item Are there any suggestions on commonality analysis?
\end{itemize}

\section{Other}
\begin{itemize}
\item We are facing difficulties such as ..., do you have suggestions?
\item How to track our tasks. e.g. using GitHub issues.
\item Equipment such as computers, test tools, OS
\end{itemize}

\end{document}