### 2020-06-21 StateOfPractice/Ao-Notes/Questions to Developers.tex
update questions to developers

\documentclass[12pt]{article}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xspace}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage[normalem]{ulem}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\title{Questions to Developers} 
\author{Ao Dong}
\date{\today}

\begin{document}

\maketitle

\section{Information about the developers and users}\citep{Jegatheesan2016}
\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item What is the typical background of a developer?
\item How large is the user group?
\item What is the typical background of a user?
\end{enumerate}

\section{Information about the software}

\ad{I drafted another version without considering our measurements}

\noindent\ad{3 aspects: dev process, software quality, user feedback}

\begin{enumerate}
\item Currently, what is the biggest obstacle against your development process?
\item In order to solve this problem, is there any improvement needed for your
development process?
\item In the past, is there any major obstacle to your development process that
has been solved? How did you solve it?
\item What is your software development model? For example, waterfall, agile, etc.
\item What is your project management process? Do you think improve this process
can tackle the current problem? Were any project management tools used?
\item What is your documentation process? Do you think improve this process can
tackle the current problem?
\item \sout{Can you name several software qualities that are the most important to
this project?}
\item Was it hard to ensure the correctness of the software? If there were any
 obstacles, what methods have been considered or practiced to improve the situation?
  If practiced, did it work?
\item When designing the software, did you consider the ease of future changes?
 For example, will it be hard to change the structure of the system, modules or code
  blocks? What measures have been taken to ensure the ease of future changes
   and maintains?
\item Do you think the users can understand your software well? Has there been
any problem to use this software?
\item Do you think the current documentation can clearly convey all necessary
knowledge to the users? If yes, how did you successfully achieve it? If no, what
improvements are needed?
\end{enumerate}

\newpage

\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}

\end{document}