### 2020-12-04 StateOfPractice/Methodology/Methodology.tex
Further work on the introduction section of the methodology

\documentclass[letterpaper,cleveref]{lipics-v2019}

\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\usepackage{xcolor}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{brown}{AD}{#1}} %Ao Dong

\newcommand{\notdone}[1]{\textcolor{red}{#1}}
\newcommand{\done}[1]{\textcolor{black}{#1}}

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\theoremstyle{definition}
\newtheorem{defn}{Definition}

\title{Methodology for Assessing the State of the Practice for Domain X} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
Olu Owojaiye and Peter Michalski and Ao Dong}

\ccsdesc{Software and its engineering}
\ccsdesc{Software and its engineering~Software product lines}
\ccsdesc{General and reference~Empirical studies}

\keywords{software quality, domain analysis, scientific computing software,
  research software, empirical measures, analytic hierarchy process}

\date{\today}

\hideLIPIcs
\nolinenumbers

\begin{document}

\maketitle

\begin{abstract}
	...
\end{abstract}

~\newpage

\tableofcontents

~\newpage

\section{Introduction} \label{SecIntroduction}

Research software uses computing to simulate mathematical models of real world
systems so that we can better understand and predict the system's behaviour.  A
small sample of some important examples of research software include the
following: designing new automotive parts, analysing the flow of blood in the
body, and determining the concentration of a pollutant released into the
groundwater.  As these examples illustrate, research software can be used for
tackling important problems that impact such areas as manufacturing, financial
planning, environmental policy, and the health, welfare and safety of
communities.

Given the importance of research software, scientists and engineers are pushing
for sustainable development practices.  This is evident from the existence of
such groups as SEI, BME.  Groups promoting sustainable practices are necessary
because the current state-of-the-practice for research software does not
incorporate state-of-the-art SE tools and methods
\citep{JohansonAndHasselbring2018}.  The lack of SE tools and methods
contributes to sustainability and reliability problems \citep{FaulkEtAl2009}.
Problems with the current state of the practice are evident from embarrassing
failures, like a retraction of derived molecular protein structures
\citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  To
improve this situation, we need to more fully understand the current state of
the practice for research software.

To understand the current quality of SCS and to assist in developing methods for
measuring qualities, we will identify several domains within SCS and endeavour
to summarize the current state of the practice within those domains.  This
``state of the practice'' exercise will build off of prior work on
measuring/assessing the state of software development practice in several SCS
domains.  We will update the work that was done previously for domains such as
Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators
\citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  We may
return to these domains and/or introduce new domains.  Potential
domains/collaborators are listed in the subsections below.

In the previous project, we measured 30 software projects for each domain, but
the measures were relatively shallow.  With the re-boot we will refine the llist
into a short list of the best projects for more detailed measurement.

In the previous state of the practice assessing exercise
\citep{SmithEtAl2018_arXivGIS}, a series of questions and simple metrics was
devised to measure the quality of the documentation and adherence to best
software development practices.  The previous set of questions is provided
\href{https://github.com/adamlazz/DomainX/blob/master/TemplateToGradeSCSoft.pdf}
{here}. As described below, the new project critically assess the previous set of questions and
revises as necessary.  In addition, the following data is also collected for each domain:

\begin{itemize}
\item Characterization of the functionality provided by the software in the
  domain.  What services does each member of the domain provide?  This
  information would be summarized in a commonality analysis document.
\item Usability testing of each software in the domain.
\item Collect empirical software engineering related data, such as the number of
  files, number of lines of code, number of open issues,
  number of closed issues, the number of lines of documentation (?). 
\end{itemize}

One of the guidelines for successfully on-boarding new developers is to
  ``Keep knowledge up to date and findable'' \citep{ShollerEtAl2019}.  The
  selected measurements should be able to assist with judging whether a project
  succeeds on this front.

The above data will be combined to rank the software in the domain.  The
specifics are yet to be determined, but in the past the Analytic Hierarchy
Process (AHP) \citep{Saaty1980} proved helpful in this.  In the past the AHP
results were calculated by a grade out of 10 from each student evaluator for
each quality for each measured product.  A better approach may be to have a
group of people familiar with the software do a pairwise comparison between each
product to come up with a number.  The number would be based on the data
collected, but not in a numerical way.  The data collected would serve the same
role as a decision support system.  It would provided insight, but human beings
would do the final pairwise comparisons, and thus the final ranking.

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we did not involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

GitHub will be used to coordinate the work of the large team of people that will
be involved in this project.  In addition to the project record left on GitHub,
the final data will be exported to Mendeley (as was done for the previous
quality measurement exercise, for instance the grades for GIS software are
summarized at \href{https://data.mendeley.com/datasets/6kprpvv7r7/1}
{https://data.mendeley.com/datasets/6kprpvv7r7/1}.)

The data for each of the different subdomains within SCS will be standardized to
facilitate a meta-analysis that combines all of the data.

The scope of this document is to conduct a current state of practice analysis of
software in a scientific computing domain for the purpose of understanding
quality and assisting in future development. We observe product, artifact, and
process quality in our analysis. We leave the assessment of scientific computing
software using performance benchmarks to other projects, such as the work of
\cite{kaagstrom1998gemm}. The methodology presented here is part of an overall
assessment of the impact of model driven engineering and code generation on the
sustainability of scientific computing software. The research proposal for this
assessment, which contains a more detailed description of this state of practice
exercise, can be found
\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ResearchProposal.pdf}
{here}. The research questions that this exercise will answer are found in
Section \ref{ResearchQuestions}. Section \ref{StepsAQDS} outlines the steps to
be taken.

Note the following formatting conventions of this document. \textcolor{red}{Red}
text denotes a link to sections of this document. \textcolor{cyan}{Cyan}
text denotes a URL link. \textcolor{blue}{Blue} text denotes a citation.

\section{Research Questions} \label{ResearchQuestions}

The following are the research questions that we wish to answer for each of our
selected domains.  In addition to answering the questions for each domain, we
also wish to combine the data from multiple domains to answer these questions
for research software in general.

\begin{enumerate}
\item What artifacts are present in current software packages? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development
  of current software packages?
\item What are the pain points for developers working on research software
  projects?  What aspects of the existing processes, methodologies and tools do
  they consider as potentially needing improvement?  How should processes,
  methodologies and tools be changed to improve software development and
  software quality?
\item For research software developers, what specific actions are taken to
  address the following:
\begin{enumerate}
\item usability
\item traceability
\item modifiability
\item maintainability
\item correctness
\item understandability
\item unambiguity
\item reproducibility
\item visibility/transparency
\end{enumerate} 
\item How does software designated as high quality by this methodology compare
  with top rated software by the community?
\end{enumerate}

%In general questions:
%
%\begin{enumerate}
%\item Comparison between domains
%\item How to measure qualities
%\item How does the quality compare for projects with the most resources to %those
%  with the fewest?
%\item What skills/knowledge are needed by future developers?
%\item How can the development process be improved?
%\item What are the common pain points?
%\end{enumerate}
%
%For each domain questions:
%
%\begin{enumerate}
%\item Best examples within the domain
%\item What software artifacts?
%\item What are the pain points?
%\item Any advice on what can be done about the pain points?
%\end{enumerate}
%
%Measure the effort invested and the reward.  Related to sustainability.
%
%Collect the data and see what conclusions follow.  For an individual domain,
%between domains.  The process isn't so much about ranking the software as it is
%about looking at the software closely and see what conclusions arise.  The
%measurements are intended to force scrutiny, from different perspectives.

\section{Overview of Steps in Assessing Quality of the Domain Software} \label{StepsAQDS}

To answer the above research questions (Section~\ref{ResearchQuestions}), we
systematically measure the quality of the software through data collection and a
series of experiments.  An overview of the measurement process is given in the
following steps, starting from determining a domain that is suitable for
measurement:

\begin{enumerate}
\item Identify the domain. (Section~\ref{SecIdentifyDomain})
\item \emph{Domain Experts}: Create a top ten list of software packages in the
  domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal,
  research questions, measurement template, survey for short list projects,
  usability tests, maintainability
  experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the
  domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages
  list. (Section~\ref{SecInitialFilter})
\item \emph{Domain Experts}: Review domain software
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item \emph{Domain Experts}: Vet domain
  analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Gather source code and documentation for each prospective software
  package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for
  deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions
    to Developers.pdf}{Questions to Developers})
\item Usability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\item Modifiability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed -
    add pairwise comparison tool}
\item Document answers for research questions.
\end{enumerate}

%\wss{The domain expert is involved in multiple steps in the process.  How best
%  to get their feedback?  The domain experts are busy and are unlikely to devote
%  significant time to the project.  We need to quickly get to the point.  Maybe
%  something around task based inspection?  Directed interview?}

\section{How to Identify the Domain} \label{SecIdentifyDomain}

A domain of research software must be identified. Research software is defined
in this exercise as ``software that is used to generate, process or analyze
results that [are intended] to appear in a publication'' \citep{hettrick2014uk}
in a scientific or engineering context.  Research software is a more general
term that covers what is often called scientific computing.  To be applicable
for the methodology described in this document, the chosen domain must have the
following properties:

\begin{enumerate}
\item The domain must have well-defined and stable theoretical underpinning.  A
  good sign of this is the existance of standard textbooks, preferably
  understandable by an upper year undergraduate student.
\item There must be a community of people studying the domain.
\item The software packages must have open source options.
\item A preliminary search, or discussion with experts, suggests that there will
  be numerous, at least 15, candidate software packages in the domain that
  qualify as `research software'.
\end{enumerate}

Some examples of domains that fit these criteria are finite element analysis
\citep{szabo1996finite}, quantum chemistry \citep{veryazov20042molcas},
seismology \citep{SmithEtAl2018}, as well as mesh generators
\citep{smith2016state}.

\section{How to Identify Candidate Software} \label{SecIdentifyCandSoft}

Once the domain of interest is identified, the candidate software for measuring
can be found through search engine queries targeting authoritative lists of
software.  Potential places to search include \href{https://github.com/}
{GitHub}, \href{https://swmath.org/} {swMATH} and domain related publications,
such as review articles. Domain experts are also asked for their suggestions and
are asked to review the initial draft of the software list.

When forming the list and reviewing the candidate software the following
properties should be considered:

\begin{enumerate}
\item The software functionality must fall within the identified domain.
\item The source code must be viewable.
\item The empirical measures listed in Section \ref{SecEmpiricalMeasures} should
  ideally be available, which implies a preference for GitHub repositories.
\item The software cannot be marked as incomplete or in an initial development
  phase.
\end{enumerate}

\section{How to Initially Filter the Software List} \label{SecInitialFilter}

If the list of software is too long (over around 30 packages), then steps need
to be taken to create a more manageable list. To reduce the length of the list,
the following filters are applied.  The filters are applied in the priority
order listed, with the filtering process stopped once the list size is
manageable.

\begin{enumerate}
\item Scope: Software is removed by narrowing what functionality is considered
  to be within the scope of the domain.
\item Usage: Software packages are eliminated if their installation procedure is
  not clear and easy to follow.
\item Age: The older software packages (age being measured by the last date when
  a change was made) are eliminated, except in the cases where an older software
  package appears to be highly recommended and currently in use.
\end{enumerate}

Copies of both the initial and filtered lists should be kept for traceability purposes.

\section{Domain Analysis} \label{SecDomainAnalysis}

Since each domain we will study will have a reasonably small scope, we
will be able to view the software as constituting a program
family.  The concept of a program family is defined by \citet{parnas1976design}
as ``a set of programs whose common properties are so extensive that it is
advantageous to study the common properties of the programs before analyzing
individual members''.  Studying the common properties within a family of related
programs is termed a domain analysis.

The domain analysis consists of a commonality analysis of
the family of software packages. Its purpose is to show the relationships
between these packages, and to facilitate an understanding of the informal
specification and development of them. \cite{weiss1997defining} defines
commonality analysis as an approach to defining a family by identifying
commonalities, variabilities, and common terminology for the
family. Commonalities are goals, theories, models, definitions and assumptions
that are common between family members.  Variabilities are goals, theories,
models, definitions and assumptions that differ between family
members. Associated with each variability are its parameters of variation, which
summarize the possible values for that variability, along with their potential
binding time.  The binding time is when the value of the variability is set.  It
could be set as specification time, build time (when the program is being
compiled) or run time (when the code is executing).

The final result of the domain analysis will be
tables of commonalities, variabilities, and parameters of variation of a program
family.   \cite{smith2008commonality} present a template for conducting a
commonality analysis, which was referred to when conducting this
work. \cite{weiss1998commonality} describes another commonality analysis
technique for deciding the members of a program family which readers are
encouraged to be familiar with. \cite{SmithAndChen2004} and
\cite{SmithMcCutchanAndCarette2017} are examples of a commonality analysis for a
family of mesh generating software and a family of material models,
respectively. The steps to produce a commonality analysis are:

\begin{enumerate}
\item Write an Introduction
\item Write the Overview of Domain
\item List Commonalities
\item List Variabilities
\item List Parameters of Variation
\item Add Terminology, Definitions, Acronyms
\end{enumerate}

A sample commonality analysis for Lattice Boltzmann Solvers can be found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Peter-Notes/Commonality-Analysis-LB-Systems.pdf}
{here}.

\section{Empirical Measures} \label{SecEmpiricalMeasures}

Some quality measurements rely on the gathering of raw and processed empirical
data. We focus on data that is reasonably easy to collect, which we combine and
analyze for this assessment. The measures that are collected relate to the
research questions (Section~\ref{ResearchQuestions}). For instance, we collect some of the
data to see how large a project is. Other measures intend to ascertain a
project’s popularity, or how active the project is.

Section \ref{rawdata} orients the reader on what raw data is collected.  Some of this data can be
observed from Git repository metrics. The rest can be collected using freeware
tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to measure
the number of binary files as well as the number of added and deleted lines in a
repository. This tool is also used to measure the number of commits over
different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc and
  Code (scc)} is used to measure the number of text based files as well as the
number of total, code, comment, and blank lines in a repository. These tools
were selected due to their installability, usability, and ability to gather the
empirical measures listed below. Details on installing and running the tools can
be found in our 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/A
  Guide to Empirical Measures.pdf} {Guide to Empirical Measures}. Section \ref{processeddata}
introduces the required processed data, which is calculated using the raw data.

\subsection{Raw Data}\label{rawdata}

The following raw data measures are extracted from repositories:

\begin{itemize}
\item Number of stars.
\item Number of forks.
\item Number of people watching the repository.
\item Number of open pull requests.
\item Number of closed pull requests.	
\item Number of developers.	
\item Number of open issues.
\item Number of closed issues.
\item Initial release date.
\item Last commit date.
\item Programming languages used.
\item Number of text-based files.
\item Number of total lines in text-based files.
\item Number of code lines in text-based files.
\item Number of comment lines in text-based files.
\item Number of blank lines in text-based files.
\item Number of binary files.  
\item Number of total lines added to text-based files.
\item Number of total lines deleted from text-based files.
\item Number of total commits.
\item Numbers of commits by year in the last 5 years. (Count from as early as
  possible if the project is younger than 5 years.)
\item Numbers of commits by month in the last 12 months.
\end{itemize}

\subsection{Processed Data}\label{processeddata}

The following measures are calculated from the raw data:

\begin{itemize}
\item Status of software package (dead or alive). Alive is defined as the
  presence of repository commits or software package version releases in the
  last 18 months.
\item Percentage of identified issues that are closed.
\item Percentage of code that is comments.
\end{itemize}

\section{Measure Using Measurement Template} \label{SecShallowMeasure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=1.0\textwidth]{measurement_template}
    \caption{Top of Measurement Template}
    \label{measurement_template_image}
  \end{center}
\end{figure}

The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}
{Measurement Template} is used to track measurements and quality scores for all
of the software packages in the domain. For each software package, we fill out one
column of the template. This process can take between 2 to 5 hours for each
package.  Project developers can be contacted for help regarding installation,
if necessary, but a cap of about 2 hours should be imposed on the entire
installation process, to keep the overall measurement time feasible.

Most of the data to be collected should be straightforward from reviewing the
measurement template.  However, in some cases some extra guidance is necessary
to eliminate ambiguity, as follows:

\begin{enumerate}
\item Initial release date: Mark the release year if an exact date is not
  available.
\item Publications about the software: A list of publications can be found
  directly on the website of some software packages. For others use Google
  Scholar or a similar index.
\item Is there evidence that performance was considered?: Search the software
  artifacts for any mention of speed, storage, throughput, performance
  optimization, parallelism, multi-core processing, or similar
  considerations. The search function on GitHub can help.
\item Getting started tutorial: Sometimes this is found within another artifact,
  like the user manual.
\item Continuous integration: Search the software artifacts for any mention of
  continuous integration. The search function on GitHub can help.
\end{enumerate}

For completing the template, the following steps should be followed:

\begin{enumerate}
\item Gather the summary information into the top section of the document
  (Figure \ref{measurement_template_image})
\item Using the GitStats tool that is described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics -
  GitStats section found near the bottom of the document
\item Using the SCC tool that is also described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - SCC
  section found near the bottom of the document
\item If the software package is found on git, gather the measurements for the
  Repo Metrics - GitHub section found near the bottom of the document
\item Review installation documentation and attempt to install the software
  package on a virtual machine
\item Gather the measurements for installability
\item Gather the measurements for correctness and verifiability
\item Gather the measurements for surface reliability
\item Gather the measurements for surface robustness
\item Gather the measurements for surface usability
\item Gather the measurements for maintainability
\item Gather the measurements for reusability
\item Gather the measurements for surface understandability
\item Gather the measurements for visibility and transparency
\item Assign a score out of ten for each quality. The score can be measured
  using the
  \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/MeasurementTemplate_ImpressionCalculator.xlsx}
  {Measurement Template Impression Calculator}. For each quality measurement,
  the file indicates the appropriate score to assign the measurement based on
  possible measurement values.
\end{enumerate}

\section{Analytic Hierarchy Process} \label{SecAHP}

The Analytical Hierarchy Process (AHP) is a decision-making technique that can
be used when comparing multiple options by multiple criteria. In our work AHP is
used for comparing and ranking the software packages of a domain using the
quality scores that are gathered in the
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}
{Measurement Template}. AHP performs a pairwise analysis between each of the
quality options using a matrix which is then used to generate an overall score
for each software package for the given criteria. \cite{SmithEtAl2016} shows how
AHP is applied to ranking software based on quality measures. We have developed
a tool for conducting this process. The tool includes an AHP JAR script and a
sensitivity analysis JAR script that is used to ensure that the software package
rankings are appropriate with respect to the uncertainty of the quality
scores. The README file of the tool is found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{here}. This
file outlines the requirements for, and configuration and usage of, the JAR
scripts. The JAR scripts, source code, and required libraries are located in the
same folder as the README file.

\section{Rank Short List} \label{SecRankShortList}

An AHP pairwise comparison between the short list packages with respect to
usability and modifiability survey results is conducted by the domain
expert. The tool can be found here. The usability and modifiability survey can
be found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}
{here}. This step is intended to start a conversation about the relative quality of
short-listed software packages.

\newpage

\appendix

\newpage

\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}

\end{document}