### 2021-01-16 StateOfPractice/Methodology/Interactions with Domain Experts.tex
updated interactions with domain experts

\documentclass[12pt]{article}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xspace}
\usepackage{xcolor}
\usepackage{fullpage}
\usepackage{hyperref}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{brown}{AD}{#1}} %Ao Dong

%% Global noindent
% \setlength\parindent{0pt}

\title{Interactions with Domain Experts} 
\author{Ao Dong}
\date{\today}

\begin{document}

\maketitle

The purpose of the document is to describe the interactions with the domain
experts, which will be through 3 steps as follows,
\begin{itemize}
\item Step 1 - Introductory Meeting in Section \ref{introductory_meeting}. In
this meeting, we will be focusing on introducing the background and objectives
of the project and vetting some details and plans.
\item Step 2 - Measuring in Section \ref{qualities_measurements}. Vetting the
Domain Analysis, usability experiments, modifiability experiments, and
developer interviews are included in this step.
\item Step 3 - Concluding Meeting in Section \ref{concluding_meeting}. Our
second and the last meeting, aiming to be
one hour in duration. Results of ranking will be discussed.
\end{itemize}

\section{Introductory Meeting}
\label{introductory_meeting}

\subsection{Prerequisites}
This section contains the tasks of preparing for the introductory meeting. These
tasks need to be finished before the meeting.

\subsubsection{Top 10 Software List}
\label{top_10_list}

The domain experts need to prepare the top 10 software packages of their choice.
Feedback can be provided in Task \ref{task_top_10_list}.

\subsection{About this Meeting}
Presenter: \ad{TBD}

The meeting will be roughly one hour. We will go
through all the steps in Section \ref{introductory_meeting}.

In the interest of time, some topics will only be discussed briefly during this
meeting, and tasks will be assigned to domain experts to finish later. For
example, Task \ref{task_usability_tests} in Section \ref{usability_tests}.

We will create issues on GitHub to track the progress of the tasks. Details can
be found in Appendix \ref{proj_mgmt_tools}. The tasks can also be found in
Appendix \ref{tasks_domain_experts}.

\subsection{Overall Objective}
\label{overall_objective}
Presenter: Dr.\ Spencer Smith and Dr.\ Jacques Carette \ad{TBD}

As described in the document 
\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ObjectivesAndResearchQuestions.pdf}{Objectives
And Research Questions}, the overall objective asks ``What is needed to produce
\{software + artifacts\} sustainably?'' The definitions of \emph{software},
\emph{artifacts} and \emph{sustainably} can be found in the same document. The
term \emph{softifacts} will be used to refer the combination of software and
artifacts.

According to the above document, to answer the above question, we will need to
find answers for these two questions,
\begin{enumerate}
\item ``What softifacts, in addition to the code, are needed?''
\item ``What software engineering principles, processes and methodologies should
be employed?''
\end{enumerate}

\subsection{Research Proposal}
\label{research_proposal}
Presenter: Dr.\ Spencer Smith \ad{TBD}

The research proposal is documented in
\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ResearchProposal.pdf}{Research
Proposal}.

A brief introduction. \ad{To be written}

\subsection{Research Questions}
\label{research_questions}
Presenter: Dr.\ Spencer Smith \ad{TBD}

The research questions are documented in the section Current State of the
Practice Research Questions of
\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ObjectivesAndResearchQuestions.pdf}{Objectives
And Research Questions}. Guided by overall objective in Section
\ref{overall_objective}, these questions focus on researching on ``i) knowledge,
ii) principles, processes
and methodologies, iii) software qualities and iv) the necessary investment of
time and energy''.

Do you have any thoughts on one of the questions - What are the ``pain points''
for developers working on SCS projects? What aspects of the existing processes,
methodologies and tools do they consider could potentially be improved?

\subsection{Vetting the List of Candidate Software}
\label{software_list}

The software list is
\href{https://docs.google.com/spreadsheets/d/122wU0v3ZtvDcqy8C4zKJ89kU-8fXAbo3Mzn6vcVXOi0/edit?usp=sharing}{Medical
	Imaging Software List} or Lattice Boltzmann Solver List. \ad{TBD}

What do you think of the list of the selected software?

Is there any software you would like to remove from or add to the list?

Feedback can be provided later in Task \ref{task_software_list}.

\subsection{Publication}
Presenter: Dr.\ Spencer Smith and Dr.\ Jacques Carette \ad{TBD}

\begin{itemize}
\item Where should we publish this paper?
\item Who are our targeted readers?
\end{itemize}

\subsection{Schedule \& Team}
Presenter: \ad{TBD}

\begin{itemize}
\item We propose to start this project from ... and finish it before ... . What
do you think?
\item Our meeting schedule is ... does it fit your plan?
\item We need ... extra members from your team, each working ... hours per week
for ... weeks. What do you think?
\item Are there any other experts you think we should keep them involved?
\end{itemize}

\subsection{Introducing the Next Steps}

\section{Qualities \& Measurements}
\label{qualities_measurements}

\subsection{Vetting the Domain Analysis}
\label{domain_analysis}
Task leader: Ao and Peter

The domain analysis can be found in Section 3.2 of
\href{https://github.com/Ao99/MISEG/blob/master/docs/SRS/SRS.pdf}{SRS for
Medical Imaging Software} or
\href{https://github.com/peter-michalski/LatticeBoltzmannSolvers/blob/master/docs/SRS/CA.pdf}{CA
of Lattice Boltzmann Solvers}.

Please review the Figure 1 in \href{https://github.com/Ao99/MISEG/blob/master/docs/SRS/SRS.pdf}{SRS for
Medical Imaging Software} or Figure/Table ? \ad{TBD. It can be a summary of the
commonalities, variabilities and parameters of variation in tables, or possibly
graphs.} in
\href{https://github.com/peter-michalski/LatticeBoltzmannSolvers/blob/master/docs/SRS/CA.pdf}{CA
of Lattice Boltzmann Solvers}.
What do you think of our analysis for the scope of the family, commonalities, and variabilities  for this domain?

Our preference is to have feedback from several graduate students. Feedback can be provided in Task \ref{task_domain_analysis}.

\subsection{Measurement Template}
\label{measurement_template}
Task leader: Ao or Peter

The template for software quality measurements can be found at
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}{Measurement
	Template}.

As described in Task \ref{task_measurement_template}, domain experts are
expected to take a look at this template. Feedback are welcome but not required.

\subsection{Survey for Short List Projects}
\label{survey_short_list_projects}
Task leader: Ao or Peter

After the first round of measurement, we would like to select a short list of
projects (about 5 to 10 of them for each domain) and directly ask questions to the
developers.

Considering the fact that some development teams may ignore our requests for
this interview, We would like to send invitations to all of the teams developing
the software packages in the list. We estimate that there will be a smaller
proportion of them accepting our request. What do you think of
this process? Feedback can be provided in Task \ref{task_review_short_list}.

The survey questions to the developers can be found at
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions%20to%20Developers.pdf}{Questions
	to Developers}.

We also would like to ask the domain experts a modified version of the above
questions. The purpose will be to learn how our domain experts and their
students create software. If we time it right, the interview with the domain
experts is also a chance to do a dry run with our questions. We can use Task
\ref{task_short_list_survey} to manage the progress of the survey with domain
experts.

\subsection{Usability Tests}
\label{usability_tests}
Task leader: Olu

A brief introduction. \ad{To be written}

The usability test can be found at ... \ad{TBD}

Feedback can be provided later in Task
\ref{task_usability_tests}. We can also use the same task to manage the progress of the tests.

\subsection{Modifiability Experiment}
\label{modifiability_experiment}
Task leader: \ad{TBD}

A brief introduction. \ad{To be written}

The modifiability experiment can be found at.. \ad{TBD}

Feedback can be provided later in Task
\ref{task_modifiability_experiment}. We can also use the same task to manage the
progress of the tests.

\section{Concluding Meeting}
\label{concluding_meeting}

\subsection{First AHP Ranking}
\label{first_AHP}
Presenter: \ad{TBD}

A brief introduction about the ranking results.

Do you have any feedback? Feedback can be provided in Task \ref{task_first_ranking}.

\subsection{Final AHP Ranking}
\label{final_AHP}
Presenter: \ad{TBD}

A brief introduction about the ranking results for the short list software packages.

Do you have any feedback? Feedback can be provided in Task \ref{task_final_ranking}.

\appendix
\section{Project Management Tools}
\label{proj_mgmt_tools}
We would like to use GitHub for task tracking and document version control. The
repository for this project is
\href{https://github.com/smiths/AIMSS}{https://github.com/smiths/AIMSS}.

Tasks will be created as issues on the repository and assigned to individual
stakeholders. An open issue usually means a task to be fulfilled, and it can be
closed once it is finished. Here is an example of an issue
\href{https://github.com/smiths/AIMSS/issues/19}{https://github.com/smiths/AIMSS/issues/19}.

\newpage

\section{Tasks for Domain Experts}
\label{tasks_domain_experts}

\subsection{Create a Top 10 List}
\label{task_top_10_list}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{top_10_list}

\subsection{Vet the Software List}
\label{task_software_list}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{software_list}

\subsection{Vet the Domain Analysis}
\label{task_domain_analysis}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\pmi{To be created}

\noindent Refer to Section \ref{domain_analysis}

\subsection{Read the Measurement Template}
\label{task_measurement_template}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{measurement_template}

\subsection{Review the Process of Creating a Short List}
\label{task_review_short_list}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{survey_short_list_projects}

\subsection{Take the Survey for the Short List Projects}
\label{task_short_list_survey}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{survey_short_list_projects}

\subsection{Participate in the Usability Tests}
\label{task_usability_tests}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{usability_tests}

\subsection{Participate in the Modifiability Experiment}
\label{task_modifiability_experiment}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\ad{To be created}

\noindent Refer to Section \ref{modifiability_experiment}

\subsection{Review the First AHP Ranking}
\label{task_first_ranking}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\pmi{To be created}

\noindent Refer to Section \ref{first_AHP}

\subsection{Review the Final Short List Ranking}
\label{task_final_ranking}
GitHub Issue:
\href{https://github.com/smiths/AIMSS/issues}{https://github.com/smiths/AIMSS/issues}
\pmi{To be created}

\noindent Refer to Section \ref{final_AHP}

%
%\section{Deleted Notes}
%\subsection{Selecting Process}
%The process of creating and shortening the list of software packages is
%documented in
%\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Ao-Notes/Ao-N%otes.pdf}{Ao's
%	Notes} or Peter's Notes \ad{TBD}
%
%What do you think of our process of selecting software candidates?
%

\end{document}