### 2021-01-16 StateOfPractice/Methodology/MeetingPresentations/Presentation for the Introductory Meeting.tex
created presentation for introductory meeting

%\documentclass[handout]{beamer} 
\documentclass[t,12pt,numbers,fleqn]{beamer}
%\documentclass[ignorenonframetext]{beamer}

\newif\ifquestions
%\questionstrue
\questionsfalse

\usepackage{pgfpages} 
\usepackage{hyperref}
\hypersetup{colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    filecolor=blue,
    urlcolor=blue,
    unicode=false}
\urlstyle{same}

\usepackage{booktabs}
\usepackage{bibentry}
\usepackage[round, authoryear]{natbib}
\bibliographystyle{plainnat}

%\usetheme{Iimenau}

\useoutertheme{split} %so the footline can be seen, without needing pgfpages

%\pgfpagesuselayout{resize to}[letterpaper,border shrink=5mm,landscape]  %if this is uncommented, the hyperref links do not work

\mode<presentation>{}

\input{./def-beamer}

\newcommand{\topic}{Introductory Meeting}

\input{./titlepage}

\begin{document}

%\nobibliography{../../../CommonFiles/ResearchProposal}

\input{./footline}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{brown}{AD}{#1}} %Ao Dong

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Interactions with Domain Experts}

\bi
\item Introductory Meeting
\item Measuring
\item Concluding Meeting
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{About this Meeting}

\bi
\item Introductory Meeting
\bi
\item Overall Objective
\item Research Proposal
\item Research Questions
\item Vetting the List of Candidate Software
\item Publication
\item Schedule \& Team
\item Introducing the Next Steps
\ei
\item Measuring
\item Concluding Meeting
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overall Objective}

``What is needed to produce \{software + artifacts\} sustainably?''

\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ObjectivesAndResearchQuestions.pdf}{Objectives
And Research Questions}

\bi
\item Definitions
\bi
\item software
\item artifacts
\item \{software + artifacts\} - softifacts
\item sustainable softifacts
\ei
\item What softifacts, in addition to the code, are needed?
\item What software engineering principles, processes and methodologies should be employed?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Research Proposal}

\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ResearchProposal.pdf}{Research
Proposal}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Research Questions}

\href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ObjectivesAndResearchQuestions.pdf}{Objectives
And Research Questions}

\bi
\item To investigate
\bi
\item knowledge
\item principles, processes and methodologies
\item software qualities
\item the necessary investment of time and energy
\ei
\item What are the ``pain points'' for
developers working on SCS projects?
\item What aspects of the existing processes, methodologies and tools do they consider could potentially be improved?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Vetting the List of Candidate Software}

\href{https://docs.google.com/spreadsheets/d/122wU0v3ZtvDcqy8C4zKJ89kU-8fXAbo3Mzn6vcVXOi0/edit?usp=sharing}{Medical Imaging Software List}

Task:
\bi
\item What do you think of the list of the selected software?
\item Is there any software you would like to remove from or add to the list?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Publication}

\bi
\item Where should we publish this paper?
\item Who are our targeted readers?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Schedule \& Team}

\bi
\item Project timeline
\item Schedule for concluding meeting
\item ... extra members from your team, each working ... hours per week for ...
weeks.
\item Any other experts?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Introducing the Next Steps}

\bi
\item Introductory Meeting
\item Measuring
\bi
\item Vetting the Domain Analysis
\item Measurement Template
\item Survey for Short List Projects
\item Usability Tests
\item Modifiability Experiment
\ei
\item Concluding Meeting
\bi
\item First AHP Ranking
\item Final AHP Ranking
\ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}