### 2021-01-16 StateOfPractice/Methodology/MeetingPresentations/titlepage.tex

%Title page for slides

\newcommand{\presenters}{Dr.~Spencer Smith, Dr.~Jacques Carette,\\Olu Owojaiye, Peter Michalski, Ao Dong} %use to switch presenters
\newcommand{\projectTitle}{Assessing the State of the Practice for Domain X}
\newcommand{\department}{Department of Computing and Software}
\newcommand{\instituteName}{Faculty of Engineering, McMaster University}

%\setbeamerfont{structure}{series=\bfseries}
%\usefonttheme[stillsansseriftext,stillsansserifmath]{serif}
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize item}[ball]

\title{
  {\normalsize \bf 
    \borange{\projectTitle}}\\[2ex]
  {\Large \bf \topic}}

\author{\presenters}

\institute{\instituteName}

\date{
\today
\bc
  \includegraphics[scale = 0.2, keepaspectratio]
  {./mcmaster-logo-full-color.jpg}
\ec
}

\renewcommand{\borange}[1] %orange is too hard to read
{
   \bred{#1}
}