### 2021-05-09 StateOfPractice/Ao-Report/01abstract.tex
created Ao's report folder

\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
\textbf{\large Abstract}
\end{center}

Abstract here