### 2021-05-09 StateOfPractice/Ao-Report/12Background.tex

\chapter{Background}
\label{ch_background}

\section{Categories and Status}

Briefly discuss open source, freeware and commercial software.

\section{Software Quality Definitions}

Refer to the Quality Definitions of Qualities created by us

\section{Analytic Hierarchy Process}