### 2021-05-09 StateOfPractice/Ao-Report/13MeasurementMethods.tex

\chapter{Measurement Methods}
\label{ch_methods}

Where to include Domain experts, in Software Product Selection?

\section{Overall Process}

Briefly introduce the following subsections.

\section{Software Product Selection}

My process of collecting and screening the packages step by step.

\section{Grading Template}

Manual measurements and empirical tools.

\section{Measuring Qualities}

e.g. virtual machines, time spent per software, where to look for docs, etc.