### 2021-05-09 StateOfPractice/Ao-Report/14MeasurementResults.tex

\chapter{Measurement Results}
\label{ch_results}

For some qualities, it might be a good idea to cross-compare with the empirical scores.

\section{Selected Software List}

\section{Installability}

\section{Correctness \& Verifiability}

\section{Reliability}

\section{Robustness}

\section{Usability}

\section{Maintainability}

\section{Reusability}

\section{Understandability}

\section{Visibility \& Transparency}