### 2021-05-09 StateOfPractice/Ao-Report/31FullGradingTemplate.tex

\chapter{Full Grading Template}
\label{ap_grading}

appendix here