### 2021-05-10 StateOfPractice/Ao-Report/13MeasurementMethods.tex
updated the outlines

\chapter{Measurement Methods}
\label{ch_methods}

Where to include Domain experts, in Software Product Selection?

It might be a good idea to make the methods general and independent to domains.

\section{Overall Process}

Briefly introduce the following subsections.

\section{Domain Selection}
How do we choose SCS domains, and point out again that the whole section should be able to apply to most domains.

\section{Software Product Selection}

My process of collecting and screening the packages step by step.

\section{Grading Template}

Manual measurements and empirical tools.

\section{Measuring Qualities}

e.g. virtual machines, time spent per software, where to look for docs, etc.

\section{Interviewee Selection}

Generally speaking, ask all the teams which we can find contacts, and continue with the ones who are willing to participate.

\section{Interview Question Selection}

The aspects we focused on.

\section{Interview Methods}

The way to ask questions, the way to transcript answers and the technologies used.