### 2021-05-10 StateOfPractice/Ao-Report/15InterviewsWithDevelopers.tex

\chapter{Interviews with Developers}
\label{ch_interviews}

\section{Summary of Answers}

\begin{itemize}
\item Start with one by one, with commonalities and interesting special cases.

\item Shorten and summarize later.
\end{itemize}

\section{Discussions}

Any conclusions?