### 2021-05-24 StateOfPractice/Ao-Report/17Recommendations.tex
updated outlines, chap 2, and chap 3

\chapter{Recommendations}
\label{ch_recommendations}

I think the recommendations can originate from both parts - measurements and interviews.