### 2021-05-24 StateOfPractice/Ao-Report/11Introduction.tex

\chapter{Introduction}
\label{ch_intro}

introduction

scientific computing (SC), also known as scientific computation or computational science

\section{Motivation}

\section{Purpose}

\section{Scope}