### 2021-05-30 StateOfPractice/Ao-Report/31FullGradingTemplate.tex
updated ao's report ch 3

\chapter{Full Grading Template}
\label{ap_grading_template}

appendix here