### 2021-06-19 StateOfPractice/Ao-Report/16ThreatToValidity.tex

\chapter{Threat to Validity}
\label{ch_validity}

\begin{itemize}
\item limited time to each software, surface quality
\item impressions, subjective
\item searched all information by ourselves, so maybe missed some info
\item mainly measured by one person, so maybe biased
\end{itemize}