### 2021-06-24 StateOfPractice/Peter-Notes/DomainExptMeeting/Presentation for the Introductory Meeting.tex
Add domain specific domain expert meeting slides

%\documentclass[handout]{beamer} 
\documentclass[t,12pt,numbers,fleqn]{beamer}
%\documentclass[ignorenonframetext]{beamer}

\newif\ifquestions
%\questionstrue
\questionsfalse

\usepackage{pgfpages} 
\usepackage{hyperref}
\hypersetup{colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    filecolor=blue,
    urlcolor=blue,
    unicode=false}
\urlstyle{same}

\usepackage{booktabs}
\usepackage{bibentry}
\usepackage[round, authoryear]{natbib}
\bibliographystyle{plainnat}

%\usetheme{Iimenau}

\useoutertheme{split} %so the footline can be seen, without needing pgfpages

%\pgfpagesuselayout{resize to}[letterpaper,border shrink=5mm,landscape]  %if this is uncommented, the hyperref links do not work

\mode<presentation>{}

\input{./def-beamer}

\newcommand{\topic}{State of the Practice for Lattice Boltzmann Solvers}

\input{./titlepage}

\begin{document}

%\nobibliography{../../../CommonFiles/ResearchProposal}

\input{./footline}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overview}

\bi
\item Goals
  \bi
  \item Understand the state of the software development practices for
    LBM Software
  \item Make recommendations for improvements
  \item A publication that is useful to the community
  \ei  
\item We have developed a standard methdology for assessing SOP for Domain X
\item The methodology requires a \textbf{domain expert} to:
  \bi
  \item Vet the preliminary results
  \item Assess the feasibility of the recommendations
  \item Navigate the publication process
  \item Answer \textbf{developer} interview questions on pain points  
  \ei
\item Today's meeting
  \bi
\item Informal
\item Questions do not have to be answered in real time, or by domain expert
  \ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overall Process}

\begin{enumerate}
%\item Identify the domain
\item \emph{Domain Expert}: Create a top ten list
\item Brief \emph{Domain Expert}
\item Initial list of candidate software packages
\item \emph{Domain Expert}: Vet domain software list
\item Domain Analysis
\item \emph{Domain Expert}: Vet domain analysis
%\item Gather source code and documentation for software
\item Collect empirical measures
\item Measure using measurement template
\item Interview developers
\item Use AHP process to rank the software packages
\item \emph{Domain Expert}: Vet AHP ranking
\item \emph{Domain Expert}: Review recommendations
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Scope}

[Ao put your scope figure here.  Peter, you can delete this slide.]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Vet Software List}

\bi
\item How does our list compare to the domain expert's list?
\item Is any software missing?
\item Is there software that should be included?
\item Any other questions/comments or concerns?  
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Software List}

[Peter and Ao, please put your list here, with hyperlinks if possible.  You can use
columns to make it all fit on one slide.]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Domain Expert Software List}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Domain Analysis}

\bi
\item So users can pick the right software for their task
\item Think of the software as a family of related software products, like:
\bi
\item Family of automobiles
\item Computer hardware
\item ...
\ei
\item Commonalities
\item Variabilities
\item Parameters of variation
\item Do the following slides capture the commonalities and variabilities?
\item We can follow up over e-mail, if that is easier
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Commonalities}

[Peter, please list your commonalities here; Ao, we'll use a scope slide
instead]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Variabilities}

[Peter, please list your variabilities, with their associated parameters of
variation here; Ao, you can delete this slide]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Summary of Measures}

\bi
\item Empirical measures
  \bi
  \item Examples
  \ei
\item Measurement template
  \bi
  \item qualities
  \ei
\item Interview questions
  \bi
  \item Examples
  \ei
  \item AHP Process with pairwise comparisons
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Thoughts on Overall Ranking?}

[Peter and Ao, give the overall ranking graphic for your domain]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Ranking Follow-Up}

\bi
\item We would like feedback on the ranking for each of the qualities
\item Is it feasible to review all 10 graphs, and the associated write-up?
\item Maybe there is a grad student that can review the rankings?
 \ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Discuss Recommendations}

\bi
\item 
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Publication}

\bi
\item Where should we publish this paper?
\item Who do you see as the targeted readers?
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions (for next time)}

\bi
\item 
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}