### 2021-06-25 StateOfPractice/Ao-Notes/DomainExptMeeting/Presentation for the Introductory Meeting.tex
updated painpoints

%\documentclass[handout]{beamer} 
\documentclass[t,12pt,numbers,fleqn]{beamer}
%\documentclass[ignorenonframetext]{beamer}

\newif\ifquestions
%\questionstrue
\questionsfalse

\usepackage{pgfpages} 
\usepackage{hyperref}
\hypersetup{colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    filecolor=blue,
    urlcolor=blue,
    unicode=false}
\urlstyle{same}

\usepackage{multicol}
\usepackage{booktabs}
\usepackage{bibentry}
\usepackage[round, authoryear]{natbib}
\bibliographystyle{plainnat}

\usepackage{tikz}
\usetikzlibrary{mindmap}
%tikstyle commands
% Define block styles
\tikzstyle{context} = [rectangle, rounded corners, text width=2.5cm, minimum
size=2.5cm,text centered, draw=black, fill=white!30]
\tikzstyle{strategy} = [trapezium, trapezium left angle=70, trapezium right
angle=110, text width=3cm, minimum height=1cm, text centered, draw=black,
fill=white!30]
\tikzstyle{subclaim} = [rectangle, text width=3cm, minimum height=1cm, text
centered, draw=black, fill=white!30]
\tikzstyle{goal} = [rectangle, text width=5cm, minimum height=1cm, text
centered, draw=black, fill=white!30]
\tikzstyle{assumption} = [ellipse, text width=3cm, minimum height=1cm, text
centered, draw=black, fill=white!30]
\tikzstyle{evidence} = [circle, text width=2cm, minimum size=2cm, text
centered,draw=black, fill=white!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

\newcommand{\famname}{MIA}
\newcommand{\progname}{MISEG}

%\usetheme{Iimenau}

\useoutertheme{split} %so the footline can be seen, without needing pgfpages

%\pgfpagesuselayout{resize to}[letterpaper,border shrink=5mm,landscape]  %if this is uncommented, the hyperref links do not work

\mode<presentation>{}

\input{./def-beamer}

\newcommand{\topic}{State of the Practice for Medical Imaging Software}

\input{./titlepage}

\begin{document}

%\nobibliography{../../../CommonFiles/ResearchProposal}

\input{./footline}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overview}

\bi
\item Goals
  \bi
  \item Understand the state of the software development practices for
    medical imaging Software
  \item Make recommendations for improvements
  \item A publication that is useful to the community
  \ei  
\item We have developed a standard methdology for assessing SOP for Domain X
\item The methodology requires a \textbf{domain expert} to:
  \bi
  \item Vet the preliminary results
  \item Assess the feasibility of the recommendations
  \item Navigate the publication process
  \item Answer \textbf{developer} interview questions on pain points  
  \ei
\item Today's meeting
  \bi
\item Informal
\item Questions do not have to be answered in real time, or by domain expert
  \ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[plain, fragile]
\frametitle{Scope}

\scalebox{0.64}
{
\begin{tikzpicture}[mindmap, grow cyclic, every node/.style=concept, concept
color=orange!40,
	level 1/.append style={level distance=5cm,sibling angle=90},
	level 2/.append style={level distance=3cm,sibling angle=90},
	level 3/.style={level distance=3cm,sibling angle=60}]
\node{\famname}
	child[concept color=teal!40] { node {Enhancement}}
	child[concept color=teal!40] { node {Analysis}
    	child[concept color=blue!30] { node {Registration}}
	    child[concept color=blue!30,] { node {Segmentation}}
	    child[concept color=blue!30] { node {Statistical Analysis}
	        child { node {Feature Extraction}}
	        child { node {Classification}}
	        child { node {Interpretation}}
	}}
	child[concept color=teal!40] { node {Simulation/\\Modeling}}
    child[concept color=teal!40] { node {Visualization}
	    child[concept color=blue!30] { node {2D Display}}
	    child[concept color=blue!30] { node {3D Rendering}}
	    child[concept color=blue!30] { node {Reformatted Views}}
    	}
;
\end{tikzpicture}
}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overall Process}

\begin{enumerate}
%\item Identify the domain
\item \emph{Domain Expert}: Create a top ten list
\item Brief \emph{Domain Expert}
\item Initial list of candidate software packages
\item \emph{Domain Expert}: Vet domain software list
\item Domain Analysis
\item \emph{Domain Expert}: Vet domain analysis
%\item Gather source code and documentation for software
\item Collect empirical measures
\item Measure using measurement template
\item Interview developers
\item Use AHP process to rank the software packages
\item \emph{Domain Expert}: Vet AHP ranking
\item \emph{Domain Expert}: Review recommendations
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Vet Software List}

\bi
\item \structure{How does our list compare to the domain expert's list?}
\item \structure{Is any software missing?}
\item \structure{Is there software that should be included?}
\item \structure{Any other questions/comments or concerns?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[plain]
%\frametitle{SOP Software List}

\begin{multicols}{2}	
	\bi
      \item 3D Slicer
      \item  Ginkgo CADx
      \item XMedCon
      \item Weasis
      \item MRIcroGL
      \item SMILI
      \item ImageJ
      \item Fiji
      \item DicomBrowser
        \item 3DimViewer
        \item Horos
        \item OsiriX Lite
        \item dwv
        \item Drishti
        \item BioImage Suite Web
        \item OHIF Viewer
          \item Slice:Drop
          \item GATE
          \item ITK-SNAP
          \item ParaView
          \item MatrixUser
          \item DICOM Viewer
          \item INVESALIUS 3
            \item medInria
            \item dicompyler
            \item MicroView
            \item Papaya
            \item AMIDE
            \item Gwyddion \ei
\end{multicols}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Common With Domain Expert List}

\bi
\item 3DSlicer
\item Horos
\item ImageJ
\item Fiji (a ``batteries-included" distribution of ImageJ)
\item Mango (on the initial list, but not the final list)
\item Papaya (the web version of Mango)
\item MRIcroGL (MRIcron development has moved to MRIcroGL)
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Only on Domain Expert List}

\bi
\item AFNI \hyperlink{https://afni.nimh.nih.gov}{https://afni.nimh.nih.gov}
\item FSL \hyperlink{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki} 
\item Freesurfer \hyperlink{https://surfer.nmr.mgh.harvard.edu}{https://surfer.nmr.mgh.harvard.edu} 
\item Tarquin \hyperlink{http://tarquin.sourceforge.net}{http://tarquin.sourceforge.net}  (for MRSpectroscopy data)
\item Diffusion Toolkit  \hyperlink{http://trackvis.org/dtk/}{http://trackvis.org/dtk/} (diffusion toolkit and trakVis)
\item MRItrix \hyperlink{https://www.mrtrix.org}{https://www.mrtrix.org}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Only on SOP List}

\begin{multicols}{2}	
\bi
\item Ginkgo CADx
\item XMedCon
\item Weasis
\item SMILI
\item DicomBrowser
\item 3DimViewer
\item OsiriX Lite
\item dwv
\item Drishti
\item BioImage Suite Web
\item OHIF Viewer
\item Slice:Drop
\item GATE
\item ITK-SNAP
\item ParaView
\item MatrixUser
\item DICOM Viewer
\item INVESALIUS 3
\item medInria
\item dicompyler
\item MicroView
\item AMIDE
\item Gwyddion
\ei
\end{multicols}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Summary of Measures}

\bi
\item Empirical measures (number of stars, forks, pull requests, lines of code, etc)
\item Measurement template
  \bi
\item Installability
\item Correctness and verifiability
\item Surface reliability
\item Surface robustness
\item Surface usability
\item Maintainability
\item Surface Understandability
  \item Visibility and transparency
  \ei
\item Developer interviews (looking for pain points)
\item AHP Process with pairwise comparisons
  \ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Thoughts on Overall Ranking?}

\bigskip

    \includegraphics[width=1.0\textwidth]{overall_scores.png}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Ranking Follow-Up}

\bi
\item \structure{We would like feedback on the ranking for each of the qualities}
\item \structure{Is it feasible to review all 10 graphs, and the associated write-up?}
\item \structure{Maybe there is a grad student that can review the rankings?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Pain Points from Developer Interviews}

\structure{Do these fit with your experience?}
\bi
\item Lack of time to 1) implement all requested new features 2) write and maintain good documents 3) review all contributions from the community
\item It is difficult to find or keep the members with experience in both software development and medical imaging
\item Lack of funding for software development (especially for maintenance)
\item No organizations to help with developing high quality software
\item Difficulty to get test data
\item Difficulty to achieve cross-platform compatibility: 1) native app - more time and work 2) web app w/o server - worse performance 3) web app w/ servers - more cost
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Recommendations}

\structure{Do these seem feasible?  What other ideas do we have?}
\bi
\item Consult with software development organization
  \bi
\item Better Scientific Software (BSSw)
\item Software Sustainability Institute
\item Software Carpentry
  \ei
\item Citations for software (Katz project)
  \item Redefine productivity to include time working on tasks like testing, continuous
    integration and documentation
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Publication}

\bi
\item \structure{Who do you see as the targeted readers?}
\item \structure{Where should we publish this paper?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions}

\bi
\item \structure{What experience do you, or your students, have with developing
    software?}
\item \structure{How big would the development group be?}
\item \structure{What is the typical background of a developer?}
\item \structure{How many users would you software typically have?}
\item \structure{What is the typical background of a user}
\item \structure{Currently, what are the most significant obstacles in your
    development process?}
\item \structure{How might you change your development process to remove or
    reduce these obstacles?}
\item \structure{How does documentation fit into your development process? Would improved
  docu- mentation help with the obstacles you typically face?}      
\item \structure{In the past, is there any major obstacle to your development process that
  has been solved? How did you solve it?}
\item \structure{What is your software development model? For example, waterfall, agile,
  etc.}
\item \structure{What is your project management process? Do you think improving this
  process can tackle the current problem? Were any project management tools
  used?}  \ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}