### 2021-06-25 StateOfPractice/Peter-Notes/DomainExptMeeting/Presentation for the Introductory Meeting.tex
Updates to LBM Domain Expert Meeting Slides

%\documentclass[handout]{beamer} 
\documentclass[t,12pt,numbers,fleqn]{beamer}
%\documentclass[ignorenonframetext]{beamer}

\newif\ifquestions
%\questionstrue
\questionsfalse

\usepackage{pgfpages} 
\usepackage{hyperref}
\hypersetup{colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    filecolor=blue,
    urlcolor=blue,
    unicode=false}
\urlstyle{same}

\usepackage{multicol}
\usepackage{booktabs}
\usepackage{bibentry}
\usepackage[round, authoryear]{natbib}
\bibliographystyle{plainnat}

%\usetheme{Iimenau}

\useoutertheme{split} %so the footline can be seen, without needing pgfpages

%\pgfpagesuselayout{resize to}[letterpaper,border shrink=5mm,landscape]  %if this is uncommented, the hyperref links do not work

\mode<presentation>{}

\input{./def-beamer}

\newcommand{\topic}{State of the Practice for Lattice Boltzmann Solvers}

\input{./titlepage}

\begin{document}

%\nobibliography{../../../CommonFiles/ResearchProposal}

\input{./footline}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overview}

\bi
\item Goals
  \bi
  \item Understand state of software development practice
  \item Make recommendations for improvements
  \item A publication that is useful to the community
  \ei  
\item We have developed a standard methdology% for assessing SOP for Domain X
\item The methodology requires a \textbf{domain expert} to:
  \bi
  \item Vet the preliminary results
  \item Assess the feasibility of the recommendations
  \item Navigate the publication process
  \item Answer \textbf{developer} interview questions on pain points  
  \ei
\item Today's meeting
  \bi
\item Informal
\item Questions do not have to be answered in real time%, or by domain expert
  \ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Overall Process}

\begin{enumerate}
%\item Identify the domain
\item \emph{Domain Expert}: Create a top ten list
%\item Brief \emph{Domain Expert}
\item Initial list of candidate software packages
\item \emph{Domain Expert}: Vet domain software list
\item Domain Analysis, Software Features
\item \emph{Domain Expert}: Vet domain analysis / features
%\item Gather source code and documentation for software
\item Collect empirical measures (stars, forks, lines of code, etc.)
\item Measure using measurement template
\item Interview developers (pain points)
\item Use AHP process to rank the software packages
\item \emph{Domain Expert}: Vet AHP ranking
\item \emph{Domain Expert}: Review recommendations %$\Leftarrow$
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Vet Software List}

\bi
\item \structure{How does our list compare to the domain expert's list?}
\item \structure{Is any software missing?}
\item \structure{Is there software that should be included?}
\item \structure{Any other questions/comments or concerns?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{SOP Software List}

\begin{multicols}{2}	
	\bi
			\item \href{https://ludwig.epcc.ed.ac.uk/}{Ludwig}
			\item \href{http://espressomd.org/wordpress/}{ESPResSo}
			\item \href{https://palabos.unige.ch/}{Palabos}
			\item \href{https://www.openlb.net/}{OpenLB}
			\item \href{https://github.com/aharwood2/LUMA/}{LUMA}
			\item \href{https://pypi.org/project/pylbm/}{pyLBM}
			\item \href{https://www.scd.stfc.ac.uk/Pages/DL_MESO.aspx}{DL\_MESO (LBE)}
			\item \href{https://www.walberla.net}{waLBerla}
			\item \href{http://sailfish.us.edu.pl/}{Sailfish}
			\item \href{https://github.com/maxlevesque/laboetie}{laboetie}
			\item \href{https://docs.tclb.io/}{TCLB}
			\item \href{http://mechsys.nongnu.org/}{MechSys}
			\item \href{https://github.com/Olllom/lettuce}{lettuce}
			\item \href{https://github.com/espressopp/espressopp}{ESPResSo++} 
			\item \href{https://github.com/carlosrosales/mplabs}{MP-LABS}		
			\item \href{http://sunlightlb.sourceforge.net/}{SunlightLB}
			\item \href{http://ccs.chem.ucl.ac.uk/lb3d}{LB3D} - 			
			\href{https://www.sciencedirect.com/science/article/pii/S0010465517301017}{paper}
			\item \href{https://code.google.com/archive/p/limbes/}{LIMBES}
			\item \href{http://faculty.fiu.edu/~sukopm/LBnD_Prime/LBnD_Prime.html}{LB2D-Prime}	
			\item \href{https://github.com/UCL/hemelb}{HemeLB} 
			\item \href{https://pypi.org/project/lbmpy/}{lbmpy}	
			\item \href{http://faculty.fiu.edu/~sukopm/LBnD_Prime/LBnD_Prime.html}{LB3D-Prime}	
			\item \href{https://github.com/UCL/LatBo.jl}{LatBo.jl}
	\ei
\end{multicols}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Domain Expert List (Common in Blue)}

\begin{multicols}{2}	
  \bi
\item \href{https://palabos.unige.ch/}{Palabos}
\item \href{https://www.openlb.net/}{OpenLB}
\item \href{https://github.com/aharwood2/LUMA/}{LUMA}
\item \href{https://www.walberla.net}{waLBerla}
\item \href{http://sailfish.us.edu.pl/}{Sailfish}
\item \href{https://github.com/UCL/hemelb}{HemeLB}
\item \href{http://ccs.chem.ucl.ac.uk/lb3d}{LB3D} - \href{https://www.sciencedirect.com/science/article/pii/S0010465517301017}{paper}
\item \href{https://github.com/carlosrosales/mplabs}{MP-LABS}
\item \href{http://sunlightlb.sourceforge.net/}{SunlightLB}
\item \href{https://pypi.org/project/pylbm/}{pyLBM}
\item  \href{https://www.scd.stfc.ac.uk/Pages/DL_MESO.aspx}{DL\_MESO (LBE)}
\item \href{https://code.google.com/archive/p/limbes/}{LIMBES}
\item \href{https://docs.tclb.io/}{TCLB} (also called CudneLB)
  \item ASL 
\item ch4-project
\item Open FSI
\item LIFE
\item LBSIM
\item Taxila LBM

\ei
\end{multicols}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Only on Domain Expert List}

%\begin{multicols}{2}	
  \bi
\item ASL (general purpose tool for PDEs, LBM a small part)
\item ch4-project (no docs, no installation guide)
\item Open FSI (uses Palabos for LBM, new project)
\item LIFE (small and new project)
\item LBSIM (dead project)
\item Taxila LBM (dead project, could not install, broken links in
  documentation)                    
	\ei
%\end{multicols}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Only on SOP List}

\begin{multicols}{2}	
  \bi
			\item \href{https://ludwig.epcc.ed.ac.uk/}{Ludwig}
			\item \href{http://espressomd.org/wordpress/}{ESPResSo}

			\item \href{https://github.com/maxlevesque/laboetie}{laboetie}
			\item \href{http://mechsys.nongnu.org/}{MechSys}
			\item \href{https://github.com/Olllom/lettuce}{lettuce}
			\item \href{https://github.com/espressopp/espressopp}{ESPResSo++}

			\item \href{http://faculty.fiu.edu/~sukopm/LBnD_Prime/LBnD_Prime.html}{LB2D-Prime}

			\item \href{https://pypi.org/project/lbmpy/}{lbmpy}	
			\item \href{http://faculty.fiu.edu/~sukopm/LBnD_Prime/LBnD_Prime.html}{LB3D-Prime}	
			\item \href{https://github.com/UCL/LatBo.jl}{LatBo.jl}
	\ei
\end{multicols}

\structure{Why is pylbm on the domain expert list twice, should one be lbmpy?}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{frame}
% \frametitle{Domain Analysis}

% \bi
% \item So users can pick the right software for their task
% \item Think of the software as a family of related software products, like:
% \bi
% \item Family of automobiles
% \item Computer hardware
% \item etc.
% \ei
% \item Commonalities
% \item Variabilities
% \item Parameters of variation
% \item \structure{Do the following slides capture the commonalities and variabilities?}
% \item We can follow up over e-mail, if that is easier
% \ei

% \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{SOP Common Features}

\bi
	\item Lattice - discretized domain within a boundary
	\item Collision operator (The Bhatnagar-Gross-Krook Collision Operator is common)
	\item Probability density function
	\item Equilibrium distribution function
	\item Boltzmann transport equation
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{SOP Features}

\bi
\item May use parallel processing (May use varied tools for this - MPI, CUDA..)
\item Different equilibrium distribution functions (compressible and incompressible)
\item Coefficients for equilibrium distribution function (based on velocity
  directions - 2, 3, 5, 9, 13, 15, 19, 27)
\item Lattice dimensions (1D, 2D, 3D)
\item Varied velocity directions, partially based on dimensions (2, 3, 5, 9, 13,
  15, 19, 27)
\item Collision operators (SRT, TRT, MRT, BGK)
\item Collision vs collision-free transport equations
\item Number of fluids that can be modeled simultaneously
\item Fluid parameters (Reynolds num., density, viscosity, etc.)
\item Lattice boundary (reflective or non-reflective)
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Domain Expert Features}

\bi
\item May use parallel processing (MPI, CUDA, OpenMP, ...)
\item Lattice dimensions (1D, 2D, 3D)
\item Turbulent (Yes, No)
\item FSI (Fluid Structure Interaction)
\item Complex Geometry
\item Non-Newtonian Fluids
\item Programming Language (C++, Python, etc)
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Potential Next Steps}

\bi
\item \structure{Can we work together on completing the following for the
    paper?}
  \bi
\item Domain analysis
\item Combined list of features
\item Matrix of software related to features \ei \ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Measurements Made}

\be
\item Installability
\item Correctness and verifiability
\item Surface reliability
\item Surface robustness
\item Surface usability
\item Maintainability
\item Reusability
\item Surface understandability
\item Visibility and transparency
 \ee

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Maintainability Measurement}

\bigskip

    \includegraphics[width=1.0\textwidth]{maintainability_chart.png}

    %\bigskip

    \structure{Does anything stand out?}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Domain Expert: Ranking Follow-Up}

\bi
\item \structure{Can you review the full ranking measurements?}
\bi
\item There are 9 measurements
\item The write up is about 10 pages long \ei \ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{From Developer Interviews: Pain Points}

\begin{enumerate}
\item Lack of time to implement new features \label{LackOfTime}
\item Team members lack software development experience
\item Lack of funding for software development
\item Lack of credit for developing software \label{LackOfCredit}
\item No organizations to help with developing high quality software \label{NoOrg}
\item Documentation could be improved, only has shallow overview of the theory
\item Setting up parallelization is challenging
\item Setting up continuous integration is challenging
\item Difficulties in ensuring correctness because of size of test cases
\item Users sometimes use tools inappropriately
\item Technical debt due to coupling between components
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Recommendations}

\structure{Do these seem feasible?  What other ideas do we have?}
\bi
\item Consult with software development organization (P\ref{NoOrg})
  \bi
\item Better Scientific Software (BSSw)
\item Software Sustainability Institute
\item Software Carpentry
  \ei
\item Citations for software (Katz project) (P\ref{LackOfCredit})
  \item Redefine productivity to include time working on tasks like testing, continuous
    integration and documentation
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Publication}

\bi
\item \structure{Who do you see as the targeted readers?}
\item \structure{Where should we publish this paper?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions}

\bi
\item \structure{What experience do you, or your students, have with developing
    software?}
\item \structure{How big would the development group be?}
\item \structure{What is the typical background of a developer?}
\item \structure{How many users would you software typically have?}
\item \structure{What is the typical background of a user?}
\item \structure{Currently, what are the most significant obstacles in your
    development process?}
\item \structure{How might you change your development process to remove or
    reduce these obstacles?}
\item \structure{How does documentation fit into your development process? Would improved
  documentation help with the obstacles you typically face?}      
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions Continued}

\bi
\item \structure{In the past, is there any major obstacle to your development process that
  has been solved? How did you solve it?}
\item \structure{What is your software development model? For example, waterfall, agile,
  etc.}
\item \structure{What is your project management process? Do you think improving this
  process can tackle the current problem? Were any project management tools
  used?}
\item \structure{Was it hard to ensure the correctness of the software? If there
    were any obstacles, what methods have been considered or practiced to
    improve the situation? If practiced, did it work?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions Continued}

\bi
\item \structure{When designing the software, did you consider the ease of
    future changes? For example, will it be hard to change the structure of
    the system, modules or code blocks? What measures have been taken to ensure
    the ease of future changes and maintainance?}
\item \structure{Provide instances where users have misunderstood the
    software. What, if any, actions were taken to address understandability
    issues?}
\item \structure{What, if any, actions were taken to address usability issues?}
\item \structure{Do you think the current documentation can clearly convey all
    necessary knowledge to the users? If yes, how did you successfully achieve
    it? If no, what improvements are needed?}
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Developer Questions Continued}

\bi
\item \structure{Do you have any concern that your computational results won't
    be reproducible in the future? Have you taken any steps to ensure
    reproducibility?}  \ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}