### 2021-07-14 StateOfPractice/Ao-Report/32FullSoftwareListBeforeFiltering.tex
ao report - update Appd B

\chapter{Full Software List Before Filtering}
\label{ap_list_before_filtering}

Table \ref{tab_list_before_filtering} lists the 48 software before filtering. We selected 29 of them to the final list, which are all open-source and in the same sub-group. We divided them into four sub-groups: Viewer (V), Tool Kit (TK), Picture Archiving and Communication System (PACS), and Convertion (C). The table also shows the sources of identifying them.

\begin{table}[]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
3D Slicer & X & X & V &  &  & X & X & X &  & X \\
Ginkgo CADx & X & X & V & X & X & X & X & X & X &  \\
XMedCon & X & X & V &  &  & X & X & X & X &  \\
Weasis & X & X & V &  &  & X & X & X & X &  \\
MRIcroGL & X & X & V &  &  & X & X & X &  & X \\
SMILI & X & X & V &  &  &  & X & X & X &  \\
ImageJ & X & X & V &  &  & X & X & X &  & X \\
Fiji & X & X & V &  &  &  &  &  &  & X \\
DicomBrowser & X & X & V &  &  &  & X & X &  &  \\
3DimViewer & X & X & V & X &  &  & X & X &  &  \\
Horos & X & X & V & X & X &  &  &  &  &  \\
OsiriX Lite & X & X & V & X & X & X &  &  &  &  \\
dwv & X & X & V & X &  & X &  &  &  &  \\
Drishti & X & X & V & X &  &  &  &  &  &  \\
\begin{tabular}[c]{@{}l@{}}BioImage Suite\\ Web\end{tabular} & X & X & V &  &  & X &  &  &  &  \\
OHIF Viewer & X & X & V & \multicolumn{7}{l}{Other source: \cite{Ziegler2020}} \\
Slice:Drop & X & X & V &  &  & X &  &  &  &  \\
\end{tabular}
\end{table}

\begin{table}[]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
GATE & X & X & V &  &  &  & X &  &  &  \\
ITK-SNAP & X & X & V &  &  &  &  &  &  & X \\
ParaView & X & X & V &  &  &  &  &  &  & X \\
MatrixUser & X & X & V &  &  &  &  & X &  &  \\
DICOM Viewer & X & X & V &  &  &  &  & X &  &  \\
INVESALIUS 3 & X & X & V &  &  &  &  & X &  &  \\
medInria & X & X & V &  &  & X &  &  &  &  \\
dicompyler & X & X & V &  &  &  &  &  & X &  \\
MicroView & X & X & V &  &  & X &  &  &  &  \\
Papaya & X & X & V &  &  &  & X & X & X &  \\
AMIDE & X & X & V &  &  &  & X & X & X &  \\
Gwyddion & X & X & V &  &  &  &  & X &  &  \\
VTK &  & X & TK &  &  &  &  &  &  & X \\
ITK &  & X & TK &  &  &  &  &  &  & X \\
DCMTK &  & X & TK &  &  &  &  & X & X &  \\
XTK &  & X & TK & \multicolumn{7}{l}{Other source: it is used by slice:drop} \\
dcm4che &  & X & TK & X &  &  & X & X &  &  \\
cornerstone &  & X & TK &  &  & X &  &  &  &  \\
\end{tabular}
\end{table}

\begin{table}[]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
orthanc &  & X & PACS & X &  &  &  & X &  &  \\
Conquest &  & X & PACS & X &  &  &  &  &  &  \\
ClearCanvas &  & X & PACS, V & X &  &  &  &  &  &  \\
dcm2niix &  & X & C &  &  &  &  & X &  &  \\
\begin{tabular}[c]{@{}l@{}}Open Dicom\\ Viewer\end{tabular} &  & X & V &  &  & X &  &  & X &  \\
MicroDicom &  &  & V & X &  &  &  &  & X &  \\
Aliza &  &  & V &  &  &  & X &  & X &  \\
jivex &  &  & V &  &  & X &  &  & X &  \\
MIPAV &  &  & V &  &  & X &  &  &  &  \\
Oviyam &  &  & V &  &  & X &  &  &  &  \\
MeVisLab &  &  & V &  &  &  &  &  &  & X \\
\begin{tabular}[c]{@{}l@{}}Sante DICOM\\ Viewer Lite\end{tabular} &  &  & V &  &  &  &  &  & X &  \\
\begin{tabular}[c]{@{}l@{}}Navegatium\\ DICOM Viewer\end{tabular} &  &  & V &  &  &  &  &  & X &  \\ \hline
\end{tabular}
\caption{\label{tab_list_before_filtering}Full software list before filtering}
\end{table}