### 2021-07-15 StateOfPractice/Ao-Report/32FullSoftwareListBeforeFiltering.tex
ao report - update Appd B

\chapter{Full Software List Before Filtering}
\label{ap_list_before_filtering}

Table \ref{tab_list_before_filtering} lists the 48 software before filtering. We selected 29 of them to the final list, which are all open-source and in the same sub-group. We divided them into four sub-groups: Viewer (V), Tool Kit (TK), Picture Archiving and Communication System (PACS), and Convertion (C). The table also shows the sources of identifying them.

\begin{table}[]
	\begin{tabular}{lllllllllll}
		\hline
		&  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
		\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
		3D Slicer \cite{Kikinis2014} & X & X & V &  &  & X & X & X &  & X \\
		Ginkgo CADx \cite{Wollny2020} & X & X & V & X & X & X & X & X & X &  \\
		XMedCon \cite{Nolf2003} & X & X & V &  &  & X & X & X & X &  \\
		Weasis \cite{Roduit2021} & X & X & V &  &  & X & X & X & X &  \\
		MRIcroGL \cite{Rorden2021} & X & X & V &  &  & X & X & X &  & X \\
		SMILI \cite{Chandra2018} & X & X & V &  &  &  & X & X & X &  \\
		ImageJ \cite{Rueden2017} & X & X & V &  &  & X & X & X &  & X \\
		Fiji \cite{Schindelin2012} & X & X & V &  &  &  &  &  &  & X \\
		DicomBrowser \cite{Archie2012} & X & X & V &  &  &  & X & X &  &  \\
		3DimViewer \cite{TESCAN2020} & X & X & V & X &  &  & X & X &  &  \\
		Horos \cite{horosproject2020} & X & X & V & X & X &  &  &  &  &  \\
		OsiriX Lite \cite{PixmeoSARL2019} & X & X & V & X & X & X &  &  &  &  \\
		dwv \cite{Martelli2021} & X & X & V & X &  & X &  &  &  &  \\
		Drishti \cite{Limaye2012} & X & X & V & X &  &  &  &  &  &  \\
		\begin{tabular}[c]{@{}l@{}}BioImage Suite\\ Web \cite{Papademetris2005} \end{tabular} & X & X & V &  &  & X &  &  &  &  \\
		OHIF Viewer \cite{Ziegler2020} & X & X & V & \multicolumn{7}{l}{Other source: \cite{Ziegler2020}} \\
		Slice:Drop \cite{Haehn2013} & X & X & V &  &  & X &  &  &  &  \\
	\end{tabular}
\end{table}

\begin{table}[]
	\begin{tabular}{lllllllllll}
		\hline
		&  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
		\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
		GATE \cite{Jan2004} & X & X & V &  &  &  & X &  &  &  \\
		ITK-SNAP \cite{Yushkevich2006} & X & X & V &  &  &  &  &  &  & X \\
		ParaView \cite{Ahrens2005} & X & X & V &  &  &  &  &  &  & X \\
		MatrixUser \cite{Liu2016} & X & X & V &  &  &  &  & X &  &  \\
		DICOM Viewer \cite{Afsar2021} & X & X & V &  &  &  &  & X &  &  \\
		INVESALIUS 3 \cite{Amorim2015} & X & X & V &  &  &  &  & X &  &  \\
		medInria \cite{Fillard2012} & X & X & V &  &  & X &  &  &  &  \\
		dicompyler \cite{Panchal2010} & X & X & V &  &  &  &  &  & X &  \\
		MicroView \cite{ParallaxInnovations2020} & X & X & V &  &  & X &  &  &  &  \\
		Papaya \cite{UTHSCSA2019} & X & X & V &  &  &  & X & X & X &  \\
		AMIDE \cite{Loening2017} & X & X & V &  &  &  & X & X & X &  \\
		Gwyddion \cite{Nevcas2012} & X & X & V &  &  &  &  & X &  &  \\
		VTK \cite{SchroederEtAl2006} &  & X & TK &  &  &  &  &  &  & X \\
		ITK \cite{McCormick2014} &  & X & TK &  &  &  &  &  &  & X \\
		DCMTK \cite{DCMTK2021} &  & X & TK &  &  &  &  & X & X &  \\
		XTK \cite{Haehn2012} &  & X & TK & \multicolumn{7}{l}{Other source: it is used by slice:drop} \\
		dcm4che \cite{dcm4che2021} &  & X & TK & X &  &  & X & X &  &  \\
		cornerstone \cite{cornerstone2020} &  & X & TK &  &  & X &  &  &  &  \\
	\end{tabular}
\end{table}

\begin{table}[]
	\begin{tabular}{lllllllllll}
		\hline
		&  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
		\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
		orthanc \cite{Jodogne2018} &  & X & PACS & X &  &  &  & X &  &  \\
		Conquest \cite{Herk2021} &  & X & PACS & X &  &  &  &  &  &  \\
		ClearCanvas \cite{ClearCanvas2015} &  & X & PACS, V & X &  &  &  &  &  &  \\
		dcm2niix \cite{Li2016} &  & X & C &  &  &  &  & X &  &  \\
		\begin{tabular}[c]{@{}l@{}}Open Dicom\\ Viewer \cite{OpenDicomViewer2011} \end{tabular} &  & X & V &  &  & X &  &  & X &  \\
		MicroDicom \cite{MicroDicom2021} &  &  & V & X &  &  &  &  & X &  \\
		Aliza \cite{Aliza2021} &  &  & V &  &  &  & X &  & X &  \\
		JiveX \cite{Visus2021} &  &  & V &  &  & X &  &  & X &  \\
		MIPAV \cite{CIT2020} &  &  & V &  &  & X &  &  &  &  \\
		Oviyam \cite{RasterImages2021} &  &  & V &  &  & X &  &  &  &  \\
		MeVisLab \cite{Heckel2009} &  &  & V &  &  &  &  &  &  & X \\
		\begin{tabular}[c]{@{}l@{}}Sante DICOM\\ Viewer Lite \cite{Santesoft2021} \end{tabular} &  &  & V &  &  &  &  &  & X &  \\
		\begin{tabular}[c]{@{}l@{}}Navegatium\\ DICOM Viewer\end{tabular} &  &  & V &  &  &  &  &  & X &  \\ \hline
	\end{tabular}
	\caption{\label{tab_list_before_filtering}Full software list before filtering}
\end{table}