### 2021-07-26 StateOfPractice/Ao-Report/32FullSoftwareListBeforeFiltering.tex

\chapter{Full Software List Before Filtering}
\label{ap_list_before_filtering}

Table \ref{tab_list_before_filtering} lists the 48 software before filtering. We selected 29 of them to the final list, which are all open-source and in Visualization (V) sub-group. We found software packages in sub-groups Tool Kit (TK) and PACS but removed them from the final list. The table also shows the sources of identifying them.

\begin{table}[H]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
3D Slicer \cite{Kikinis2014} & X & X & V &  &  & X & X & X &  & X \\
Ginkgo CADx \cite{Wollny2020} & X & X & V & X & X & X & X & X & X &  \\
XMedCon \cite{Nolf2003} & X & X & V &  &  & X & X & X & X &  \\
Weasis \cite{Roduit2021} & X & X & V &  &  & X & X & X & X &  \\
MRIcroGL \cite{Rorden2021} & X & X & V &  &  & X & X & X &  & X \\
SMILI \cite{Chandra2018} & X & X & V &  &  &  & X & X & X &  \\
ImageJ \cite{Rueden2017} & X & X & V &  &  & X & X & X &  & X \\
Fiji \cite{Schindelin2012} & X & X & V &  &  &  &  &  &  & X \\
DicomBrowser \cite{Archie2012} & X & X & V &  &  &  & X & X &  &  \\
\end{tabular}
\end{table}

\begin{table}[H]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
3DimViewer \cite{TESCAN2020} & X & X & V & X &  &  & X & X &  &  \\
Horos \cite{horosproject2020} & X & X & V & X & X &  &  &  &  &  \\
OsiriX Lite \cite{PixmeoSARL2019} & X & X & V & X & X & X &  &  &  &  \\
dwv \cite{Martelli2021} & X & X & V & X &  & X &  &  &  &  \\
Drishti \cite{Limaye2012} & X & X & V & X &  &  &  &  &  &  \\
\begin{tabular}[c]{@{}l@{}}BioImage Suite\\ Web \cite{Papademetris2005} \end{tabular} & X & X & V &  &  & X &  &  &  &  \\
OHIF Viewer \cite{Ziegler2020} & X & X & V & \multicolumn{7}{l}{Other source: \cite{Ziegler2020}} \\
Slice:Drop \cite{Haehn2013} & X & X & V &  &  & X &  &  &  &  \\
GATE \cite{Jan2004} & X & X & V &  &  &  & X &  &  &  \\
ITK-SNAP \cite{Yushkevich2006} & X & X & V &  &  &  &  &  &  & X \\
ParaView \cite{Ahrens2005} & X & X & V &  &  &  &  &  &  & X \\
MatrixUser \cite{Liu2016} & X & X & V &  &  &  &  & X &  &  \\
DICOM Viewer \cite{Afsar2021} & X & X & V &  &  &  &  & X &  &  \\
INVESALIUS 3 \cite{Amorim2015} & X & X & V &  &  &  &  & X &  &  \\
medInria \cite{Fillard2012} & X & X & V &  &  & X &  &  &  &  \\
dicompyler \cite{Panchal2010} & X & X & V &  &  &  &  &  & X &  \\
MicroView \cite{ParallaxInnovations2020} & X & X & V &  &  & X &  &  &  &  \\
Papaya \cite{UTHSCSA2019} & X & X & V &  &  &  & X & X & X &  \\
AMIDE \cite{Loening2017} & X & X & V &  &  &  & X & X & X &  \\
Gwyddion \cite{Nevcas2012} & X & X & V &  &  &  &  & X &  &  \\
VTK \cite{SchroederEtAl2006} &  & X & TK &  &  &  &  &  &  & X \\
ITK \cite{McCormick2014} &  & X & TK &  &  &  &  &  &  & X \\
\end{tabular}
\end{table}

\begin{table}[H]
\begin{tabular}{lllllllllll}
\hline
 &  &  &  & \multicolumn{7}{c}{Source} \\ \cline{5-11} 
\multirow{-2}{*}{Software} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Final\\ list\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Open-\\ source\end{tabular}} & \multirow{-2}{*}{\begin{tabular}[c]{@{}l@{}}Sub-\\ group\end{tabular}} & \cite{Bjorn2017} & \cite{Bruhschwein2019} & \cite{Haak2015} & \cite{Emms2019} & \cite{Hasan2020} & \cite{Mu2019} & \cite{Samala2014} \\ \hline
DCMTK \cite{DCMTK2021} &  & X & TK &  &  &  &  & X & X &  \\
XTK \cite{Haehn2012} &  & X & TK & \multicolumn{7}{l}{Other source: it is used by slice:drop} \\
dcm4che \cite{dcm4che2021} &  & X & TK & X &  &  & X & X &  &  \\
cornerstone \cite{cornerstone2020} &  & X & TK &  &  & X &  &  &  &  \\
dcm2niix \cite{Li2016} &  & X & TK &  &  &  &  & X &  &  \\
orthanc \cite{Jodogne2018} &  & X & PACS & X &  &  &  & X &  &  \\
Conquest \cite{Herk2021} &  & X & PACS & X &  &  &  &  &  &  \\
ClearCanvas \cite{ClearCanvas2015} &  & X & PACS, V & X &  &  &  &  &  &  \\
\begin{tabular}[c]{@{}l@{}}Open Dicom\\ Viewer \cite{OpenDicomViewer2011} \end{tabular} &  & X & V &  &  & X &  &  & X &  \\
MicroDicom \cite{MicroDicom2021} &  &  & V & X &  &  &  &  & X &  \\
Aliza \cite{Aliza2021} &  &  & V &  &  &  & X &  & X &  \\
JiveX \cite{Visus2021} &  &  & V &  &  & X &  &  & X &  \\
MIPAV \cite{CIT2020} &  &  & V &  &  & X &  &  &  &  \\
Oviyam \cite{RasterImages2021} &  &  & V &  &  & X &  &  &  &  \\
MeVisLab \cite{Heckel2009} &  &  & V &  &  &  &  &  &  & X \\
\begin{tabular}[c]{@{}l@{}}Sante DICOM\\ Viewer Lite \cite{Santesoft2021} \end{tabular} &  &  & V &  &  &  &  &  & X &  \\
\begin{tabular}[c]{@{}l@{}}Navegatium\\ DICOM Viewer\end{tabular} &  &  & V &  &  &  &  &  & X &  \\ \hline
\end{tabular}
\caption{\label{tab_list_before_filtering}Full software list before filtering}
\end{table}