### 2021-07-28 StateOfPractice/Ao-Report/01abstract.tex
MI report - added abstract

\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
\textbf{\large Abstract}
\end{center}

We present a method to assess the state of the practice for software in Scientific Computing (SC) domains and apply the method to the Medical Imaging (MI) domain. This method guided us to select 29 MI software projects from 48 candidates, assess nine software qualities by answering 103 questions for each software, and interview eight of the 29 development teams. The results helped us with revealing the current status of MI software development, such as the merits, drawbacks, and pain points during the development process and the software qualities. Based on our findings, we suggested guidelines and recommendations for future SC software development. \newline

\noindent\textbf{Keywords:} Medical Imaging, Scientific Computing, software engineering, software quality, Analytic Hierarchy Process, developer interview