### 2021-08-10 StateOfPractice/Ao-Report/33InterviewAnswers.tex

\chapter{Other Interview Answers}
\label{ap_interview}

We asked 20 interview questions to the nine interviewees from eight software projects. We discuss the answers to interview questions 5, 9, 10, 11, 12, 13, 14, and 19 in Section \ref{ch_interview}, and summarize the answers to the other questions in this section.

\begin{description}
\item[Q1.] Interviewees’ current position/title? degrees?
\end{description}

Six of the nine interviewees revealed their position/title, such as CEO of a company, endowed chair and professor in universities, software engineers in a commercial company and a hospital.

Most of them answered their backgrounds and degrees. Table \ref{tab_q1_degrees} shows the highest academic degrees the participants have, and Table \ref{tab_q1_majors} shows what majors they studied. Many of the interviewees studied in multiple majors.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Highest degree & Num ans. \\ \hline
PHD & 4 \\
Master & 3 \\
Bachelor & 0 \\
Unspecified academic degree & 2 \\ \hline
\end{tabular}
\caption{\label{tab_q1_degrees}Interviewees' highest academic degrees}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Major & Num ans. \\ \hline
Computer Science & 4 \\
Physics & 2 \\
Biomedical Engineering & 1 \\
Neuroimaging & 1 \\
Geology (image analysis) & 1 \\
Media Arts and Sciences & 1 \\
Mechanical Engineering & 1 \\
Materials Engineering & 1 \\
Psychology & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q1_majors}Interviewees' majors at university}
\end{table}

\begin{description}
\item[Q2.] Interviewees’ contribution to/relationship with the software?
\end{description}

Table \ref{tab_q2_roles} shows the interviewees’ roles and responsibilities in the projects. One of the participants did not explicitly mention his role, but implicitly revealed that he was a primary contributor to the project.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Role in the projects & Num ans. \\ \hline
Chief Architect & 2 \\
Lead Developer & 1 \\
Core Developer & 5 \\
Unspecified & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q2_roles}Interviewees' roles in the projects}
\end{table}

\begin{description}
\item[Q3.] Length of time the interviewee has been involved with this software?
\end{description}

Table \ref{tab_q3_years} shows the distribution of the lengths of time the interviewees had worked on the projects.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Years & Num ans. \\ \hline
0-1 & 1 \\
2-5 & 0 \\
6-10 & 2 \\
11-15 & 3 \\
16-20 & 2 \\
21-25 & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q3_years}Lengths of time that the interviewees worked in the projects}
\end{table}

\begin{description}
\item[Q4.] How large is the development group?
\end{description}

The size of each group grows and shrinks over the years. Most teams mentioned that the team members join and leave. Some teams said that when there was sufficient funding, they could afford more developers.

Table \ref{tab_q4_members} shows the numbers of active members at the time of interviews. The members include people working on development and project management.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Num mbrs & Num ans. \\ \hline
1-3 & 5 \\
4-6 & 3 \\ \hline
\end{tabular}
\caption{\label{tab_q4_members}Numbers of current members in the projects}
\end{table}

As shown in the table, no team had a vast number of members. Some projects had more developers, such as \textit{3D Slicer}; on the other hand, some teams such as \textit{dwv} had only one primary developer, plus a maximum of two or three developers occasionally.

\textit{3D Slicer} is a special case, because it supports third-party extensions. So there have been community members developing and maintaining these extensions. Table \ref{tab_q4_members} does not include these members.

\begin{description}
\item[Q6.] What is the typical background of a developer?
\end{description}

Not all interviewees could clearly answer this question. Many of them talked about the backgrounds of members with who they were familiar. Table \ref{tab_q6_dev_backgrounds} shows the number of times all interviewees mentioned a background.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Background of a developer & Num ans. \\ \hline
\begin{tabular}[c]{@{}l@{}}Computer Science, Information\\ Technology, and Software Development \end{tabular} & 6 \\
Imaging & 2 \\
Medical Imaging & 1 \\
Mathematics & 1 \\
Biomedical Engineering & 1 \\
Computer Aided Medical Procedures & 1 \\
Physician & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q6_dev_backgrounds}Backgrounds of developers by the numbers of interviewees with the answers}
\end{table}

\begin{description}
\item[Q7.] What is your estimated number of users? How did you come up with that estimate?
\end{description}

None of the interviewees knew the exact number of users. Some of them provided estimations based on different facts. However, we do not think these numbers are comparable to each other.

\begin{table}[H]
\centering
\hspace*{-2cm}\begin{tabular}{lll}
\hline
Software & Rough estimation & Considered facts \\ \hline
3D Slicer & 100,000 & \begin{tabular}[c]{@{}l@{}} 1) search results on Google Scholar;\\ 2) number of new posts per year on slicer.org; \\ 3) number of downloads. \end{tabular} \\
INVESALIUS 3 & 75,000 & Number of random IDs created by new installation.\\
dwv & No estimation & About 20 companies integrated \textit{dwv} in their products. \\
BioImage Suite Web & 100 active users &  Only counted the active users from several Universities. \\
ITK-SNAP & 10,000 plus & Number of downloads. \\
MRIcroGL & No estimation & The top 1 downloaded software on the NITRC list. \\
Weasis & 10,000 incl. one-time users & Number of profiles. \\
OHIF & About 5000 & Some platforms integrated \textit{OHIF}; hard to estimate. \\ \hline
\end{tabular}
\caption{\label{tab_q7_num_users}Rough Estimations for the Number of Users}
\end{table}

Table \ref{tab_q7_num_users} shows the estimations and how the interviewees made them. It is clear that some estimated only the active users, and some counted users who had used only once. So we do not compare these numbers.

\begin{description}
\item[Q8.] What is the typical background of a user?
\end{description}

All interviewees provided several different user backgrounds, and all of them mentioned medical researchers or medical professionals. Table \ref{tab_q8_user_backgrounds} shows the number of times all interviewees mentioned a background.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Background of a user & Num ans. \\ \hline
Medical Researchers & 6 \\
Doctors/Health care professionals/Surgeons & 5 \\
Student Researchers & 4 \\
Patients & 3 \\
Paleontologist & 1 \\
Biomechanical Engineers & 1 \\
Imaging Researchers & 1 \\
Mechanical Engineers & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q8_user_backgrounds}Backgrounds of users by the numbers of interviewees with the answers}
\end{table}