### 2021-09-22 StateOfPractice/QDefOfQualities/QDefOfQualities.tex

\documentclass[letterpaper, cleveref]{lipics-v2019}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\hypersetup{
colorlinks=true,        % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with linkbordercolor)
citecolor=blue,         % color of links to bibliography
filecolor=magenta,      % color of file links
urlcolor=cyan           % color of external links
}
\usepackage{bbding}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\newcommand{\notdone}[1]{\textcolor{red}{#1}}
\newcommand{\done}[1]{\textcolor{black}{#1}}

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\theoremstyle{definition}
\newtheorem{defn}{Definition}

\title{Quality Definitions of Qualities} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
Olu Owojaiye and Peter Michalski and Ao Dong}

\date{\today}

\hideLIPIcs
\nolinenumbers

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007</concept_id>
<concept_desc>Software and its engineering</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering}

\keywords{software qualities, installability, correctness, verfiability,
  validatability, testability, reliability, robustness, performance,
  usability, maintainability, reusability, portability, understandability,
  interoperability, visibility/transparency, reproducibility, productivity,
  sustainability, completness, consistency, modifiability, traceability,
  unambiguity, verifiability, abstract}

\begin{document}
\maketitle

\begin{abstract}
	Existing definitions of software qualities (also called quality
  characteristics) are collected and analyzed.  For each quality a
  ``best'\rq{} definition is recommended based on the following criteria:
  unambiguity, simplicity, consistency, ...
\end{abstract}

\tableofcontents

\section{Introduction} \label{SecIntroduction}

% Purpose

In the course of conducting and presenting our software engineering related
research we have often desired clear definitions for relevant software
qualities, like maintainability, usability, productivity, sustainability, etc.
Unfortunately, the definitions of qualities that we found in the literature were
not always satisfying.  The inadequacy of existing definitions became
particularly clear when we recently undertook an exercise measuring the
\lq\lq{}state of the practice\rq\rq{} in different software domains, including
the domains of seismology software \citep{SmithEtAl2018}, Geographic Information
Systems \citep{SmithEtAl2018_arXivGIS} and mesh generators
\citep{smith2016state}. When determining methods to quickly assess quality
across large samples of existing software, we found communication amongst
ourselves and with others challenging. We craved unambiguous and standardized
definitions to enable and inspire methods for measuring software qualities.  To
help ground our future research, and possibly help others in the same situation,
we have collected an exhaustive list of definitions.  From this list we have
produced our set of recommended software quality definitions.

In the current work, quality is defined as a measure of the excellence or worth
of an entity.  As is standard practice, we do not think of quality as a single
measure, but rather as a set of measures.  In the overview of managing software
quality provided by \citet[p 110]{VanVliet2000} the different measures are
called quality attributes, quality factors, and quality criteria.  In
\citet{ISO9126} they are called quality characteristics, subcharacteristics and
attributes.  To keep our presentation simple, and to not imply a classification
hierarchy that we don't intend, we will just use the term qualities to refer to
the different measures of quality.

Many taxonomies have been proposed for software qualities, as in
\citet{McCallEtAl1977, boehm1976quantitative, ISO9126}. Rather than aiming for a
complete coverage of the history of qualities, the wide variety of potential
qualities, or the relationships between them, we will instead take a practical
approach; we will focus on the qualities that we have found most relevant in our
research.  This means we will include the usual software engineering suspects,
like usability and maintainability.  Furthermore, since we also do research on
scientific computing software, we will also include scientifically relevant
qualities, like validatability and reproducibility relevant.

Given our respect for \citet[p.\ 15--33]{GhezziEtAl2003}, we used their list of
qualities as a starting point. Our list, shown in
Table~\ref{Tab_SoftwareQualities}, includes all of their qualities, except for
timeliness, which we excluded because the already included quality of
productivity implies the importance of timeliness.  To the list from
\citet{GhezziEtAl2003} we have added the practical quality of installability,
since our aforementioned studies on the state of the practice in different
software domains (for instance \citet{SmithEtAl2018_arXivGIS}) have shown us
that none of the other qualities matter if the software cannot be properly
installed.  Given our interest in scientific computing, we also added the
qualities of validatability and reproducibility. Since we were not able to find
much information in the literature on the quality of verifiability, we also
added the closely related quality of testability to our List.  The quality of
sustainability was added since this term has emerged as a popular goal for
software projects. The order of the qualities in
Table~\ref{Tab_SoftwareQualities} follow the order used in
\cite{GhezziEtAl2003}, which puts related qualities (like correctness and
reliability) together.  The order is roughly the order in which qualities are
considered.  We followed these same guidelines when inserting our additional
qualities into the list.

\begin{table}[h!]
  \begin{center}
  \begin{tabular}{lllll}
  \toprule
  & External & Internal & Product & Process \\  
  \midrule
  Installability & \Checkmark & & \Checkmark & \\
  Correctness & \Checkmark & \Checkmark & \Checkmark & \Checkmark \\
  Reliability & \Checkmark & & \Checkmark & \Checkmark \\
  Robustness & \Checkmark & & \Checkmark & \Checkmark \\     
  Performance & \Checkmark & & \Checkmark & \\
  Usability & \Checkmark & & \Checkmark & \Checkmark \\
  Verifiability & & \Checkmark & \Checkmark & \Checkmark \\
  Validatability & \Checkmark & & \Checkmark & \\     
  Testability & & \Checkmark & \Checkmark & \\
  Maintainability &  & \Checkmark & \Checkmark & \Checkmark\\
  Reusability & & \Checkmark & \Checkmark & \Checkmark\\
  Portability & & \Checkmark & \Checkmark & \\
  Understandability & & \Checkmark &  & \\     
  Interoperability & \Checkmark & & \Checkmark & \\
  Productivity &  &  &  & \Checkmark \\
  Visibility &  & &  & \Checkmark \\
  Reproducibility & \Checkmark & & \Checkmark & \\
  Sustainability & & \Checkmark & \Checkmark & \\      
  \bottomrule
\end{tabular}
\end{center}
\caption{Software Qualities}
\label{Tab_SoftwareQualities}
\end{table}

Table~\ref{Tab_SoftwareQualities} makes a distinction between external and
internal qualities.  External qualities are the qualities that are relevant to
the user of the software while internal qualities are those that are relevant to
developers \citep[p.\ 16]{GhezziEtAl2003}.  Although the users are not typically
interested in internal qualities, it is the internal qualities that influence
external quality.  External qualities are assessed using the produced software
and any user relevant documentation.  Internal qualities reference such things
as the code, documentation, test cases, and the software development process.
Correctness is shown as both an internal and external quality because it is
possible to judge the correctness of software and of the various internal
artifacts produced while developing software.  The external quality of
reliability is influenced by the internal quality of verifiability.  Since
validatability is related to getting the right requirements, it is shown as a
external quality.  Although one could judge the usability of internal
documentation, the quality of usability is considered as external, with the
quality of understandability introduced to cover the internal equivalent.
Filling in the table the way we want to use it.

Some qualities focus on the product, others on the process, and yet others on
both.  This is shown in Table~\ref{Tab_SoftwareQualities}.  Define process.
Define product. Process quality improves product quality.  The check marks are
for qualities that can be applied to a process, not the qualities that a process
improves.  Examples.  Performance isn't for process.  For process use
productivity.

% Scope

The topic of quality has been visited and revisited many times through the
course of software engineering.  Starting with McCall, Boehm.  Iso standards.
Taxonomies, relationships between qualities.  Rather than introduce distinctions
that aren't relevant, we'll just focus on qualities of interest. We are not
endeavoring to review all of the available literature, nor are we trying to
provide a complete framework for all qualities and the relationships between
them.  We are taking a more practical approach.  We have identified the
qualities that come up most in our research.  Not specifically emphasizing
scientific computing, but when a specific context is needed for a definition,
the scientific context will be used.  For instance, the ISO definition of
accuracy is defined in terms of precision, but in scientific computing accuracy
and precision are two different concepts.  Some are product and some are process
and some are both.  Some are internal and some are external.

\wss{List of definitions don't include textbooks that simply repeat
definitions from the literature.}

% Methodology
Methodology:

\begin{itemize}

  \item Collect the list of qualities to define, based on the qualities that are
  listed in software engineering texts, such as Ghezzi Et al and Van Hliet.

  \item For each quality find as many definitions as possible

  \item Based on the existing definitions, propose a high quality definition of
  each quality.  The proposed definition may use one of the existing, or combine
  several of the options.  We may also make other modifications.  The goal is
  for the proposed definition to address our acceptance criteria.

\end{itemize}

software products, software artifacts and software development processes, and
ii) qualities that are considered important for good specifications.  The
specification could be a specification of requirements, design or a test plan.

The pattern in the presentation for each of the qualities is the same.  First we
summarize all of the definitions that we could find in the literature.  To keep
the clutter of quotation marks down, we have adopted the convention that each
definition is given verbatim from the cited source, but without showing
quotation marks.  In cases where the definition has to be rephrased, quotation
marks are used to show those portions that are taken verbatim for the original
source.  After the summary of the existing definitions, we propose the
definition that we would like to work with going forward.  This definition can
either be our preference from the existing definitions, or a new definition,
which is often found by combining existing definitions.  Following the proposed
definition is an explanation for the reasoning that led to this choice.

\wss{introduce the terminology of quality characteristics, as in ISO9126?}

\wss{tell the reader that the definitions will be sorted in the order of their
publication.}

In determining our recommended definition for each quality, we used the
following criteria:

\begin{description}

  \item[Unambiguous]

  \item[Consistency] The final list of definitions should be consistent in
  the terminology used.  For instance, we do not use synonyms for variety;
  we use the same word to mean the same concept throughout the final list of
  definitions.  This is why measurement is given by ``the effort required.''
  and not the various synonyms for this that are seen in the literature.
  \wss{We may change this phrase we use for measuring something.  I just
  wanted to put this in here as a placeholder.}

  \item[Generality]

  \item[Measurability]

  \item[Etc.]

\end{description}

Not necessary to have every quality measurable.  Analogy to the definition of
true error in scientific computing.

Roadmap.
\subsection{Background}

Quality definitions history: McCall, Ghezzi et al, VanHleit, Boehm, ISO standards.

\subsection{Additional Definitions}

Some standards give additional definitions for terms that are used a lot.
These are reproduced here, to aid in the interpretation of the definitions
below (and their refinement).

\begin{defn} 
\textsf{Effectiveness}:
The accuracy and completeness with which users achieve specified
goals. \cite{ISO9241:11}
\end{defn}

\begin{defn} 
\textsf{Efficiency}: 
The resources expended in relation to the accuracy and
completeness with which users achieve goals. \cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Satisfaction}: 
Freedom from discomfort, and positive attitude to the use of the product.
\cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Context of use}:
characteristics of the users, tasks and the organizational and physical
environments. \cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Goal}: intended outcome. \cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Task}: activities required to achieve a goal. \cite{ISO9241:11}
\end{defn}

\section{Qualities of Software Products, Artifacts and Processes}
\label{SecQualities}

\wss{add a definition related to capability?  The word seems to come up in  many definitions.}

To assess the current state of software development, and to understand how
future changes impact software development, we need a clear definition of what
we mean by quality.  The concept of quality is decomposed into a set of separate
components that together make up ``quality''. Unfortunately, these are called
\emph{qualities}. These are associated to the software product, the software
artifacts (documentation, test cases, etc) and to the software development
process itself, and combinations thereof.

Our analysis is centred around a set of software qualities.  Quality is not
considered as a single measure, but a collection of different qualities, often
called ``ilities.''  These qualities highlight the desirable nonfunctional
properties for software artifacts, which include both documentation and
code. Some qualities, such as visibility and productivity, apply to the process
used for developing the software. The following list of qualities is based on
\cite{GhezziEtAl2003}. To the list from \cite{GhezziEtAl2003}, we have added
three qualities important for SC: installability, reproducibility and
sustainability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Installability \oo{owner}}

\begin{defn}
  The capability of the software product to be installed in a specified
  environment \citep{ISO9126}. %2001
\end{defn}

\begin{defn} \label{Installability_Selected2}
	The degree of effectiveness and efficiency with which a product or system can
	be successfully installed and/or uninstalled in a specified environment
	\citep{ISO/IEC25010}. %2011
\end{defn}

\begin{defn} \label{Installability_Selected1} Installability refers to the cost
  or effort required for the installation, given an installation is possible to
  begin with \citep{lenhard2013measuring}.
\end{defn}

\noindent \textbf{Proposed Definition}

Combined definitions \ref{Installability_Selected2} and
\ref{Installability_Selected1} rephrased: Installability refers to the effort
required for the installation, uninstallation or reinstallation of a software or
product in a specified environment.

\wss{Remove the repetition of the word being defined in the definition?  Remove
  reinstallation?  This quality characteristic is a product quality.  It is an
  external quality.  Installability should be something that can be measured
  directly.}

\noindent \textbf{Reasoning}

 The definition captures a vital metric of measure for this quality - effort,
 and considers not just the installation but uninstallation and re-installation
 which are all related to installation activities.

 \wss{remove re-installation - it doesn't seem different from installation.
 Connection between installability and portability.  Use the terminology effort
 (without cost) to be consistent}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \subsection{Correctness \oo{owner}}

\begin{defn} 
  The extent to which a program satisfies its specifications and fulfills the
  user's mission objectives \citep{McCallEtAl1977}. 
\end{defn}

\begin{defn}
  The ability of software products to perform their exact tasks, as defined by
  their specification \citep{meyer1988object}.
\end{defn}

\begin{defn}  
  The degree to which a system or component is free from faults in its
  specification, design and implementation \citep{IEEEComputerDictionary1991}.
\end{defn}

\begin{defn} \label{Correctness_Selected}
  A program is functionally correct if it behaves according to its stated
  functional specifications \citep{GhezziEtAl2003}.  	
\end{defn}

\begin{defn}
  The degree to which a system is free from faults in its specification, design, and 
  implementation \citep{mcconnell2004code}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition~\ref{Correctness_Selected} rephrased: A program is correct if it
behaves according to its stated specifications.

\noindent \textbf{Reasoning}

To determine whether a program is correct, we must be able to refer to both its
specification, design and implementation to understand the intended purpose and
this definition captures it all.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Reliability} \oo{owner}}

\begin{defn}
  Code possesses the characteristic reliability to the extent that it can be
  expected to perform its intended functions satisfactorily
  \citep{boehm1976quantitative}.
\end{defn}

\begin{defn}
  Extent to which a program can be expected to perform its intended function
  with required precision \citep{McCallEtAl1977}.
\end{defn}

\begin{defn} \label{reliabilitySelected1} It is the probability of failure-free
  operation of a computer program in a specified environment for a specified
  time \citep{musa1987software}.
\end{defn}

\begin{defn}
  The ability of a system or component to perform its required function under
  stated conditions for a specified period of time
  \citep{IEEEComputerDictionary1991}.
\end{defn}

\begin{defn}
 The capability of the software product to maintain a specified level of
 performance when used under specified conditions \cite{ISO9126}. %2001
\end{defn}

\begin{defn}
  Informally, software is reliable if the user can depend on it
  \citep{GhezziEtAl2003}.
\end{defn}

\begin{defn} \label{reliabilitySelected2} Reliability is concerned with measuring
	the probability of occurence of faliure - that is, the probability of the
	observable effects of errors \citep{GhezziEtAl2003}.
\end{defn}

\begin{defn}
  The ability of a system to perform its required functions under stated
  conditions whenever required—having a long mean time between failures
  \citep{mcconnell2004code}.
\end{defn}

\begin{defn}
  Informally, the reliability of a system is the probability, over a given
  period of time, that the system will correctly deliver services as expected by
  the user \citep{sommerville}. %2011
\end{defn}

\begin{defn}
  Reliability expresses the ability of the software to maintain a specified
  level of fault tolerance, when used under specified condition
  \citep{singh2013different}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{reliabilitySelected1} and \ref{reliabilitySelected2} rephrased:
Probability of failure-free operation of a computer program in a specified
environment for a specified time, i.e. the average time interval between two
failures also known as the mean time to failure (MTTF).

\wss{Needs to be completed.}\oo{Done}

\noindent \textbf{Reasoning}

Reliability is defined such that the measurement metrics are visible - the
specified environment and time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Robustness \pmi{owner}}

\begin{defn}
	The degree to which a system or component can function correctly in the
	presence of invalid inputs or stressful environmental conditions
	\citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn} \label{RobustnessDefnSelected}
	A program is robust if it behaves ``reasonably'', even in circumstances that
	were not anticipated in the requirements specification - for example, when it
	encounters incorrect input data or some hardware malfunction
	\citep{ghezzi1991fundamentals}. \wss{Use Ghezzi 2003 instead?}
\end{defn}

\begin{defn}
	The quality can be further informally refined as the ability of a software
  to keep an acceptable behaviour, expressed in terms of robustness
  requirements, in spite of exceptional or unforeseen execution conditions
  (such as the unavailability of system resources, communication failures,
  invalid or stressful inputs, etc.) \citep{fernandez2005model}.
\end{defn}

\begin{defn} \label{RobustnessPartialDefn}
	Code possesses the characteristic of robustness to the extent that it can
	continue to perform despite some violation of the assumptions in its
	specification \citep{boehm2007software}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{RobustnessPartialDefn} and Definition
\ref{RobustnessDefnSelected} rephrased: Software possesses the characteristic of
robustness if it behaves ``reasonably'' in two situations: i) when it encounters
circumstances not anticipated in the requirements specification; and ii) when
the assumptions in its requirements specification are violated.

\noindent \textbf{Reasoning}

This definition indicates that robustness is related to the quality of
correctness (both within and outside of it).

\pmi{Still may want to refine what we mean by reasonable}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Performance} \pmi{owner}}

\begin{defn} \label{PerformanceDefnSelected}
  The degree to which a system or component accomplishes its designated
  functions within given constraints, such as speed, accuracy, or memory
  usage \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
	In software engineering we often equate performance with efficiency. A
  software system is efficient if it uses computing resources economically
  \citep{ghezzi1991fundamentals}.  \wss{Change to 2003 reference?}
\end{defn}

\begin{defn} \label{PerformanceDefnSelected2}
	How well or how rapidly the system must perform specific functions. Performance
	requirements encompass speed (database response times, for instance), throughput
	(transactions per second), capacity (concurrent usage loads), and timing (hard
	real-time demands) \citep{wiegers2003softreq}.
\end{defn}

\noindent \textbf{Proposed Definition}

Combined definition \ref{PerformanceDefnSelected} and
\ref{PerformanceDefnSelected2}: The degree to which a system or component
accomplishes its designated functions within given constraints, such as speed
(database response times, for instance), throughput (transactions per second),
capacity (concurrent usage loads), and timing (hard real-time
demands). \wss{What is the difference between speed and timing?}

\noindent \textbf{Reasoning}

This definition offers a comprehensive list of constraints that are commonly
associated with software performance, such as speed, throughput, capacity, and
timing.

\pmi{I will check if Wiegers discusses the difference between speed and
timing.}

\pmi{Wiegers did not discuss this difference.}

\pmi{Speed - change over time; timing - related to specification: when
something should happen in relation to something else - state transitions}

\wss{If we need both speed and timing, we should define them.  We aren't
  including accuracy from the IEEE definition.  Is that on purpose?}

\pmi{We have decided not to include accuracy. SS will find difference b/w
speed and timing from a colleague.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Usability} \jc{owner}}

\begin{defn}
  ``The effort required to learn, operate, prepare input, and interpret output
  of a program'' \citep{McCallEtAl1977}.
\end{defn}

\begin{defn}
  ``A software is usable - or user friendly - if its human users find it easy
  to use'' \citep{ghezzi1991fundamentals}. \wss{switch to 2003 ref?}
\end{defn}

\begin{defn}
  The extent to which a product can be used by specified users to achieve
  specified goals with effectiveness, efficiency, and satisfaction in a
  specified context of use. \cite{ISO16982:2002} and \cite{ISO9241:11}
\end{defn}

Nielsen and (separately) Schneidermann have defined usability as part of
usefulness and is composed of:
\begin{defn}
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the first
  time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they
  perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the
economics sense of the word).
\cite{NielsenNorman} %2012
\end{defn}

\begin{defn}
Desirable outcomes of usability:
\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in
using the user interface during the first contact with the website. For
example, if a travel agent's website is a well-designed one, the user should
be able to move through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
  website. If a user has the goal of booking a flight, a good design will guide
  him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
  subsequent visits. So, a good design on the travel agent's site means the user
  should learn from the first time and book a second ticket just as easily.
\end{enumerate}
From the Interaction Design Foundation
\url{https://www.interaction-design.org/literature/topics/usability}.
\end{defn}

One core reference, for definitions and metrics, is
\citet{bevan1995measuring}.

\begin{defn}
  The capability of the software to be understood, learned, used and liked by
  the user, when used under specified conditions \cite{ISO9126}.
\end{defn}
\noindent \textbf{Proposed Definition}

\begin{defn}
  The extent to which a product can be used by specified users to achieve
  specified goals with effectiveness, efficiency, and satisfaction in a
  specified context of use. \cite{ISO16982:2002} and \cite{ISO9241:11}
\end{defn}

\noindent \textbf{Reasoning}

There has been a lot of thought put behind the ISO definition, from experts.
The standard has been revisited, but the definition has not changed, which 
means that it has stood the test of time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Verifiability} \oo{owner}}

\begin{defn}
	 Verification is the process of evaluating a system or component to
   determine whether the products of a given development phase satisfy the
   conditions imposed at the start of that phase
   \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
	Verification involves solving the equations right \citep[p.~23]{Roache1998}.
\end{defn}

\begin{defn}
  A software is verifiable if its properties can be verified easily
  \citep{GhezziEtAl2003}.  \wss{Is this a direct quote?} \oo{yes}
\end{defn}

\begin{defn}
	By verification, we mean all activities that are undertaken to ascertain
  that the software meets its objectives \citep{GhezziEtAl2003}.
\end{defn}

\begin{defn}
	Verifiability refers to how well software components or the integrated
  product can be evaluated to demonstrate whether the system functions as
  expected \citep{wiegers2003softreq}.
\end{defn}

\begin{defn} \label{Defn_Verifiability1}
  Verifiability ``means that you should be able to write a set of tests that
  can demonstrate that the delivered system meets each specified
  requirement'' \citep{sommerville}. %2011
\end{defn}

\begin{defn}
	The evaluation of whether or not a product, service, or result complies
  with a regulation, requirement, specification, or imposed condition
  \citep{project2017guide} \oo {verification definition}
\end{defn}

\noindent \textbf{Proposed Definition} Definition~\ref{Defn_Verifiability1}
rephrased: Verifiability is the extent to which a set of tests can be
written and executed, to demonstrate that the delivered system meets the
specification.

\noindent \textbf{Reasoning}

Definition is concise and measurable.  Verifiability involves solving the
equations right \citep[p.~23]{Roache1998} and the definition states that
system needs to be verifiable based on the specifications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Validatability} \oo{owner}}

\begin{defn} \label{Defn_Validatability}
  The process of evaluating a system or component during or at the end of the
  development process to determine whether it satisfies specified requirements
  \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
  Validation means solving the right equations \citep[p.~23]{Roache1998}.
\end{defn}

\begin{defn}
  Software validation is achieved through a series of tests that demonstrate
  conformity with requirements \citep{pressman2005software}.
\end{defn}

\begin{defn}
	Validation is concerned with establishing that the product fulfills its
	intended use \citep{van2008software}.
\end{defn}

\begin{defn}
  Software validation is the process of checking that the system conforms to
  its specification and that it meets the real needs of the users of the
  system \citep{sommerville}. %2011
\end{defn}

\begin{defn}
	Software validation, where the software is checked to ensure that it is
  what the customer requires \citep{sommerville}. %2011
\end{defn}

\begin{defn}
  The assurance that a product, service, or result meets the needs of the
  customer and other identified stakeholders \citep{project2017guide}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{Defn_Validatability} rephrased:  The effort required in
evaluating a system or component during or at the end of the development
process to determine whether it satisfies specified requirements

\noindent \textbf{Reasoning}

Definition describes that system should be measurable through the
development life cycle, It is concise and can be measured.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Testability} \oo{owner}}

\begin{defn}
	Code possesses the characteristic testability to the extent that it
  facilitates the establishment of verification criteria and supports
  evaluation of its performance \citep{boehm1976quantitative}
\end{defn}

\begin{defn}
	Effort required to test a program to ensure it performs
	its intended function \citep{McCallEtAl1977}.
\end{defn}

\begin{defn} \label{Defn_Testability}
	 The degree to which a system or component facilitates the establishment of
	 test criteria and the performance of tests to determine whether those
	 criteria have been met \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
	The degree to which you can unit-test and system-test a system; the degree
  to which you can verify that the system meets its requirements
  \citep{mcconnell2004code}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{Defn_Testability}: The degree to which a system or component
 facilitates the establishment of test criteria and the performance of tests
 to determine whether those criteria have been met.

\noindent \textbf{Reasoning}

Definition is concise and measurable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Maintainability} \pmi{owner}}

\begin{defn}
  Effort required to locate and fix an error in an operational program
  \citep{McCallEtAl1977}.

  \wss{It looks like \citet{pressman2005software} was using
  \citet{McCallEtAl1977} for their definition.} \wss{PM - can you please
  look into whether \citet{pressman2005software} cite
  \citet{McCallEtAl1977}?  Do they give a reason for dropping the word
  operational?}  \pmi{Yes, \citet{pressman2005software} does cite the
  relevant section of the textbook with \citet{McCallEtAl1977}. While they
  do not give a reason for specifically dropping the word operational, they
  do indicate that their given definition ``is a very limited definition''.
  Perhaps their intent was breadth.} \wss{Peter, please verify the McCall
  reference definition, and then just cite that.  Olu has found the McCall
  reference, so you shouldn't have to search for it.  We can remove both the
  VanVliet and Pressman reference in preference of the McCall definition.}

  \pmi{I have verified the McCall definition and removed the Pressman and
  van Vliet references above. I will leave these comments until all similar
  definitions in this document have been edited.}
\end{defn}

\begin{defn} \label{MaintainabilityDefnSelected1} 
  The ease with which a software system or component can be modified to correct
  faults, improve performance or other attributes, or adapt to a changed
  environment \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
  We will view maintainability as two separate qualities: repairability and
  evolvability. Software is repairable if it allows the fixing of defects;
  it is evolvable if it allows changes that enable it to satisfy new
  requirements \citep{ghezzi1991fundamentals}. \wss{update to 2003 ref?}
\end{defn}

\begin{defn}
  The capability of the software to be modified.  Modifications may include
  corrections, improvements or adaptation of the software to changes in
  environment, and in requirements and functional specifications.
  \cite{ISO9126} %2001
\end{defn}

\begin{defn}
  A set of attributes that bear on the effort needed to make specified
  modifications (which may include corrections, improvements, or adaptations of
  software to environmental changes and changes in the requirements and
  functional specifications) \citep{pfleeger2006software}.
\end{defn}

\begin{defn} \label{MaintainabilityDefnSelected2}
  Code possesses the characteristic of maintainability to the extent that it
  facilitates updating to satisfy new requirements or to correct deficiencies
  \citep{boehm2007software}.
\end{defn}

\begin{defn}
  ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
  efficiency with which a product or system can be modified by the intended
  maintainers \citep{ISO/IEC25010}. %2011
\end{defn}

\noindent \textbf{Proposed Definition}

Combined altered Definition \ref{MaintainabilityDefnSelected1} and Definition
\ref{MaintainabilityDefnSelected2}: The effort with which a software system or
component can be modified to:

\begin{enumerate}
	\item correct faults
	\item improve performance or other
	attributes
	\item satisfy new requirements
\end{enumerate}

\noindent \textbf{Reasoning}

This definition offers a comprehensive list of potential reasons for modifying
software, such as correcting faults, improving performance or other attributes,
or satisfying new requirements.

\wss{What are the other attributes?  What is an attribute?  Is it the same
as a quality?  We are calling performance a quality in this document.}

\pmi{The term quality attribute has come up often in the literature -
qualities appear to be a subset of attributes. Quality attributes seem to be
associated with non-functional requirements. The term resource attribute was
encountered in a journal and implied an association with a functional
requirement. The two might differ along those lines of NFR and FR.}

\pmi{I think this is done}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Reusability \pmi{owner}}

\begin{defn}
  Extent to which a program can be used in other applications - related to
  the packaging and scope of the functions that programs perform
  \citep{McCallEtAl1977}.
\end{defn}

\begin{defn}
  The degree to which a software module or other work product can be used in
  more than one software system \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
  A product is reusable if we can ``use the product - perhaps with minor changes
  - to build another product'' \citep{ghezzi1991fundamentals}.  \wss{Update
  to 2003?}
\end{defn}

\begin{defn} \label{ReusabilityDefnSelected}
  The extent to which a software component can be used with or without
  adaptation in a problem solution other than the one for which it was
  originally developed \citep{kalagiakos2003non}.
\end{defn}

\begin{defn}
  Reusability is the likelihood a segment of source code that can be used
  again to add new functionalities with slight or no modification
  \citep{sandhu2010survey}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{ReusabilityDefnSelected}: The extent to which a software
component can be used with or without adaptation in a problem solution other
than the one for which it was originally developed.

\noindent \textbf{Reasoning}

This definition highlights the possible but not necessary adaptation of the
software component(s) being transferred.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{{Portability} \pmi{owner}}

\begin{defn} \label{PortabilityDefnSelected}
  Effort required to transfer a program from one hardware configuration
  and/or software system environment to another \citep{McCallEtAl1977}.
\end{defn}

\begin{defn}
  The ease with which a system or component can be transferred from one
  hardware or software environment to another \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
  An application is portable across a class of environments to the degree
  that the effort required to transport and adapt it to a new environment in
  the class is less than the effort of redevelopment
  \citep{mooney1990strategies}.
\end{defn}

\begin{defn}
  Portability refers to the ability to run a system on different hardware
  platforms \citep{ghezzi1991fundamentals}.  \wss{update to 2003?}
\end{defn}

\begin{defn}
  The capability of software to be transferred from one environment to
  another \cite{ISO9126}. %2001
\end{defn}

\begin{defn}
  A set of attributes that bear on the ability of software to be transferred
  from one environment to another (including the organizational, hardware, of
  software environment) \citep{pfleeger2006software}.
\end{defn}

\begin{defn}
  Code possesses the characteristic of portability to the extent that it can
  be operated easily and well on computer configurations other than its
  current one. This implies that special function features, not easily
  available at other facilities, are not used, that standard library
  functions and subroutines are selected for universal applicability, and so
  on \citep{boehm2007software}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{PortabilityDefnSelected} rephrased: Effort required to transfer
a program between system environments (including hardware and software).

\noindent \textbf{Reasoning}

This is measurable and succinct.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Understandability} \jc{owner}}

\begin{defn}
  The effort ``to uncover the logic of the application''
  \citep{ghezzi1991fundamentals}.
\end{defn}

\begin{defn}
  The capability of the software product to enable the user to understand
  whether the software is suitable, and how it can be used for particular tasks
  and conditions of use \cite{ISO9126}. %2001
\end{defn}

Understandability is artifact-dependent.  What it means for a user-interface
(graphical or otherwise) to be understandable is wildly different than what
it means for the code, and even the user documentation.

The literature here is thin and scattered.  More work will need to be done
to find something useful.

Interestingly, the business literature seems to have taken more care to
define this.  Here we encounter:

\begin{quote}
  Understandability is the concept that X should be presented so that a reader
  can easily comprehend it.
\end{quote}

At least this brings in the idea that the \emph{reader} is actively involved,
and indirectly that the reader's knowledge may be relevant, as well as the
``clarity of exposition'' of X.

Section 11.2 of \citet{adams2015nonfunctional} does have a full definition.

\noindent \textbf{Proposed Definition}

\wss{Still needs to be completed}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Interoperability} \ad{owner}}

\begin{defn}
  Effort required to couple one system with another \citep{McCallEtAl1977}.
  \wss{Ao, if this is from McCall, we should just cite the primary source
  and remove the Van Vliet reference.} \pmi{Yes this is almost exactly from
  McCall, except he did not have the word "The" at the beginning.}
  \ad{Someone with access to the book added it. I found the book in the pub
  folder and fixed it.}
\end{defn}

\begin{defn}
  Interoperability is the ability of two or more systems or components to
  exchange information and to use the information that has been exchanged
  \citep{IEEEComputerDictionary1991}.
\end{defn}

\begin{defn}
  The ability of a system to coexist and cooperate with other systems.
  \citep{ghezzi1991fundamentals}.  \wss{2003 ref?}
\end{defn}

\begin{defn}
  The capability of the software to interact with one or more specified
  systems \cite{ISO9126}. %2001
\end{defn}

\begin{defn}
  The capability to communicate, execute programs, and transfer data among
  various functional units in a manner that requires the user to have little
  or no knowledge of the unique characteristics of those units
  \citep{ISO/IEC/IEEE24765}. %2010
\end{defn}

\begin{defn}
  \label{InteroperabilitySelected} The degree to which two or more systems,
  products or components can exchange information and use the information
  that has been exchanged \citep{ISO/IEC25010}. %2011
\end{defn}

\begin{defn}
  Interoperability is a characteristic of a product or system, whose
  interfaces are completely understood, to work with other products or
  systems, present or future, in either implementation or access, without
  any restrictions \citep{AFUL2019}.
\end{defn}

\begin{defn}
  Interoperability is the ability of different information systems, devices and
  applications (‘systems’) to access, exchange, integrate and cooperatively use
  data in a coordinated manner, within and across organizational, regional and
  national boundaries, to provide timely and seamless portability of information
  and optimize the health of individuals and populations globally. Health data
  exchange architectures, application interfaces and standards enable data to be
  accessed and shared appropriately and securely across the complete spectrum of
  care, within all applicable settings and with relevant stakeholders, including
  by the individual \citep{HIMSS2019}.

  Four Levels of Interoperability:
  \begin{itemize}
  \item Foundational (Level 1) – establishes the inter-connectivity requirements
    needed for one system or application to securely communicate data to and
    receive data from another

  \item Structural (Level 2) – defines the format, syntax, and organization of
    data exchange including at the data field level for interpretation

  \item Semantic (Level 3) – provides for common underlying models and
    codification of the data including the use of data elements with
    standardized definitions from publicly available value sets and coding
    vocabularies, providing shared understanding and meaning to the user

  \item Organizational (Level 4) – includes governance, policy, social, legal
    and organizational considerations to facilitate the secure, seamless and
    timely communication and use of data both within and between organizations,
    entities and individuals. These components enable shared consent, trust and
    integrated end-user processes and workflows
  \end{itemize}
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{InteroperabilitySelected}.

\noindent \textbf{Reasoning}

This definition is concise and also detailed enough to show the concept not
only on system, but also on products and components. It also covered the
concept with more than 2 systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Productivity} \ad{owner}}

\begin{defn}
  The best definition of the productivity of a process is
	\[\text{Productivity} = \dfrac{\text{Outputs produced by the
              process}}{\text{Inputs consumed by the process}}\]

Thus, we can improve the productivity of the software process by increasing its
outputs, decreasing its inputs, or both. However, this means that we need to
provide meaningful definitions of the inputs and outputs of the software
process. \wss{Where did this definition come from?  Is this all from Boehm?
  This is quite a long quote.  We might actually want to paraphrase.}
\ad{all from Boehm. I feel the best way is deleting everything after the
equation}

\textbf{Defining inputs.} For the software process, providing a meaningful
definition of inputs is a nontrivial but generally workable problem. Inputs to
the software process generally comprise labor, computers, supplies, and other
support facilities and equipment. However, one has to be careful which of
various classes of items are to be counted as inputs. For example:
\begin{itemize}
\item Phases (just software development, or should we include system
engineering, soft-ware requirements analysis, installation, or post development
support?)
\item Activities (to include documentation, project management, facilities
management, conversion, training, database administration?)
\item Personnel (to include secretaries, computer operators, business managers,
contract administrators, line management?)
\item Resources (to include facilities, equipment, communications, current
versus future dollar payments?)
\end{itemize}
An organization can usually reach an agreement on which of the above are
meaningful as inputs in their organizational context. Frequently, one can use
present-value dollars as a uniform scale for various classes of resources.

\textbf{Defining outputs.} The big problem in defining software productivity is
defining outputs. Here we find a defining delivered source instructions (DSI) or
lines of code as the output of the software process is totally inadequate, and
they argue that there are a number of deficiencies in using DSI. However,
	most organizations doing practical productivity measurement still use DSI as
	their primary metric \citep{Boehm1987}. \wss{Is this a direct quote from
		\citet{Boehm1987}?  The sentences seem incomplete?}
\ad{I added some deleted parts, now it is a direct quote. After the ending of
the quote, the following 2 pages discuss the flaws of DCI and a list of
alternatives to DCI, so I ended the quote here.}
\end{defn}

\begin{defn}
  A quality of the software production process, referring to its efficiency
  and performance \citep{ghezzi1991fundamentals}.  \wss{2003 ref?}
\end{defn}

\begin{defn}
	\label{ProductivitySelected2} The number of lines of new code developed per
  person-day (an imperfect measure of productivity but one that could be
  measured consistently) \citep{MacCormack2003}.
\end{defn}

\begin{defn}
	\label{ProductivitySelected1} Productivity is the amount of output (what is
  produced) per unit of input used. In general, productivity is difficult to
  measure because outputs and inputs are typically quite diverse and are
  often themselves difficult to measure. In the context of software,
  productivity measurement is usually based on a simple ratio of product
  size to project effort. Thus,If we can measure the size of the software
  product and the effort required to develop the product, we have:
	\begin{align}
	\text{productivity} = \text{size}/\text{effort}
	\end{align}
  Equation (1) assumes that size is the output of the software production
  process and effort is the input to the process. This can be contrasted
  with the viewpoint of software cost models where we use size as an
  independent variable (i.e., an input) to predict effort which is treated
  as an output. Equation (1) is simple to operationalize if we have a single
  dominant size measure, for example, product size measured in lines of code
  \citep{Kitchenham2004}.
\end{defn}

\noindent \textbf{Proposed Definition}

The revision of the combination of Definition \ref{ProductivitySelected1}
and Definition \ref{ProductivitySelected2}: Productivity is the amount of
output per unit of input used, which can be measured by the summation of all
output (such as the number of lines of new code, the number of pages of new
documents and the number of new test cases) produced per person-day.

\noindent \textbf{Reasoning}

It is concise and measurable. \wss{What is the output?  What is the input?
I think the definition needs to give more information on these.  In
particular, the above definitions focus on code as the output, but
documentation, test cases etc should also be part of the output.  If we are
going to measure this, we need a better idea of what we are measuring for
outputs and inputs.} \ad{I added another def and made it more measurable}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Visibility/Transparency} \ad{owner}}

\begin{defn}
  \label{VisibilitySelected} A software development process is visible if
  all of its steps and its current status are documented clearly. Another
  term used to characterize this property is transparency
  \citep{ghezzi1991fundamentals}. \wss{change to 2003 ref?}
\end{defn}

\begin{defn}
  Business process visibility, also called process visibility, is the ability to
  accurately and completely view the processes, transactions and other
  activities operating within an enterprise \citep{Rouse2013}.
\end{defn}

\begin{defn}
  Visibility provides transparency into the development process. It is the
  ability to see progress at any point and determine the distance to completion
  of a goal.  Visibility provides status of not only the progress of the
  project, but the product itself \citep{GSA2019}.
\end{defn}

\begin{defn}
  Process transparency refers to the ability to look inside. The “look inside”
  provides an in-depth and clear visibility into the business processes and how
  these operate \citep{PRIME2019}.
\end{defn}

\begin{defn}
The degree to which something is seen by the public
\citep{CambridgeVisibility2019}.

The degree to which something is seen or known about
\citep{CambridgeVisibility2019}.
\end{defn}

\noindent \textbf{Proposed Definition}

Definition \ref{VisibilitySelected} rephrased: The extent to which all of
the steps of a software development process and the current status of it are
conveyed clearly. \wss{I wonder if rather than ``documented clearly'', we
should be more abstract and talk about how these things can be easily
determined.  Documentation is probably how the information will be conveyed,
  but it doesn't have to be how it is done.}  \ad{Adjusted}

\noindent \textbf{Reasoning}

Definition \ref{VisibilitySelected} points out that documentation is the way to
improve visibility. It is rephrased because the original one might refer to
binary status - ``visible'' or ``invisible''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Reproducibility} \wss{owner}}

Reproducibility is a required component of the scientific method
\citep{Davison2012}.  Although QA has, ``a bad name among creative
scientists and engineers'' \citep[p.~352]{Roache1998}, the community need to
recognize that participating in QA management also improves reproducibility.
Reproducibility, like QA, benefits from a consistent and repeatable
computing environment, version control and separating code from
configuration/parameters \citep{Davison2012}.

Reproducibility is defined as:

\begin{defn}
  A result is said to be reproducible if another researcher can take the
  original code and input data, execute it, and re-obtain the same result
  (Peng, Dominici, and Zeger, 2006), as cited in
  \citet{BenureauAndRougier2017}.
\end{defn}

The related concept of replicable is defined as:

\begin{defn}
  Documentation achieves replicability if the description it provides of the
  algorithms is sufficiently precise and complete for an independent researcher
  to re-obtain the results it presents.  \citep{BenureauAndRougier2017}
\end{defn}

It would be worthwhile to look for some additional definitions.

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed}

The importance of reproducibility is seen by retractions of influential
papers, like the review of \href{https://www.theguardian
.com/commentisfree/2020/jun/05/lancet-had-to-do-one-of-the-biggest
-retractions-in-modern-history-how-could-this-happen} {hydroxychloroquine
trials}

\noindent \textbf{Reasoning}

\wss{Reproducibility literature review is given in \citet{FeinbergEtAl2020}.
Includes examples where replicability was found to be lacking.  Has a good
litereature review in Section 2.}

\wss{Needs to be completed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Sustainability} \wss{owner}}

One of the original definitions of sustainability (for systems, not software
specific), and still often quoted, is:

\begin{defn}
  \noindent The ability to meet the needs of the present without
  compromising the ability of future generations to meet their own needs
  \citep{Brundtland1987}.
\end{defn}

This is the definition used by \citet{IISD2019}.

To make it more useful, this definition is often split into three dimensions:
social, economic and environmental. \wss{cite UN paper [9] in
  \citet{PenzenstadlerAndHenning2013}} To this list Penzenstadler and Henning
(2013) have added technical sustainability \citep{PenzenstadlerAndHenning2013}.
Where technical sustainability for software is defined as:

\begin{defn}
  \noindent Technical sustainability has the central objective of long-time
  usage of systems and their adequate evolution with changing surrounding
  conditions and respective requirements \citep{PenzenstadlerAndHenning2013}.
\end{defn}

The fourth dimension of technical sustainability is also added by
\citep{WolframEtAl2017}.  Technical sustainability is the focus on the
thesis by \citet{Hygerth2016}.

\begin{defn}
  \noindent Sustainable development is a mindset (principles) and an
  accompanying set of practices that enable a team to achieve and maintain
  an optimal development pace indefinitely \citep{Tate2005}.
\end{defn}

Parnas discusses as software aging \citep{Parnas1994a}.

SCS specific definitions:

\begin{defn}
  The concept of sustainability is based on three pillars: the ecological, the
  economical and the social. This means that for a software to be sustainable,
  we must take all of its effects -- direct and indirect -- on the environment,
  the economy and the society into account. In addition, the entire life cycle
  of a software has to be considered: from planning and conception to
  programming, distribution, installation, usage and disposal \citep{Heine2017}.
\end{defn}

Software Sustainability Institute proposal:

\begin{defn}
Capacity of the software to endure
\end{defn}

\citet{Katz2016} builds on this definition.

\begin{defn}
  \noindent The capacity of the software to endure. In other words,
  sustainability means that the software will continue to be available in
  the future, on new platforms, meeting new needs \citep{Katz2016}.
\end{defn}

\href{https://collegeville.github.io/CW3S19/WorkshopResources/Presentations/4-5-Katz_data-driven-software-sustainability.pdf}
{Katz Presentation}

\href{https://www.software.ac.uk/blog/2019-09-12-making-open-source-research-software-visible-path-better-sustainability}
{Neil's blog}

Definition from Neil Chue Hong:
\begin{defn}
	Sustainable software is software which is:
	-- Easy to evolve and maintain
	-- Fulfils its intent over time
	-- Survives uncertainty
	-- Supports relevant concerns (Political, Economic, Social, Technical,
	Legal, Environmental) \citep{Katz2016}.
\end{defn}

\begin{defn}
Sustainability encompasses cost efficient maintainability and evolvability
\citep{SehestedtEtAl2014}.
\end{defn}

\citet{SehestedtEtAl2014} goes on to say that sustainability can be observed
by evaluating the four criteria of the architectural model: completeness,
consistency, correctness and clarity.

Definitions for sustainability are often built by combining other definitions.
\href{https://collegeville.github.io/CW3S19/WorkshopResources/Presentations/1-4-2019CollegevilleWorkshopWillenbring.pdf}
{Willenbring.pdf}
lists sustainability factors: extensible, interoperable, maintainable,
portable, reusable, scalable and usable.  \wss{Sounds like they are listing
almost all software qualities.  It seems that sustainability is at least in
part achieved by having high quality software.}

From
\href{https://figshare.com/articles/Defining_Sustainability_through_Developers_Eyes_Recommendations_from_an_Interview_Study/1111925/1}
{Rosado de Souza, et al.} there are two categories of software sustainability:

\begin{description}
\item[Intrinsic] Pertaining to characteristics of the software
\item[Extrinsic] Pertaining to the software development environment
\end{description}

Find paper that combines nonfunctional qualities into sustainability.

Sounds like definition of maintainability.

Paper critical of a lack of a definition \citep{VentersEtAl2014}.

Sustainability depends on the software artifacts AND the software team AND the
development process.

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

\section{Desirable Qualities of Good Specifications} \label{SecDesirableQs}

To achieve the qualities listed in Section~\ref{SecQualities}, the
documentation should achieve the qualities listed in this section.  All but
the final quality listed (abstraction), are adapted from the IEEE
recommended practise for producing good software requirements
\citep{IEEE1998}.  Abstraction means only revealing relevant details, which
in a requirements document means stating what is to be achieved, but
remaining silent on how it is to be achieved. Abstraction is an important
software development principle for dealing with complexity \citep[p
.~40]{GhezziEtAl2003}.  Correctness was in the above list, so it is not
repeated here.  \citet{SmithAndKoothoor2016} present further details on the
qualities of documentation for SCS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Completeness} \ad{owner}}

\begin{defn}
  Those attributes of the software that provide full implementation of the
  functions required. \citep{McCallEtAl1977}. \wss{Ao, if this is from McCall,
  we should just cite the primary source and remove the Van Vliet reference.}
  \pmi{McCall's definition verbatim: Those attributes of the software that
  provide full implementation of the functions required.} \ad{Originally added
  by someone else so I didn't have access to the book. But now it's available
  and I double checked, Peter got it accurately.}
\end{defn}

\begin{defn}
  \label{CompletenessSelected} A specification is complete to the extent
  that all of its parts are present and each part is fully developed. A
  software specification must exhibit several properties to assure its
  completeness \citep{Boehm1984}:
  \begin{itemize}
  \item No TBDs. TBDs are places in the specification where decisions have been
    postponed by writing "To be Determined" or "TBD."
  \item No nonexistent references. These are references in the specification to
    functions, inputs, or outputs (including databases) not defined in the
    specification.
  \item No missing specification items. These are items that should be present
    as part of the standard format of the specification, but are not present.
  \item No missing functions. These are functions that should be part of the
    software product but are not called for in the specification.
  \item No missing products. These are products that should be part of the
    delivered software but are not called for in the specification.
  \end{itemize}
\end{defn}

\begin{defn}
	An SRS is complete if, and only if, it includes the following elements:
\begin{itemize}
\item All significant requirements, whether relating to functionality,
performance, design constraints, attributes, or external interfaces. In
particular any external requirements imposed by a system specification should be
acknowledged and treated.
\item Definition of the responses of the software to all realizable classes of
input data in all realizable classes of situations. Note that it is important to
specify the responses to both valid and invalid input values.
\item Full labels and references to all figures, tables, and diagrams in the SRS
and definition of all terms and units of measure \citep{IEEE1998}.
	\end{itemize}
\end{defn}

\begin{defn}
The quality of being whole or perfect and having nothing missing.
\citep{CambridgeCompleteness2019}.
\end{defn}

\wss{Were there any other definitions of completeness?  You could add the
definition of completeness from \citet[p.\ 5--6]{IEEE1998}.  This definition
is for requirements, but maybe there is something we can generalize from the
definition?  We could also look for definitions outside of software
development.} \ad{added IEEE and dictionary} \ad{\url{https://annals-csis
.org/proceedings/2016/pliks/468.pdf} has a method of measuring completeness
and consistency, but I haven't summarized any def from it yet}

\noindent \textbf{Proposed Definition}

The first sentence of Definition \ref{CompletenessSelected}: A specification is
complete to the extent that all of its parts are present and each part is fully
developed.

\noindent \textbf{Reasoning}

It is concise and measurable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Consistency} \ad{owner}}

\begin{defn}
  Those attributes of the software that provide uniform design and
  implementation techniques and notation \citep{McCallEtAl1977}.  \wss{Ao,
    if this is from McCall, we should just cite the primary source and
    remove the Van Vliet reference.} \pmi{McCall's definition verbatim:
  Those attributes of the software that provide uniform design and
  implementation techniques and notation.}\ad{Originally added by someone
  else so I didn't have access to the book. But now it's available and I
  double checked, Peter got it accurately.}
\end{defn}

\begin{defn}
  \label{ConsistencySelected} A specification is consistent to the extent
  that its provisions do not conflict with each other or with governing
  specifications and objectives. Specifications require consistency in
  several ways \citep{Boehm1984}.
  \begin{itemize}
  \item Internal consistency. Items within the specification do not conflict
    with each other.
  \item External consistency. Items in the specification do not conflict with
    external specifications or entities.
  \item Traceability. Items in the specification have clear antecedents in
    earlier specifications or statements of system objectives.
  \end{itemize}
\end{defn}

\begin{defn}
Consistency refers to internal consistency. If an SRS does not agree with some
higher-level document, such as a system requirements specification, then it is
not correct \citep{IEEE1998}.
\end{defn}

\begin{defn}
 An SRS is internally consistent if, and only if, no subset of individual
 requirements described in it conflict. The three types of likely conflicts
 in an SRS are as follows \citep{IEEE1998}:
 \begin{enumerate}[a)]
 \item The specified characteristics of real-world
 objects may conflict. For example,
 \begin{enumerate}[1)]
 \item The format of an output report may be described
 in one requirement as tabular but in another as textual.
 \item One
 requirement may state that all lights shall be green while another may
 state that all lights shall be blue.
 \end{enumerate}
 \item There may be logical or temporal conflict between two specified actions.
 For example,
 \begin{enumerate}[1)]
 \item One requirement may specify that the program will add two inputs and
 another may specify that the program will multiply them.
 \item One requirement may state that “A” must always follow “B,” while another
 may require that “A and B” occur simultaneously.
 \end{enumerate}
 \item Two or more requirements may describe the same real-world object but use
 different terms for that object. For example, a program’s request for a user
 input may be called a “prompt” in one requirement and a “cue” in another. The
 use of standard terminology and definitions promotes consistency.
 \end{enumerate}
\end{defn}

\begin{defn}
  Consistency requires that no two or more requirements in a specification
  contradict each other. It is also often regarded as the case where words
  and terms have the same meaning throughout the requirements specifications
  (consistent use of terminology). These two views of consistency imply that
  mutually exclusive statements and clashes in terminology should be avoided
  \citep{ZOWGHI2003}.
\end{defn}

\begin{defn}
  Consistency: 1. the degree of uniformity, standardization, and freedom
  from contradiction among the documents or parts of a system or component 2
  . software attributes that provide uniform design and implementation
  techniques and notations \citep{ISO/IEC/IEEE24765}.
\end{defn}

\begin{defn}
The state or condition of always happening or behaving in the same way
\citep{CambridgeConsistency2019}.
\end{defn}

\wss{The definition from \citet{IEEE1998} might again be useful.  We could also
  look for definitions outside of software development.  Even a dictionary
  definition could be helpful.}
\ad{added IEEE and dictionary}

\noindent \textbf{Proposed Definition}

The first sentence of Definition \ref{ConsistencySelected}: A specification is
consistent to the extent that its provisions do not conflict with each other or
with governing specifications and objectives.

\noindent \textbf{Reasoning}

It is concise and measurable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Modifiability} \jc{owner}}

Here we do seem to have a simple, if somewhat uninformative, definition:

\begin{defn}
  Modifiability is the degree of ease at which changes can be made to a
  system, and the flexibility with which the system adapts to such changes.
\end{defn}

IEEE Standard 610 seems to speak about this. (which is superseded?)

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

\subsection{\notdone{Traceability} \jc{owner}}

Here the Wikipedia page~\url{https://en.wikipedia.org/wiki/Traceability} is
actually rather informative, especially as it also lists how this concept is
used in other domains.  A generic definition that is still quite useful is

\begin{defn}
  The capability (and implementation) of keeping track of a given set or
  type of information to a given degree, or the ability to chronologically
  interrelate uniquely identifiable entities in a way that is verifiable.
\end{defn}

By specializing the above to software artifacts, ``interrelate'' to ``why is
this here'' (for forward tracing from requirements), this does indeed give
what is meant in SE.

Various standards (DO178C, ISO 26262, and IEC61508) explicitly mention it.

24765-2017 - ISO/IEC/IEEE International Standard - Systems and software
engineering--Vocabulary has a full definition, namely
\begin{enumerate}
\item the degree to which a relationship can be established between two or more
  products of the development process, especially products having a
  predecessor-successor or master-subordinate relationship to one another;
\item the identification and documentation of derivation paths (upward) and
  allocation or flow-down paths (downward) of work products in the work product
  hierarchy;
\item the degree to which each element in a software development product
  establishes its reason for existing; and discernible association among two or
  more logical entities, such as requirements, system elements, verifications,
  or tasks.
\end{enumerate}

\begin{defn}
  The ability to link software components to requirements
  \citep{McCallEtAl1977}. (As summarized in \citet{VanVliet2000}.)
\end{defn} \wss{Once we have verified McCall, we can remove the Van Vliet
reference.}  \pmi{McCall's definition verbatim: Those attributes of the
software that provide a thread from the requirements to the implementation
with respect to the specific development and operational environment.}

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed.}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Unambiguity} \wss{owner}}

A specification is unambiguous when it has a unique interpretation.  If
there is a possibility that two readers will have two different
interpretations, than the specification is ambiguous.  \wss{When I get the
Ghezzi text back from Olu, I'll check to see if they have anything to add to
this definition.}

A Software Requirements Specification (SRS) is unambiguous if, and only if,
every requirement stated therein has only one interpretation \citep{IEEE1998}.

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed.}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Verifiability} \wss{owner}}

\begin{itemize}

\item Verification - Are we building the product right?  Are we implementing
the requirements correctly (internal) \item Validation - Are we building the
right product? Are we getting the right requirements (external) \item
According to  \href{https://en.wikipedia.org/wiki/Software_verification_and_validation}
{Capability Maturity Model (CMM)}
  \begin{itemize}
  \item Software Verification: The process of evaluating software to determine
    whether the products of a given development phase satisfy the conditions
    imposed at the start of that phase. [IEEE-STD-610] \wss{Need a proper
      citation}
  \item Software Validation: The process of evaluating software during or at the
    end of the development process to determine whether it satisfies specified
    requirements. [IEEE-STD-610] \wss{Need a proper citation}
  \end{itemize}
\end{itemize}

``An SRS is verifiable if, and only if, every requirement stated therein is
verifiable. A requirement is verifiable if, and only if, there exists some
finite cost-effective process with which a person or machine can check that the
software product meets the requirement. In general any ambiguous requirement is
not verifiable.'' \citep{IEEE1998}

Verifiability is related to testability, which is defined by McCall et al.\
as ``The effort required to test a program to ensure that it performs its
intended function'' \citep{VanVliet2000}. \wss{Need to verify McCall
reference, then delete VanVliet.} \pmi{McCall's definition verbatim: Effort
required to test a program to insure it performs its intended function.}

\wss{When I get the Ghezzi text back from Olu, I'll check to see if they have
  anything to add to this definition.}

\begin{defn}
  A software system is verifiable if its properties can be verified easily
  \citep{ghezzi1991fundamentals}.
\end{defn}

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\notdone{Abstract} \wss{owner}}

\begin{defn}
  Documented requirements are said to be abstract if they state what the
  software must do and the properties it must possess, but do not speak about
  how these are to be achieved \citep{GhezziEtAl2003}.
\end{defn}

\begin{defn}
  ``An abstraction for a software artifact is a succinct description that
  suppresses the details that are unimportant to a software developer and
  emphasizes the information that is important.'' \citep{Krueger1992}
\end{defn}

\begin{defn}
  ``Abstraction means that we concentrate on the essential features and ignore,
  abstract from, details that are not relevant at the level we are currently
  working.''  \citep[p.\ 296]{VanVliet2000}
\end{defn}

\begin{defn}
  ``Abstraction in mathematics is the process of extracting the underlying
  essence of a mathematical concept, removing any dependence on real world
  objects with which it might originally have been connected, and generalizing
  it so that it has wider applications or matching among other abstract
  descriptions of equivalent phenomena.''
  \href{https://en.wikipedia.org/wiki/Abstraction_(mathematics)} {Wikipedia
    Definition}
\end{defn}

Abstraction is related to reusability (and other qualities).

\wss{When I get the Ghezzi text back from Olu, I'll check to see if they have
  anything to add to this definition.}

\noindent \textbf{Proposed Definition}

\wss{Needs to be completed}

\noindent \textbf{Reasoning}

\wss{Needs to be completed}

\newpage

\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}

\end{document}