### 2021-10-06 StateOfPractice/DomainMeasurements/LBM-Peter/Commonality-Analysis-Template.tex
Rearrange domain measurement reports to have more consistent naming and locations

\documentclass[12pt]{article}

\usepackage{titling}
\usepackage{blindtext}
\usepackage{fullpage}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=blue,
    filecolor=black,
    linkcolor=red,
    urlcolor=blue
}
\newcommand{\colAwidth}{0.24\textwidth}
\newcommand{\colBwidth}{0.76\textwidth}

\newcommand{\colCwidth}{0.28\textwidth}
\newcommand{\colDwidth}{0.72\textwidth}

\newcommand{\colEwidth}{0.33\textwidth}
\newcommand{\colFwidth}{0.67\textwidth}

\newcounter{comnum} %Commonality Number
\newcommand{\cthecomnum}{C\thecomnum}
\newcommand{\cref}[1]{C\ref{#1}}

\newcounter{varnum} %Variability Number
\newcommand{\vthevarnum}{V\thevarnum}
\newcommand{\vref}[1]{V\ref{#1}}

\newcounter{parnum} %Parameter of Variation Number
\newcommand{\ptheparnum}{P\theparnum}
\newcommand{\pref}[1]{P\ref{#1}}

\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\begin{document}

\title{\bf{Commonality Analysis for Lattice Boltzmann Systems}} 
\author{Peter Michalski, \\Department of Computing and Software\\McMaster University}
\date{\today}

\begin{titlepage}
	\clearpage\maketitle
	\thispagestyle{empty}
	\begin{abstract}
		This report presents a commonality analysis for Lattice Boltzmann systems using a commonality analysis template based on \cite{SmithAndChen2004}. The document reviews both the methodology of commonality analysis and the details of Lattice Boltzmann systems. The commonality analysis itself consists of the following:
		\textit{i}) terminology and definitions; \textit{ii}) commonalities, or features that are common to all potential family members; \textit{iii}) variabilities, or features and characteristics that may vary among family members; and, \textit{iv}) parameters of variation, or the potential values that can be assigned to the variabilities. 
		The documentation of the above items for Lattice Boltzmann systems is clarified by decomposing each item into subsections on Lattice Boltzmann methods, input, output, nonfunctional requirements, and, where appropriate, system constraints.\\

		\noindent\textbf{Keywords:} Commonality analysis, Lattice Boltzmann methods, program family.

	\end{abstract}
\end{titlepage}

\pagenumbering{roman}

\newpage

\tableofcontents

\newpage

\pagenumbering{arabic}

\section{Introduction}
Mesh generating systems are well suited to development as a program family because they fit Parnas’s definition of a program family: “a set of programs whose common properties are so extensive that it is advantageous to study the common properties of the programs before analyzing individual members” \citep{parnas1976design}. Developing mesh generating systems as a program family is advantageous because mesh generators share many common features, or commonalities. Furthermore, when there are differences between systems, the variabilities between them can be systematically considered. The purpose of this document
is to record these commonalities and variabilities and show the relationships between them, and thus facilitate the specification and development of mesh generator program family members. This document will be valuable in all future phases of system development and analysis. For instance, the requirements documentation for any mesh generator will use the commonality analysis, since the requirements should refine the commonalities, which are shared requirements of all mesh generators. Moreover, the design of any future system will use this documentation to facilitate consideration of the variabilities, so that likely
changes can be made to the system with a minimal amount of work.

The scope of the commonality analysis presented here can be considered both from the point of view of mesh generating systems and from the point of view of software engineering methodologies. From the mesh generator perspective, the starting point for the current document is the commonality analysis conducted by \cite{chen2003}. However, the scope of the current document is broader than that of \cite{chen2003}, which was restricted to two-dimensional meshes and did not consider non-conformal, hybrid or mixed meshes. The current commonality analysis is intended to cover all mesh generators that are targeted at finite element applications. Meshes for cartography or other uses are not explicitly
considered, although much of the information in this document does overlap with other mesh uses. With respect to software engineering methodologies, the scope of the current report is restricted to informal methods, with the intention that the informal requirements will form a starting point for later development and refinement by formal methods.

The first section below provides an overview of the program family of mesh generators, by reviewing what is involved in a commonality analysis and the basics of mesh generating systems. After this, the basic terminology and definitions necessary for understanding the remainder of the document are provided. The definitions include terms used in describing a commonality analysis and terms that are used in defining the characteristics and properties of meshes. The next three sections consist of the lists of commonalities, variabilities and parameters of variation, respectively. These three sections form the heart of the documentation and include an extensive set of cross-references to demonstrate the relationships between the different items. The final section provides information on unresolved issues.

\newpage
\section{Overview}
This section provides both an overview of the process of commonality analysis and of mesh generation. The subsection on commonality analysis briefly introduces this topic, along with references that can be searched for further information. The subsection on mesh generators outlines the following: \textit{i}) the basics of mesh generation, \textit{ii}) the scope of the current analysis, \textit{iii}) the overall philosophy that has been adopted for the commonality analysis, and \textit{iv}) the arguments in favour of the development of mesh generators as a program family.

\subsection{Commonality Analysis}
In some situations it is advantageous to develop a collection of related software products as a program family. The idea is that if the software products are similar enough, then it should be possible to predict what the products have in common, what differs between them and then reuse the common aspects and thus support rapid development of the family members. The idea of program families was introduced by \cite{dijkstra1972} and later investigated by \cite{parnas1976design,parnas1979designing}. More recently, \cite{weiss1997defining, weiss1998commonality, ardis1997defining} has considered the concept of a program family in the context of what he terms Family oriented Abstraction, Specification and Translation (FAST) \citep{cuka1997specifying,weiss1999software}.

In the approach advocated by Weiss, the first step is a commonality analysis. This analysis consists of systematically identifying and documenting the commonalities that all program family members share, the variabilities between family members and the terminology used in describing the family. A commonality analysis provides a systematic way of gaining confidence that a family is worth building and of deciding what the family members will be. A commonality analysis document provides the following benefits \cite{weiss1997defining, weiss1998commonality}:

\begin{enumerate}
	\item A starting point for the design of a domain specific languages (DSL): Once a DSL, or application modelling language, is developed the program family members can be rapidly generated by specifying a given family member using the language.
	\item A basis for a common design for all family members: When the software engineers come to designing the individual family members, they can take advantage of the commonalities to reuse code. Moreover, the variabilities can be considered in the design so that they can be easily accommodated. One approach to the design may be to decompose the system into components that can each be customized by specification of values for its various parameters, where the parameters correspond to the parameters of variation identified in the commonality analysis document.
	\item A historical reference: This document records the important issues concerning the scope and the nature of the family (as well as some unsolved issues) to facilitate the involvement of the participants in maintaining and evolving the family.
	\item A basis for re-engineering a domain: Existing projects may not have been developed using software engineering methodologies. The projects can be systematically reorganized and redesigned with the aid of a commonality analysis to unify the existing products.
	\item A basic training reference for new software developers: This document provides the necessary basic information for a new team-member to understand the family. 
\end{enumerate}

The next section will show that the above uses of the commonality analysis document will be beneficial for the development of a program family of mesh generators. The commonality analysis will benefit all subsequent stages in the software development process. For instance, as mentioned in the introduction, the commonalities will act as requirements that will be the starting point for writing a software requirements specification. The commonalities will be refined into specific requirements by fixing the value of their associated variabilities. The change in the values of the variabilities then corresponds to the change from one program family member to another.

As commonalities and variabilities are requirements, they should express “What” functionalities and qualities the system should have, and not mention “How” these requirements are to be accomplished. That is, the commonalities and variabilities should not involve design decisions. The design decisions will be made after the requirements for a family member have been specified. The one exception to this is system constraints, which are requirements that explicitly make design decisions.

Besides the “What” versus “How” test, there are other tests that can be used to review commonalities and variabilities, as proposed by \cite{weiss1997defining}. One such test is the “what is ruled out” test. This test determines if a commonality or variability actually makes a decision because if no alternatives are ruled out then no decision has really been made. Another test is the “negation” test. If the negation of a decision represents a position that someone might argue for, then the original decision is likely to be meaningful. For instance, the statement that “the software should be reliable” has a negation that no one would likely argue for and thus the statement does not represent a good characterization of a goal for the system.

In \cite{weiss1997defining} the stages of a commonality analysis are described in a systematic way. The stages include the following: prepare, plan, analyze, quantify and review. The stages are completed through the aid of a moderator and a series of meeting and preliminary documents and documentation reviews. Although the systematic approach advocated by Weiss has its advantages, for the case of writing a commonality analysis document for mesh generating systems it was decided that a less structured approach is feasible. Mesh generators are simpler than other software systems in the sense that they have fewer interactions with the environment. Also, the theory of mesh generation has a solid mathematical basis that can be used to remove some of the ambiguity that Weiss’s approach is aimed at reducing. Therefore, the approach adopted here is to revise the original commonality analysis document produced by \cite{chen2003} and then make the new document available to others for review. The new document will be maintained in a concurrent versioning system repository so that multiple authors can work on it, and more importantly, so that the documentation’s revision history can be tracked and the documentation can be rolled back to an earlier version if necessary.

\subsection{Lattice Boltzmann Systems}
\newpage
\section{Terminology and Definitions}
This section is divided into two subsections. The first discusses the terminology that comes from the software engineering field, while the second presents the definitions used in mesh generation. Common acronyms are also listed in this section. The lists are not intended to be read sequentially, but rather to be consulted for reference purposes; therefore, the terms are ordered alphabetically, with the consequence that some terms that appear early in the list depend on the definitions of later terms.

\subsection{Software Engineering Related Definitions and Acronyms}
\noindent\textbf{Commonality}: A requirement or goal common to all family members.\\

\noindent\textbf{Goal}: “Goals capture, at different levels of abstraction, the various objectives the system under consideration should achieve.” \cite{van2001goal}\\

\noindent\textbf{Program Family}: A set of programs that are analyzed and designed together starting from the initial stages of the software development life-cycle.\\

\noindent\textbf{Requirements}: A software requirement is: \textit{i}) a condition or capability needed by a user to solve a problem or achieve an objective; \textit{ii}) a condition or capability that must be met or possessed by a system or system component to satisfy a contract, standard, specification, or other formally imposed document; or, \textit{iii}) a documented representation of a condition or capability as in the above two definitions. \cite{thayer2000ieee}\\

\noindent\textbf{Variability}: A requirement or goal that varies between family members.

\subsection{Lattice Boltzmann Related Definitions and Acronyms}
\newpage
\section{Commonalities}\label{comm_sec}

This section lists all the common features among all the potential family members. The commonalities are organized using the following abstraction of the system, which can be used to describe all mesh generators: input information, then generate the mesh and finally output the results. Section \ref{comm_lbm} describes the commonalities for the mesh generation step, which includes the discretization of the domain, as well as other information on the problem such as the boundary conditions, material properties, \textit{etc}. Section \ref{comm_in} highlights the input information that is required for all mesh generators, such as the geometry of the domain that is going to be discretized. The next section, Section \ref{comm_out}, shows the common features for the output of mesh generators, such as the requirement that mesh information be written to files. (Although the mesh information could simply be written to the computer’s memory, in all practical applications it is desirable to have a persistent record of the mesh that was created.) The final section covers qualities of the system that cannot be classified as input, mesh generation or output. These commonalities are termed nonfunctional requirements of the system. For instance, all systems will have the goal that the response time to a user’s request is small enough to allow the user to focus on his/her problem and to maintain his/her train of thought, without being distracted by excessive waiting time. The commonality in this case is refined by a later variability because the specific requirement on the response time will depend on the intended usage of the mesh generating system.

Each commonality below uses the same structure. All of the commonalities are assigned a unique item number, which takes the form of a natural number with the prefix ``C''. Following this, a description of the commonality is provided along with a list of related variabilities, which are given as hyperlinks that allow navigation of the document to the text describing the variability. Finally, each commonality ends with a summary of the history, including the date of creation and any dates of modification, along with a brief description of the modification. If necessary, a previous version of the document can be obtained by using the concurrent versioning system where the files are stored.

The commonalities listed in this section are for a mesh generator that is acting as a preprocessor for a finite element analysis program, which means that the mesh generator will produce data files with information on the discretization of the domain, the boundary conditions, the material properties, the system properties, etc. In some cases the mesh generating systems only focus on the discretization of the domain and leave it to another system to provide the other information necessary for the finite element analysis. For systems that only focus on the discretization of the domain the following subset of the commonalities applies: .

\subsection{Lattice Boltzmann Method Solvers}\label{comm_lbm}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Input}\label{comm_in}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Output}\label{comm_out}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Nonfunctional Requirements}\label{comm_nfr}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline

\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}
		\hline
		\bf Item Number& C\refstepcounter{comnum}\thecomnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colAwidth} | p{\colBwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Variability & \vref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\newpage
\section{Variabilities}\label{var_sec}
This section provides a list of characteristics that may vary among family members. As in Section \ref{comm_sec}, the first three subsections on variabilities are organized into the following sublists: Mesh Generation, Input and Output. The final two subsections list variabilities that can be characterized as system constraints and as nonfunctional requirements.

As for the commonalities, each variability is labelled with a unique item number. In this case the numbers are prepended with the letter ``V''. The other four headings provided for each variability are: Description, Related Commonality, Related Parameter and History. The related commonalities and parameters are given as a set of identifiers that respectively refer back to the previous section on commonalities or refer forward to the next section on parameters of variation.

\subsection{Lattice Boltzmann Method Solvers}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Input}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Output}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{System Constraints}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Nonfunctional Requirements}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}
		\hline
		\bf Item Number& V\refstepcounter{varnum}\thevarnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colCwidth} | p{\colDwidth}|}		
		\hline
		\bf Description & \\
		\hline
		\bf Related Commonality & \cref{}\\
		\hline
		\bf Related Parameter & \pref{}\\
		\hline
		\bf History & \\
		\hline
	\end{tabular}
\end{minipage}\\
\newpage
\section{Parameters of Variation}
This section specifies the parameters of variation for the variabilities listed in Section \ref{var_sec}. They are organized into the same five subcategories as employed previously: Mesh Generation, Input, Output, System Constraints, Nonfunctional Requirements.

Each parameter of variation is given a unique identifier of the form “P” followed by a natural number. The corresponding variability is listed and a hyperlink is provided that allows navigation back to the appropriate item in Section \ref{var_sec}. The final entry for each parameter of variation is the binding time, which is the time in the software lifecycle when the variability is fixed. The binding time could be during specification, or during building
of the system (compile time), or during execution of the system (run time). It is possible to have a mixture of binding times. For instance, a parameter of variation could have a binding time of “specification or building” to represent that the parameter could be set at specification time, or it could be postponed until the given family member is built. The choice of postponing the decision until the build would be associated with the presence of a domain specific language that would allow postponing decisions on the values of the parameter of variation.

\subsection{Lattice Boltzmann Method Solvers}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Input}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Output}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{System Constraints}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
\subsection{Nonfunctional Requirements}
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
~\newline\\
~\newline
\noindent
\begin{minipage}{\textwidth}
	\renewcommand*{\arraystretch}{1.5}
	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}
		\hline
		\bf Item Number& P\refstepcounter{parnum}\theparnum \\
		\hline
	\end{tabular}\\

	\begin{tabular}{| p{\colEwidth} | p{\colFwidth}|}		
		\hline
		\bf Corresponding Variability & \vref{}\\
		\hline
		\bf Range of Parameters & \\
		\hline
		\bf Binding Time & \\
		\hline
	\end{tabular}
\end{minipage}\\
\newpage
\section{Issues}
\newpage
\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}
\newpage
\section{Appendix}
\end{document}