### 2021-10-06 StateOfPractice/DomainMeasurements/MedImaging-Ao/Empirical Measures Template.tex

\documentclass[12pt]{article}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xspace}
\usepackage{fullpage}
\usepackage{hyperref}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\def\CC{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}\xspace}

\begin{document}

\title{Empirical Measures Template} 
\author{Ao Dong}
\date{\today}

\maketitle

The tools for Empirical Measures:
\begin{itemize}
\item \href{https://github.com/tomgi/git_stats}{git\_stats}
\item \href{https://github.com/boyter/scc}{scc}
\end{itemize}

\ad{For now it only contains raw metrics. We'll have to consider how to use them to calculate more meaningful metrics.}

\begin{longtable}{p{16cm}}
  \caption{Grading Template}   \label{table:TemplateFull}\\
  \toprule
  \textbf{Raw Metrics} (Measured via git\_stats)\\
  \midrule
  Number of text-based files. (number)\\
  Number of binary files. (number)\\
  Number of total lines in text-based files. (number)\\
  Number of total lines added to text-based files. (number)\\
  Number of total lines deleted from text-based files. (number)\\
  Number of total commits. (number)\\
Numbers of commits by year in the last 5 years. (Count from as early as possible
if the project is younger than 5 years.) (list of numbers)\\
  Numbers of commits by month in the last 12 months. (list of numbers)\\

  \midrule
  \textbf{Raw Metrics} (Measured via scc)\\
  \midrule
  Number of text-based files. (number)\\
  Number of total lines in text-based files. (number)\\
  Number of code lines in text-based files. (number)\\
  Number of comment lines in text-based files. (number)\\
  Number of blank lines in text-based files. (number)\\

  \midrule
  \textbf{Raw Metrics} (Measured via GitHub API)\\
  \midrule
Numbers of commits, added lines, and deleted lines by author by week.
(2-dimensional list of numbers) \ad{I know we're not likely to measure this.
Just a note that we can get this data. It's only available for GitHub repos, and
I have to test tools for parsing JSON data (e.g. to Excel table).}\\
  \bottomrule

\end{longtable}

\end{document}