### 2021-10-20 StateOfPractice/Methodology/Methodology.tex
Updates to Methodology document

\documentclass[letterpaper,cleveref]{lipics-v2019}

\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{paralist}
\usepackage{totcount}
\usepackage{booktabs}

\hypersetup{
colorlinks=true,       % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,       % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan           % color of external links
}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\oo}[1]{\authornote{magenta}{OO}{#1}} %Olu Owojaiye
\newcommand{\pmi}[1]{\authornote{green}{PM}{#1}} %Peter Michalski
\newcommand{\ad}[1]{\authornote{brown}{AD}{#1}} %Ao Dong

\newcommand{\notdone}[1]{\textcolor{red}{#1}}
\newcommand{\done}[1]{\textcolor{black}{#1}}

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm

\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf
+}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}}

\newcounter{totHours} %to track the estimate of total hours to assess SOP
\setcounter{totHours}{0}
\regtotcounter{totHours}

\title{Methodology for Assessing the State of the Practice for Domain X} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
Olu Owojaiye and Peter Michalski and Ao Dong}

\ccsdesc{Software and its engineering}
\ccsdesc{Software and its engineering~Software product lines}
\ccsdesc{General and reference~Empirical studies}

\keywords{software quality, domain analysis, scientific computing software,
  research software, empirical measures, analytic hierarchy process}

\date{\today}

\hideLIPIcs
\nolinenumbers

\begin{document}

\maketitle

\begin{abstract}
	To improve software development methods and tools for research software, we
	first need to understand the current state of the practice.  Therefore, we
	have developed a methodology for assessing the state of the software
	development practices for a given research software domain.  The methodology
	is applied to one domain at a time in recognition that software development in
	different domains is likely to have adopted different best practices.
	Moreover, providing a means to measure different domains facilitates
	comparison of development practices between domains.  For each domain we wish
	to answer questions such as: 
  \begin{inparaenum}[i)]
    \item What artifacts (documents, code, test cases, etc.) are present?
    \item What tools are used?
    \item What principles, process and methodologies are used?
    \item What are the pain points for developers?
    \item What actions are used to improve qualities like maintainability and
    reproducibility?
  \end{inparaenum} 
  To answer these questions, our methodology prescribes the following steps: 
  \begin{inparaenum}[i)] 
    \item Identify the domain;
    \item Identify a list of candidate software packages;
    \item Filter the list to a length of about 30 packages;
    \item Gather source code and documentation for each package;
    \item Collect raw data on each software repo, like number of stars, number
    of open issues, number of lines of code, and so on;
    \item Fill in the measurement template (the template consists of 108 questions
    to assess 9 qualities (like installability, usability and visibility));
    \item Interview developers (the interview consists of 20 questions and takes
    about an hour);
    \item Rank the software using the Analytic Hierarchy Process (AHP); and,
    \item Analyze the data to answer the questions posed above.
  \end{inparaenum}
  A domain expert should be engaged throughout the process, to ensure that
  implicit information about the domain is properly represented and to assist
  with conducing an analysis of the commonalities and variabilities between the
  30 selected packages.  Using our methodology, spreadsheet templates and AHP
  tool, we estimate (based on our experience with using the process) the time to
  complete an assessment for a given domain at \total{totHours} person hours.
\end{abstract}

~\newpage

\tableofcontents

~\newpage

\section{Introduction} \label{SecIntroduction}

Research software uses computing to simulate mathematical models of real world
systems so that we can better understand and predict those systems's behaviour.  A
small set of examples of important research software includes the
following: designing new automotive parts, analyzing the flow of blood in the
body, and determining the concentration of a pollutant released into the
groundwater.  As these examples illustrate, research software can be used for
tackling important problems that impact such areas as manufacturing, financial
planning, environmental policy, and the health, welfare and safety of
communities.

Given the importance of research software, scientists and engineers are pushing
for methods and tools to sustainably develop high quality software.  This is
evident from the existence of such groups as the Software Sustainability
Institute (\href{https://www.software.ac.uk/} {SEI}) and Better Scientific
Software (\href{https://bssw.io/} {BSS}).  Sustainability promoting groups such
as these are necessary because unfortunately the current state-of-the-practice
for research software often does not incorporate state-of-the-art Software
Engineering (SE) tools and methods \citep{JohansonAndHasselbring2018}.  The lack
of SE tools and methods contributes to sustainability and reliability problems
\citep{FaulkEtAl2009}.  Problems with the current state of the practice are
evident from embarrassing failures, like a retraction of derived molecular
protein structures \citep{Miller2006}, false reproduction of sonoluminescent
fusion \citep{PostAndVotta2005}, and fixing and then reintroducing the same
error in a large code base three times in 20 year
\citep{MilewiczAndRaybourn2018}.  To improve this situation, we need to first
fully understand the current state of the practice for research software
development.

The purpose of our proposed methodology is to understand how software quality is
impacted by the software development principles, processes and tools currently
used within research software communities.  Since research software is so broad
a category, we will reduce the scope of our methodology to focus on one specific
research software domain at a time.  To emphasize that the method is generic, we
will label the specific domain as X within this document.

This ``state of the practice for domain X'' exercise builds off of prior work on
measuring/assessing the state of software development practice in several
research domains.  We have updated the work that was done previously for domains
such as Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh
Generators \citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  Initial
tests of the new methodology have been done for medical image analysis software
\citep{Dong2021} and for Lattice Boltzmann Method (LBM) software
\citep{Michalski2021}.

In the previous ``state of the practice'' project, we measured 30 software
projects for each domain, but the measures were relatively shallow.  With this
re-boot we still target about 30 software examples from each domain, but we are
now collecting more data.  In keeping with the previous project, we still have
the constraint that the work load for applying the methodology to a given domain
needs to be feasible for a team as small as one individual, and for a time that
is short, ideally under a person month per domain.\footnote{A person month is
considered to be $20$ working days ($4$ weeks in a month, with $5$ days of work
per week) at $8$ person hours per day, or $20 \cdot 8 = 160$ person hours.}

To begin our re-boot of the previous methodology we critically assessed, and
subsequently modified, our
\href{https://github.com/adamlazz/DomainX/blob/master/TemplateToGradeSCSoft.pdf}
{previous set of measures}.  In addition, the following data has been added to
the new methodology:

\begin{itemize}
\item Characterization of the functionality provided by the software in the
  domain via a commonality analysis.
\item Project repository related data, such as the number of files, number of
  lines of code, percentage of issues that are closed, etc. 
\item Interviews with software developers in domain X.
%\item Usability testing for a sample of software in the domain.  
\end{itemize}

Unlike the previous measurement process, the new methodology involves and
engages a domain expert partner throughout.  We did not previously engage the
domain expert with the rationale that we wished to eliminate potential bias.
However, this advantage is not worth the inability to evaluate the functionality
of the software.  Moreover, not having an expert makes navigating the on-line
resources difficult, since some information will be implicit to domain experts,
and not every statement found on-line will be accurate   The importance of the
domain expert is particularly noteworthy when it comes time to publication and
dissemination of the state of the practice assessment.  Throughout this document
the person (or persons) that provide domaine expertise will be designated as the
\textit{Domain Expert}.

In the proposed methodology, the collected data is combined to rank the software
within each domain using the Analytic Hierarchy Process (AHP) \citep{Saaty1980}.
As in the previous measurement exercise, we use AHP to develop a list of
software ranked by quality.  However, in the new process we do not stop with
this list.  The Domain Expert is consulted to verify the ordering, and to
discuss the decisions that led to the ranking.  The AHP process is used to
facilitate a conversation with the Domain Expert as a means to deepen our
understanding of the software in the domain, and the needs of typical
developers.

So that the collected data for a given domain can benefit the scientific
community, our recommendation is that all collected data be made public.  For
instance, the data collection for each domain could be put on a GitHub
repository.  In addition to the project record left on GitHub, the final data
can be exported to \href{https://data.mendeley.com/}{Mendeley Data}.  As an
example, the measurements for the
\href{https://data.mendeley.com/datasets/6kprpvv7r7/1} {state of practice on GIS
software} are available on Mendeley.  Ideally the full analysis of the state of
the practice for domain X will be published in a suitable journal.

The scope of this methodology includes observations on product, artifact and
process quality for research softwares. We leave the assessment of research
software using performance benchmarks to other projects, such as the work of
\cite{kaagstrom1998gemm}.  Currently we are also leaving experiments to measure
usability and modifiability as future work, as discussed in
Section~\ref{SecFutureWork}.

Within this document the following typographic conventions have been adopted:
\begin{inparaenum}[i)] 
  \item \textcolor{red}{Red} text denotes a link to sections of this document. 
  \item \textcolor{cyan}{Cyan} text denotes a URL link.
  \item \textcolor{blue}{Blue} text denotes a citation.
\end{inparaenum}

The full methodology is presented in the sections that follow.
Section~\ref{SecResearchQuestions} highlights the research questions that are to
be answered for each measured domain. These questions are answered by the data
collected using the process outlined in Section~\ref{StepsAQDS}.  The major
steps in the process are outlined in
Sections~\ref{SecIdentifyDomain}--\ref{SecAHP}.  Following this, the time
required for assessing a single domain is given in
Section~\ref{SecEstTimeRequired}.

\section{Research Questions} \label{SecResearchQuestions}

The following are the research questions that we wish to answer for each of our
selected domains.  In addition to answering the questions for each domain, we
also wish to combine the data from multiple domains to answer these questions
for research software in general.

\begin{enumerate}
\item What artifacts are present in current software packages? 
\item What tools (development, dependencies, project management) are used by
current software packages?
\item What principles, processes, and methodologies are used in the development
  of current software packages?
\item What are the pain points for developers working on research software
  projects?  What aspects of the existing processes, methodologies and tools do
  they consider as potentially needing improvement?  How should processes,
  methodologies and tools be changed to improve software development and
  software quality?
\item For research software developers, what specific actions are taken to
  address the following:
\begin{enumerate}
\item usability
\item traceability
\item modifiability
\item maintainability
\item correctness
\item understandability
\item unambiguity
\item reproducibility
\item visibility/transparency
\end{enumerate} 
\item How does software designated as high quality by this methodology compare
  with top rated software by the community?
\end{enumerate}

\section{Overview of Steps in Assessing Quality of the Domain Software}
\label{StepsAQDS}

To answer the above research questions (Section~\ref{SecResearchQuestions}), we
systematically measure the quality of the software through data collection and a
series of experiments.  An overview of the measurement process is given in the
following steps, starting from determining a domain that is suitable for
measurement:

\begin{enumerate}
\item Identify the domain (X). (Section~\ref{SecIdentifyDomain})
\item Ask the Domain Expert to create a top ten list of software packages in the
  domain. (Section~\ref{SecDomainExpert})
\item Meet with the Domain Expert to brief them on the overall objective,
  research proposal, research questions, measurement template, and developer
  interviews. (Section~\ref{SecDomainExpert}) \label{StepMeeting}
\item Identify broad list of candidate software packages in the domain.
  (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list.
  (Section~\ref{SecInitialFilter})
\item Review software list with Domain Expert. (Section~\ref{SecDomainExpert})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item Ask Domain Expert to vet domain analysis.
  (Section~\ref{SecDomainExpert})
\item Gather source code and documentation for each prospective software
  package.
\item Collect repository based information. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecMeasureTemplate})
\item Survey developers (Section~\ref{SecSurveyQuestions})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
% \item Identify a short list of top software packages, typically four to six, for
%   deeper exploration according to the AHP rankings of the measurements.
\item Ask Domain Expert to vet AHP ranking. (Section~\ref{SecDomainExpert})
%\item Rank short list. (Section~\ref{SecRankShortList})
\item Answer research questions (from Section~\ref{SecResearchQuestions}) and
document answers.
\end{enumerate}

%\wss{The domain expert is involved in multiple steps in the process.  How best
%  to get their feedback?  The domain experts are busy and are unlikely to devote
%  significant time to the project.  We need to quickly get to the point.  Maybe
%  something around task based inspection?  Directed interview?}

\section{How to Identify the Domain?} \label{SecIdentifyDomain}

A domain of research software must be identified. Research software is defined
in this exercise as ``software that is used to generate, process or analyze
results that [are intended] to appear in a publication'' \citep{hettrick2014uk}
in a scientific or engineering context.  Research software is a more general
term for what is often called scientific computing.  To be applicable for the
methodology described in this document, the chosen domain must have the
following properties:

\begin{enumerate}
\item The domain must have well-defined and stable theoretical underpinning.  A
  good sign of this is the existence of standard textbooks, preferably
  understandable by an upper year undergraduate student.
\item There must be a community of people studying the domain.
\item The software packages must have open source options.
\item A preliminary search, or discussion with experts, suggests that there will
  be numerous, at least 15, candidate software packages in the domain that
  qualify as `research software'.
\end{enumerate}

Some examples of domains that fit these criteria are finite element analysis
\citep{szabo1996finite}, quantum chemistry \citep{veryazov20042molcas},
seismology \citep{SmithEtAl2018}, as well as mesh generators
\citep{smith2016state}.

\section{Interaction With Domain Expert} \label{SecDomainExpert}

As mentioned in the introduction (Section~\ref{SecIntroduction}), the Domain
Expert is an important member of the state of the practice assessment team.
Pitfalls exist if non-domain experts attempt to acquire an authoritative list of
software, or perform a commonality analysis or try to rank the software. The
main source of problems for non-domain experts is that they can only rely on
information that is available on-line, but on-line data has two potential
problems:
\begin{inparaenum}[i)]
  \item the on-line resources could have false or inaccurate information; and,
  \item the on-line resources could leave out relevant information that is so
in-grained with experts that nobody thinks to explicitly record the information.
\end{inparaenum}

Domain experts may be recruited from academia or industry.  The only
requirements are knowledge of the domain and a willingness to be engaged in the
assessment process.  The domain expert does not have to be a software developer,
but they should be a user of domain software.  Given that the domain experts are
likely to be busy people, the measurement process needs to benefit from their
expertise, without putting too much of a burden on their time.

In advance of the first meeting with the Domain Expert (Step~\ref{StepsAQDS} in
Section~\ref{StepsAQDS}) the expert is asked to create a top ten list of
software packages in the domain.  This is done to help the expert get in the
right mind set in advance of the first meeting.  Moreover, by doing the exercise
in advance, we avoid the potential pitfall of the expert approving the
discovered list of software without giving it adequate thought.  The emphasis
during the first meeting is for the Domain Expert to learn what is expected of
them.  The discussion should also cover avenues for publication and
dissemination.

The Domain Experts provide are asked to vet the collected data and analysis.  In
particular, they are asked to vet the proposed list of software packages, the
domain analysis and the AHP ranking.  These interactions can be done either
electronically or with in-person (or virtual) meetings.

\section{How to Identify Candidate Software?} \label{SecIdentifyCandSoft}

Once the domain of interest is identified, the candidate software for measuring
can be found through search engine queries targeting authoritative lists of
software.  Potential places to search include \href{https://github.com/}
{GitHub}, \href{https://swmath.org/} {swMATH} and domain related publications,
such as review articles. Domain Experts are also asked for their suggestions and
are asked to review the initial draft of the software list.

When forming the list and reviewing the candidate software the following
properties should be considered:

\begin{enumerate}
\item The software functionality must fall within the identified domain.
\item The source code must be viewable.
\item The empirical measures listed in Section \ref{SecEmpiricalMeasures} should
  ideally be available, which implies a preference for GitHub-style
  repositories.
\item The software cannot be marked as incomplete, or in an initial development
  phase.
\end{enumerate}

\section{How to Initially Filter the Software List?} \label{SecInitialFilter}

If the list of software is too long (over around 30 packages), then steps need
to be taken to create a more manageable list. To reduce the length of the list,
the following filters are applied.  The filters are applied in the priority
order listed, with the filtering process stopped once the list size is
manageable.

\begin{enumerate}
\item Scope: Software is removed by narrowing what functionality is considered
  to be within the scope of the domain.
\item Usage: Software packages are eliminated if their installation procedure is
  not clear and easy to follow.
\item Age: The older software packages (age being measured by the last date when
  a change was made) are eliminated, except in the cases where an older software
  package appears to be highly recommended and currently in use.
\end{enumerate}

Copies of both the initial and filtered lists, along with the rationale for
shortening the list, should be kept for traceability purposes.

\section{Domain Analysis} \label{SecDomainAnalysis}

Since each domain we will study will have a reasonably small scope, we will be
able to view the software as constituting a program family.  The concept of a
program family is defined by \citet{parnas1976design} as ``a set of programs
whose common properties are so extensive that it is advantageous to study the
common properties of the programs before analyzing individual members''.
Studying the common properties within a family of related programs is termed a
domain analysis.

The domain analysis consists of a commonality analysis of the family of software
packages. Its purpose is to show the relationships between these packages, and
to facilitate an understanding of the informal specification and development of
them. \cite{weiss1997defining} defines commonality analysis as an approach to
defining a family by identifying commonalities, variabilities, and common
terminology for the family. Commonalities are goals, theories, models,
definitions and assumptions that are common between family members.
Variabilities are goals, theories, models, definitions and assumptions that
differ between family members. Associated with each variability are its
parameters of variation, which summarize the possible values for that
variability, along with their potential binding time.  The binding time is when
the value of the variability is set.  It could be set as specification time,
build time (when the program is being compiled) or run time (when the code is
executing).

The final result of the domain analysis will be tables of commonalities,
variabilities, and parameters of variation of a program family.
\cite{smith2008commonality} present a template for conducting a commonality
analysis, which was referred to when conducting this work.
\cite{weiss1998commonality} describes another commonality analysis technique for
deciding the members of a program family. \cite{SmithAndChen2004} and
\cite{SmithMcCutchanAndCarette2017} are examples of a commonality analysis for a
family of mesh generating software and a family of material models,
respectively. The steps to produce a commonality analysis are:

\begin{enumerate}
\item Write an Introduction
\item Write an Overview of the Domain Knowledge
\item List Commonalities
\item List Variabilities
\item List Parameters of Variation
\item Add Terminology, Definitions, Acronyms
\end{enumerate}

% A sample commonality analysis for Lattice Boltzmann Solvers can be found
% \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Peter-Notes/Commonality-Analysis-LB-Systems.pdf}
% {here}.

\section{Repository Based Measures} \label{SecEmpiricalMeasures}

Some quality measurements rely on gathering raw and processed data from software
repositories. We focus on data that is reasonably easy to collect, which we
combine and analyze. The measures that are collected relate to the research
questions (Section~\ref{SecResearchQuestions}). For instance, we collect data to
see how large a project is, to ascertain a project’s popularity, and to
determine whether to project is being actively developed.

Section~\ref{rawdata} lists the raw data that is collected.  Some of this data
can be observed from GitHub repository metrics. The rest can be collected using
freeware tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to
measure the number of binary files as well as the number of added and deleted
lines in a repository. This tool is also used to measure the number of commits
over different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc
and Code (scc)} is used to measure the number of text based files as well as the
number of total, code, comment, and blank lines in a repository. These tools
were selected due to their installability, usability, and ability to gather the
empirical measures listed below. Details on installing and running the tools can
be found in Appendix~\ref{SecRepoTools}.  Section \ref{processeddata} introduces
the required processed data, which is calculated using the raw data.

\subsection{Raw Data}\label{rawdata}

The following raw data measures are extracted from repositories:

\begin{itemize}
\item Number of stars.
\item Number of forks.
\item Number of people watching the repository.
\item Number of open pull requests.
\item Number of closed pull requests.	
\item Number of developers.	
\item Number of open issues.
\item Number of closed issues.
\item Initial release date.
\item Last commit date.
\item Programming languages used.
\item Number of text-based files.
\item Number of total lines in text-based files.
\item Number of code lines in text-based files.
\item Number of comment lines in text-based files.
\item Number of blank lines in text-based files.
\item Number of binary files.  
\item Number of total lines added to text-based files.
\item Number of total lines deleted from text-based files.
\item Number of total commits.
\item Numbers of commits by year in the last 5 years. (Count from as early as
  possible if the project is younger than 5 years.)
\item Numbers of commits by month in the last 12 months.
\end{itemize}

\subsection{Processed Data}\label{processeddata}

The following measures are calculated from the raw data:

\begin{itemize}
\item Status of software package as either dead or alive, where alive is defined
  as the presence of repository commits or software package version releases in
  the last 18 months.
\item Percentage of identified issues that are closed.
\item Percentage of code that is comments.
\end{itemize}

\section{Measure Using Measurement Template} \label{SecMeasureTemplate}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=1.0\textwidth]{measurement_template}
    \caption{Top of Measurement Template}
    \label{measurement_template_image}
  \end{center}
\end{figure}

The Measurement Template is found in
Appendix~\ref{SecGradingTemplate}.  This template is used to track measurements
and quality scores for all of the software packages in the domain. For each
software package, we fill out one column of the template. This process can take
between 2 to 5 hours for each package.  Project developers can be contacted for
help regarding installation, if necessary, but a cap of about 2 hours should be
imposed on the entire installation process, to keep the overall measurement time
feasible.

Most of the data to be collected should be straightforward from reviewing the
measurement template.  However, in some cases some extra guidance is necessary
to eliminate ambiguity, as follows:

\begin{enumerate}
\item Initial release date: Mark the release year if an exact date is not
  available.
\item Publications about the software: A list of publications can be found
  directly on the website of some software packages. For others use Google
  Scholar or a similar index.
\item Is there evidence that performance was considered?: Search the software
  artifacts for any mention of speed, storage, throughput, performance
  optimization, parallelism, multi-core processing, or similar
  considerations. The search function on GitHub can help.
\item Getting started tutorial: Sometimes this is found within another artifact,
  like the user manual.
\item Continuous integration: Search the software artifacts for any mention of
  continuous integration. The search function on GitHub can help.
\end{enumerate}

For completing the template, the following steps should be followed:

\begin{enumerate}
\item Gather the summary information into the top section of the document
  (Figure \ref{measurement_template_image})
\item Using the GitStats tool that is described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics -
  GitStats section found near the bottom of the document
\item Using the SCC tool that is also described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - SCC
  section found near the bottom of the document
\item If the software package is found on git, gather the measurements for the
  Repo Metrics - GitHub section found near the bottom of the document
\item Review installation documentation and attempt to install the software
  package on a virtual machine
\item Gather the measurements for installability
\item Gather the measurements for correctness and verifiability
\item Gather the measurements for surface reliability
\item Gather the measurements for surface robustness
\item Gather the measurements for surface usability
\item Gather the measurements for maintainability
\item Gather the measurements for reusability
\item Gather the measurements for surface understandability
\item Gather the measurements for visibility and transparency
\item Assign a score out of ten for each quality. The score can be measured
  using the Measurement Template Impression Calculator, found in
  Appendix~\ref{SecImpressionCalculator}. For each quality measurement, the file
  indicates the appropriate score to assign the measurement based on possible
  measurement values.
\end{enumerate}

\section{Analytic Hierarchy Process} \label{SecAHP}

The Analytical Hierarchy Process (AHP) is a decision-making technique that can
be used when comparing multiple options by multiple criteria. In our work AHP is
used for comparing and ranking the software packages of a domain using the
quality scores that are gathered in the Measurement Template
(Appendix~\ref{SecGradingTemplate}). AHP performs a pairwise analysis between
each of the quality options using a matrix which is used to generate an overall
score for each software package for the given criteria. \cite{SmithEtAl2016}
shows how AHP is applied to ranking software based on quality measures. We have
developed a
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{tool}
for conducting this process. The tool includes an AHP JAR script and a
sensitivity analysis JAR script that is used to ensure that the software package
rankings are appropriate with respect to the uncertainty of the quality scores.
The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{README
file} outlines the requirements for, and configuration and usage of, the JAR
scripts. The JAR scripts, source code, and required libraries are located in the
same folder as the README file.

% \section{Rank Short List} \label{SecRankShortList}

% An AHP pairwise comparison between the short list packages with respect to
% usability and modifiability experiment results is conducted, and then reviewed
% by the domain expert. The tool can be found here. The usability and
% modifiability survey can be found
% \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}
% {here}. This step is intended to start a conversation about the relative quality
% of short-listed software packages.

\section{Estimate of Time Required} \label{SecEstTimeRequired}

estimate the time required in person hours.  Add a table.  Assume reviewed
material, using provided spreadsheets and AHP tool.  Assume domain has been
identified and domain expert engaged.  Following steps given in
Section~\ref{StepsAQDS}.

\begin{table}[h]
  \caption{Estimated Person Hours for Assessing the State of Practice for Domain
  X} \label{TabPersonHours}
  \centering
  \begin{tabular}{p{10cm} l}
    \toprule
    \textbf{Task} & \textbf{Hours} \\
    \midrule

    Initial 1 hour meeting with domain expert plus meeting prep & 10
    \addtocounter{totHours}{10} \\

    Identify candidate software & 10 \addtocounter{totHours}{10} \\

    Filter software list & 5 \addtocounter{totHours}{5} \\

    Domain analysis & 20 \addtocounter{totHours}{20} \\

    Interaction with domain expert(s) (initial meeting, review software list,
    vet domain analysis, vet AHP ranking) & 10 \addtocounter{totHours}{10} \\

    Collect raw data and fill-in measurement template & 30
    \addtocounter{totHours}{30} \\

    Solicit developers for interviews & 2 \addtocounter{totHours}{2} \\

    Conduct interviews & 15 \addtocounter{totHours}{15} \\

    AHP ranking & 2 \addtocounter{totHours}{2} \\

    Analyze data and answer research questions & 30 \addtocounter{totHours}{30}
    \\

    \midrule
    \textbf{Total} & \textbf{\thetotHours} \\
    \bottomrule
  \end{tabular}

\end{table}

\section{Future Work} \label{SecFutureWork}

- usability and modifiability experiments

Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})

Modifiability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})

- not just for research software

- use it to inform research on Drasil - what artifacts, what process dev
currently use, how could the new process promoted by Drasil fit into the current
development processes?

- meta analysis of collected data.  Already have GIS, mesh generator.  Currently
working on medical image analysis and LBM.

\section{Concluding Remarks} \label{SecConcludingRemarks}

- summary

\newpage

\appendix

\section{A Guide to Repository-Based Measurement Tools} \label{SecRepoTools}

This appendix covers the tools used for collecting repository based data
(Section~\ref{SecEmpiricalMeasures}).  The two tools covered are git\_stats and
scc.  The tools do not have to be used in any particular order.

\subsection{git\_stats}

\subsubsection{Introduction}

Source Code: \href{https://github.com/tomgi/git_stats}{GitHub repo}

\subsubsection{User Manual} \label{git_stats_manual}

Official Manual: \href{https://github.com/tomgi/git_stats}{GitHub repo}

\subsubsection{Demo of Installation and Running the Tool}

The installation steps on your machine may be different from this section.
Please refer to the user manual mentioned in Section \ref{git_stats_manual}, if
necessary.  The steps shown here were executed on a virtual machine with 8 cores
and 16 GB RAM running Debian GNU/Linux 9.11

\begin{enumerate}
\item Install ruby/gem environment
\begin{lstlisting}
apt-get install ruby ruby-nokogiri ruby-nokogiri-diff ruby-nokogumbo
\end{lstlisting}

Check the installation:
\begin{lstlisting}
gem --version
\end{lstlisting}

\item Install the tool
\begin{lstlisting}
sudo gem install git_stats
\end{lstlisting}

\item Prepare the target repo

Make sure the target repo (the repo to be analyzed, not the repo of this tool)
is on your machine.
In this demo, the target repo is downloaded from a GitHub repo:
\begin{lstlisting}
# change [git path] to the url of your target repo
git clone [git path]
# e.g. git clone https://github.com/nroduit/Weasis.git
\end{lstlisting}

\item Generate analytics
\begin{lstlisting}
# make sure [repo path] is the target repo path
# the [output path] can be anywhere you desire
git_stats generate -p [repo path] -o [output path]
# e.g. git_stats generate -p /home/user/git-stats/Weasis -o 
# /home/user/git-stats/Weasis-analytics
\end{lstlisting}

\item View the analytics

View the analytic results by open [output path]/index.html with any browser or
other software supporting HTML web page format.
\begin{figure}[h!]
\includegraphics[scale=0.4]{git-stats-demo.JPG}
\end{figure}

\item Download the data

On most of the taps of this web page, the data can be downloaded for more
analytics by clicking the menu button beside the data-range section.

\end{enumerate}

\subsection{scc}

\subsubsection{Introduction}

Source Code: \href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{User Manual} \label{scc_manual}

Official Manual: \href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{Demo of Installation and Running the Tool}

The installation steps on your machine may be different from this section.
Please refer to the user manual mentioned in Section \ref{scc_manual}, if
necessary.  These steps were executed on a virtual machine with 8 cores and 16
GB RAM running Debian GNU/Linux 9.11

\begin{enumerate}
\item Install Golang

Follow the \href{https://golang.org/doc/install}{official instructions}, or the
following demo,

download the installation package:
\begin{lstlisting}
wget https://dl.google.com/go/go1.14.3.linux-amd64.tar.gz
\end{lstlisting}

unpack to /usr/local:
\begin{lstlisting}
sudo tar -C /usr/local -xzf go1.14.3.linux-amd64.tar.gz
\end{lstlisting}

use a text editor to open \textasciitilde/.profile, e.g.:
\begin{lstlisting}
nano ~/.profile
\end{lstlisting}

add the following lines to the end of this file:
\begin{lstlisting}
export GOPATH=$HOME/go
export PATH=$PATH:/usr/local/go/bin:$GOPATH/bin
\end{lstlisting}

save the file, and load the commands into the current shell instance:
\begin{lstlisting}
source ~/.profile
\end{lstlisting}

check the installation:
\begin{lstlisting}
go version
\end{lstlisting}

\item Install the tool
\begin{lstlisting}
go get -u github.com/boyter/scc/
\end{lstlisting}

\item Prepare the target repo

Make sure the target repo (the repo to be analyzed, not the repo of this tool)
is on your machine.
In this demo, the target repo is downloaded from a GitHub repo:
\begin{lstlisting}
# change [project path] to your desired folder
cd [project path]
git clone https://github.com/nroduit/Weasis.git
\end{lstlisting}

\item Generate analytics
\begin{lstlisting}
# make sure [repo path] is the target repo path
cd [repo path]
# use scc to generate analytics
scc
\end{lstlisting}

\item View the analytics

The results will be appear something like the following.

\begin{figure}[h!]
\includegraphics[scale=0.5]{scc-demo.JPG}
\end{figure}

\end{enumerate}

\newpage

\section{Measurement Template} \label{SecGradingTemplate}

The table below lists the full set of measures that are assessed for each
software product.  The measures are grouped under headings for each quality, and
one for summary information.  Following each measure, the type for a valid
result is given in brackets.  Many of the types are given as enumerated sets.
For instance, the response on many of the questions is one of ``yes,'' ``no,''
or ``unclear.''  The type ``number'' means natural number, a positive integer.  The
types for date and url are not explicitly defined, but they are what one would
expect from their names.  In some cases the response for a given question is not
necessarily limited to one answer, such as the question on what platforms are
supported by the software product.  Case like this are indicated by ``set of''
preceding the type of an individual answer.  The type in these cases are then
the power set of the individual response type.  In some cases a superscript $^*$
is used to indicate that a response of this type should be accompanied by
explanatory text.  For instance, if problems were caused by uninstall, the
reviewer should note what problems were caused.  An (I) precedes the question
description when its measurement requires a successful installation.

The table below lists the set of measures that are used to assess each software
product. The first set identifies summary information, followed by 9 sets for
software qualities and 3 sets for raw metrics. Each measure is followed by the
type for a valid result. A superscript $^*$ indicate that a response of this type
needs to be accompanied by explanatory text.

\newpage

\captionof{table}{Measurement Template}\label{measurementtemplate}
\def\arraystretch{1.22}
\begin{tabular}{p{14cm}}
	\hline
	\textbf{Summary Information}\\
	\hline
	Software name? (string)\\
	URL? (URL)\\
	Affiliation (institution(s)) (string or {N/A})\\
	Software purpose (string)\\
	Number of developers (all developers that have contributed at least one commit
	to the project) (use repo commit logs) (number)\\
	How is the project funded? (unfunded, unclear, funded$^*$) where $^*$ requires a
	string to say the source of funding\\
	Initial release date? (date)\\
	Last commit date? (date)\\
	Status? (alive is defined as presence of commits in the last 18 months)
	({alive, dead, unclear})\\
	License? ({GNU GPL, BSD, MIT, terms of use, trial, none, unclear, other$^*$}) $^*$
	given via a string \\
	Platforms? (set of {Windows, Linux, OS X, Android, other$^*$}) $^*$ given via
	string\\
	Software Category? The concept category includes software that does not have
	an officially released version. Public software has a released version in the
	public domain. Private software has a released version available to authorized
	users only. ({concept, public, private})\\
	Development model? ({open source, freeware, commercial, unclear})\\
	Publications about the software? Refers to publications that have used or
	mentioned the software. (number or {unknown})\\
	Source code URL? ({set of url, n/a, unclear})\\
	Programming language(s)? (set of {FORTRAN, Matlab, C, C++, Java, R, Ruby,
	Python, Cython, BASIC, Pascal, IDL, unclear, other$^*$}) $^*$ given via string \\
	Is there evidence that performance was considered? Performance refers to
	either speed, storage, or throughput. ({yes$^*$, no})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have) \\
	\hline
\end{tabular}

\def\arraystretch{1.5}
\begin{tabular}{p{14cm}}
	\hline		
	\textbf{Installability (Measured via installation on a virtual machine.) }\\
	\hline
	Are there installation instructions? ({yes, no})\\
	Are the installation instructions in one place? Place referring to a single
	document or web-page. ({yes, no, n/a})\\
	Are the installation instructions linear? Linear meaning progressing  in a
	single series of steps. ({yes, no, n/a})\\
	Are the instructions written as if the person doing the installation has none
	of the dependent packages installed? ({yes, no, unclear})\\
	Are compatible operating system versions listed? ({yes, no})\\
	Is there something in place to automate the installation (makefile, script,
	installer, etc)? ({yes$^*$, no})\\
	If the software installation broke, was a descriptive error message displayed?
	({yes, no, n/a})\\
	Is there a specified way to validate the installation? ({yes$^*$, no})\\
	How many steps were involved in the installation? (Includes manual steps like
	unzipping files) Specify OS. (number, OS)\\
	What OS was used for the installation? ({Windows, Linux, OS X, Android, other$^*$
	}) $^*$given via string\\
	How many extra software packages need to be installed before or during
	installation? (number)\\
	Are required package versions listed? ({yes, no, n/a})\\
	Are there instructions for the installation of required packages /
	dependencies? ({yes, no, n/a})\\
	Run uninstall, if available. Were any obvious problems caused? ({yes$^*$ , no,
	unavail})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline
\end{tabular}

\def\arraystretch{1.33}
\begin{tabular}{p{14cm}}
	\hline		
	\textbf{Correctness and Verifiability}\\
	\hline
	Any reference to the requirements specifications of the program or theory
	manuals? ({yes$^*$ , no, unclear})\\
	What tools or techniques are used to build confidence of correctness?
	({literate programming, automated testing, symbolic execution, model checking,
	assertions used in the code, Sphinx, Doxygen, Javadoc, confluence, unclear,
	other$^*$}) $^*$ given via string\\
	If there is a getting started tutorial? ({yes, no})\\
	Are the tutorial instructions linear? ({yes, no, n/a})\\
	Does the getting started tutorial provide an expected output? ({yes, no$^*$,
	n/a})\\
	Does your tutorial output match the expected output? ({yes, no, n/a})\\
	Are unit tests available?  ({yes, no, unclear})\\
	Is there evidence of continuous integration? (for example mentioned in
	documentation, Jenkins, Travis CI, Bamboo, other) ({yes$^*$, no, unclear})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have) \\
	\hline	
	\textbf{Surface Reliability}\\
	\hline
	Did the software “break” during installation? ({yes$^*$ , no})\\
	If the software installation broke, was the installation instance recoverable?
	({yes, no, n/a})\\
	Did the software “break” during the initial tutorial testing? ({yes$^*$, no,
	n/a})\\
	If the tutorial testing broke, was a descriptive error message displayed?
	({yes, no, n/a})\\
	If the tutorial testing broke, was the tutorial testing instance recoverable?
	({yes, no, n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
	\hline		
	\textbf{Surface Robustness}\\
	\hline
	Does the software handle unexpected/unanticipated input (like data of the
	wrong type, empty input, missing files or links) reasonably? (a reasonable
	response can include an appropriate error message.) ({yes, no$^*$ })\\
	For any plain text input files, if all new lines are replaced with new lines
	and carriage returns, will the software handle this gracefully? ({yes, no$^*$,
	n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
	\textbf{Surface Usability}\\
	\hline
	Is there a getting started tutorial? ({yes, no})\\
	Is there a user manual? ({yes, no})\\
	Are expected user characteristics documented? ({yes, no})\\
	What is the user support model? FAQ? User forum? E-mail address to direct
	questions? Etc. (string)\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
		\hline	
	\textbf{Maintainability}\\
	\hline
	What is the current version number? (number)\\
	Is there any information on how code is reviewed, or how to contribute?
	({yes$^*$, no})\\
	Are artifacts available? (List every type of file that is not a code file –
	for examples please look at the ‘Artifact Name’ column of
	https://gitlab.cas.mcmaster.ca/SEforSC/se4sc/-/blob/git-svn/GradStudents/Olu/ResearchProposal/Artifacts\_MiningV3.xlsx)
	({yes$^*$, no, unclear}) $^*$list via string\\
	What issue tracking tool is employed? (set of {Trac, JIRA, Redmine, e-mail,
	discussion board, sourceforge, google code, git, BitBucket, none, unclear,
	other$^*$}) $^*$ given via string\\
	What is the percentage of identified issues that are closed? (percentage)\\
	What percentage of code is comments? (percentage)\\
	Which version control system is in use? ({svn, cvs, git, github, unclear,
	other$^*$}) $^*$ given via string\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
	\textbf{Reusability}\\
	\hline
	How many code files are there? (number)\\
	Is API documented? ({yes, no, n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
		\hline		
	\textbf{Surface Understandability (Based on 10 random source files)}\\
	\hline
	Consistent indentation and formatting style? ({yes, no, n/a})\\
	Explicit identification of a coding standard? ({yes$^*$, no, n/a})\\
	Are the code identifiers consistent, distinctive, and meaningful? ({yes, no$^*$ ,
	n/a})\\
	Are constants (other than 0 and 1) hard coded into the program? ({yes, no$^*$ ,
	n/a})\\
	Comments are clear, indicate what is being done, not how? ({yes, no$^*$ , n/a})\\
	Parameters are in the same order for all functions? ({yes, no$^*$ , n/a})\\
	Is the name/URL of any algorithms used mentioned? ({yes, no$^*$ , n/a})\\
	Is code modularized? ({yes, no$^*$ , n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
	\textbf{Visibility/Transparency}\\
	\hline
	Is the development process defined? If yes, what process is used. ({yes$^*$, no,
	n/a})\\
	Are there any documents recording the development process and status?  ({yes$^*$,
	no}))\\
	Is the development environment documented? ({yes$^*$, no})\\
	Are there release notes? ({yes$^*$, no})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
		\hline		
	\textbf{Raw Metrics (Measured via git\_stats)}\\
	\hline
	Number of text-based files. (number)\\
	Number of binary files. (number)\\
	Number of total lines in text-based files. (number)\\
	Number of total lines added to text-based files. (number)\\
	Number of total lines deleted from text-based files. (number)\\
	Number of total commits. (number)\\
	Numbers of commits by year in the last 5 years. (Count from as early as
	possible if the project is younger than 5 years.) (list of numbers)\\
	Numbers of commits by month in the last 12 months. (list of numbers)\\
	\hline
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
		\hline		
	\textbf{Raw Metrics (Measured via scc)}\\
	\hline
	Number of text-based files. (number)\\
	Number of total lines in text-based files. (number)\\
	Number of code lines in text-based files. (number)\\
	Number of comment lines in text-based files. (number)\\
	Number of blank lines in text-based files. (number)\\
	\hline
	\textbf{Repo Metrics (Measured via GitHub)}\\
	\hline
	Number of stars. (number)\\
	Number of forks. (number)\\
	Number of people watching this repo. (number)\\
	Number of open pull requests. (number)\\
	Number of closed pull requests. (number)\\
	\hline
\end{tabular}

\newpage

\section{Measurement Template Impression Calculator} \label{SecImpressionCalculator}

The table below lists how each quality measure of the measurement template is
used to calculate an overall impression in each software quality set.

\newpage

\captionof{table}{Measurement Template}
\def\arraystretch{1.5}
\begin{tabular}{p{14cm}}
	\hline		
	\textbf{Installability (Measured via installation on a virtual machine.) }\\
	\hline
	Are there installation instructions? ({yes=1, no=-1})\\
	Are the installation instructions in one place? Place referring to a single
	document or web-page. ({yes=1, no=0, n/a=0})\\
	Are the installation instructions linear? Linear meaning progressing  in a
	single series of steps. ({yes=1, no=0, n/a=0})\\
	Are the instructions written as if the person doing the installation has none
	of the dependent packages installed? ({yes=1, no=0, unclear=0})\\
	Are compatible operating system versions listed? ({yes=1, no=0})\\
	Is there something in place to automate the installation (makefile, script,
	installer, etc)? ({yes$^*$=1, no=-1})\\
	If the software installation broke, was a descriptive error message displayed?
	({yes=0, no=-2, n/a=1})\\
	Is there a specified way to validate the installation? ({yes$^*$=1, no=0})\\
	How many steps were involved in the installation? (Includes manual steps like
	unzipping files) Specify OS. ($<$10 = 1)\\
	What OS was used for the installation? (does not count)\\
	How many extra software packages need to be installed before or during
	installation? ($<$10 = 1)\\
	Are required package versions listed? ({yes=1, no=0, n/a=1})\\
	Are there instructions for the installation of required packages /
	dependencies? ({yes=1, no=0, n/a=1})\\
	Run uninstall, if available. Were any obvious problems caused? ({yes$^*$=0, no=1,
	unavail=1})\\
	Overall impression? (a sum of $>$10 is rounded down to 10)\\
	\hline
\end{tabular}

$\mathit{ab} \mathit{val}$

\def\arraystretch{1.33}
\begin{tabular}{p{14cm}}
	\hline		
	\textbf{Correctness and Verifiability}\\
	\hline
	Any reference to the requirements specifications of the program or theory
	manuals? ({yes$^*$=2, no=0, unclear=0})\\
	What tools or techniques are used to build confidence of correctness? (any=1,
	unclear=0)\\
	If there is a getting started tutorial? ({yes=2, no=0})\\
	Are the tutorial instructions linear? ({yes=1, no=0, n/a=0})\\
	Does the getting started tutorial provide an expected output? ({yes=1, no$^*$=0,
	n/a=0})\\
	Does your tutorial output match the expected output? ({yes=1, no=0, n/a=0})\\
	Are unit tests available?  ({yes=1, no=0, unclear=0})\\
	Is there evidence of continuous integration? (for example mentioned in
	documentation, Jenkins, Travis CI, Bamboo, other) ({yes$^*$=1, no=0,
	unclear=0})\\
	\hline	
	\textbf{Surface Reliability}\\
	\hline
	Did the software “break” during installation? ({yes$^*$=0, no=5})\\
	If the software installation broke, was the installation instance recoverable?
	({yes=5, no=0, n/a=0})\\
	Did the software “break” during the initial tutorial testing? ({yes$^*$=0, no=5,
	n/a=0})\\
	If the tutorial testing broke, was a descriptive error message displayed?
	({yes=2, no=0, n/a=0})\\
	If the tutorial testing broke, was the tutorial testing instance recoverable?
	({yes=3, no=0, n/a=0})\\
	\hline		
	\textbf{Surface Robustness}\\
	\hline
	Does the software handle unexpected/unanticipated input (like data of the
	wrong type, empty input, missing files or links) reasonably? (a reasonable
	response can include an appropriate error message.) ({yes=5, no$^*$=0})\\
	For any plain text input files, if all new lines are replaced with new lines
	and carriage returns, will the software handle this gracefully? ({yes=5,
	no$^*$=0, n/a=5})\\
	\hline		
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
		\hline		
	\textbf{Surface Usability}\\
	\hline
	Is there a getting started tutorial? ({yes=3, no=0})\\
	Is there a user manual? ({yes=4, no=0})\\
	Are expected user characteristics documented? ({yes=1, no=0})\\
	What is the user support model? FAQ? User forum? E-mail address to direct
	questions? Etc. (one=1, two+=2, none=0)\\
	\hline
	\textbf{Maintainability}\\
	\hline
	What is the current version number? (provided=1, nothing=0)\\
	Is there any information on how code is reviewed, or how to contribute?
	({yes$^*$=1, no=0})\\
	Are artifacts available? (List every type of file that is not a code file –
	for examples please look at the ‘Artifact Name’ column of
	https://gitlab.cas.mcmaster.ca/SEforSC/se4sc/-/blob/git-svn/GradStudents/Olu/ResearchProposal/Artifacts\_MiningV3.xlsx)
	(Rate 0 – 2 depending on how many and perceived quality)\\
	What issue tracking tool is employed? (nothing=0, email of other private=1,
	anything public or accessible by all devs (eg git) = 2)\\
	What is the percentage of identified issues that are closed? (50$\%$+=1,
	$<$50$\%$=0)\\
	What percentage of code is comments? (10$\%$+=1, $<$10$\%$=0)\\
	Which version control system is in use? (anything=2, nothing=0)\\
	\hline		
	\textbf{Reusability}\\
	\hline
	How many code files are there? (0-9=0, 10-49=1, 50-99=3, 100-299=4, 300-599=5,
	600-999=6, 1000+=8)\\
	Is API documented? ({yes=2, no=0, n/a=0})\\
	\hline
\end{tabular}

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
\hline	
\textbf{Surface Understandability (Based on 10 random source files)}\\
\hline
Consistent indentation and formatting style? ({yes=1, no=0, n/a=0})\\
Explicit identification of a coding standard? ({yes$^*$=1, no=0, n/a=0})\\
Are the code identifiers consistent, distinctive, and meaningful? ({yes=2,
no$^*$=0, n/a=0})\\
Are constants (other than 0 and 1) hard coded into the program? ({yes=1, no$^*$=0,
n/a=0})\\
Comments are clear, indicate what is being done, not how? ({yes=2, no$^*$=0,
n/a=0})\\
Is the name/URL of any algorithms used mentioned? ({yes=1, no$^*$=0, n/a=0})\\
Parameters are in the same order for all functions? ({yes=1, no$^*$=0, n/a=0})\\
Is code modularized? ({yes=1, no$^*$=0, n/a=0})\\
\hline		
\textbf{Visibility/Transparency}\\
\hline
Is the development process defined? If yes, what process is used. ({yes$^*$=3,
no=0, n/a=0})\\
Are there any documents recording the development process and status?  ({yes$^*$=3,
no=0}))\\
Is the development environment documented? ({yes$^*$=2, no=0})\\
Are there release notes? ({yes$^*$=2, no=0})\\
\hline
\end{tabular}

\newpage

\section{Survey Questions for Developers} \label{SecSurveyQuestions}

Introduction.  Roadmap.  Anything about traceability to research questions.

Information about these interview questions:  This gives you an idea what I
would like to learn about the development of {domain} software. Interviews will
be one-to-one and will be open-ended (not just “yes or no” answers). Because of
this, the exact wording may change a little. Sometimes I will use other short
questions to make sure I understand what you told me or if I need more
information when we are talking such as: “So, you are saying that ...?), to get
more information (“Please tell me more?”), or to learn what you think or feel
about something (“Why do you think that is...?”).

\subsection{Information about the developers and users} \citep{Jegatheesan2016}
\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item Do you have a defined process for accepting new contributions into your team?
\item What is the typical background of a developer?
\item What is your estimated number of users? How did you come up with that estimate?
\item What is the typical background of a user?
\end{enumerate}

\subsection{Information about the software}

Square brackets are used for traceability to qualities and relevant research
questions found in Section ...

\begin{enumerate}
\item Currently, what are the most significant obstacles in your development
process?
\item How might you change your development process to remove or reduce these
obstacles?
\item How does documentation fit into your development process? Would improved
documentation help with the obstacles you typically face? [traceability -
research question 5b, visibility/transparency - research question 5i]
\item In the past, is there any major obstacle to your development process that
has been solved? How did you solve it?
\item What is your software development model? For example, waterfall, agile,
etc.
\item What is your project management process? Do you think improve this process
can tackle the current problem? Were any project management tools used?
\item Was it hard to ensure the correctness of the software? If there were any
 obstacles, what methods have been considered or practiced to improve the
 situation? If practiced, did it work? [correctness - research question 5e]
\item When designing the software, did you consider the ease of future changes?
 For example, will it be hard to change the structure of the system, modules or
 code blocks? What measures have been taken to ensure the ease of future changes
 and maintains? [maintainability - research question 5d, modifiability -
 research question 5c]
\item Provide instances where users have misunderstood the software. What, if
any, actions were taken to address understandability issues? [understandability
- research question 5f]
\item What, if any, actions were taken to address usability issues? [usability -
research question 5a]
\item Do you think the current documentation can clearly convey all necessary
knowledge to the users? If yes, how did you successfully achieve it? If no, what
improvements are needed? [unambiguity - research question 5g]
\item Do you have any concern that your computational results won't be
reproducible in the future? Have you taken any steps to ensure reproducibility?
[reproducibility - research question 5h]
\end{enumerate}

\newpage

\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}

\end{document}