### 2022-10-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/ResubmissionResponseToReviewer.tex
Start of resubmission letter to the editor

\documentclass[12pt]{casletter}

\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 180mm
\textheight 200mm
%\renewcommand\baselinestretch{0.85}

\usepackage{xcolor}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 24131}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail:
                                  %  smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
    ~\\
    ~\\
    ~\\
    Dr.\ Eugenio O\~{n}ate\\
    Editor-in-Chief\\
    Archives of Computational Methods in Engineering\\
    ~\\
    {\bf Re: Revisions to ARCO-D-22-00308R1, State of the Practice for Lattice
    Boltzmann Method Software}}

  \opening {Dear Dr.\ O\~{n}ate:}

  Thank you and the reviewer for the feedback on our submission.  In response to
  your e-mail, dated August 26, 2022, we have revised the paper to incorporate
  the requested revisions.  We provide a summary below.  We have also included
  with the revised submission a version of the paper showing all additions and
  deletions.

  We would like to thank you and the reviewer for their thoughtful and
  constructive comments.

  \textbf{Reviewer \#1}

  \begin{enumerate}
  \item The anonymization of the authors poses some challenges in evaluating
  the paper because the citations for the methodology and the raw data
  underlying the study have been anonymized away. \medskip

    \emph{We have removed the anonymization in the resubmission.  \smallskip}

  \item I think it would be useful to have slightly more by way of explanation
  here -- to provide a bit more understanding of the process followed for
  readers who choose not to pursue the details provided in the reference.  Two
  key steps to put a little more flesh around would be (1) how are the entries
  of the measurement template (Fig 2) turned into a single numerical score for
  the section, and (2) how is the aggregate AHP score obtained?  As to the first
  point, for example, I look at the Installability Overall Impression scores,
  which I initially took to be a numerical score for installability, but in the
  AHP Installability scores shown in Fig 3, there is no correlation between the
  number in the measurement template and the final score. \medskip

    \emph{Response. \smallskip}

  \item [I]n Fig 2, I see responses for "Descriptive error messages?" of yes, no,
  and n/a.  I have a hard time imagining what n/a could mean in this context.
  Presumably the installer either provides descriptive error messages or it does
  not?  Is this explained in one of the anonymized references or in the raw
  dataset, which I can't access either due to anonymization. \medskip

    \emph{Response. \smallskip}

  \item AHP is left completely to outside references, and I think it would be
  beneficial to have at least a brief description here. \medskip

    \emph{Response. \smallskip}

  \item I think there was a statement that AHP weights all of the qualities
  equally, but in the graphs of the scores for various qualities (e.g., Figs
  3-6), there are dramatic differences in the ranges of the scores given to the
  packages -- anywhere from 0.06 as the top score to nearly 0.16.  The highest
  overall score is ~0.075 (Fig 12). So what is the *theoretical* top score in
  any category and overall?  Are there some categories in which even the
  highest-scoring packages are quite low on the absolute scale?  If so, I would
  think such situations deserve some discussion. \medskip

    \emph{Response. \smallskip}

  \end{enumerate}

  \closing{Best regards,~\newline} \vspace{-29mm}
  \includegraphics[scale=0.75]{Signature.pdf}

\end {letter}

\end{document}