### 2024-06-13 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
done up to section 2.3. Basically down 2 pages.

%\documentclass[11pt]{article}
\documentclass[sn-mathphys-num]{sn-jnl}

\usepackage{booktabs}

\usepackage{amssymb}
\usepackage{amsmath,amsthm}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{pbox}

\newcounter{rqnum} %research question number
\newcommand{\rqtherqnum}{RQ`'\therqnum}
\newcommand{\rqref}[1]{RQ\ref{#1}}

\newcounter{pnum} %pain point number
\newcommand{\ppthepnum}{P`'\thepnum}
\newcommand{\ppref}[1]{P\ref{#1}}

\newcounter{qnum} %quality number
\newcommand{\qthepnum}{Q`'\theqnum}
\newcommand{\qref}[1]{Q\ref{#1}}

\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf
+}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}}

%\pgfplotsset{compat=1.13}

% prepared for: Journal of Imaging Informatics in Medicine
% https://link.springer.com/journal/10278

\begin{document}

\title[SOP for MI Software]{State of the Practice for Medical Imaging Software
Based on Open Source Repositories}

\author*[1]{\fnm{W.\ Spencer} \sur{Smith}}\email{smiths@mcmaster.ca}

\author[1]{\fnm{Ao} \sur{Dong}}

\author[1]{\fnm{Jacques} \sur{Carette}}\email{carette@mcmaster.ca}

\author[2]{\fnm{Michael} \sur{Noseworthy}}\email{nosewor@mcmaster.ca}

\affil*[1]{\orgdiv{Computing and Software Department}, \orgname{McMaster University}, \orgaddress{\street{1280 Main Street West}, \city{Hamilton}, \postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

\affil[2]{\orgdiv{Electrical and Computer Engineering Department}, \orgname{McMaster University}, \orgaddress{\street{1280 Main Street West}, \city{Hamilton}, \postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

% \author[1,*]{W.\ Spencer Smith}
% \author[1]{Ao Dong}
% \author[1]{Jacques Carette}
% \author[2]{Michael D.\ Noseworthy}

% \affil[1]{McMaster University, Computing and Software Department, {Canada}}
% \affil[2]{McMaster University, Electrical \& Computer Engineering Department,
% {Canada}}
% \affil[*]{Corresponding Author}

\abstract{
We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 
projects from 48 candidates and assessed 9 software qualities (installability,
correctness/ verifiability, reliability, robustness, usability, maintainability,
reusability, understandability, and visibility/transparency) by answering 108
questions for each. Using the Analytic Hierarchy Process (AHP) on the
quantitive data, we ranked the MI software.  The top five are \textit{3D
Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{OHIF Viewer}, and
\textit{ParaView}.  This is consistent with the community's, with four of
these also appearing in the top five using GitHub metrics (stars-per-year).  
In general, the quality and quantity of documentation present in the
project correlates quite well with its popularity.
%We observed 88\% of the documentation artifacts
%recommended by research software development guidelines.  However, the current
%state of the practice deviates from the existing guidelines because of the
%rarity of some recommended artifacts (like test plans, requirements
%specification, code of conduct, code style guidelines, product roadmaps, and
%Application Program Interface (API) documentation).
}

\keywords{
	medical imaging, research software, software engineering, software
	quality, analytic hierarchy process}

\maketitle

% RQ1 to RQ4
% Some of Section 1 (Introduction)
% Sections 2 to 6 (includes looking at artifacts observed in repos)
% Some of Sections 11 to 13 (Threats to Validity, Future Work, Conclusions)

\section{Introduction} \label{ch_intro}

We study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide information for diagnostic, analytic, and medical
applications. Given its importance, we want to understand the merits and
drawbacks of the current development processes, tools, and methodologies. We
use a software engineering lens to assess the quality of existing MI software
and provide recommendations for future development.

\subsection{Research Questions} \label{sec_motivation}

As well as state of the practice (SOP) for MI software, we would like to
understand the impact of the often cited gap (chasm!) between recommended
software engineering practices and that used for most research 
software~\cite{Storer2017}. Although scientists spend
a substantial proportion of their working times on software development
\cite{Hannay2009, Prabhu2011}, few are formally trained~\cite{Hannay2009}.

%We have a standard set of questions designed to assess the qualities of any
%research software project \cite{SmithEtAl2021, SmithAndMichalski2022}. We
%particularly focus on \emph{software artifactcs}, i.e.
%the documents, scripts and code that constitutes a software development
%project. Example artifacts include requirements, specifications, user manuals,
%unit tests, system tests, usability tests, build scripts, API (Application
%Programming Interface) documentation, READMEs, license documents, process
%documents, and code.  %Once we have learned what MI developers do, we then put
%% this information in context by contrasting MI software against the trends shown
%% by developers in other research software communities.

Our investigation is based on the following four research questions:
\begin{enumerate}
\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
\textbf{open source} software projects exist? (Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Establish
quantitative measurements of those project's development practices. From
these measures, which of those projects follow best practices?
(Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
similar are the top projects identified in \rqref{RQ_HighestQuality}
to the most popular projects as viewed by the community?
(Section~\ref{Sec_VsCommunityRanking})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
do MI projects compare to general research software with respect to the
artifacts present in their repositories?
(Section~\ref{Sec_CompareArtifacts})
\end{enumerate}

\subsection{Scope} \label{sec_scope}

We only cover MI visualization software.  We exclude
other categories of MI software, including
Segmentation, Registration, Visualization, Enhancement, Quantification,
Simulation, plus MI archiving and telemedicine systems (Compression, Storage,
and Communication).  We also exclude Statistical Analysis and Image-based
Physiological Modelling and Feature Extraction, Classification, and
Interpretation. Software that provides MI support functions is also out of
scope; therefore, we have not assessed the toolkit libraries VTK and ITK.
Finally, Picture Archiving and Communication System (PACS), which helps users to
economically store and conveniently access images, are considered out of scope.

\subsection{Methodology} \label{SecMethodology}

We have a standard set of questions designed to assess the qualities of any
research software project~\cite{SmithEtAl2021, SmithAndMichalski2022}.
This has been applied to MI software and Lattice Boltzmann Solvers
\cite{SmithEtAl2024}.  This builds off prior work to assess the
state of the practice for such domains as Geographic Information Systems
\cite{smith2018state}, Mesh Generators \cite{smith2016state}, Seismology
software \cite{Smith2018Seismology}, and Statistical software for psychology
\cite{smith2018statistical}.  We maintain the previous constraint that the
work load for measuring a given domain should take around one person-month's
worth of effort ($20$ working days at $8$ person-hours per day).

We identify a list of potential packages (through online searches) which is
then filtered and vetted by a domain expert. We aim for roughly $30$
packages. For each remaining package, we measure its qualities by filling in a
grading template~\cite{SmithEtAl2021}.  This data is used to rank the projects
with AHP.  Further details on interaction with the domain expert, software
qualities, grading the software and AHP are summarizes below and in longer form
in~\cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Domain Expert} \label{sec_vet_software_list}

The Domain Expert vets the proposed list. For too many domains, there are
online resources that are wildly inaccurate, which only an expert can correct.
(Details of our domain expert removed for double-blind).
%The Domain Experts vet the proposed list of software packages and the AHP
%ranking.  Doing these tasks with non-experts runs the risk of trusting
%inaccurate resources and overlooking the tacit knowledge of an expert. For the
%current assessment, our Domain Expert (and paper co-author) is Dr.\ Michael
%Noseworthy, Professor of Electrical and Computer Engineering at McMaster
%University, Co-Director of the McMaster School of Biomedical Engineering, and
%Director of Medical Imaging Physics and Engineering at St.\ Joseph's Healthcare,
%Hamilton, Ontario, Canada.

In advance of the first meeting with the Domain Expert, they are asked to
independently create a list of top software packages in the domain.  This helps
get the expert's knowledge refreshed in advance of the meeting.

\subsubsection{Software Qualities} \label{sec_software_quality}

Quality is defined as a measure of the excellence or worth of an entity.  As is
common practice, we do not think of quality as a single measure, but rather as
a set of measures.  That is, quality is a collection of different qualities,
often called ``ilities.''  For this study we selected 9 qualities to measure:
installability, correctness/ verifiability, reliability, robustness, usability,
maintainability, reusability, understandability, and visibility/transparency.
With the exception of installability, all the qualities are defined in Ghezzi
et al. (2003) \cite{GhezziEtAl2003}. Installability is defined as the effort
required for the installation and/or uninstallation of software in a specified
environment \cite{ISO/IEC25010}.

\subsubsection{Grading} \label{sec_grading_software}

We use an existing template~\cite{SmithEtAl2021} that is designed to measure
the aforementioned qualities. To stay within our given measurement time frame,
each package gets up to five hours of time.  Project developers can be
contacted for help regarding installation, if necessary, but we impose a cap of
about two hours on the installation process.
Figure~\ref{fg_grading_template_example} shows an excerpt of the measurement
spreadsheet.  It includes a column for each measured software package.
% \cite{Dong2021-Data} provides the full set of measurement data.
(The full data is available on Mendeley; link will be provided after
refereeing.)

\begin{figure}[!ht]
\includegraphics[scale=0.53]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure}

The full template consists of 108 questions over 9 qualities.  These
questions are designed to be unambiguous, quantifiable, and measurable
with constrained time and domain knowledge.
%We
%designed the questions to be unambiguous, quantifiable, and measurable with
%limited time and domain knowledge. We group the measures under headings for each
%quality, and one for summary information. The summary information (shown in
%Figure~\ref{fg_grading_template_example}) is the first section of the template.
%This section summarizes general information, such as the software name, purpose,
%platform, programming language, publications about the software, the first
%release and the most recent change date, website, source code repository of the
%product, number of developers, etc.  We only measure open source software that
%is ready for public use by other researchers (following the definition of the
%public category from \cite{GewaltigAndCannon2012}). Information in the summary
%section sets the context for the project, but it does not directly affect the
%grading scores.

%For measuring each quality, we ask several questions and the typical answers are
%among the collection of ``yes'', ``no'', ``n/a'', ``unclear'', a number, a
%string, a date, a set of strings, etc. 
The grader, after answering questions for each quality assigned an overall
score (between 1 and 10) based on the answers.  Several of the qualities use
the word ``surface'' to highlight that these particular qualities are a shallow
measure.  For example, we do not do user studies for measuring usability.
Instead, we are look signs that the developers considered usability.
We use two freeware tools to collect repository related data:
\href{https://github.com/tomgi/git_stats}{GitStats} and
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}.  Further
details on quality measurement are provided in~% Smith et al.\ (2024)
\cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Analytic Hierarchy Process (AHP)} \label{sec_AHP}

Developed by Saaty in the 1970s, AHP is widely used to 
analyze multiple criteria decisions~\cite{VaidyaEtAl2006}. AHP organizes
multiple criteria in a hierarchical structure and uses pairwise comparisons
between alternatives to calculate relative ratios~\cite{Saaty1990}. AHP works
with sets of $n$ \textit{options} and $m$ \textit{criteria}.  In our project
$n=29$ and $m=9$ since there are 29 options (software products) and 9 criteria
(qualities). With AHP the sum of the grades (scores) for all products for a
given quality will be 1.0.  We rank the software for each of the qualities, and
then we combine the quality rankings into an overall ranking based on the
relative priorities between qualities.

\section{Review} \label{ch_results}

We initially identified 48 candidate software projects from
the literature \cite{Bjorn2017, Bruhschwein2019, Haak2015}, on-line articles
\cite{Emms2019, Hasan2020, Mu2019}, and forum discussions \cite{Samala2014}.
Then we filtered as follows:

\begin{enumerate}

\item Removed the packages with no source code available, such as
\textit{MicroDicom}, \textit{Aliza}, and \textit{jivex}.

\item Focused on MI software that provides visualization functions.
We removed seven packages
that were toolkits or libraries, such as \textit{VTK}, \textit{ITK}, and
\textit{dcm4che}, and another three that were for PACS.

\item Removed \textit{Open Dicom Viewer} as it has not received any
updates since 2011.

\end{enumerate}

The Domain Expert provided a list 12 software packages.
We found 6 packages were on both lists: \textit{3D
Slicer}, \textit{Horos}, \textit{ImageJ}, \textit{Fiji}, \textit{MRIcron} 
(we use its descendant \textit{MRIcroGL}) and \textit{Mango} (we
use the web version \textit{Papaya}).  The remaining six packages were on our
out-of-scope list. The Domain Expert agreed with our final choice of 29
packages -- see Table~\ref{tab_final_list}. This contains summary data
collected in the year 2020.

The projects are sorted in descending order of lines of code.
We found the initial release dates (Rlsd) for most
projects and marked the two unknown dates with ``?''. The date of the last
update is the date of the latest update, at the time of measurement. We found
funding information (Fnd) for only eight projects.  For the Number Of
Contributors (NOC) we considered anyone who made at least one accepted commit as
a contributor. The NOC is not usually the same as the number of long-term
project members, since many projects received change requests and code from the
community.  With respect to the OS, 25 packages work on all three OSs: Windows
(W), macOS (M), and Linux (L). Although the usual approach to cross-platform
compatibility was to work natively on multiple OSes, five projects achieved
platform-independence via web applications. The full measurement data for all
packages is available on \href{https://data.mendeley.com/datasets/k3pcdvdzj2/1}
{Mendeley Data}.

\begin{table}[!ht]
\centering
\begin{tabular}{p{3.7cm}lllllllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cmidrule{7-9}
 &  &  &  &  &  & W & M & L &  \\ \midrule
ParaView \cite{Ahrens2005} & 2002 & 2020-10 & \checkmark & 100 & 886326 & \checkmark & \checkmark & \checkmark & \checkmark \\
Gwyddion \cite{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & \checkmark & \checkmark & \checkmark &  \\
Horos \cite{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & \checkmark &  &  \\
OsiriX Lite \cite{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & \checkmark &  &  \\
3D Slicer \cite{Kikinis2014} & 1998 & 2020-08 & \checkmark & 100 & 501451 & \checkmark & \checkmark & \checkmark &  \\
Drishti \cite{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & \checkmark & \checkmark & \checkmark &  \\
Ginkgo CADx \cite{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & \checkmark & \checkmark & \checkmark &  \\
GATE \cite{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & \checkmark & \checkmark &  \\
3DimViewer \cite{TESCAN2020} & ? & 2020-03 & \checkmark & 3 & 178065 & \checkmark & \checkmark &  &  \\
medInria \cite{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & \checkmark & \checkmark & \checkmark &  \\
BioImage Suite Web \cite{Papademetris2005} & 2018 & 2020-10 & \checkmark & 13 & 139699 &
\checkmark & \checkmark & \checkmark & \checkmark \\
Weasis \cite{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & \checkmark & \checkmark & \checkmark &  \\
AMIDE \cite{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & \checkmark & \checkmark & \checkmark &  \\
XMedCon \cite{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & \checkmark & \checkmark & \checkmark &  \\
ITK-SNAP \cite{Yushkevich2006} & 2006 & 2020-06 & \checkmark & 13 & 88530 & \checkmark & \checkmark & \checkmark &  \\
Papaya \cite{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & \checkmark & \checkmark & \checkmark &  \\
OHIF Viewer \cite{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & \checkmark & \checkmark & \checkmark & \checkmark \\
SMILI \cite{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & \checkmark & \checkmark & \checkmark &  \\
INVESALIUS 3 \cite{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & \checkmark & \checkmark & \checkmark &  \\
dwv \cite{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & \checkmark & \checkmark & \checkmark & \checkmark \\
DICOM Viewer \cite{Afsar2021} & 2018 & 2020-04 & \checkmark & 5 & 30761 & \checkmark & \checkmark & \checkmark &  \\
MicroView \cite{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & \checkmark & \checkmark & \checkmark &  \\
MatrixUser \cite{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & \checkmark & \checkmark & \checkmark &  \\
Slice:Drop \cite{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & \checkmark & \checkmark & \checkmark & \checkmark \\
dicompyler \cite{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & \checkmark & \checkmark &  &  \\
Fiji \cite{Schindelin2012} & 2011 & 2020-08 & \checkmark & 55 & 10833 & \checkmark & \checkmark & \checkmark &  \\
ImageJ \cite{Rueden2017} & 1997 & 2020-08 & \checkmark & 18 & 9681 & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL \cite{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & \checkmark & \checkmark & \checkmark &  \\
DicomBrowser \cite{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & \checkmark & \checkmark & \checkmark &  \\ \bottomrule
\end{tabular}
\caption{Final software list (sorted in descending order of the number of Lines
Of Code (LOC))}
\label{tab_final_list}
\end{table}

The programming languages used in order of decreasing popularity are \CC,
JavaScript, Java, C, Python, Pascal, Matlab.  The most popular language is \CC,
for 11 of 29 projects; Pascal and Matlab were each used for a single project.

\subsection{Installability} \label{sec_result_installability}

Figure \ref{fg_installability_scores} lists the installability scores.  We found
installation instructions for 16 projects, but two did not need them
(\textit{BioImage Suite Web} and \textit{Slice:Drop})
as they are web applications. 10 of the projects required extra
dependencies: Five depend on a specific browser; \textit{dwv}, \textit{OHIF
Viewer}, and \textit{GATE} needs extra libraries to build; \textit{ImageJ} and
\textit{Fiji} need an unzip tool; \textit{MatrixUser} needs Matlab;
\textit{DICOM Viewer} needs a Nextcloud platform.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{installability_scores.pdf}
\caption{AHP installability scores}
\label{fg_installability_scores}
\end{figure}

The scores are based on the ease of following the installation instructions,
and automated installation and uninstallation process. There were no issues
for all but the bottom four, just various degrees of ease and automation.

\textit{GATE}, \textit{dwv}, and \textit{DICOM Viewer} showed severe
installation problems. We were not able to install them, even after a reasonable
amount of time (2 hours).  For \textit{dwv} and \textit{GATE} we failed to build
from the source code, but we were able to proceed with measuring other qualities
using a deployed on-line version for \textit{dwv}, and a VM version for
\textit{GATE}. For \textit{DICOM Viewer} we could not install the NextCloud
dependency, and thus could not measure reliability nor robustness for it.

\textit{MatrixUser} depends on Matlab, whose installation is not easy.
For users who already have Matlab, this score should be higher.

\subsection{Correctness \& Verifiability} \label{sec_result_correctness_verifiability}

The packages with higher scores for correctness and verifiability (see
Figure~\ref{fg_correctness_verifiability_scores}) used a wider array of
techniques to improve correctness, and had better documentation to witness
this. For instance, we looked for evidence of unit testing, and only evidence
for about half of the projects. We identified five projects using CI/CD
tools: \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{dwv}, and
\textit{OHIF Viewer}.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{correctness_verifiability_scores.pdf}
\caption{AHP correctness \& verifiability scores}
\label{fg_correctness_verifiability_scores}
\end{figure}

Even for some projects with well-organized documentation, requirements
specifications and theory manuals were still missing.
The only requirements-related document we found was a road
map of \textit{3D Slicer}, which contained design requirements for upcoming
changes.

\subsection{Surface Reliability} \label{sec_result_reliability}

Figure~\ref{fg_reliability_scores} shows the our results.  We were able to
follow the steps in the tutorials that existed (seven packages had them.)
However
\textit{GATE} could not open macro files and became unresponsive several times,
without any descriptive error message. We found that \textit{Drishti} crashed
when loading damaged image files, without showing any descriptive error message.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{reliability_scores.pdf}
\caption{AHP surface reliability scores}
\label{fg_reliability_scores}
\end{figure}

\subsection{Surface Robustness} \label{sec_result_robustness}

Figure \ref{fg_robustness_scores} presents the scores for surface robustness.
The packages with higher scores elegantly handled unexpected/unanticipated
inputs, typically showing a clear error message. We may have underscored
\textit{OHIF Viewer}, since we needed further customization to load
data.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{robustness_scores.pdf}
\caption{AHP surface robustness scores}
\label{fg_robustness_scores}
\end{figure}

According to their documentation, all 29 software packages should support the
DICOM standard. To test robustness, we prepared two types of image files:
correct and incorrect formats (with the incorrect format created by relabelled a
text file to have the ``.dcm'' extension).  All software packages loaded the
correct format image, except for \textit{GATE}, which failed for unknown
reasons.  For the broken format, \textit{MatrixUser}, \textit{dwv}, and
\textit{Slice:Drop} ignored the incorrect format of the file and loaded it
regardless. They did not show any error message and displayed a blank image.
\textit{MRIcroGL} behaved similarly except that it showed a meaningless image.
\textit{Drishti} successfully detected the broken format of the file, but the
software crashed as a result.

\subsection{Surface Usability} \label{sec_result_usability}

Figure~\ref{fg_usability_scores} shows the AHP scores for surface usability. The
software with higher scores usually provided both comprehensive documented
guidance and a good user experience. \textit{INVESALIUS 3} provided an excellent
example of a detailed and precise user manual. \textit{GATE} also provided
numerous documents, but unfortunately we had difficulty understanding and using
them. We found getting started tutorials for only 11 projects, but a user manual
for 22 projects. \textit{MRIcroGL} was the only project that explicitly
documented expected user characteristics.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

\subsection{Maintainability} \label{sec_score_maintainability}

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. Only a few of the 29 projects had a
product, developer's manual, or API documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents. Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%).
Table~\ref{tab_maintainability_docs} shows which projects had these documents,
in descending order of their maintainability scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lccc}
\toprule
\multicolumn{1}{c}{Software} & Prod.\ Roadmap & Dev.\ Manual & API Doc. \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark \\
Weasis &  & \checkmark &  \\
OHIF Viewer &  & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark \\
ParaView & \checkmark &  &  \\
SMILI &  &  & \checkmark \\
medInria &  & \checkmark &  \\
INVESALIUS 3 & \checkmark &  &  \\
dwv &  &  & \checkmark \\
BioImage Suite Web &  & \checkmark &  \\
Gwyddion &  & \checkmark & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with the maintainability documents (listed in descending order of 
maintainability score)}
\label{tab_maintainability_docs}
\end{table}

Twenty-seven of the 29 projects used git as the version control tool, with 24 of
these using GitHub. \textit{AMIDE} used Mercurial and \textit{Gwyddion} used
Subversion. \textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used
SourceForge. \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

\subsection{Reusability} \label{sec_result_reusability}

Figure~\ref{fg_reusability_scores} shows the AHP results for reusability. As
described in Section~\ref{sec_grading_software}, we gave higher scores to the
projects with API documentation. As shown in
Table~\ref{tab_maintainability_docs}, seven projects had API documents. We also
assumed that projects with more code files and less LOC per code file are more
reusable. Table \ref{tab_loc_per_file} shows the number of text-based files by
project, which we used to approximate the number of code files. The table also
lists the total number of lines (including comments and blanks), LOC, and
average LOC per file. We arranged the items in descending order of their
reusability scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (sorted in descending order of reusability
scores)}
\label{tab_loc_per_file}
\end{table}

\subsection{Surface Understandability} \label{sec_result_understandability}

Figure~\ref{fg_surface_understandability_scores} shows the scores for surface
understandability. All projects had a consistent coding style with parameters in
the same order for all functions, modularized code, and, clear comments that
indicate what is done, not how. However, we only found explicit identification
of a coding standard for 3 out of the 29: \textit{3D Slicer}, \textit{Weasis},
and \textit{ImageJ}. We also found hard-coded constants (rather than symbolic
constants) in \textit{medInria}, \textit{dicompyler}, \textit{MicroView}, and
\textit{Papaya}. We did not find any reference to the algorithms used in
projects \textit{XMedCon}, \textit{DicomBrowser}, \textit{3DimViewer},
\textit{BioImage Suite Web}, \textit{Slice:Drop}, \textit{MatrixUser},
\textit{DICOM Viewer}, \textit{dicompyler}, and \textit{Papaya}.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

\subsection{Visibility/Transparency} \label{sec_result_visibility_transparency}

Figure~\ref{fg_visibility_transparency_scores} shows the AHP scores for
visibility/transparency. Generally speaking, the teams that actively documented
their development process and plans scored higher.
Table~\ref{tab_Visibility/Transparency_docs} shows the projects that had
documents for the development process, project status, development environment,
and release notes, in descending order of their visibility/transparency
scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
Software & Dev.\ Process & Proj.\ Status & Dev.\ Env. & Rls.\ Notes \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL &  &  &  & \checkmark \\
Weasis &  &  & \checkmark & \checkmark \\
ParaView &  & \checkmark &  &  \\
OHIF Viewer &  &  & \checkmark & \checkmark \\
DICOM Viewer &  &  & \checkmark & \checkmark \\
medInria &  &  & \checkmark & \checkmark \\
SMILI &  &  &  & \checkmark \\
Drishti &  &  &  & \checkmark \\
INVESALIUS 3 &  &  &  & \checkmark \\
OsiriX Lite &  &  &  & \checkmark \\
GATE &  &  &  & \checkmark \\
MicroView &  &  &  & \checkmark \\
MatrixUser &  &  &  & \checkmark \\
BioImage Suite Web &  &  & \checkmark &  \\
ITK-SNAP &  &  &  & \checkmark \\
Horos &  &  &  & \checkmark \\
dwv &  &  &  & \checkmark \\
Gwyddion &  &  &  & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with visibility/transparency related documents (listed in
descending order of visibility/transparency score)}
\label{tab_Visibility/Transparency_docs}
\end{table}

\subsection{Overall Scores} \label{Sec_OverallQ}

As described in Section~\ref{sec_AHP}, for our AHP measurements, we have nine
criteria (qualities) and 29 alternatives (software packages). In the absence of
a specific real world context, we assumed all nine qualities are equally
important. Figure~\ref{fg_overall_scores} shows the overall scores in descending
order. Since we produced the scores from the AHP process, the total sum of the
29 scores is precisely 1.0.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

\label{fg_overall_scores}
\end{figure}

The top four software products \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, and \textit{OHIF Viewer} have higher scores in most criteria.
\textit{3D Slicer} has a score in the top two for all qualities; \textit{ImageJ}
ranks near the top for all qualities, except for correctness \& verifiability.
\textit{OHIF Viewer} and \textit{Fiji} have similar overall scores, with
\textit{Fiji} doing better in installability and \textit{OHIF Viewer} doing
better in correctness \& verifiability.  Given the installation problems, we may
have underestimated the scores on reliability and robustness for \textit{DICOM
Viewer}, but we compared it equally for the other seven qualities.

\section{Discussion}

In Section~\ref{Sec_VsCommunityRanking}, we compare our ranking to the MI
software community's ranking, as a way to judge our results, and as a way to
understand what the MI community currently values. To understand the state of
the practice for MI software relative to other software, we compare MI software
to other research software in Section~\ref{Sec_CompareArtifacts}.  The
comparison is based on the artifacts our measurement exercise found in the MI
repositories, versus what is typically recommended for research software.  We
provide details on artifacts that are recommended but rarely observed for MI
software. Section~\ref{sec_threats_to_validity} presents threats to the validity
of our data and conclusions.

\subsection{Comparison to Community Ranking} \label{Sec_VsCommunityRanking}

To answer~\rqref{RQ_CompareHQ2Popular} about how our ranking compares to the
popularity of projects as judged by the scientific community, we compare our
ranking (from Section~\ref{ch_results}) with the community ratings on GitHub, as
shown by GitHub stars, number of forks, and number of people watching the
projects.

Table~\ref{tab_ranking_vs_GitHub} shows our ranking of the 29 MI projects, and
their GitHub metrics, if applicable. As mentioned in
Section~\ref{sec_score_maintainability}, 24 projects used GitHub. Since GitHub
repositories have different creation dates, we collect the number of months each
stayed on GitHub, and calculate the average number of new stars, people
watching, and forks per 12 months. The items in
Table~\ref{tab_ranking_vs_GitHub} are listed in descending order of the average
number of new stars per year.  The non-GitHub items are listed in the order of
our ranking.  We collected all GitHub statistics in July 2021.

Generally speaking, most of the top-ranking MI software projects also received
greater attention and popularity on GitHub. Between our ranking and the GitHub
stars-per-year ranking, four of the top five software projects appear in both
lists. Our top five packages are scattered among the first eight positions on the
GitHub list. However, as discussed below, there are discrepancies between the two
lists.

In some cases projects are popular in the community, but were assigned a low
rank by our methodology.  This is the case for \textit{dwv}. The reason for the
low ranking is that, as mentioned in Section~\ref{sec_result_installability}, we
failed to build it locally, and used the test version on its websites for the
measurements. We followed the instructions and tried to run the command
\texttt{yarn run test} locally, which did not work. In addition, the test
version did not detect a broken DICOM file and displayed a blank image as
described in Section~\ref{sec_result_robustness}. We might underestimate the
scores for \textit{dwv} due to uncommon technical issues.

We also ranked \textit{DICOM Viewer} much lower than its popularity. As
mentioned in Section~\ref{sec_result_installability}, it depended on the
NextCloud platform that we could not successfully install. Thus, we might
underestimate the scores of its surface reliability and surface robustness.

Further reason for discrepancies between our ranking and the community's ranking
is that we weighted all qualities equally. This is not likely how users
implicitly rank the different qualities. As a result, some projects with high
community popularity may have scored lower with our method because of a
relatively higher (compared to the scientific community's implicit ranking)
weighting of the poor scores for some qualities. A further explanation for
discrepancies between our measures and the star measures may also be due to
inaccuracy with using stars to approximate popularity.  Stars are not an ideal
measure because stars represent the community's feeling in the past more than
they measure current preferences \cite{Szulik2017}.  The issue with stars is
that they tend only to be added, not removed.  A final reason for
inconsistencies between our ranking and the community's ranking is that, as for
consumer products, more factors influence popularity than just quality.

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table}[!ht]
\centering
\begin{tabular}{llllll}
\toprule
Software & Comm.\ Rank & Our Rank & Stars/yr & Watches/yr & Forks/yr \\ 
\midrule
3D Slicer & 1 & 1 & 284 & 19 & 128 \\
OHIF Viewer & 2 & 4 & 277 & 19 & 224 \\
dwv & 3 & 19 & 124 & 12 & 51 \\
ImageJ & 4 & 2 & 84 & 9 & 30 \\
ParaView & 5 & 5 & 67 & 7 & 28 \\
Horos & 6 & 12 & 49 & 9 & 18 \\
Papaya & 7 & 17 & 45 & 5 & 20 \\
Fiji & 8 & 3 & 44 & 5 & 21 \\
DICOM Viewer & 9 & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 8 & 40 & 4 & 17 \\
Weasis & 11 & 7 & 36 & 5 & 19 \\
dicompyler & 12 & 26 & 35 & 5 & 14 \\
OsiriX Lite & 13 & 11 & 34 & 9 & 24 \\
MRIcroGL & 14 & 18 & 24 & 3 & 3 \\
GATE & 15 & 24 & 19 & 6 & 26 \\
Ginkgo CADx & 16 & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 17 & 6 & 18 & 5 & 7 \\
Drishti & 18 & 27 & 16 & 4 & 4 \\
Slice:Drop & 19 & 21 & 10 & 2 & 5 \\
ITK-SNAP & 20 & 13 & 9 & 1 & 4 \\
medInria & 21 & 9 & 7 & 3 & 6 \\
SMILI & 22 & 10 & 3 & 1 & 2 \\
MatrixUser & 23 & 28 & 2 & 0 & 0 \\
MicroView & 24 & 15 & 1 & 1 & 1 \\
Gwyddion & 25 & 16 & n/a & n/a & n/a \\
XMedCon & 26 & 20 & n/a & n/a & n/a \\
DicomBrowser & 27 & 22 & n/a & n/a & n/a \\
AMIDE & 28 & 23 & n/a & n/a & n/a \\
3DimViewer & 29 & 25 & n/a & n/a & n/a \\ 
\bottomrule
\end{tabular}
\caption{Software ranking by our methodology versus the community (Comm.)\
ranking using GitHub metrics (Sorted in descending order of community
popularity, as estimated by the number of new stars per year)}
\label{tab_ranking_vs_GitHub}
\end{table}
\endgroup

Although our ranking and the estimate of the community's ranking are not perfect
measures, they do suggest a correlation between best practices and popularity.
We do not know which comes first, the use of best practices or popularity, but
we do know that the top ranked packages tend to incorporate best practices. The
next sections will explore how the practices of the MI community compare to the
broader research software community.

\subsection{Comparison Between MI and Research Software for Artifacts}
\label{Sec_CompareArtifacts}

As part of filling in the measurement template (from
Section~\ref{sec_grading_software}), we summarized the artifacts observed in
each MI package. Table~\ref{artifactspresent} groups the artifacts by frequency
into categories of common (20 to 29 ($>$67\%) packages), uncommon (10 to 19
(33-67\%) packages), and rare (1 to 9 ($<$33\%) packages).
Tables~\ref{tab_maintainability_docs} and~\ref{tab_Visibility/Transparency_docs}
show the details on which projects use which types of artifacts for documents
related to maintainability and visibility, respectively.

\begin{table}[ht!]
    %\begin{center}
    \begin{tabular}{ p{3.1 cm} p{5.4 cm} p{3.5 cm}}
    \toprule
    Common & Uncommon & Rare \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    User manual (22) & Authors (14) & Dependency list (7)\\
    Release info. (22) & Frequently Asked Questions (FAQ) (14) & Troubleshooting guide (6)\\
     & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design documentation (5)\\
     & Citation (11) & Code style guide (3)\\
     & & Code of conduct (1)\\
     & & Requirements (1)\\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency (The number 
    in brackets is the number of occurrences)}
    \label{artifactspresent}
    %\end{center}
\end{table}

We answer~\rqref{RQ_CompareArtifacts} by comparing the artifacts that we
observed in MI repositories to those observed and recommended for research
software in general. Our comparison may point out areas where some MI software
packages fall short of current best practices. This is not intended to be a
criticism of any existing packages, especially since for practical reasons not
every project has the resources to achieve the highest possible quality.
However, rather than delve into the nuances of which software can justify
compromising which practices, we will write our comparison under the ideal
assumption that every project has sufficient resources to match best practices.

Table~\ref{Tbl_Guidelines} (based on data from \cite{SmithAndMichalski2022})
shows that MI artifacts generally match the recommendations found in nine
current research software development guidelines:
\begin{itemize}
\item United States Geological Survey Software Planning Checklist
\cite{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\cite{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\cite{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \cite{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \cite{SmithAndRoscoe2018},
\item Trilinos Developers Guide \cite{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \cite{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \cite{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \cite{OrvizEtAl2017}.
\end{itemize}

In Table~\ref{Tbl_Guidelines} each row corresponds to an artifact.  For a given
row, a checkmark in one of the columns means that the corresponding guideline
recommends this artifact.  The last column shows whether the artifact appears in
the measured set of MI software, either not at all (blank), commonly (C),
uncommonly (U) or rarely (R).  We did our best to interpret the meaning of each
artifact consistently between guidelines and specific MI software, but the
terminology and the contents of artifacts are not standardized.  The challenge
even exists for the ubiquitous README file.  The content of README files shows
significant variation between projects \cite{PranaEtAl2018}.  Although some
content is reasonably consistent, with 97\% of README files contain at least one
section describing the `What' of the repository and 89\% offering some `How'
content, other categories are more variable.  For instance, information on
`Contribution', `Why', and `Who', appear in 28\%, 26\% and 53\% of the analyzed
files, respectively \cite{PranaEtAl2018}.

The frequency of checkmarks in Table~\ref{Tbl_Guidelines} indicates the
popularity of recommending a given artifact, but it does not imply that the most
commonly recommended artifacts are the most important artifacts. Just because a
guideline does not explicitly recommend an artifact, does not mean the guideline
authors do not value it.  They may have excluded it because it is out of the
scope of their recommendations, or outside their experience.  For instance, an
artifact related to uninstall is only explicitly mentioned by one set of
guidelines \cite{vanGompelEtAl2016}, but other guideline authors would likely
see its value.  They may simply feel that uninstall is implied by install, or
they may have never asked themselves whether they need separate uninstall
instructions.

\begin{table}[h]
%\begin{center}
\begin{tabular}{ p{2.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} &
\cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} &
\cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
%\end{center}
\end{table}

Two of the items that appear in Table~\ref{artifactspresent} do not appear in
the software development guidelines shown in Table~\ref{Tbl_Guidelines}:
Troubleshooting guide and Developer's manual.  Although the guidelines do not
name these two artifacts, the information contained within them overlaps with
the recommended artifacts.  A Troubleshooting guideline contains information
that would typically be found in a User manual.  A Developer's guide overlaps
with information from the README, INSTALL, Uninstall, Dependency List, Release
Information, API documentation and Design documentation.  In our current
analysis, we have identified artifacts by the names given by the software
guidelines and MI examples.  In the future, a more in-depth analysis would look
at the knowledge fragments that are captured in the artifacts, rather than
focusing on the names of the files that collect these fragments together.

Although the MI community shows examples of 88\% (23 of 26) of the practices we
found in research software guidelines (Table~\ref{Tbl_Guidelines}), we did not
observe three recommended artifacts: i) Uninstall, ii) Test plans, and iii)
Requirements.  Uninstall is likely an omission caused by the focus on installing
software. Given the storage capacity of current hardware, developers and users
are not generally concerned with uninstall.  Moreover, as mentioned above,
uninstall is not particularly emphasized in existing recommendations.

Test plans describe the scope, approach, resources, and the schedule of planned
test activities \cite{VanVliet2000}.  The plan should cover details such as the
makeup of the testing team, automated testing tools and technology to employ,
the testing process, system tests, integration tests and unit tests. We did not
observe test plans for MI software, but that doesn't mean plans weren't created;
it means that the plans are not under version control. Test plans would have to
at least be implicitly created, since we observed test cases with reasonable
frequency for MI software (test cases are categorized as uncommon).

The other apparently neglected document is the requirements specification, which
records the functionalities, expected performance, goals, context, design
constraints, external interfaces and other quality attributes of the software
\cite{IEEE1998}.  If developers write requirements they typically are based on a
template, which provide documentation structure, guidelines, and rules.
Although there is no universally accepted template, examples exist
\cite{IEEE1998, NASA1989, RobertsonAndRobertson1999Vol}.

MI software is like other research software in its neglect of requirements
documentation.  Although requirements documentation is recommended by some
\cite{TobiasEtAl2018, HerouxEtAl2008, SmithAndKoothoor2016}, in practice
research software developers often do not produce a proper requirements
specification \cite{HeatonAndCarver2015}. Sanders and Kelly
\cite{SandersAndKelly2008} interviewed 16 scientists from 10 disciplines and
found that none of the scientists created requirements specifications, unless
regulations in their field mandated such a document.  Requirements are the least
commonly produced type of documentation for research software in general
\cite{Nguyen-HoanEtAl2010}. When looking at the pain points for research
software developers, software requirements and management is the software
engineering discipline that most hurts scientific developers, accounting for
23\% of the technical problems \cite{WieseEtAl2019}.  The lack of support for
requirements is likely due to the perception that up-front requirements are
impossible for research software \cite{CarverEtAl2007, SegalAndMorris2008}, but
if we drop the insistence on ``up-front'' requirements, allowing instead for the
requirements to be written iteratively and incrementally, requirements are
feasible \cite{Smith2016}.  Smith et al. provide a requirements template
tailored to research software \cite{SmithEtAl2007}.

Table~\ref{Tbl_Guidelines} shows several recommended artifacts that are rarely
observed in practice.  A theme among these rare artifacts is that, except for
the user-focused getting started manual, they are developer-focused.  The
dependency list, which is a list of software library dependencies, was rarely
observed, but this information is still likely present, just embedded in build
scripts. The other developer-focused and rare artifacts are as follows:

\begin{itemize}

\item \textbf{A Contributing file} provides new contributors with the
information that they need to start adding/modifying the repository's files.
Abdalla \cite{Abdalla2016} provides a simple template for creating an
open-source contributor guideline.

\item \textbf{A Developer Code of Conduct} explicitly states the expectations
for developers on how they should treat one another \cite{TouraniEtAl2017}.
Three popular codes of conduct are \cite{TouraniEtAl2017}:
\href{https://www.contributor-covenant.org/version/2/1/code_of_conduct/}
{Contributor Covenant}, \href{https://ubuntu.com/community/code-of-conduct}
{Ubuntu Code of Conduct}, and \href{https://www.djangoproject.com/conduct/}
{Django Code of Conduct}. A code of conduct can improve inclusivity, which
brings the benefit of a wider pool of contributors.  For example, a code of
conduct can improve the participation of women \cite{SinghEtAl2021}.

\item \textbf{Code Style Guidelines} present standards for writing code. Style
guides codify such elements as formatting, commenting, naming identifiers, best
practices and dangers to avoid \cite{Carty2020}.  Three sample style guides are:
\href{https://google.github.io/styleguide/javaguide.html} {Google Java Style
Guide},
\href{http://cnl.sogang.ac.kr/cnlab/lectures/programming/python/PEP8_Style_Guide.pdf}
{PEP8 Style Guide for Python}, and
\href{https://google.github.io/styleguide/cppguide.html} {Google \CC Style
Guide}.  Linting tools, like \href{https://pypi.org/project/flake8/}{flake8} for
Python, can be used to enforced coding styles, like the PEP8 standard.

\item \textbf{A Product Roadmap} explains the vision and direction for a product
offering \cite{MunchEtAl2019}.  Although they have different forms, all
roadmaps cover the following: \begin{inparaenum}[i)]
	\item Where are we now?,
	\item Where are we going?, and
	\item How can we get there? \end{inparaenum} \cite{PhaalEtAl2005}. A product
roadmap provides the following benefits: continuity of purpose, facilitation of
collaboration, and assistance with prioritization \cite{Pichler2012}.

\item \textbf{Design Documentation} explicitly states the design goals and
priorities, records the likely changes, decomposes the complex system into
separate modules, and specifies the relationship between the modules. Design
documentation shows the syntax of the interface for the modules, and in some
cases also documents the semantics.  Some potential elements of design
documentation include representation of the system design and class design using
Unified Modelling Language class diagrams \cite{Gamma1995}, or rigorous
documentation of the system design following the template for a Module Guide
(MG) \cite{ParnasEtAl1984}.  An MG organizes the modules in a hierarchy by their
secrets.

\item \textbf{API Documentation} shows developers the services or data provided
by the software application (or library) through such resources as its methods
or objects \cite{MengEtAl2018}.  Understandability is improved by API
documentation \cite{MengEtAl2018}. API documentation can be generated using
tools like Doxygen, pydoc, and javadoc.

\end{itemize}

The rare artifacts for MI software are similar to the rare artifacts for Lattice
Boltzmann solvers \cite{SmithEtAl2024}, except LBM software is more likely to
have developer related artifacts, like Contributing, Dependency list, and Design
documentation.

To improve MI software in the future, an increased use of checklists could help.
Developers can use checklists to ensure they follow best practices.  Some
examples include checklists merging branches into master \cite{Brown2015},
checklists for saving and sharing changes to the project \cite{WilsonEtAl2016},
checklists for new and departing team members \cite{HerouxAndBernholdt2018},
checklists for processes related to commits and releases \cite{HerouxEtAl2008}
and checklists for overall software quality \cite{ThielEtAl2020, SSI2022}.

The above discussion shows that, taken together, MI projects fall somewhat short
of recommended best practices for research software.  However, MI software is
not alone in this.  Many, if not most, research projects fall short of best
practices.  A gap exists in research software development practices and software
engineering recommendations \cite{Storer2017}.  Johanson and Hasselbring observe
that the state-of-the-practice for research software in industry and academia
does not incorporate state-of-the-art SE tools and methods
\cite{JohansonAndHasselbring2018}.  This causes sustainability and reliability
problems \cite{FaulkEtAl2009}. Rather than benefit from capturing and reusing
previous knowledge, projects waste time and energy ``reinventing the wheel''
\cite{deSouzaEtAl2019}.

\subsection{Threats to Validity} \label{sec_threats_to_validity}

Our categories for threat to validity come from an analysis of software
engineering secondary studies by Ampatzoglou et al.\ \cite{AmpatzoglouEtAl2019},
where a secondary study analyzes the data from a set of primary studies.
Ampatzoglou et al.\ is appropriate because a common example of a secondary study
is a systematic literature review. Our methodology is a systematic software
review --- the primary studies are the software packages, and our work collects
and analyzes these primary studies.  We identified similar threats to validity
in our assessment of the state of the practice of Lattice Boltzmann Solvers
\cite{SmithEtAl2024}.

\subsubsection{Reliability}

A study is reliable if repetition of the study by different researchers using
the original study's methodology would lead to the same results
\cite{RunesonAndHost2009}. Reliability means that data and analysis are
independent of the specific researcher(s) doing the study.  For the current
study the identified reliability related threats are as follows:

\begin{itemize}
\item One individual does the manual measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experiences, and biases.
\item The manual measurements for the full set of packages took several months.
Over this time the software repositories may have changed and the reviewer's
judgement may have drifted.
\end{itemize}

We reduced concern over the reliability risk associated with the reviewer's
judgement by demonstrating that the measurement process is reasonably
reproducible.  In previous work \cite{SmithEtAl2016} we graded five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

\subsubsection{Construct Validity}

Construct validity is when the adopted metrics represent what they are intended
to measure \cite{RunesonAndHost2009}. Our construct threats are often related to
how we assume our measurements influences the various software qualities, as
summarized in Section~\ref{sec_grading_software}.  Specifically, our construct
validity related threats include the following:

\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item As mentioned in Section~\ref{sec_result_installability}, we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We used
a deployed on-line version for \textit{dwv}, a VM version for \textit{GATE}, but
no alternative for \textit{DICOM Viewer}. We might underestimate their rank due
to these technical issues.
\item Measuring software robustness only involved two pieces of data. This is
likely part of the reason for limited variation in the robustness scores
(Figure~\ref{fg_robustness_scores}). We could add more robustness data by
pushing the software to deal with more unexpected situations, like a broken
Internet connection, but this would require a larger investment of measurement
time. 
\item We may have inaccurately estimated maintainability by assuming a higher
ratio of comments to source code improves maintainability. Moreover, we assumed
that maintainability is improved if a high percentage of issues are closed, but
a project may have a wealth of open issues, and still be maintainable.
\item We assess reusability by the number of code files and LOC per file. This
measure is indicative of modularity, but it does not necessarily mean a good
modularization. The modules may not be general enough to be easily reused, or
the formatting may be poor, or the understandability of the code may be low.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item As discussed in Section~\ref{Sec_OverallQ}, our overall AHP ranking makes
the unrealistic assumption of equal weighting.
\item We approximated popularity by stars and watches
(Section~\ref{Sec_VsCommunityRanking}), but this assumption may not be valid. 
\item In building Table~\ref{Tbl_Guidelines} some judgement was necessary on our
part, since not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

\subsubsection{Internal Validity} \label{Sec_InternalValidity}

Internal validity means that discovered causal relations are trustworthy and
cannot be explained by other factors \cite{RunesonAndHost2009}. In our
methodology the internal validity threats include the following:

\begin{itemize}
\item In our search for software packages (Section~\ref{ch_results}), we may
have missed a relevant package.
\item Our methodology assumes that all relevant software development activities
will leave a trace in the repositories, but this is not necessarily true. For
instance, we saw little evidence of requirements
(Section~\ref{Sec_CompareArtifacts}), but teams might keep this kind of
information outside their repos, possibly in journal papers or technical
reports.
\end{itemize}

\subsubsection{External Validity}

If the results of a study can be generalized (applied) to other
situations/cases, then the study is externally valid \cite{RunesonAndHost2009}.
We are confident that our search was exhaustive.  We do not believe that we
missed any highly popular MI examples.  Therefore, the bulk of our validity
concerns are internal (Section~\ref{Sec_InternalValidity}).  However, our hope
is that the trends observed, and the lessons learned for MI software can be
applied to other research software.  With that in mind we identified the
following threat to external validity: We cannot generalize our results if the
development of MI software is fundamentally different from other research
software.

Although there are differences, like the importance of data privacy for MI data,
we found the approach to developing LBM software \cite{SmithEtAl2024} and MI
software to be similar.  Except for the domain specific aspects, we believe that
the trends observed in the current study are externally valid for other research
software.

\section{Conclusions} \label{ch_conclusions}

We analyzed the state of the practice for the MI domain with the goal of
understanding current practice by answering four research questions
(Section~\ref{sec_motivation}).  Our methods in Section~\ref{SecMethodology} form a
general process to evaluate domain-specific software, that we apply to the
specific domain of MI software. We identified 48 MI software candidates, then,
with the help of the Domain Expert selected 29 of them for our final list.

Section~\ref{ch_results} lists our measurement results for ranking the 29
projects for nine software qualities. Our ranking results appear credible since
they are mostly consistent with the ranking from the scientific community
implied by the GitHub stars-per-year metric. As discussed in
Section~\ref{Sec_VsCommunityRanking}, four of the top five software projects
appear in both our list and in the GitHub popularity list.  Moreover, our top
five packages appear among the first eight positions on the GitHub list.  The
noteworthy discrepancies between the two lists are for the packages that we were
unable to install (\textit{dwv} and \textit{Dicom Viewer}). Our findings suggest
that the community's top ranked packages tend to be those that incorporate best
practices.

Based on our grading scores \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}
and \textit{OHIF Viewer} are the top four software performers.  However, the
separation between the top performers and the others is not extreme.  Almost all
packages do well on at least a few qualities, as shown in
Table~\ref{topperformerstable}, which summarizes the packages ranked first and
second for each quality. Almost 70\% (20 of 29) of the software packages appear
in the top two for at least two qualities.  The only packages that do not appear
in Table~\ref{topperformerstable}, or only appear once, are \textit{Papaya},
\textit{MatrixUser}, \textit{MRIcroGL}, \textit{XMedCon}, \textit{dicompyler},
\textit{DicomBrowser}, \textit{AMIDE}, \textit{3DimViewer}, and
\textit{Drishti}. The shortness of this list suggests parity with respect to
adoption of best practices for MI software overall.

\begin{table}[ht!]
	%\begin{center}
	\renewcommand{\arraystretch}{1.8}
	\begin{tabular}{ p{2.3cm}p{9.9cm} }
		\toprule

		Quality & Ranked 1st or 2nd\\

		\midrule

		Installability & 3D Slicer, BioImage Suite Web, Slice:Drop, INVESALIUS\\

		\pbox{3.0cm}{Correctness \\ and Verifiability} & OHIF Viewer, 3D
		Slicer, ImageJ\\

		Reliability & SMILI, ImageJ, Fiji, 3D Slicer, Slice:Drop, OHIF
		Viewer\\

		Robustness & XMedCon, Weasis, SMILI, ParaView, OsiriX Lite,
		MicroView, medInria, ITK-SNAP, INVESALIUS, ImageJ, Horos, Gwyddion,
		Fiji, dicompyler, DicomBrowser, BioImage Suite Web, AMIDE, 3DimViewer,
		3D Slicer, OHIF Viewer, DICOM Viewer\\

		Usability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, ParaView,
		INVESALIUS, Ginkgo CADx, SMILI, OsiriX Lite, BioImage Suite Web,
		ITK-SNAP, medInria, MicroView, Gwyddion\\

		Maintainability & 3D Slicer, Weasis, ImageJ, OHIF Viewer, ParaView\\

		Reusability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, SMILI, dwv, BioImage
		Suite Web, GATE, ParaView\\

		\pbox{3.0cm}{Understandability} & 3D Slicer, ImageJ, Weasis,
		Fiji, Horos, OsiriX Lite, dwv, Drishti, OHIF Viewer, GATE, ITK-SNAP,
		ParaView, INVESALIUS\\

		\pbox{3.0cm}{Visibility and \\Transparency} & ImageJ, 3D Slicer, Fiji\\

		Overall Quality & 3D Slicer, ImageJ\\

		\bottomrule		
	\end{tabular}
	\caption{Top performers for each quality (sorted by order of quality
	measurement)} \label{topperformerstable}
	%\end{center}
\end{table}

When it comes to recommended software artifacts, the state of open source MI
projects is healthy, with our surveyed examples showing 88\% of the
documentation artifacts recommended by research software development guidelines
(Section~\ref{Sec_CompareArtifacts}).  However, we did notice areas where
practice seems to lag behind the research software development guidelines.  For
instance, the guidelines recommend three artifacts that were not observed:
uninstall instructions, test plans, and requirements documentation. We observed
the following recommended artifacts, but only rarely: contributing file,
developer code of conduct, code style guidelines, product roadmap, design
documentation, and API documentation (Section~\ref{Sec_CompareArtifacts}).
Future MI projects may wish to consider incorporating more artifacts to
potentially improve software quality.  A deeper understanding of the needs of
the MI community will require data beyond what is available in open source
repositories.  Future work should involve interviewing MI developers to more
deeply understand their software development pain points.

\section*{Acknowledgements}

We would like to thank Peter Michalski and Oluwaseun Owojaiye for fruitful
discussions on topics relevant to this paper.

\section*{Conflict of Interest}

On behalf of all authors, the corresponding author states that there is no
conflict of interest.

\bibliography{SOP_MI_OSR}

\end{document}