### 2024-10-31 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Complete draft of the paper

\documentclass[final, 12pt, 3p, times]{elsarticle}

\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{pbox}
\usepackage{paralist}

\newcounter{rqnum} %research question number
\newcommand{\rqtherqnum}{RQ`'\therqnum}
\newcommand{\rqref}[1]{RQ\ref{#1}}

\newcounter{pnum} %pain point number
\newcommand{\ppthepnum}{P`'\thepnum}
\newcommand{\ppref}[1]{P\ref{#1}}

\newcounter{qnum} %quality number
\newcommand{\qthepnum}{Q`'\theqnum}
\newcommand{\qref}[1]{Q\ref{#1}}

\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf
+}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}}

% prepared for: Journal of Imaging Informatics in Medicine
% https://link.springer.com/journal/10278

\journal{Critical Reviews in Biomedical Engineering}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{State of the Practice for Medical Imaging Software Based on Open Source
Repositories}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

%\author{Anonymous}
\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[ECE]{Michael Noseworthy}

\affiliation[CAS]{organization={McMaster University, Computing and Software
Department}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\affiliation[ECE]{organization={McMaster University, Electrical and Computer Engineering}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\begin{abstract}

We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 projects from 48
candidates and assessed 9 software qualities (installability, correctness/
verifiability, reliability, robustness, usability, maintainability, reusability,
understandability, and visibility/transparency) by answering 108 questions for
each. Using the Analytic Hierarchy Process (AHP) on the quantitative data, we
ranked the MI software.  The top five are \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, \textit{OHIF Viewer}, and \textit{ParaView}.  This is consistent
with the community's view, with four of these also appearing in the top five
using GitHub metrics (stars-per-year).  The quality and quantity of
documentation present in a project correlates quite well with its popularity.
Generally, MI software is in a healthy state as shown by the following: in the
repositories we observed 88\% of the documentation artifacts recommended by
research software development guidelines, 100\% of MI projects use version
control tools, and developers appear to use the common quasi-agile research
software development process. However, the current state of the practice
deviates from the existing guidelines because of the rarity of some recommended
artifacts (like a test plan, requirements specification, code of conduct, and
code style guidelines), low usage of continuous integration (17\% of the
projects), low use of unit testing (about 50\% of projects), and room for
improvement with documentation. From interviewing the developers, we identified
6 concerns: lack of development time, lack of funding, technology hurdles,
correctness, usability, maintainability, and reproducibility. Recommendations
for improving the state of the practice include the following: increase
documentation, increase testing by enriching datasets, increase continuous
integration usage, move to web applications, employ linters, use peer reviews,
and design for change.

\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
	medical imaging, research software, software engineering, software
	quality, analytic hierarchy process
\end{keyword}

\end{frontmatter}

%% \linenumbers

\section{Introduction} \label{ch_intro}

We study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide critical information for diagnostic, analytic, and medical
applications. Given its importance, we want to understand the merits and
drawbacks of the current development processes, tools, and methodologies. We
use a software engineering lens to assess the quality of existing MI software.

\subsection{Research Questions} \label{sec_motivation}

As well as state of the practice (SOP) for MI software, we would like to
understand the impact of the often cited gap (chasm!) between recommended
software engineering practices and that used for most research 
software~\cite{Storer2017}. Although scientists spend
a substantial proportion of their working times on software development
\cite{Hannay2009, Prabhu2011}, few are formally trained~\cite{Hannay2009}.

Our investigation is based on the following four research questions:
\begin{enumerate}
\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
{open source} software projects exist? (Section~\ref{SecWhatProjects})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Based on
quantitative measurements of each project's development practices, which
projects follow best practices? (Section~\ref{SecMeasurementResults})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
similar are the top projects identified in \rqref{RQ_HighestQuality}
to the most popular projects as viewed by the community?
(Section~\ref{Sec_VsCommunityRanking})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
do MI projects compare to general research software with respect to the
artifacts (documents, scripts and code) present in their repositories?
(Section~\ref{Sec_CompareArtifacts})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:] How
do MI projects compare to research software in general with respect to the use
of tools (Section~\ref{Sec_CompareTools}) for development
(\rqref{RQ_CompareToolsProjMngmnt}.a); and, project management
(\rqref{RQ_CompareToolsProjMngmnt}.b)?
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
How do MI projects compare to research software in general with respect to
principles, processes, and methodologies used?
(Section~\ref{Sec_CompareMethodologies})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
the pain points for developers working on MI software projects?
(Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
do the pain points of developers from MI compare to the pain points
for research software in general? (Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
developers what specific best practices are taken to address the pain points
and software quality concerns? (Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
What research software development practice could potentially address the
pain point concerns identified in \rqref{RQ_PainPoints}?
(Section~\ref{ch_recommendations})
\end{enumerate}

\subsection{Scope} \label{sec_scope}

We only cover MI visualization software.  We exclude other categories of MI
software, including segmentation, registration, visualization, enhancement,
quantification, simulation, plus MI archiving and telemedicine systems
(compression, storage, and communication).  We also exclude statistical analysis
and image-based physiological modelling and feature extraction, classification,
and interpretation. Software that provides MI support functions is also out of
scope; therefore, we have not assessed the toolkit libraries VTK and ITK.
Finally, Picture Archiving and Communication System (PACS), which helps users to
economically store and conveniently access images, are considered out of scope.

\subsection{Methodology} \label{SecMethodology}

We have a standard set of questions designed to assess the qualities of any
research software project.  [blind review redacted details on the author's
previous studies using the same methodology.]
%% DOUBLE BLIND
% ~\cite{SmithEtAl2021, SmithAndMichalski2022}.  This has
% been applied to MI software and Lattice Boltzmann Solvers \cite{SmithEtAl2024}.
% This builds off prior work to assess the state of the practice for such domains
% as Geographic Information Systems \cite{smith2018state}, Mesh Generators
% \cite{smith2016state}, Seismology software \cite{Smith2018Seismology}, and
% Statistical software for psychology \cite{smith2018statistical}.  
We maintain the previous constraint that the work load for measuring a given
domain should take around one person-month's worth of effort ($20$ working days
at $8$ person-hours per day).

We identify a list of potential packages (through online searches) which is then
filtered and vetted by a domain expert. We aim for roughly $30$ packages. For
each remaining package, we measure its qualities by filling in a grading
template [citation redacted for double blind]. %%DOUBLE BLIND
%~\cite{SmithEtAl2021}.  
This data is used to rank the projects with the
Analytic Hierarchy Process (AHP).  We summarize further details on the
interaction with the domain expert, software qualities, grading the software and
AHP below and in longer form in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024)~\cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Domain Expert} \label{sec_vet_software_list}

The Domain Expert vets the proposed list because online resources can be
inaccurate.  The expert also vets the AHP ranking.  For the current assessment,
our Domain Expert is [details of our domain expert removed for double-blind].
%%DOUBLE BLIND %our Domain Expert (and paper co-author) is Dr.\ Michael
%Noseworthy, Professor of Electrical and Computer Engineering at McMaster
%University, Co-Director of the McMaster School of Biomedical Engineering, and
%Director of Medical Imaging Physics and Engineering at St.\ Joseph's
%Healthcare, Hamilton, Ontario, Canada.

In advance of the first meeting with the Domain Expert, they are asked to
independently create a list of top software packages in the domain.  This helps
get the expert's knowledge refreshed in advance of the meeting.

\subsubsection{Software Qualities} \label{sec_software_quality}

Quality is defined as a measure of the excellence or worth of an entity.  As is
common practice, we do not think of quality as a single measure, but rather as
a set of measures.  That is, quality is a collection of different qualities,
often called ``ilities.''  For this study we selected 9 qualities to measure:
installability, correctness/ verifiability, reliability, robustness, usability,
maintainability, reusability, understandability, and visibility/transparency.
With the exception of installability, all the qualities are defined in Ghezzi
et al. (2003) \cite{GhezziEtAl2003}. Installability is defined as the effort
required for the installation and/or uninstallation of software in a specified
environment \cite{ISO/IEC25010}.

\subsubsection{Grading} \label{sec_grading_software}

We use an existing template [citation redacted]
%% DOUBLE BLIND
%~\cite{SmithEtAl2021} 
that is designed to measure the
aforementioned qualities. To stay within our given measurement time frame, each
package gets up to five hours of time.  Project developers can be contacted for
help regarding installation, if necessary, but we impose a cap of about two
hours on the installation process.  Figure~\ref{fg_grading_template_example}
shows an excerpt of the measurement spreadsheet.  The rows are the measures and
the columns correspond to the software packages.  [The full data is available on
Mendeley; link will be provided after refereeing.] 
% \cite{Dong2021-Data}
%provides the full set of measurement data.   %%DOUBLE BLIND

\begin{figure*}[!ht]
\includegraphics[scale=0.67]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure*}

The full template consists of 108 questions over 9 qualities.  These
questions are designed to be unambiguous, quantifiable, and measurable
with constrained time and domain knowledge.

The grader, after answering questions for each quality assigns an overall
score (between 1 and 10) based on the answers.  Several of the qualities use
the word ``surface'' to highlight that these particular qualities are a shallow
measure.  For example, usability is not measured using user studies.
Instead, we look for signs that the developers considered usability.
We use two freeware tools to collect repository related data:
\href{https://github.com/tomgi/git_stats}{GitStats} and
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}.  Further
details on quality measurement are provided in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024) \cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Analytic Hierarchy Process (AHP)} \label{sec_AHP}

Developed by Saaty in the 1970s, AHP is widely used to analyze multiple criteria
decisions~\cite{VaidyaEtAl2006}. AHP organizes multiple criteria in a
hierarchical structure and uses pairwise comparisons between alternatives to
calculate relative ratios~\cite{Saaty1990}. AHP works with sets of $n$
\textit{options} and $m$ \textit{criteria}.  In our project $n=29$ and $m=9$
since there are 29 options (software products) and 9 criteria (qualities). With
AHP the sum of the grades (scores) for all products for a given quality will be
1.0.  We rank the software for each of the qualities, and then we combine the
quality rankings into an overall ranking based on the relative priorities
between qualities.

\subsubsection{Interview Methods} \label{sec_interview_methods}

The repository-based measurements are incomplete because they don't generally
capture the development process, the developer pain points, the perceived
threats to software quality, and the developers' strategies to address these
threats.  Therefore, part of our methodology involves interviewing developers.
We based our interviews on a list of 20 questions, which can be found in
[citation redacted for double blind].
%Smith et al. \cite{SmithEtAl2021}. 
Some questions are about the background of the software, the development teams,
the interviewees, and how they organize their projects.  We also ask about the
developer's understanding of the users. Some questions focus on the current and
past difficulties, and the solutions the team has found, or plan to try. We also
discuss documentation, both with respect to how it is currently done, and how it
is perceived. A few questions are about specific software qualities, such as
maintainability, understandability, usability, and reproducibility. The
interviews are semi-structured based on the question list.

\section{In-Scope Open-Source MI Software} \label{SecWhatProjects}

We initially identified 48 candidate software projects from the literature
\cite{Bjorn2017, Bruhschwein2019, Haak2015}, on-line articles \cite{Emms2019,
Hasan2020, Mu2019}, and forum discussions \cite{Samala2014}.  Then we filtered
as follows:

\begin{enumerate}

\item Removed the packages with no source code available, such as
\textit{MicroDicom}, \textit{Aliza}, and \textit{jivex}.

\item Focused on MI software that provides visualization functions.  We removed
seven packages that were toolkits or libraries, such as \textit{VTK},
\textit{ITK}, and \textit{dcm4che}, and another three that were for PACS.

\item Removed \textit{Open Dicom Viewer} as it has not received any
updates since 2011.

\end{enumerate}

The Domain Expert provided a list 12 software packages.  We found 6 packages
were on both lists: \textit{3D Slicer}, \textit{Horos}, \textit{ImageJ},
\textit{Fiji}, \textit{MRIcron} (we use its descendant \textit{MRIcroGL}) and
\textit{Mango} (we use the web version \textit{Papaya}).  The remaining six
packages were on our out-of-scope list. The Domain Expert agreed with our final
choice of 29 packages.  Table~\ref{tab_final_list} summarizes the in-scope
open-source MI software that is available at the time of measurement (the year
2020), thus answering \rqref{RQ_WhatProjects}.

\begin{table*}[!ht]
\centering
\begin{tabular}{p{3.7cm}lllllllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cmidrule{7-9}
 &  &  &  &  &  & W & M & L &  \\ \midrule
ParaView \cite{Ahrens2005} & 2002 & 2020-10 & \checkmark & 100 & 886326 & \checkmark & \checkmark & \checkmark & \checkmark \\
Gwyddion \cite{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & \checkmark & \checkmark & \checkmark &  \\
Horos \cite{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & \checkmark &  &  \\
OsiriX Lite \cite{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & \checkmark &  &  \\
3D Slicer \cite{Kikinis2014} & 1998 & 2020-08 & \checkmark & 100 & 501451 & \checkmark & \checkmark & \checkmark &  \\
Drishti \cite{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & \checkmark & \checkmark & \checkmark &  \\
Ginkgo CADx \cite{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & \checkmark & \checkmark & \checkmark &  \\
GATE \cite{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & \checkmark & \checkmark &  \\
3DimViewer \cite{TESCAN2020} & ? & 2020-03 & \checkmark & 3 & 178065 & \checkmark & \checkmark &  &  \\
medInria \cite{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & \checkmark & \checkmark & \checkmark &  \\
BioImage Suite Web \cite{Papademetris2005} & 2018 & 2020-10 & \checkmark & 13 & 139699 &
\checkmark & \checkmark & \checkmark & \checkmark \\
Weasis \cite{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & \checkmark & \checkmark & \checkmark &  \\
AMIDE \cite{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & \checkmark & \checkmark & \checkmark &  \\
XMedCon \cite{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & \checkmark & \checkmark & \checkmark &  \\
ITK-SNAP \cite{Yushkevich2006} & 2006 & 2020-06 & \checkmark & 13 & 88530 & \checkmark & \checkmark & \checkmark &  \\
Papaya \cite{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & \checkmark & \checkmark & \checkmark &  \\
OHIF Viewer \cite{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & \checkmark & \checkmark & \checkmark & \checkmark \\
SMILI \cite{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & \checkmark & \checkmark & \checkmark &  \\
INVESALIUS 3 \cite{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & \checkmark & \checkmark & \checkmark &  \\
dwv \cite{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & \checkmark & \checkmark & \checkmark & \checkmark \\
DICOM Viewer \cite{Afsar2021} & 2018 & 2020-04 & \checkmark & 5 & 30761 & \checkmark & \checkmark & \checkmark &  \\
MicroView \cite{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & \checkmark & \checkmark & \checkmark &  \\
MatrixUser \cite{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & \checkmark & \checkmark & \checkmark &  \\
Slice:Drop \cite{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & \checkmark & \checkmark & \checkmark & \checkmark \\
dicompyler \cite{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & \checkmark & \checkmark &  &  \\
Fiji \cite{Schindelin2012} & 2011 & 2020-08 & \checkmark & 55 & 10833 & \checkmark & \checkmark & \checkmark &  \\
ImageJ \cite{Rueden2017} & 1997 & 2020-08 & \checkmark & 18 & 9681 & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL \cite{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & \checkmark & \checkmark & \checkmark &  \\
DicomBrowser \cite{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & \checkmark & \checkmark & \checkmark &  \\ \bottomrule
\end{tabular}
\caption{Final software list (sorted in descending order of the number of Lines
Of Code (LOC))}
\label{tab_final_list}
\end{table*}

In Table~\ref{tab_final_list} the projects are sorted in descending order of
lines of code.  We found the initial release dates (Rlsd) for most projects and
marked the two unknown dates with ``?''. The date of the last update is the date
of the latest update, at the time of measurement. We found funding information
(Fnd) for only eight projects.  For the Number Of Contributors (NOC) we
considered anyone who made at least one accepted commit as a contributor. The
NOC is not usually the same as the number of long-term project members, since
many projects received change requests and code from the community.  With
respect to the OS, 25 packages work on all three OSs: Windows (W), macOS (M),
and Linux (L). Although the usual approach to cross-platform compatibility was
to work natively on multiple OSes, five projects achieved platform-independence
via web applications. The full measurement data for all packages is available on
[removed for blind review]
%\href{https://data.mendeley.com/datasets/k3pcdvdzj2/1} {Mendeley Data}.  
%% DOUBLE BLIND

The programming languages used in order of decreasing popularity are \CC,
JavaScript, Java, C, Python, Pascal, Matlab.  The most popular language is \CC,
for 11 of 29 projects; Pascal and Matlab were each used for a single project.

\section{Which Projects Follow Best Practices?} \label{SecMeasurementResults}

To answer \rqref{RQ_HighestQuality} we measured the software as described in
Section~\ref{SecMethodology}.  In the absence of a specific real world context,
we assumed all nine qualities are equally important.
Figure~\ref{fg_overall_scores} shows the overall AHP scores in descending order.

\begin{figure*}[!ht]
\includegraphics[scale=0.47]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

\label{fg_overall_scores}
\end{figure*}

The top four software products \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, and \textit{OHIF Viewer} have higher scores in most criteria.
\textit{3D Slicer} has a score in the top two for all qualities; \textit{ImageJ}
ranks near the top for all qualities, except for correctness \& verifiability.
\textit{OHIF Viewer} and \textit{Fiji} have similar overall scores, with
\textit{Fiji} doing better in installability and \textit{OHIF Viewer} doing
better in correctness \& verifiability.  Given the installation problems, we may
have underestimated the scores on reliability and robustness for \textit{DICOM
Viewer}, but we compared it equally for the other seven qualities.

The overall score is based on the measurement of the identified qualities. Full
details of how the projects compare for each quality can be found in [citation
redacted for blind review].
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2024_MI_SOP}
The highlights are as follows:

\begin{description}

\item [Installability] We found installation instructions for 16 projects, but
two did not need them (\textit{BioImage Suite Web} and \textit{Slice:Drop}) as
they are web applications. 10 of the projects required extra dependencies: Five
depend on a specific browser; \textit{dwv}, \textit{OHIF Viewer}, and
\textit{GATE} needs extra libraries to build; \textit{ImageJ} and \textit{Fiji}
need an unzip tool; \textit{MatrixUser} needs Matlab; \textit{DICOM Viewer}
needs a Nextcloud platform. \textit{GATE}, \textit{dwv}, and \textit{DICOM
Viewer} showed severe installation problems. We were not able to install them,
even after a reasonable amount of time (2 hours).

\item [Correctness \& Verifiability] The packages with higher scores for
correctness and verifiability used a wider array of techniques to improve
correctness, and had better documentation to witness this.  For instance, we
looked for evidence of unit testing, and found evidence for only about half of
the projects. We identified five projects using continuous integration tools:
\textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{dwv}, and
\textit{OHIF Viewer}. The only requirements-related document we found was a road
map of \textit{3D Slicer}, which contained design requirements for upcoming
changes.

\item [Surface Reliability] We were able to follow the steps in the tutorials
that existed (seven packages had them.) However, \textit{GATE} could not open
macro files and became unresponsive several times, without any descriptive error
message. We found that \textit{Drishti} crashed when loading damaged image
files, without showing any descriptive error message.

\item [Surface Robustness] According to their documentation, all 29 software
packages should support the DICOM standard. To test robustness, we prepared two
types of image files: correct and incorrect formats (with the incorrect format
created by relabelling a text file to have the ``.dcm'' extension).  All
software packages loaded the correct format image, except for \textit{GATE},
which failed for unknown reasons.  For the broken format, \textit{MatrixUser},
\textit{dwv}, and \textit{Slice:Drop} ignored the incorrect format, did not show
any error message and displayed a blank image.  \textit{MRIcroGL} behaved
similarly except that it showed a meaningless image.  \textit{Drishti}
successfully detected the broken format of the file, but the software crashed as
a result.

\item [Surface Usability] The software with higher usability scores usually
provided both comprehensive documented guidance and a good user experience.
\textit{INVESALIUS 3} provided an excellent example of a detailed and precise
user manual. \textit{GATE} also provided numerous documents, but unfortunately
we had difficulty understanding and using them. We found getting started
tutorials for only 11 projects, but a user manual for 22 projects.
\textit{MRIcroGL} was the only project that explicitly documented expected user
characteristics.

\item [Maintainability] We gave \textit{3D Slicer} the highest score for
maintainability because we found it had the most comprehensive artifacts. Only a
few of the 29 projects had a product, developer's manual, or API (Application
Programming Interface) documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents.  Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%). Twenty-seven of the 29
projects used git for version control, with 24 of these using GitHub.
\textit{AMIDE} used Mercurial and \textit{Gwyddion} used Subversion.
\textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used SourceForge.
\textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

\item [Reusability] We have assumed that smaller code files are
likely more reusable -- see Table \ref{tab_loc_per_file} for the details.

\item [Surface Understandability] All projects had a consistent coding style
with parameters in the same order for all functions, modularized code, and,
clear comments that indicate what is done, not how. However, we only found
explicit identification of a coding standard for 3 out of the 29: \textit{3D
Slicer}, \textit{Weasis}, and \textit{ImageJ}. We also found hard-coded
constants (rather than symbolic constants) in \textit{medInria},
\textit{dicompyler}, \textit{MicroView}, and \textit{Papaya}. We did not find
any reference to the algorithms used in projects \textit{XMedCon},
\textit{DicomBrowser}, \textit{3DimViewer}, \textit{BioImage Suite Web},
\textit{Slice:Drop}, \textit{MatrixUser}, \textit{DICOM Viewer},
\textit{dicompyler}, and \textit{Papaya}.

\item [Visibility/Transparency] Generally speaking, the teams that actively
documented their development process and plans scored higher.  \textit{3D
Slicer} and \textit{ImageJ} were the only projects to include documentation for
all of the following: development process, project status, development
environment and release notes.

\end{description}

\begin{table*}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (by reusability scores)}
\label{tab_loc_per_file}
\end{table*}

\section{Comparison to Community Ranking} \label{Sec_VsCommunityRanking}

To answer \rqref{RQ_CompareHQ2Popular} we use GitHub stars, number of forks and
number of people watching the projects are proxies for community ranking.
Table~\ref{tab_ranking_vs_GitHub} show the statistics for data collected in July
2021.  Recall that 24 projects use GitHub (Section~\ref{SecMeasurementResults}).

Our ranking and GitHub popularity, at least for the top five projects, seems to
line up fairly well. However, we ranked some popular packages fairly low, such
as \textit{dwv}. This is because we were unable to build it locally, even though
we followed its installation instructions. However, we were able to use its web
version for the rest of the measurements. Additionally, this version did not
detect a broken DICOM file and instead displayed a blank image
(Section~\ref{SecMeasurementResults}).  \textit{DICOM Viewer} ranked low as we
were unable to install the NextCloud platform.

Another likely reason for discrepancies is that we weighted all qualities
equally. This is not likely how users implicitly rank the different qualities.
This would require a broader user study to properly assess. Furthermore our
measures of popularity are only \emph{proxies} which are biased towards
past rather than current preferences~\cite{Szulik2017}, as these are
monotonically increasing quantities. Finally there are often more factors than
just quality that influence the popularity of ``consumer'' products.

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table*}[!ht]
\centering
\begin{tabular}{llllll}
\toprule
Software & Comm.\ Rank & Our Rank & Stars/yr & Watches/yr & Forks/yr \\ 
\midrule
3D Slicer & 1 & 1 & 284 & 19 & 128 \\
OHIF Viewer & 2 & 4 & 277 & 19 & 224 \\
dwv & 3 & 19 & 124 & 12 & 51 \\
ImageJ & 4 & 2 & 84 & 9 & 30 \\
ParaView & 5 & 5 & 67 & 7 & 28 \\
Horos & 6 & 12 & 49 & 9 & 18 \\
Papaya & 7 & 17 & 45 & 5 & 20 \\
Fiji & 8 & 3 & 44 & 5 & 21 \\
DICOM Viewer & 9 & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 8 & 40 & 4 & 17 \\
Weasis & 11 & 7 & 36 & 5 & 19 \\
dicompyler & 12 & 26 & 35 & 5 & 14 \\
OsiriX Lite & 13 & 11 & 34 & 9 & 24 \\
MRIcroGL & 14 & 18 & 24 & 3 & 3 \\
GATE & 15 & 24 & 19 & 6 & 26 \\
Ginkgo CADx & 16 & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 17 & 6 & 18 & 5 & 7 \\
Drishti & 18 & 27 & 16 & 4 & 4 \\
Slice:Drop & 19 & 21 & 10 & 2 & 5 \\
ITK-SNAP & 20 & 13 & 9 & 1 & 4 \\
medInria & 21 & 9 & 7 & 3 & 6 \\
SMILI & 22 & 10 & 3 & 1 & 2 \\
MatrixUser & 23 & 28 & 2 & 0 & 0 \\
MicroView & 24 & 15 & 1 & 1 & 1 \\
Gwyddion & 25 & 16 & n/a & n/a & n/a \\
XMedCon & 26 & 20 & n/a & n/a & n/a \\
DicomBrowser & 27 & 22 & n/a & n/a & n/a \\
AMIDE & 28 & 23 & n/a & n/a & n/a \\
3DimViewer & 29 & 25 & n/a & n/a & n/a \\ 
\bottomrule
\end{tabular}
\caption{Software ranking by our methodology versus the community (Comm.)\
ranking using GitHub metrics (Sorted in descending order of community
popularity, as estimated by the number of new stars per year)}
\label{tab_ranking_vs_GitHub}
\end{table*}
\endgroup

Although both rankings are imperfect measures, they nevertheless suggest a
correlation between best practices and popularity. We don't know if this is
causal, in either direction (i.e. if best practices enable popularity or if
popularity increases the need for using more software development best
practices).

\section{Compare to Recommended Software Artifacts} \label{Sec_CompareArtifacts}

For \rqref{RQ_CompareArtifacts} we use nine research software development
guidelines to compare recommended software artifacts versus those present in MI
software.  These guidelines are:
\begin{itemize}
\item United States Geological Survey Software Planning Checklist
\cite{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\cite{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\cite{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \cite{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \cite{SmithAndRoscoe2018},
\item Trilinos Developers Guide \cite{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \cite{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \cite{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \cite{OrvizEtAl2017}.
\end{itemize}

\begin{table*}[!ht]
\begin{center}
\begin{tabular}{ p{3.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} &
\cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} &
\cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table*}

In Table~\ref{Tbl_Guidelines} each row corresponds to an artifact.  For a given
row, a checkmark in one of the columns means that the corresponding guideline
recommends this artifact.  The last column shows whether the artifact appears in
the measured set of MI software, either not at all (blank), commonly (C),
uncommonly (U) or rarely (R).  We did our best to interpret the meaning of each
artifact consistently between guidelines and specific MI software, but the
terminology and the contents of artifacts are not standardized.  The challenge
even exists for the ubiquitous README file.  The content of README files shows
significant variation between projects \cite{PranaEtAl2018}.  Although some
content is reasonably consistent, with 97\% of README files contain at least one
section describing the `What' of the repository and 89\% offering some `How'
content, other categories are more variable.  For instance, information on
`Contribution', `Why', and `Who', appear in 28\%, 26\% and 53\% of the analyzed
files, respectively \cite{PranaEtAl2018}.

\begin{table*}[!ht]
    \begin{center}
    \begin{tabular}{ p{3.8 cm} p{6.4 cm} p{4.5 cm}}
    \toprule
    Common & Uncommon & Rare \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    User manual (22) & Authors (14) & Dependency list (7)\\
    Release info. (22) & Frequently Asked Questions (14) & Troubleshooting guide (6)\\
     & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design documentation (5)\\
     & Citation (11) & Code style guide (3)\\
     & & Code of conduct (1)\\
     & & Requirements (1)\\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency (The number 
    in brackets is the number of occurrences)}
    \label{artifactspresent}
    \end{center}
\end{table*}

Table~\ref{artifactspresent} presents our measurements for MI software.
The table groups the artifacts by frequency
into categories of common (20 to 29 ($>$67\%) packages), uncommon (10 to 19
(33-67\%) packages), and rare (1 to 9 ($<$33\%) packages).

Note that ``popularity'' in Table~\ref{Tbl_Guidelines} does no imply that these
oft recommended artifacts are the most important. Guidelines are often brief,
to encourage adoption, and thus even guidelines that mention the need for
installation instructions rarely mention uninstallation instructions.
Two items in Table~\ref{artifactspresent} do not appear in any guidelines:
\emph{Troubleshooting guide} and \emph{Developer's manual}.  However the
information within these documents overlaps with
the recommended artifacts.  Troubleshooting information often can be found in
a User Manual, while the information in a ``Developer's Manual'' is often
scattered amongst many other documents.

Three of the 26 recommended artifacts were never observed in the MI software:
i) Uninstall, ii) Test plans, and iii) Requirements. It is possible that
some of these were created but never put under version control.

Neglecting requirements documentation is unfortunately common for research
software, and MI software is no exception to this trend.  Although such
documentation is recommended by some \cite{TobiasEtAl2018, HerouxEtAl2008,
SmithAndKoothoor2016}, in practice this is rare~\cite{HeatonAndCarver2015}.
Sanders and Kelly \cite{SandersAndKelly2008} interviewed 16 scientists from 10
disciplines and found that none of the scientists created requirements
specifications, unless regulations in their field mandated such a document.
Requirements are the least commonly produced type of documentation for research
software in general \cite{Nguyen-HoanEtAl2010}.

This is unfortunate as when scientific developers are surveyed on their pain
points, Wiese et al.~\cite{WieseEtAl2019} found that software requirements and
management is the software engineering discipline that most hurts them,
accounting for 23\% of the technical problems reported by study participants.
Further adding to the misfortune, there is a widespread perception that up-front
requirements are impossible for research software \cite{CarverEtAl2007,
SegalAndMorris2008}. Fortunately a more agile approach to requirements is
feasible~\cite{Smith2016}, and research-software specific templates
exist~\cite{SmithEtAl2007}.

A theme emerges amongst the artifacts rarely observed in practice: they
are developer-focused (a list of library dependencies, a contributor's guide,
a developer Code of Conduct, coding style guidelines, product roadmap, design
documentation and API documentation).

Other communities use checklists to help with best practices. Examples include
checklists for merging branches~\cite{Brown2015}, for saving and sharing changes
to the project~\cite{WilsonEtAl2016}, for new and departing team
members~\cite{HerouxAndBernholdt2018}, for processes related to commits and
releases~\cite{HerouxEtAl2008} and for overall software
quality~\cite{ThielEtAl2020, SSI2022}.

MI projects fall somewhat short of recommended best practices, but are not alone
amongst research software projects. This gap has been documented
before~\cite{Storer2017,JohansonAndHasselbring2018}, and is known to cause
sustainability and reliability problems \cite{FaulkEtAl2009}, and to waste
development effort~\cite{deSouzaEtAl2019}.

\section{Comparison of Tool Usage Between MI and Other Research Software}
\label{Sec_CompareTools}

We use data from interviews, as described in Section~\ref{sec_interview_methods}
to answer \rqref{RQ_CompareToolsProjMngmnt}, \rqref{RQ_CompareMethodologies}
and \rqref{RQ_PainPoints}. Requests were sent to all 29 projects.  Nine
developers from eight of the projects agreed to participate: \textit{3D Slicer},
\textit{INVESALIUS 3}, \textit{dwv}, \textit{BioImage Suite Web},
\textit{ITK-SNAP}, \textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We
spent about 90 minutes for each interview. The full interview answers can be
found in [citation redacted for blind review].
%\cite{Dong2021}.

Developers use software tools for user support, version control, Continuous
Integration and Deployment (CI/CD), documentation and project management. To
answer \rqref{RQ_CompareToolsProjMngmnt}, we summarize the tool usage in these
categories, and compare this to the usage by the research software community.

Table~\ref{tab_user_support_model} summarizes the user support models by the
number of projects using each model (projects may use more than one support
model). The prevalent usage of GitHub issues may be by design, or it might just
naturally arise as users seek help.

\begin{table}[!ht]
\centering
\begin{tabular}{lc}
\toprule
\multicolumn{1}{c}{User Support Model} & Num.\ Projects \\
\midrule
GitHub issue & 24 \\
Frequently Asked Questions (FAQ) & 12 \\
Forum & 10 \\
E-mail address & 9 \\
GitLab issue, SourceForge discussions & 2 \\
Troubleshooting & 2 \\
Contact form & 1 \\ 
\bottomrule
\end{tabular}
\caption{\label{tab_user_support_model}User support models by number of projects}
\end{table}

Most teams interview (eight of nine) mentioned using GitHub and pull requests to manage
contributions from the community. The interviewees generally gave very positive
feedback on using GitHub. Some teams previously used a different approach to
version control and eventually transferred to git and GitHub.

The common use of version control for MI software illustrates considerable
improvement from the poor adoption of version control tools that Wilson lamented
in 2006 \cite{Wilson2006}.  The proliferation of version control tools for MI
matches the increase in the broader research software community.  A little over
10 years ago Nguyen-Hoan et al.\ \cite{Nguyen-HoanEtAl2010} estimated that only
50\% of research software projects used version control, but even at that time
there was an increase from previous usage levels. A survey in 2018 shows 81\% of
developers use a version control system \cite{AlNoamanyAndBorghi2018}.
\cite{Smith2018} has similar results, showing version control usage for alive
projects in mesh generation, geographic information systems and statistical
software for psychiatry increasing from 75\%, 89\% and 17\% (respectively) to
100\%, 95\% and 100\% (respectively) over a four-year period ending in 2018. A
recent survey by \cite{CarverEtAl2022} shows version control use among
practitioners at over 95\%. All but one of the software guides cited in
Section~\ref{Sec_CompareArtifacts} recommends version control. The high usage of
version control tools in MI matches the general trend in research software.

As mentioned in Section~\ref{SecMeasurementResults}, we identified 17\% of MI
projects using CI/CD tools. We found which projects used CI/CD by examining the
documentation and source code. The count of CI/CD usage may be higher, since
traces of CI/CD may not always appear in a repository.  This was the case for a
study of Lattice Boltzmann Method (LBM) software, where interviews with
developers showed that more projects used CI/CD than was evident from repository
artifacts alone [citation redacted for blind review].  %\cite{Michalski2021}.  
The 17\% utilization for MI software contrasts with the high frequency with
which research software development guidelines recommend continuous integration
\cite{BrettEtAl2021, Brown2015, ThielEtAl2020, Zadka2018, vanGompelEtAl2016}.
Although there is currently little data available on CI/CD utilization for
research software, our impression is that CI/CD is not yet common practice,
despite its recommendation.  This is certainly the case for LBM software, where
usage numbers sit around 12.5\% [citation redacted for blind review]
%\cite{Michalski2021}.  
The survey of \cite{CarverEtAl2022} suggests higher use of CI/CD with 54\%
(54/100) of respondents indicating that they use it. However, that study surveys
practitioners, rather than assessing projects.  Additional information on CI/CD
is given in the recommendations (Section~\ref{Sec_ContinuousIntegration}).

For documentation tools and methods mentioned by the interviewees, the most
popular (mentioned by about 30\% of developers) were forum discussions and
videos.  The second most popular options (mentioned by about 20\% of developers)
were GitHub, wiki pages, workshops, and social media. The least frequently
mentioned options (about 10\% of developers) included writing books, and google
forms.  In contrasting MI software with LBM software, the most significant
documentation tool difference is that LBM software often uses document
generation tools, like doxygen and sphinx [citation redacted for blind review].
%\cite{Michalski2021}

The interviewees talked about two types of project management tools:
\begin{inparaenum}[i)]
\item trackers, including GitHub, issue trackers, bug trackers and Jira; and,
\item documentation tools, including GitHub, Wiki page, Google Doc, and
Confluence.
\end{inparaenum}
Of the specifically named tools, interviewees mentioned GitHub 3 times, and each
of the other tools once each.

Based on information provided by \cite{JungEtAl2022}, tool utilization for MI
software has much in common with tool utilization for ocean modelling software.
Both use tools for editing, compiling, code management, testing, building, and
project management.  From the data available, ocean modelling differs from MI
software in its use of Kanban boards for project management.

\section[Comparison to Other Research Software]{Comparison of Principles, Process, and
Methodologies to Research Software in General} \label{Sec_CompareMethodologies}

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes, and methodologies used for MI software to what can be
gleaned from the literature on research software in general. In our interviews
with developers the responses about development model were vague, with only two
interviewees following a definite development model. In some cases the
interviewees felt their process was similar to an existing development model.
Three teams (about 38\%) either followed agile, or something similar to agile.
Two teams (25\%) either followed a waterfall process, or something similar.
Three teams (about 38\%) explicitly stated that their process was undefined or
self-directed.

Our observations of an informally defined process, with elements of agile
methods, matches observations for research software in general.
Scientific developers naturally use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process \cite{Kelly2015}.  A waterfall-like process can work for research
software \cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rationale design process \cite{parnas1986rational}.

No interviewee introduced any strictly defined project management process. The
most common approach was following the issues, such as bugs and feature
requests. The \textit{3D Slicer} team had weekly meetings to
discuss the goals for the project. The \textit{INVESALIUS 3} team relied on the
GitHub process for their project management while the \textit{ITK-SNAP} team had a
fixed six-month release pace. Only the interviewee from \textit{OHIF} 
mentioned the team having a project manager. The \textit{3D Slicer} team and
\textit{BioImage Suite Web} team do nightly builds and tests. The \textit{OHIF}
developer believes that a better project management process can improve junior
developer efficiency while also improving internal and external communication.

We identified the use of unit testing in less than half of the 29 projects even
though the interviewees believed that testing (including usability tests with
users) was the top approach to improve correctness, usability, and
reproducibility.  This level of testing matches what was observed for LBM
software [citation redacted for blind review]
%\cite{Michalski2021} 
and is apparently greater than the level of
testing for ocean modelling software \cite{JungEtAl2022}.

Interviewees from all of the eight projects thought that documentation was
essential to their projects, and most of them (7 of 8) said that it could save
their time to answer questions from users and developers. Half of them (4 of 8)
saw the need to improve their documentation, and only three of them thought that
their documentations conveyed information clearly enough.

As Table~\ref{Tbl_Guidelines} suggests, an emphasis on documentation, especially
for new developers, is echoed in research software guidelines. Multiple
guidelines recommend a document explaining how to contribute to a project, often
named CONTRIBUTING. Guidelines also recommend tutorials, user guides and quick
start examples. Smith and Roscoe \cite{SmithAndRoscoe2018} suggest including instructions
specifically for on-boarding new developers. For open-source software in general
(not just research software), \cite{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos, and screenshots.

\section{Developer Pain Points} \label{painpoints}

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer three research questions (first
mentioned in Section~\ref{sec_motivation}): \rqref{RQ_PainPoints}) What are the
pain points for developers working on research software projects?;
\rqref{RQ_ComparePainPoints}) How do the pain points of developers from MI
compare to the pain points for research software in general?; and
\rqref{RQ_Concerns}) For MI developers what specific best practices are taken to
address the pain points and software quality concerns?

Our interviews identified pain points related to a lack of time and funding,
technology hurdles, improving correctness, and improving usability.  In this
section, we go through each pain point and contrast the MI experience with
observations from other domains.  We also cover potential ways to address the
pain points, as promoted by the community.  In addition to pain points, we summarize
MI developer strategies for improving maintainability and reproducibility.
Although the interviewees did not explicitly identify these two qualities as
pain points, we did discuss threats to these qualities and ways to improve them
as part of our interview process \cite{SmithEtAl2021}.  The interviewee's
practices for addressing pain points and improving quality can potentially be
emulated by other MI developers. Moreover, these practices may provide examples
that can be followed by other research software domains.

Pinto et al.\ \cite{PintoEtAl2018} list some pain points that did not come up in
our conversations with MI developers: interruptions while coding, scope bloat,
lack of user feedback, hard to collaborate on software projects, and aloneness.
Wiese et al.\ \cite{WieseEtAl2019} mention two additional pain points that did
not explicitly arise in our interviews: reproducibility, and software scope
determination.  To the list of pain points not discussed for MI, our study of
LBM software [citation redacted for blind review] %\cite{SmithEtAl2024} 
adds lack of software experience for the developers, technical debt, and
documentation. We did not observe any pain points for MI that were not also
observed for LBM. From the pain points mentioned above, although the topics of
reproducibility and technical debt did not come up in our MI interviews, we
covered these two topics as part of the discussion of software qualities, as
summarized at the end of this section.  Although previous studies show pain
points that were not mentioned by MI developers, we cannot conclude that these
pain points are not relevant for MI software development, since we only
interviewed nine developers.

\begin{enumerate}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} Many interviewees thought lack of time, along with lack of
funding (discussed next), were their most significant obstacles. Other domains
of research software also experience the lack of time pain point
\cite{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}. Our study of LBM software [citation redacted for blind review]
%\cite{SmithEtAl2024} 
also highlighted lack of time as a significant pain point.

Potential and proven solutions suggested by the interviewees include:

\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for CI/CD (Continuous Integration and Continuous
Delivery.)
\end{itemize}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Funding:} Developers felt the pain of having to attract funding to develop and
maintain their software. For instance, the interviewees from \textit{3D Slicer}
and \textit{OHIF} said getting funding for software maintenance is more
challenging than finding funding for research. The interviewee from 
\textit{ITK-SNAP} thought more funding was a way to solve the lack of time
problem, because they could hire more dedicated developers. On the other hand,
the interviewee from \textit{Weasis} did not feel that funding could
solve the same problem, since they would still need time to supervise the project.

Funding challenges have also been noted by others \cite{GewaltigAndCannon2012,
Goble2014, KaterbowAndFeulner2018, SmithEtAl2024}. Researchers that devote time
to software have the additional challenge that funding agencies do not always
count software when they are judging the academic excellence of the applicant.
Wiese et al.\ \cite{WieseEtAl2019} reported developer pains related to
publicity, since publishing norms have historically made it difficult to get
credit for creating software.  As studied by \cite{HowisonAndBullard2016},
research software (specifically biology software, but the trend likely applies
to other research software domains) is infrequently cited. There is a lack of a
formal reward system for research software \cite{PintoEtAl2018}.

An interviewee proposed an idea for increasing funding: Licensing the software
to commercial companies to integrate it into their products.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} The technology hurdles mentioned by MI developers
include: hard to keep up with changes in OS and libraries, difficult to transfer
to new technologies, hard to support multiple OSes, and hard to support lower-end
computers. Developers expressed difficulty balancing between four factors:
cross-platform compatibility, convenience to development and maintenance,
performance, and security.

Another pain point survey \cite{WieseEtAl2019} highlights that technology
hurdles are an issue for research software in general.  Some technical-related
problems mentioned include dependency management, cross-platform compatibility,
CI, hardware issues and operating system issues \cite{WieseEtAl2019,
PintoEtAl2018}. Technology pain points for LBM developers include setting up
parallelization and CI [citation redacted for blind review].
%\cite{SmithEtAl2024}.

Five solutions proposed by the MI developers are:

\begin{itemize}
\item Adopting a web-based approach with backend servers, to better support
lower-end computers;
\item Using memory-mapped files to consume less computer memory, to better
support lower-end computers; 
\item Using computing power from the computers GPU for web applications;
\item Maintaining better documentations to ease the development and maintenance
processes;
\item Improving performance via more powerful computers, which one interviewee
pointed out has already happened.
\end{itemize}

As the above list shows, developers perceive that web-based applications will
address the technology hurdle.  Most of the 29 teams (83\%) chose to develop
native applications. For the eight teams we interviewed, three of them were
building web applications, and the \textit{MRIcroGL} team was considering a
web-based solution.

The advantage for native applications is higher performance, while web
applications have the advantage of cross-platform compatibility and a simpler
build process.  These web advantages mirror the native disadvantages of
difficulty with cross-platform compatibility and a complex build process.  The
lower performance disadvantage of web applications can be improved with a server
backend, but in this case there are disadvantages for privacy protection and
server costs.  These issues are discussed further in the recommendations
(Section~\ref{sec_webapps}).

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Ensuring Correctness:} Interviewees identified multiple threats to
correctness.  The most frequently mentioned threat was complexity.  Complexity
enters the software by various means, including the large variety of data formats,
complicated data standards, differing outputs between medical imaging machines,
and the addition of (non-viewing related) functionality.  Other threats to
correctness identified include the following:

\begin{itemize}
\item Lack of real world image data for testing, in part because of patient
privacy concerns (the pain point of privacy concerns also arises for research
software in general \cite{WieseEtAl2019});
\item Tests are expensive and time-consuming because of the need for huge datasets;
\item Software releases are difficult to manage;
\item No systematic unit testing; and,
\item No dedicated quality assurance team.
\end{itemize}

As implied by the above threats to correctness, testing was the most often
mentioned strategy for MI developers for ensuring correctness.  Seven teams
mentioned test related activities, including test-driven development, component
tests, integration tests, smoke tests, regression tests, self tests and
automated tests.  With the common emphasis on testing to improve correctness, MI
software is ahead of some other scientific domains.  For research software in
general, there is a problem of insufficient testing \cite{PintoEtAl2018} and
more developers think testing is important than the number that believe they
have a sufficient understanding of testing concepts \cite{HannayEtAl2009}.  Our
study of LBM software suggests that this domain shares the challenges of
insufficient testing and insufficient understanding of testing concepts
[citation redacted for double blind]. %\cite{SmithEtAl2024}.  
Automated testing
is a specific challenge for LBM software since free testing services do not
offer adequate facilities for large amounts of data.
Although not specifically mentioned during our interviews, the large data sets
for MI likely also cause a challenge for using free testing services, like
GitHub Actions.

Research software in general often struggles with the oracle problem for testing
because for many potential test cases the developer doesn't have a means to
judge the correctness of their calculated solutions \cite{HannayEtAl2009,
KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}.  The MI developers did not
allude to this challenge, likely because for a give image (test case) it is
possible to determine, potentially by using other software, the expected
analysis results.

A frequently cited strategy for building confidence in correctness (mentioned by
3 interviewees) is a two state development process with stable releases and
nightly builds.  Other strategies for ensuring correctness that came up during
the interviews include CI/CD, using de-identified copies of medical images for
debugging, sending beta versions to medical workers who can access the data to
do the tests, and collecting/maintaining a dataset of problematic images.  Some
additional strategies used by MI developers include:

\begin{itemize}
\item Using open datasets.
\item If (part of) the team belongs to a medical school or a hospital, using the
datasets they can access;
\item If the team has access to MRI scanners, self-building sample images for
testing;
\item If the team has connections with MI equipment manufacturers, asking for
their help on data format problems;
\end{itemize}

The feedback from the interviewees makes it clear that increased connections
between the development team and medical professionals/institutions could ease
the pain of ensuring correctness via testing.

\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:] \textbf{Usability:}
The discussion with the developers focused on usability issues for two classes
of users: the end users and other developers.  The threats to usability for end
users include an unintuitive user interface, inadequate feedback from the
interface (such as lack of a progress bar), users being unable to determine the
purpose of the software, not all users knowing if the software includes certain
features, not all users understanding how to use the command line tool, and not
all users understanding that the software is a web application. For developers
the threats to usability include not being able to find clear instructions on
how to deploy the software, and the architecture being difficult for new
developers to understand.

At least to some extent the problems for MI software users are due to holes in
their background knowledge.  For research software in general, users do not
always have the expertise required to install or use the software
\cite{WieseEtAl2019}. A similar pattern has been observed during interviews with
LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

To improve the usability of MI software, the most common strategies mentioned by
developers are as follows:

\begin{itemize}
    \item Use documentation (user manuals, mailing lists, forums) (mentioned by
    4 developers)
    \item Usability tests and interviews with end users; and, (mentioned by 3
    developers)
    \item Adjusting the software according to user feedback. (mentioned by 3
    developers)
\end{itemize}

Other suggested and practiced strategies include a graphical user interface,
testing every release with active users, making simple things simple and
complicated things possible, focusing on limited number of functions, icons with
clear visual expressions, designing the software to be intuitive, having a UX
(User eXperience) designer, dialog windows for important notifications,
providing an example for users to follow, downsampling images to consume less
memory, and providing an option to load only part of the data to boost
performance.  The last two points recognize that an important component of
usability is performance, since poor performance frustrates users.

\end{enumerate}

Up to this point, we have covered the pain points that came up in interviews
with MI developers, along with a summary of the techniques that are currently
used to address these pain points.  Although the developers did not explicitly
identify the qualities of maintainability and reproducibility as pain points in
our interviews, as part of our interview questions
(Section~\ref{sec_interview_methods}) they did share their approaches for
improving these qualities, as discussed below.

\begin{enumerate} \item[Q\refstepcounter{qnum}\theqnum
\label{Q_Maintainability}:] \textbf{Maintainability:}  Maintainability has been
rated as the third most important software quality for research software
\cite{Nguyen-HoanEtAl2010}. The push for sustainable software
\cite{deSouzaEtAl2019} is motivated by the pain that past developers have had
with accumulating too much technical debt \cite{KruchtenEtAl2012}.  For LBM
software, technical debt has been identified as one of the developer pain points
[citation redacted for blind review]. %\cite{SmithEtAl2024}.

To improve maintainability, the most popular (with five out of nine interviewees
mentioning it) strategy is to use a modular approach, with often repeated
functions in a library.  Other strategies that were mentioned for improving
maintainability include supporting third-party extensions, an easy-to-understand
architecture, a dedicated architect, starting from simple solutions, and
documentation.  The \textit{3D Slicer} team used a well-defined structure for
the software, which they named as an ``event-driven MVC pattern''.  \textit{3D
Slicer} discovers and loads necessary modules at runtime. The \textit{BioImage
Suite Web} team had designed and re-designed their software multiple times in
the last 10+ years. They found that their modular approach effectively supports
maintainability \cite{Joshi2011}.

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  Although the MI developers did not mention
reproducibility explicitly as a pain point, they did mention the need to improve
documentation.  Good documentation does not just address the pain points of lack
of developer time (\ppref{P_LackDevTime}), technology hurdles
(\ppref{P_TechnologyHurdles}), usability \ppref{P_Usability}, and
maintainability.  Documentation is also necessary for reproducibility. The
challenges of inadequate documentation are a known problem for research software
\cite{PintoEtAl2018, WieseEtAl2019} and for non-research software
\cite{LethbridgeEtAl2003}.

In our interviews, we discussed threats to reproducibility and strategies for
improving it.  The threats that were mentioned include closed-source software,
no user interaction tests, no unit tests, the need to change versions of some
common libraries, variability between CPUs, and misinterpretation of how
manufacturers create medical images.

The most commonly cited (by 6 teams) strategy to improve reproducibility was
testing (regression tests, unit tests, having good tests). The second most
common strategy (mentioned by 5 teams) is making code, data, and documentation
available, possibly by creating open-source libraries.  Other ideas that were
mentioned include running the same tests on all platforms, a dockerized version
of the software to insulate it from the OS environment, using standard
libraries, monitoring the upgrades of the library dependencies, clearly
documenting the version information, bringing along the exact versions of all
the dependencies with the software, providing checksums of the data, and
benchmarking the software against other software that overlaps in functionality.
Specifically one interviewee suggested using \textit{3D Slicer} as the benchmark
to test their reproducibility.

\end{enumerate}

\section{Recommendations} \label{ch_recommendations}

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are suggestions for consideration in the future. We
expand on some of the ideas that came out of our interviews with developers
(Section~\ref{painpoints}), including continuous integration, moving to web
applications, and enriching the test data sets. We also bring in new like
employing linters, peer review, and design for change.  Our aim is to mention
ideas that are at least somewhat beyond conventional best practices. The ideas
listed here have the potential to become best practices in the medium to
long-term. We list the ideas roughly in the order of increasing implementation
effort.

\subsection{Use Continuous Integration} \label{Sec_ContinuousIntegration}

Continuous integration involves frequent pushes to a code repository.  With
every push the software is built and tested \cite[p.\ 13]
{HumbleAndFarley2010}, \cite{ShahinEtAl2017, Fowler2006}.  CI can take
significant time and effort to set up and integrate into a team's workflow, but
the benefits are impressive, as follows:

\begin{itemize}
	\item Elimination of headaches associated with a separate integration phase
	\cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
	\item Detection and removal of bugs \cite{Fowler2006} via
	automated testing.  
	\item Everyone is always working on a stable base, since the rejection of
	inadequate commits means that the main branch will always be working.  A
	stable base will always pass all tests.  If the CI system uses generators
	and linters, it will also have current documentation and standard compliant
	code.  A stable base improves developer productivity, allowing them to focus
	on coding, testing, and documentation.
\end{itemize}

Continuous integration and delivery helps with addressing several pain points.
For instance, CI/CD helps reduce development time (\ppref{P_LackDevTime}) by
removing the need for a time-consuming integration stage and by automating
regression testing.  Automated regression tests also help with ensuring
correctness (\ppref{P_Correctness}) and the quality of reproducibility
(\qref{Q_Reproducibility}). Details on CI for MI can be found in [citation
removed for blind review] %Smith et al. \cite{SmithEtAl2024_MI_SOP} 
, including an overview of the elements of CI, how to enable incorporation of CI
into a team's workflow, how to set-up a CI system, and challenges for CI.

\subsection{Move To Web Applications} \label{sec_webapps}

Section~\ref{painpoints} describes the pain point of technology hurdles
(\ppref{P_TechnologyHurdles}), which motivates considering the use of web
applications. Here we give further advice to help with deciding whether to adopt
a web application. The decision will be based on whether, on balance, the web
application improves the four factors identified by developers: compatibility,
maintainability, performance, and security. To enable decision-making, a team
will need to prioritize between these factors, based on their objectives and
experience. The suggestions are intended to provide ideas and avenues for
exploration; a web application will not be the right fit for all projects and
all teams.

\begin{itemize}

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React}, and \href{https://elm-lang.org/}{Elm}.

\item \textbf{Backend servers can potentially deliver high performance.} Web
applications with backend servers may perform even better than native
applications. If a team needs to support lower-end computers, it is good to use
back-end servers for heavy computing tasks.  For backend servers where traffic
and latency is not an issue, options include
\href{https://www.django-rest-framework.org/} {Django}, \href{https://laravel.com/} {Laravel} and
\href{https://nodejs.org/en/} {Node.js}.  The advantage of Django is that it provides access to Python
libraries.  For backend servers where traffic and latency is an issue,
\href{https://github.com/gin-gonic/gin} {Gin} is an option.

\item \textbf{Backend servers can have low costs.} Serverless solutions from
major cloud service providers (like Amazon Web Services (AWS) and Google Cloud
Platform) may be worth exploring. Serverless solutions still use a server, but
the server provider only charges the team when they use the server. The solution
is event-driven, and costs the team by the number of requests processed. Thus,
serverless can be very cost-effective for less intensively used functions.

\item \textbf{Web transmission may diminish security.} Transferring sensitive
data on-line can be a problem for projects requiring high security. Regulations
for some MI applications may forbid doing web transmissions. In this case, a web
application with a backend may not be an option.

\end{itemize}

\subsection{Enrich the Testing Datasets} 
\label{sec_recommendations_testing_dataset}

As described in Section~\ref{painpoints}, ensuring correctness
(\ppref{P_Correctness}) via testing can be problematic because of limited access
to real-world medical imaging datasets.  We build on the suggestions we heard
from our interviewees as follows:

\begin{itemize}
\item \textbf{Build and maintain good connections to datasets.} A team can build
connections with professionals working in the medical domain, who may have
access to private datasets and can perform tests for the team.

\item \textbf{Collect and maintain datasets over time.} A team may face problems
caused by various unique inputs over the years of software development. This
data should be collected and maintained over time to form a good, comprehensive,
dataset for testing.

\item \textbf{Search for open data sources.} In general, there are many open MI
datasets.  For instance, there are
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \cite{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\cite{PriorEtAl2017}, \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \cite{Smirniotopoulos2014}, and datasets for liver
\cite{BilicEtAl2019} and brain \cite{MenzeEtAl2015} tumor segmentation
benchmarks.

\item \textbf{Create sample data for testing.} If a team can access tools
creating sample data, they may also self-build datasets for testing. For
example, an MI software development team can use an MRI scanner to create images
of objects, animals, and volunteers. The team can build the images based on
specific testing requirements.

\item \textbf{Remove privacy from sensitive data.} For data with sensitive
information, a team can ask the data owner to remove such information or add
noise to protect privacy. One example is using de-identified copies of medical
images for testing.

\end{itemize}

\subsection{Employ Linters} \label{Sec_Linters}

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \cite{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a CI system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  Despite the lack of mention in the guidelines, we believe that
linters have the potential to improve code quality at a relatively low cost.

Linters have the following benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues \cite{SourceLevel2022_Lint}. Most popular programming languages have an
accompanying linter.  For example, Python has the options of PyLint, flake8 and
Black \cite{Zadka2018}.

Linters address several MI pain points.  For instance, linters address the lack
of development time (\ppref{P_LackDevTime}) by increasing the developer's
productive time via guarding against frustrating, time-consuming, and mundane
mistakes.  Moreover, since a linter can include rules that capture the wisdom of
senior programmers, it can help newer developers avoid common mistakes. With
respect to the technology hurdle pain point (\ppref{P_TechnologyHurdles}),
linters can assist with the move toward web applications
(Section~\ref{sec_webapps}).  For instance, ESLint in React is a pluggable
linter that lets the developer know if they have imported something and not used
it, if a function could be short-handed, if there are indentation
inconsistencies, etc. By insisting on code
standardization, linters can reduce technical debt and thus improve
maintainability (\qref{Q_Maintainability}). Although linters are tools for code
analysis, the idea of statically checking for adherence to basic rules can be
extended to check documentation. Tools to enforce documentation standards
partially explains the relatively higher quality of statistical tools that are
part of the Comprehensive R Archive Network (CRAN) [citation redacted for blind
review].  %\cite{SmithEtAl2018_StatSoft}.

\subsection{Conduct a Mix of Rigorous and Informal Peer Reviews} \label{Sec_PeerReview}

We advocate incorporating peer review into the development process, as
frequently recommended for research software \cite{HerouxEtAl2008,
OrvizEtAl2017, USGS2019}. In most cases a modern, lightweight review, should be
adequate.  Modern code review is informal, tool-based, asynchronous, and focused
on reviewing code changes \cite{SadowskiEtAl2018}. Managing a project via GitHub
pull requests is an example of a modern approach to reviewing code.  Software
development organizations have moved to this lightweight style of code review
because of the inefficiencies of rigorous inspections \cite{RigbyAndBird2013}.
However, for important parts of the code, developers may benefit from mixing in
a more rigorous approach.

Fagan \cite{Fagan1976} began work on rigorous review via code inspection.
Elements of a typical inspection include reviewing the code against a checklist
(checking the consistency of variable names, look for terminating loops, etc.),
performing specific review tasks (such as summarizing the code's purpose,
cross-referencing the code to the technical manual, creating a data dictionary
for a given module, etc.) Rigorous inspection finds 60-65\% of latent defects on
average, and often tops 85\% in defect removal efficiency \cite{Jones2008}. The
success rate of code inspection is generally higher than most forms of testing,
which average between 30 --- 35\% for defect removal efficiency
\cite{EbertAndJones2009, Jones2008}. For research software, Kelly and Shepard
\cite{KellyAndShepard2000} show a task based inspection approach can be
effective.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

\subsection{Design For Change} \label{Sec_DesForChange}

We recommend that MI developers use the principle of design for change for their
modularizations.  Although the advice to modularize research software to handle
complexity is common \cite{WilsonEtAl2014, StewartEtAl2017, Storer2017},
specific guidelines on how to divide the software into modules is less
prevalent.  Not every decomposition is a good design for supporting change, as
shown by \cite{Parnas1972a}.  For instance, a design with low cohesion and high
coupling \cite[p.\ 48]{GhezziEtAl2003} will make change difficult. Especially in
research software, where change is inevitable, designers need to produce a
modularization that supports change.  Ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code
\cite{JungEtAl2022}.

Developers can accomplish design for change by first identifying likely changes,
either implicitly or explicitly, and second by hiding each likely change behind
a well-defined module interface \cite{Parnas1972a}.

\section{Threats to Validity} \label{sec_threats_to_validity}

We follow Ampatzoglou et al.'s \cite{AmpatzoglouEtAl2019} analysis of
threats to validity in software engineering secondary studies.

\subsection{Reliability}

A study is reliable if repeating it by another researcher, using the same
methodology, would lead to the same results~\cite{RunesonAndHost2009}.
We identify the following threats:

\begin{itemize}
\item One individual does the measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{itemize}

The measurement process used has previously been shown to be reasonably
reproducible.  [citation redacted] 
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
reports grading five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

\subsection{Construct Validity}

Construct validity is when the adopted metrics represent what they are intended
to measure~\cite{RunesonAndHost2009}. We have identified the following
potential issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item We could not install or build \textit{dwv}, \textit{GATE}, and
\textit{DICOM Viewer}. We used a deployed on-line version for \textit{dwv}, a
virtual machine version for \textit{GATE}, but no alternative for \textit{DICOM
Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting.
\item We approximated popularity by stars and watches.
\item Table~\ref{Tbl_Guidelines} required judgement as
not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

\subsection{Internal Validity} \label{Sec_InternalValidity}

Internal validity means that discovered causal relations are trustworthy and
cannot be explained by other factors~\cite{RunesonAndHost2009}. We identify
the following:

\begin{itemize}
\item Our search (Section~\ref{SecWhatProjects}) could have missed a relevant
package.
\item Our methodology assumes that development activities
will leave a trace in the repositories, but this is not necessarily true. For
instance, we saw little evidence of requirements
(Section~\ref{Sec_CompareArtifacts}), but teams might keep this kind of
information outside their repos.
\item We interviewed a relatively small sample of 8 teams. Their pain points
(Section~\ref{painpoints}) may not be representative of the rest of their
community.
\end{itemize}

\subsection{External Validity}

If the results of a study can be generalized to other situations, then the
study is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if the development of MI software is fundamentally
different from other research software.  Although there are differences, like
the importance of data privacy for MI data, we found the approach to developing
[other software (citation redacted)]
%% DOUBLE BLIND
%LBM software \cite{SmithEtAl2024} 
and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

\section{Conclusions} \label{ch_conclusions}

Our analysis of the state of the practice for MI domain along nine software
qualities strongly indicates that ``higher quality'' is consistent with
community ranking proxies. Although our quality measures are rather shallow, we
see this as an advantage. The shallow measures are a proxy for the importance of
\emph{first impressions} for software adoption.

\begin{table*}[!ht]
	%\begin{center}
	\renewcommand{\arraystretch}{1.8}
	\begin{tabular}{ p{3cm}p{12.5cm} }
		\toprule

		Quality & Ranked 1st or 2nd\\

		\midrule

		Installability & 3D Slicer, BioImage Suite Web, Slice:Drop, INVESALIUS\\

		\pbox{3.0cm}{Correctness \\ and Verifiability} & OHIF Viewer, 3D
		Slicer, ImageJ\\

		Reliability & SMILI, ImageJ, Fiji, 3D Slicer, Slice:Drop, OHIF
		Viewer\\

		Robustness & XMedCon, Weasis, SMILI, ParaView, OsiriX Lite,
		MicroView, medInria, ITK-SNAP, INVESALIUS, ImageJ, Horos, Gwyddion,
		Fiji, dicompyler, DicomBrowser, BioImage Suite Web, AMIDE, 3DimViewer,
		3D Slicer, OHIF Viewer, DICOM Viewer\\

		Usability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, ParaView,
		INVESALIUS, Ginkgo CADx, SMILI, OsiriX Lite, BioImage Suite Web,
		ITK-SNAP, medInria, MicroView, Gwyddion\\

		Maintainability & 3D Slicer, Weasis, ImageJ, OHIF Viewer, ParaView\\

		Reusability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, SMILI, dwv, BioImage
		Suite Web, GATE, ParaView\\

		\pbox{3.0cm}{Understandability} & 3D Slicer, ImageJ, Weasis,
		Fiji, Horos, OsiriX Lite, dwv, Drishti, OHIF Viewer, GATE, ITK-SNAP,
		ParaView, INVESALIUS\\

		\pbox{3.0cm}{Visibility and \\Transparency} & ImageJ, 3D Slicer, Fiji\\

		Overall Quality & 3D Slicer, ImageJ\\

		\bottomrule		
	\end{tabular}
	\caption{Top performers for each quality (sorted by order of quality
	measurement)} \label{topperformerstable}
	%\end{center}
\end{table*}

Our grading scores indicate that \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji} and \textit{OHIF Viewer} are the overall top four.  However, the
separation between the top performers and the others is not extreme.  Almost all
packages do well on at least a few qualities, as shown in
Table~\ref{topperformerstable}, which summarizes the packages ranked first and
second for each quality. Almost 70\% (20 of 29) of the software packages appear
in the top two for at least two qualities.  The only packages that do not appear
in Table~\ref{topperformerstable}, or only appear once, are \textit{Papaya},
\textit{MatrixUser}, \textit{MRIcroGL}, \textit{XMedCon}, \textit{dicompyler},
\textit{DicomBrowser}, \textit{AMIDE}, \textit{3DimViewer}, and
\textit{Drishti}.

While we did find a reasonable amount of documentation, especially when consider
all MI projects, there were definitely some holes. Some important documentation
(test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation)
(Section~\ref{Sec_CompareArtifacts}) were rare.

Our interviewees proposed strategies to improve the state of the practice, to
address the identified pain points, and to improve software quality. To their
list (Section~\ref{painpoints}) we added some of our own recommended strategies
(Section~\ref{ch_recommendations}). The strategies that emerged include
increasing documentation, increasing testing by enriching datasets, increasing
modularity, using continuous integration, moving to web applications, employing
linters, increasing peer review, and using design for change to guide
modularization.

A deeper understanding of the needs of the MI community will require data beyond
what is available in repositories.  Future work should involve interviewing MI
developers to better understand their ``pain points''.

\section*{Acknowledgements}

[redacted for blind review]
%% DOUBLE BLIND
% We would like to thank Peter Michalski and Oluwaseun Owojaiye for fruitful
% discussions on topics relevant to this paper.

\section*{Conflict of Interest}

On behalf of all authors, the corresponding author states that there is no
conflict of interest.

\bibliographystyle{Med_Bibliography_Style}
\bibliography{SOP_MI_CritRev}

\end{document}