### 2024-11-08 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
partway done section 8

\documentclass[doubleblind,12pt, 3p, times]{elsarticle}

\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{pbox}
\usepackage{paralist}
\usepackage[inline]{enumitem}

\newcounter{rqnum} %research question number
\newcommand{\rqtherqnum}{RQ`'\therqnum}
\newcommand{\rqref}[1]{RQ\ref{#1}}

\newcounter{pnum} %pain point number
\newcommand{\ppthepnum}{P`'\thepnum}
\newcommand{\ppref}[1]{P\ref{#1}}

\newcounter{qnum} %quality number
\newcommand{\qthepnum}{Q`'\theqnum}
\newcommand{\qref}[1]{Q\ref{#1}}

\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf
+}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}}

% prepared for: Journal of Imaging Informatics in Medicine
% https://link.springer.com/journal/10278

\journal{Critical Reviews in Biomedical Engineering}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{State of the Practice for Medical Imaging Software Based on Open Source
Repositories}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

%\author{Anonymous}
\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[ECE]{Michael Noseworthy}

\affiliation[CAS]{organization={McMaster University, Computing and Software
Department}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\affiliation[ECE]{organization={McMaster University, Electrical and Computer Engineering}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\begin{abstract}

We present the state of the practice for the development of Medical Imaging
(MI) software based on data available in open source repositories. We selected
29 projects from 48 candidates and assessed 9 software qualities
(installability, correctness/ verifiability, reliability, robustness,
usability, maintainability, reusability, understandability, and
visibility/transparency) by answering 108 questions for each. Using the
Analytic Hierarchy Process (AHP) on the quantitative data, we ranked the MI
software.  The top five are \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji},
\textit{OHIF Viewer}, and \textit{ParaView}.  This is consistent with the
community's view, with four of these also appearing in the top five using
GitHub metrics (stars-per-year).  The quality and quantity of documentation
present in a project correlates quite well with its popularity.  Generally, MI
software is in a healthy state: in the repositories
we observed 88\% of the documentation artifacts recommended by research
software development guidelines, 100\% of MI projects use version control
tools, and developers appear to use the common quasi-agile research software
development process. However, the current state of the practice deviates from
existing guidelines as some recommended artifacts are rarely present
(like a test plan, requirements specification, code of conduct, and code style
guidelines), low usage of continuous integration (17\% of the projects), low
use of unit testing (about 50\% of projects), and room for improvement with
documentation. From developper interviews , we identified 7 concerns: lack
of development time, lack of funding, technology hurdles, correctness,
usability, maintainability, and reproducibility. We recommend:
increase effort on documentation,
increase testing by enriching datasets, increase continuous integration usage,
move to web applications, employ linters, use peer reviews, and design for
change.

\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
    medical imaging, research software, software engineering, software
    quality, analytic hierarchy process
\end{keyword}

\end{frontmatter}

%% \linenumbers

\section{Introduction} \label{ch_intro}

We study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide critical information for diagnostic, analytic, and medical
applications. Given its importance, we want to understand the merits and
drawbacks of the current development processes, tools, and methodologies. We
use a software engineering lens to assess the quality of existing MI software.

\subsection{Research Questions} \label{sec_motivation}

As well as state of the practice (SOP) for MI software, we would like to
understand the impact of the often cited gap between recommended
software engineering practices and that used for most research 
software~\cite{Storer2017}. Although scientists spend
a substantial proportion of their working times on software development
\cite{Hannay2009, Prabhu2011}, few are formally trained~\cite{Hannay2009}.

Our investigation is based on the following research questions:
\begin{enumerate}
\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
{open source} software projects exist? (Section~\ref{SecWhatProjects})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Based on
quantitative measurements of each project's development practices, which
projects follow best practices? (Section~\ref{SecMeasurementResults})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
similar are the top projects identified in \rqref{RQ_HighestQuality}
to the most popular projects as viewed by the community?
(Section~\ref{Sec_VsCommunityRanking})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
do MI projects compare to general research software with respect to the
artifacts (documents, scripts and code) present in their repositories?
(Section~\ref{Sec_CompareArtifacts})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:] How
do MI projects compare to research software in general with respect to the use
of tools (Section~\ref{Sec_CompareTools}) for development
(\rqref{RQ_CompareToolsProjMngmnt}.a); and, project management
(\rqref{RQ_CompareToolsProjMngmnt}.b)?
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
How do MI projects compare to research software in general with respect to
principles, processes, and methodologies used?
(Section~\ref{Sec_CompareMethodologies})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
the pain points for developers working on MI software projects?
(Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
do the pain points of developers from MI compare to the pain points
for research software in general? (Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
developers what specific best practices are taken to address the pain points
(Section~\ref{painpoints})
and software quality concerns? (Section~\ref{sec:sq})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
What research software development practice could potentially address the
pain point concerns identified in \rqref{RQ_PainPoints}?
(Section~\ref{ch_recommendations})
\end{enumerate}

\subsection{Scope} \label{sec_scope}

We only cover MI visualization software.  We exclude other categories of MI
software, including segmentation, registration, visualization, enhancement,
quantification, simulation, plus MI archiving and telemedicine systems
(compression, storage, and communication).  We also exclude statistical analysis
and image-based physiological modelling and feature extraction, classification,
and interpretation. Software that provides MI support functions is also out of
scope; therefore, we have not assessed the toolkit libraries VTK and ITK.
Finally, Picture Archiving and Communication System (PACS), which helps users to
economically store and conveniently access images, are considered out of scope.

\subsection{Methodology} \label{SecMethodology}

We have a standard set of questions designed to assess the qualities of any
research software project.  [blind review redacted details on the author's
previous studies using the same methodology.]
%% DOUBLE BLIND
% ~\cite{SmithEtAl2021, SmithAndMichalski2022}.  This has
% been applied to MI software and Lattice Boltzmann Solvers \cite{SmithEtAl2024}.
% This builds off prior work to assess the state of the practice for such domains
% as Geographic Information Systems \cite{smith2018state}, Mesh Generators
% \cite{smith2016state}, Seismology software \cite{Smith2018Seismology}, and
% Statistical software for psychology \cite{smith2018statistical}.  
We maintain the previous constraint that the work load for measuring a given
domain should take around one person-month's worth of effort ($20$ working days
at $8$ person-hours per day).

We identify a list of potential packages (through online searches) which is then
filtered and vetted by a domain expert. We aim for roughly $30$ packages. For
each remaining package, we measure its qualities by filling in a grading
template [citation redacted for double blind]. %%DOUBLE BLIND
%~\cite{SmithEtAl2021}.  
This data is used to rank the projects with the
Analytic Hierarchy Process (AHP).  We summarize further details on the
interaction with the domain expert, software qualities, grading the software and
AHP below and in longer form in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024)~\cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Domain Expert} \label{sec_vet_software_list}

The Domain Expert vets the proposed list because online resources can be
inaccurate.  The expert also vets the AHP ranking.  For the current assessment,
our Domain Expert is [details of our domain expert removed for double-blind].
%%DOUBLE BLIND %our Domain Expert (and paper co-author) is Dr.\ Michael
%Noseworthy, Professor of Electrical and Computer Engineering at McMaster
%University, Co-Director of the McMaster School of Biomedical Engineering, and
%Director of Medical Imaging Physics and Engineering at St.\ Joseph's
%Healthcare, Hamilton, Ontario, Canada.

In advance of the first meeting with the Domain Expert, they are asked to
independently create a list of top software packages in the domain.  This helps
get the expert's knowledge refreshed in advance of the meeting.

\subsubsection{Software Qualities} \label{sec_software_quality}

Quality is defined as a measure of the excellence or worth of an entity.  As is
common practice, we do not think of quality as a single measure, but rather as
a set of measures.  That is, quality is a collection of different qualities,
often called ``ilities.''  For this study we selected 9 qualities to measure:
installability, correctness/ verifiability, reliability, robustness, usability,
maintainability, reusability, understandability, and visibility/transparency.
With the exception of installability, all the qualities are defined in Ghezzi
et al. (2003) \cite{GhezziEtAl2003}. Installability is defined as the effort
required for the installation and/or uninstallation of software in a specified
environment \cite{ISO/IEC25010}.

\subsubsection{Grading} \label{sec_grading_software}

We use an existing template [citation redacted]
%% DOUBLE BLIND
%~\cite{SmithEtAl2021} 
that is designed to measure the
aforementioned qualities. To stay within our given measurement time frame, each
package gets up to five hours of time.  Project developers can be contacted for
help regarding installation, if necessary, but we impose a cap of about two
hours on the installation process.  Figure~\ref{fg_grading_template_example}
shows an excerpt of the measurement spreadsheet.  The rows are the measures and
the columns correspond to the software packages.  [The full data is available on
Mendeley; link will be provided after refereeing.] 
% \cite{Dong2021-Data}
%provides the full set of measurement data.   %%DOUBLE BLIND

\begin{figure*}[ht!]
\includegraphics[scale=0.67]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure*}

The full template consists of 108 questions over 9 qualities.  These
questions are designed to be unambiguous, quantifiable, and measurable
with constrained time and domain knowledge.

The grader, after answering questions for each quality assigns an overall
score (between 1 and 10) based on the answers.  Several of the qualities use
the word ``surface'' to highlight that these particular qualities are a shallow
measure.  For example, usability is not measured using user studies.
Instead, we look for signs that the developers considered usability.
We use two freeware tools to collect repository related data:
\href{https://github.com/tomgi/git_stats}{GitStats} and
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}.  Further
details on quality measurement are provided in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024) \cite{SmithEtAl2024_MI_SOP}.

\subsubsection{Analytic Hierarchy Process (AHP)} \label{sec_AHP}

Developed by Saaty in the 1970s, AHP is widely used to analyze multiple criteria
decisions~\cite{VaidyaEtAl2006}. AHP organizes multiple criteria in a
hierarchical structure and uses pairwise comparisons between alternatives to
calculate relative ratios~\cite{Saaty1990}. AHP works with sets of $n$
\textit{options} and $m$ \textit{criteria}.  In our project $n=29$ and $m=9$
since there are 29 options (software products) and 9 criteria (qualities). With
AHP the sum of the grades (scores) for all products for a given quality will be
1.0.  We rank the software for each of the qualities, and then we combine the
quality rankings into an overall ranking based on the relative priorities
between qualities.

\subsubsection{Interview Methods} \label{sec_interview_methods}

The repository-based measurements are incomplete because they don't generally
capture the development process, the developer pain points, the perceived
threats to software quality, and the developers' strategies to address these
threats.  Therefore, part of our methodology involves interviewing developers.
We based our interviews on a list of 20 questions, which can be found in
[citation redacted for double blind].
%Smith et al. \cite{SmithEtAl2021}. 
Some questions are about the background of the software, the development teams,
the interviewees, and how they organize their projects.  We also ask about the
developer's understanding of the users. Some questions focus on the current and
past difficulties, and the solutions the team has found, or plan to try. We also
discuss documentation, both with respect to how it is currently done, and how it
is perceived. A few questions are about specific software qualities, such as
maintainability, understandability, usability, and reproducibility. The
interviews are semi-structured based on the question list.

\section{In-Scope Open-Source MI Software} \label{SecWhatProjects}

We initially identified 48 candidate software projects from the literature
\cite{Bjorn2017, Bruhschwein2019, Haak2015}, on-line articles \cite{Emms2019,
Hasan2020, Mu2019}, and forum discussions \cite{Samala2014}.  Then we filtered
as follows:

\begin{enumerate}

\item Removed the packages with no source code available, such as
\textit{MicroDicom}, \textit{Aliza}, and \textit{jivex}.

\item Focused on MI software that provides visualization functions.  We removed
seven packages that were toolkits or libraries, such as \textit{VTK},
\textit{ITK}, and \textit{dcm4che}, and another three that were for PACS.

\item Removed \textit{Open Dicom Viewer} as it has not received any
updates since 2011.

\end{enumerate}

The Domain Expert provided a list 12 software packages.  We found 6 packages
were on both lists: \textit{3D Slicer}, \textit{Horos}, \textit{ImageJ},
\textit{Fiji}, \textit{MRIcron} (we use its descendant \textit{MRIcroGL}) and
\textit{Mango} (we use the web version \textit{Papaya}).  The remaining six
packages were on our out-of-scope list. The Domain Expert agreed with our final
choice of 29 packages.  Table~\ref{tab_final_list} summarizes the in-scope
open-source MI software that is available at the time of measurement (the year
2020), thus answering \rqref{RQ_WhatProjects}.

\begin{table*}[ht!]
\centering
\begin{tabular}{p{3.7cm}lllllllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cmidrule{7-9}
 &  &  &  &  &  & W & M & L &  \\ \midrule
ParaView \cite{Ahrens2005} & 2002 & 2020-10 & \checkmark & 100 & 886326 & \checkmark & \checkmark & \checkmark & \checkmark \\
Gwyddion \cite{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & \checkmark & \checkmark & \checkmark &  \\
Horos \cite{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & \checkmark &  &  \\
OsiriX Lite \cite{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & \checkmark &  &  \\
3D Slicer \cite{Kikinis2014} & 1998 & 2020-08 & \checkmark & 100 & 501451 & \checkmark & \checkmark & \checkmark &  \\
Drishti \cite{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & \checkmark & \checkmark & \checkmark &  \\
Ginkgo CADx \cite{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & \checkmark & \checkmark & \checkmark &  \\
GATE \cite{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & \checkmark & \checkmark &  \\
3DimViewer \cite{TESCAN2020} & ? & 2020-03 & \checkmark & 3 & 178065 & \checkmark & \checkmark &  &  \\
medInria \cite{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & \checkmark & \checkmark & \checkmark &  \\
BioImage Suite Web \cite{Papademetris2005} & 2018 & 2020-10 & \checkmark & 13 & 139699 &
\checkmark & \checkmark & \checkmark & \checkmark \\
Weasis \cite{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & \checkmark & \checkmark & \checkmark &  \\
AMIDE \cite{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & \checkmark & \checkmark & \checkmark &  \\
XMedCon \cite{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & \checkmark & \checkmark & \checkmark &  \\
ITK-SNAP \cite{Yushkevich2006} & 2006 & 2020-06 & \checkmark & 13 & 88530 & \checkmark & \checkmark & \checkmark &  \\
Papaya \cite{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & \checkmark & \checkmark & \checkmark &  \\
OHIF Viewer \cite{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & \checkmark & \checkmark & \checkmark & \checkmark \\
SMILI \cite{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & \checkmark & \checkmark & \checkmark &  \\
INVESALIUS 3 \cite{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & \checkmark & \checkmark & \checkmark &  \\
dwv \cite{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & \checkmark & \checkmark & \checkmark & \checkmark \\
DICOM Viewer \cite{Afsar2021} & 2018 & 2020-04 & \checkmark & 5 & 30761 & \checkmark & \checkmark & \checkmark &  \\
MicroView \cite{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & \checkmark & \checkmark & \checkmark &  \\
MatrixUser \cite{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & \checkmark & \checkmark & \checkmark &  \\
Slice:Drop \cite{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & \checkmark & \checkmark & \checkmark & \checkmark \\
dicompyler \cite{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & \checkmark & \checkmark &  &  \\
Fiji \cite{Schindelin2012} & 2011 & 2020-08 & \checkmark & 55 & 10833 & \checkmark & \checkmark & \checkmark &  \\
ImageJ \cite{Rueden2017} & 1997 & 2020-08 & \checkmark & 18 & 9681 & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL \cite{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & \checkmark & \checkmark & \checkmark &  \\
DicomBrowser \cite{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & \checkmark & \checkmark & \checkmark &  \\ \bottomrule
\end{tabular}
\caption{Final software list (sorted in descending order of the number of Lines
Of Code (LOC))}
\label{tab_final_list}
\end{table*}

In Table~\ref{tab_final_list} the projects are sorted in descending order of
lines of code.  We found the initial release dates (Rlsd) for most projects and
marked the two unknown dates with ``?''. The date of the last update is the date
of the latest update, at the time of measurement. We found funding information
(Fnd) for only eight projects.  For the Number Of Contributors (NOC) we
considered anyone who made at least one accepted commit as a contributor. The
NOC is not usually the same as the number of long-term project members, since
many projects received change requests and code from the community.  With
respect to the OS, 25 packages work on all three OSs: Windows (W), macOS (M),
and Linux (L). Although the usual approach to cross-platform compatibility was
to work natively on multiple OSes, five projects achieved platform-independence
via web applications. The full measurement data for all packages is available on
[removed for blind review]
%\href{https://data.mendeley.com/datasets/k3pcdvdzj2/1} {Mendeley Data}.  
%% DOUBLE BLIND

The programming languages used in order of decreasing popularity are \CC,
JavaScript, Java, C, Python, Pascal, Matlab.  The most popular language is \CC,
for 11 of 29 projects; Pascal and Matlab were each used for a single project.

\section{Which Projects Follow Best Practices?} \label{SecMeasurementResults}

We measured the software as described in
Section~\ref{SecMethodology}.  In the absence of a specific real world context,
we assumed all nine qualities are equally important.
Figure~\ref{fg_overall_scores} shows the overall AHP scores in descending order.

\begin{figure*}[ht!]
\includegraphics[scale=0.47]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

\label{fg_overall_scores}
\end{figure*}

The top four software products \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, and \textit{OHIF Viewer} have higher scores in most criteria.
\textit{3D Slicer} has a score in the top two for all qualities; \textit{ImageJ}
ranks near the top for all qualities, except for correctness \& verifiability.
\textit{OHIF Viewer} and \textit{Fiji} have similar overall scores, with
\textit{Fiji} doing better in installability and \textit{OHIF Viewer} doing
better in correctness \& verifiability.  Given the installation problems, we may
have underestimated the scores on reliability and robustness for \textit{DICOM
Viewer}, but we compared it equally for the other seven qualities.

The overall score is based on the measurement of the identified qualities. Full
details of how the projects compare for each quality can be found in [citation
redacted for blind review].
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2024_MI_SOP}
The highlights are as follows:

\begin{description}

\item [Installability] We found installation instructions for 16 projects, but
two did not need them (\textit{BioImage Suite Web} and \textit{Slice:Drop}) as
they are web applications. 10 of the projects required extra dependencies: Five
depend on a specific browser; \textit{dwv}, \textit{OHIF Viewer}, and
\textit{GATE} needs extra libraries to build; \textit{ImageJ} and \textit{Fiji}
need an unzip tool; \textit{MatrixUser} needs Matlab; \textit{DICOM Viewer}
needs a Nextcloud platform. We were not able to install \textit{GATE},
\textit{dwv}, \textit{DICOM Viewer} even after trying for 2 hours (each).

\item [Correctness \& Verifiability] The packages with higher scores for
correctness and verifiability used a wider array of techniques to improve
correctness, and had better documentation to witness this.  For instance, we
looked for evidence of unit testing, and found evidence for only about half of
the projects. We identified five projects using continuous integration tools:
\textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{dwv}, and
\textit{OHIF Viewer}. The only requirements-related document we found was a road
map of \textit{3D Slicer}, which contained design requirements for upcoming
changes.

\item [Surface Reliability] We were able to follow the steps in the tutorials
that existed (seven packages had them.) However, \textit{GATE} could not open
macro files and became unresponsive several times, without any descriptive error
message. We found that \textit{Drishti} crashed when loading damaged image
files, without showing any descriptive error message.

\item [Surface Robustness] According to their documentation, all 29 software
packages should support the DICOM standard. To test robustness, we prepared two
types of image files: correct and incorrect formats (with the incorrect format
created by relabelling a text file to have the ``.dcm'' extension).  All
software packages loaded the correct format image, except for \textit{GATE},
which failed for unknown reasons.  For the broken format, \textit{MatrixUser},
\textit{dwv}, and \textit{Slice:Drop} ignored the incorrect format, did not show
any error message and displayed a blank image.  \textit{MRIcroGL} behaved
similarly except that it showed a meaningless image.  \textit{Drishti}
successfully detected the broken format, but the software then crashed.

\item [Surface Usability] The software with higher usability scores usually
provided both comprehensive documented guidance and a good user experience.
\textit{INVESALIUS 3} provided an excellent example of a detailed and precise
user manual. \textit{GATE} also provided numerous documents, but unfortunately
we had difficulty understanding and using them. We found getting started
tutorials for only 11 projects, but a user manual for 22 projects.
\textit{MRIcroGL} was the only project that explicitly documented expected user
characteristics.

\item [Maintainability] We gave \textit{3D Slicer} the highest score for
maintainability because we found it had the most comprehensive artifacts. Only a
few of the 29 projects had a product, developer's manual, or API (Application
Programming Interface) documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents.  Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%). Twenty-seven of the 29
projects used git for version control, with 24 of these using GitHub.
\textit{AMIDE} used Mercurial and \textit{Gwyddion} used Subversion.
\textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used SourceForge.
\textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

\item [Reusability] We have assumed that smaller code files are
likely more reusable -- see Table \ref{tab_loc_per_file} for the details.

\item [Surface Understandability] All projects had a consistent coding style
with parameters in the same order for all functions, modularized code, and,
clear comments that indicate what is done, not how. However, we only found
explicit identification of a coding standard for 3 out of the 29: \textit{3D
Slicer}, \textit{Weasis}, and \textit{ImageJ}. We also found hard-coded
constants (rather than symbolic constants) in \textit{medInria},
\textit{dicompyler}, \textit{MicroView}, and \textit{Papaya}. We did not find
any reference to the algorithms used in projects \textit{XMedCon},
\textit{DicomBrowser}, \textit{3DimViewer}, \textit{BioImage Suite Web},
\textit{Slice:Drop}, \textit{MatrixUser}, \textit{DICOM Viewer},
\textit{dicompyler}, and \textit{Papaya}.

\item [Visibility/Transparency] Generally speaking, the teams that actively
documented their development process and plans scored higher.  \textit{3D
Slicer} and \textit{ImageJ} were the only projects to include documentation for
all of the following: development process, project status, development
environment and release notes.

\end{description}

\begin{table*}[ht!]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (by reusability scores)}
\label{tab_loc_per_file}
\end{table*}

\section{Comparison to Community Ranking} \label{Sec_VsCommunityRanking}

We use GitHub stars, number of forks and
number of people watching the projects as proxies for community ranking.
Table~\ref{tab_ranking_vs_GitHub} show the statistics for data collected in July
2021.

Our ranking and GitHub popularity, at least for the top five projects, seems to
line up fairly well. However, we ranked some popular packages fairly low, such
as \textit{dwv}. This is because we were unable to build it locally, even though
we followed its installation instructions. However, we were able to use its web
version for the rest of the measurements. Additionally, this version did not
detect a broken DICOM file and instead displayed a blank image
(Section~\ref{SecMeasurementResults}).  \textit{DICOM Viewer} ranked low as we
were unable to install the NextCloud platform.

We weighted all qualities equally, which is not the likely (implicit) weight
that users implicitly use. To properly assess this would require a broad user
study.  Furthermore our measures of popularity are only \emph{proxies} which
are biased towards past rather than current preferences~\cite{Szulik2017}, as
these are monotonically increasing quantities. Finally there are often more
factors than just quality that influence the popularity of ``consumer''
products.

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table*}[ht!]
\centering
\begin{tabular}{llllll}
\toprule
Software & Comm.\ Rank & Our Rank & Stars/yr & Watches/yr & Forks/yr \\ 
\midrule
3D Slicer & 1 & 1 & 284 & 19 & 128 \\
OHIF Viewer & 2 & 4 & 277 & 19 & 224 \\
dwv & 3 & 19 & 124 & 12 & 51 \\
ImageJ & 4 & 2 & 84 & 9 & 30 \\
ParaView & 5 & 5 & 67 & 7 & 28 \\
Horos & 6 & 12 & 49 & 9 & 18 \\
Papaya & 7 & 17 & 45 & 5 & 20 \\
Fiji & 8 & 3 & 44 & 5 & 21 \\
DICOM Viewer & 9 & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 8 & 40 & 4 & 17 \\
Weasis & 11 & 7 & 36 & 5 & 19 \\
dicompyler & 12 & 26 & 35 & 5 & 14 \\
OsiriX Lite & 13 & 11 & 34 & 9 & 24 \\
MRIcroGL & 14 & 18 & 24 & 3 & 3 \\
GATE & 15 & 24 & 19 & 6 & 26 \\
Ginkgo CADx & 16 & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 17 & 6 & 18 & 5 & 7 \\
Drishti & 18 & 27 & 16 & 4 & 4 \\
Slice:Drop & 19 & 21 & 10 & 2 & 5 \\
ITK-SNAP & 20 & 13 & 9 & 1 & 4 \\
medInria & 21 & 9 & 7 & 3 & 6 \\
SMILI & 22 & 10 & 3 & 1 & 2 \\
MatrixUser & 23 & 28 & 2 & 0 & 0 \\
MicroView & 24 & 15 & 1 & 1 & 1 \\
Gwyddion & 25 & 16 & n/a & n/a & n/a \\
XMedCon & 26 & 20 & n/a & n/a & n/a \\
DicomBrowser & 27 & 22 & n/a & n/a & n/a \\
AMIDE & 28 & 23 & n/a & n/a & n/a \\
3DimViewer & 29 & 25 & n/a & n/a & n/a \\ 
\bottomrule
\end{tabular}
\caption{Software ranking by our methodology versus the community (Comm.)\
ranking using GitHub metrics (Sorted in descending order of community
popularity, as estimated by the number of new stars per year)}
\label{tab_ranking_vs_GitHub}
\end{table*}
\endgroup

Although both rankings are imperfect measures, they nevertheless suggest a
correlation between best practices and popularity. We don't know if this is
causal, in either direction (i.e. if best practices enable popularity or if
popularity increases the need for best practices).

\section{Comparing with Recommended Software Artifacts} \label{Sec_CompareArtifacts}

We use a set of nine research software development
guidelines to compare recommended software artifacts versus those present in MI
software:
\begin{itemize}
\item United States Geological Survey Software Planning Checklist
\cite{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\cite{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\cite{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \cite{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \cite{SmithAndRoscoe2018},
\item Trilinos Developers Guide \cite{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \cite{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \cite{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \cite{OrvizEtAl2017}.
\end{itemize}

\begin{table*}[ht!]
\begin{center}
\begin{tabular}{ p{3.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm}p{0.5cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} &
\cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} &
\cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table*}

We show the recommended artifacts in Table~\ref{Tbl_Guidelines}, one per row.
For a given row, a checkmark in one of the columns means that the corresponding
guideline recommends this artifact.  The last column shows whether the artifact
appears in the measured set of MI software, either not at all (blank), commonly
(C), uncommonly (U) or rarely (R).  We did our best to interpret the meaning of
each artifact consistently between guidelines and specific MI software, but the
terminology and the contents of artifacts are not standardized.  The challenge
even exists for the ubiquitous README file.  The content of README files shows
significant variation between projects~\cite{PranaEtAl2018}.  Some content is
reasonably consistent, with 97\% of README files contain at least one section
describing the `What' of the repository and 89\% offering some `How' content,
other categories are more variable.  For instance, information on
`Contribution', `Why', and `Who', appear in 28\%, 26\% and 53\% of the analyzed
files, respectively~\cite{PranaEtAl2018}.

\begin{table*}[ht!]
    \begin{center}
    \begin{tabular}{ p{3.8 cm} p{6.4 cm} p{4.5 cm}}
    \toprule
    Common & Uncommon & Rare \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    User manual (22) & Authors (14) & Dependency list (7)\\
    Release info. (22) & Frequently Asked Questions (14) & Troubleshooting guide (6)\\
     & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design documentation (5)\\
     & Citation (11) & Code style guide (3)\\
     & & Code of conduct (1)\\
     & & Requirements (1)\\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency
    (number of occurrences)}
    \label{artifactspresent}
    \end{center}
\end{table*}

Table~\ref{artifactspresent} presents our measurements for MI software.
The table groups the artifacts by frequency
into categories of common (20 to 29 ($>$67\%) packages), uncommon (10 to 19
(33-67\%) packages), and rare (1 to 9 ($<$33\%) packages).

Note that ``popularity'' in Table~\ref{Tbl_Guidelines} does not imply that these
oft recommended artifacts are the most important. Guidelines are often brief,
to encourage adoption, and thus even guidelines that mention the need for
installation instructions rarely mention uninstallation instructions.
Two items in Table~\ref{artifactspresent} do not appear in any guidelines:
\emph{Troubleshooting guide} and \emph{Developer's manual}.  However the
information within these documents overlaps with
the recommended artifacts.  Troubleshooting information often can be found in
a User Manual, while the information in a ``Developer's Manual'' is often
scattered amongst many other documents.

Three of the 26 recommended artifacts were never observed in the MI software:
i) Uninstall, ii) Test plans, and iii) Requirements. It is possible that
some of these were created but never put under version control.

Neglecting requirements documentation is unfortunately common for research
software, and MI software is no exception to this trend.  Although such
documentation is recommended by some \cite{TobiasEtAl2018, HerouxEtAl2008,
SmithAndKoothoor2016}, in practice this is rare~\cite{HeatonAndCarver2015}.
Sanders and Kelly \cite{SandersAndKelly2008} interviewed 16 scientists from 10
disciplines and found that none of the scientists created requirements
specifications, unless regulations in their field mandated such a document.
Requirements are the least commonly produced type of documentation for research
software in general \cite{Nguyen-HoanEtAl2010}.

This is unfortunate as when scientific developers are surveyed on their pain
points, Wiese et al.~\cite{WieseEtAl2019} found that software requirements and
management is the software engineering discipline that most hurts them,
accounting for 23\% of the technical problems reported by study participants.
Further adding to the misfortune, there is a widespread perception that up-front
requirements are impossible for research software \cite{CarverEtAl2007,
SegalAndMorris2008}. Fortunately a more agile approach to requirements is
feasible~\cite{Smith2016}, and research-software specific templates
exist~\cite{SmithEtAl2007}.

A theme emerges amongst the artifacts rarely observed in practice: they
are developer-focused (a list of library dependencies, a contributor's guide,
a developer Code of Conduct, coding style guidelines, product roadmap, design
documentation and API documentation).

Other communities use checklists to help with best practices. Examples include
checklists for merging branches~\cite{Brown2015}, for saving and sharing changes
to the project~\cite{WilsonEtAl2016}, for new and departing team
members~\cite{HerouxAndBernholdt2018}, for processes related to commits and
releases~\cite{HerouxEtAl2008} and for overall software
quality~\cite{ThielEtAl2020, SSI2022}.

MI projects fall somewhat short of recommended best practices, but are not alone
amongst research software projects. This gap has been documented
before~\cite{Storer2017,JohansonAndHasselbring2018}, and is known to cause
sustainability and reliability problems \cite{FaulkEtAl2009}, and to waste
development effort~\cite{deSouzaEtAl2019}.

\section{Comparison of Tool Usage Between MI and Other Research Software}
\label{Sec_CompareTools}

We use data from interviews, as described in Section~\ref{sec_interview_methods}
to answer \rqref{RQ_CompareToolsProjMngmnt} (tool usage),
\rqref{RQ_CompareMethodologies} (methodology) and \rqref{RQ_PainPoints} (pain
points). Requests were sent to all 29 projects.  Nine developers from eight of
the projects agreed to participate: \textit{3D Slicer}, \textit{INVESALIUS 3},
\textit{dwv}, \textit{BioImage Suite Web}, \textit{ITK-SNAP},
\textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We spent about 90
minutes for each interview. The full interview answers can be
found in [citation redacted for blind review].
%\cite{Dong2021}.

Developers use software tools for user support, version control (VCS),
Continuous Integration and Deployment (CI/CD), documentation and project
management.  We summarize the tool usage in each of these categories, and
compare this to the usage by the research software community.

\paragraph{User support} Table~\ref{tab_user_support_model} summarizes the user
support models by the number of projects using each model (projects may use
more than one support model).

\begin{table}[ht!]
\centering
\begin{tabular}{lc}
\toprule
\multicolumn{1}{c}{User Support Model} & Num.\ Projects \\
\midrule
GitHub issue & 24 \\
Frequently Asked Questions (FAQ) & 12 \\
Forum & 10 \\
E-mail address & 9 \\
GitLab issue, SourceForge discussions & 2 \\
Troubleshooting & 2 \\
Contact form & 1 \\ 
\bottomrule
\end{tabular}
\caption{\label{tab_user_support_model}User support models by number of projects}
\end{table}

\paragraph{Version control} Most teams interviewed (eight of nine) use GitHub
for version control, which makes it unsurprising that they also use pull
requests for managing community contributions. The experience of using GitHub
was reported as generally positive; some teams had actively migrated to GitHub
from other VCS.  This represents a considerable uptake since its rare use in
2006~\cite{Wilson2006}, which matches the uptake in other communities: 10 years
ago Nguyen et al~\cite{Nguyen-HoanEtAl2010} estimated that only 50\% of
research software projects used version control, and by 2018 this had jumped to
over 80\%~\cite{AlNoamanyAndBorghi2018} with many communities~\cite{Smith2018}
reaching 100\% already in 2018. In 2022, version control use sits at over
95\%~\cite{CarverEtAl2022}.

\paragraph{CI/CD} 
Only 17\% of MI projects use CI/CD tools. We estimated this using artifacts
present in the repository -- actual usage may be higher as some tools are
entirely external. This was the case for a study of Lattice Boltzmann Method
(LBM) software [citation redacted for blind review].  %\cite{Michalski2021}.
This contrasts with the high frequency of guidelines recommending continuous
integration
\cite{BrettEtAl2021, Brown2015, ThielEtAl2020, Zadka2018, vanGompelEtAl2016}.
A recent survey~\cite{CarverEtAl2022} of practioners (rather than projects)
suggests higher use of CI/CD with 54\% of respondents indicating that they use
it.

\paragraph{Documentation} The most popular (mentioned by about 30\% of
developers) were forum discussions and videos.  The second most popular options
(20\%) were GitHub, wiki pages, workshops, and social media. The least
frequently mentioned options (about 10\% of developers) included writing books,
and Google forms.  As a point of contrast, LBM software most often uses API
document generation tools, like doxygen and sphinx [citation redacted for blind
review].
%\cite{Michalski2021}

\paragraph{Project management} Two types of tools were mentionned:
\begin{inparaenum}[i)]
\item trackers, including GitHub, issue trackers, bug trackers and Jira; and,
\item documentation tools, including GitHub, Wiki page, Google Doc, and
Confluence.
\end{inparaenum}
Of the named tools, interviewees mentioned GitHub 3 times, and each
of the other tools once each.

Based on~\cite{JungEtAl2022}, tool use for MI
software is similar to tool use for ocean modelling software.
Both use tools for editing, compiling, code management, testing, building, and
project management, but differ
in the latter's use of Kanban boards for project management.

\section[Comparison to Other Research Software]{Comparison of Principles, Process, and
Methodologies to Research Software in General} \label{Sec_CompareMethodologies}

In our interviews, developers' responses to questions about development model
were vague, with only two interviewees following a definite model, with others
feeling their process was ``similar'' to an existing model.  Three teams
followed agile or alike, two teams  followed waterfall or alike, and three
teams explicitly stated that their process was undefined or self-directed.

This matches observations for research software in general.
Scientific developers naturally use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process~\cite{Kelly2015}.  A waterfall-like process can work for research
software~\cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rational design process~\cite{parnas1986rational}.

No interviewee mentioned a strictly defined project management process. 
% JC: I don't think this level of per-project detail makes sense in this
% review article - it should be about stats and trends, not reporting on
% minute details. We could refer the reader to Ao's thesis for the details.
% 
% The
% most common approach was following the issues, such as bugs and feature
% requests. The \textit{3D Slicer} team had weekly meetings to discuss
% the goals for the project. The \textit{INVESALIUS 3} team relied on the
% GitHub process for their project management while the \textit{ITK-SNAP} 
% team had a fixed six-month release pace. Only the interviewee from 
% \textit{OHIF} 
% mentioned the team having a project manager. The \textit{3D Slicer} team and
% \textit{BioImage Suite Web} team do nightly builds and tests. The \textit{OHIF}
% developer believes that a better project management process can improve junior
% developer efficiency while also improving internal and external communication.

Less than half of the projects used unit testing, even
though the interviewees believed that testing (including usability tests) was
the top approach to improve correctness, usability, and reproducibility.  This
level of testing matches what was observed for LBM
software [citation redacted for blind review]
%\cite{Michalski2021} 
and is apparently greater than the level of
testing for ocean modelling software~\cite{JungEtAl2022}.

All Interviewees thought that documentation was
essential to their projects, and most of them (7 of 8) said that it could save
their time to answer questions from users and developers. Half of them (4 of 8)
saw the need to improve their documentation, and only three of them thought that
their documentations conveyed information clearly enough.

\section{Developer Pain Points} \label{painpoints}

We draw out the point points the points experienced by MI
software developers, compare them to those of general research software
developers, and what practices are used to address these pain points.

Interviews identified the following issues:
\begin{enumerate*}
\item lack of time and funding,
\item technology hurdles, 
\item correctness, and 
\item usability.  
\end{enumerate*}
We first detail each pain point, and contrast this experience with that
from other domains. We then separately cover potential way to mitigate these
issues.

Pinto et al.~\cite{PintoEtAl2018} list some pain points that did not come up in
our conversations: interruptions while coding, scope bloat, lack of user
feedback, hard to collaborate on software projects, and aloneness.  Wiese et
al.~\cite{WieseEtAl2019} mentions two more: reproducibility, and software scope
determination.  [citation redacted for blind review] %\cite{SmithEtAl2024} 
adds lack of software experience, technical debt, and documentation.
We cannot conclude that these are not also issues with MI software due to our
small response size.

\begin{enumerate}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} This was felt to be the most significant obstacle, and
is not uncommon
\cite{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.
%\cite{SmithEtAl2024}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Funding:} As research software is not yet seen as essential infrastructure,
there is an ongoing struggle to attract funding and, even harder, maintain
research software. This is a common complaint%
\cite{GewaltigAndCannon2012, Goble2014, KaterbowAndFeulner2018, SmithEtAl2024}.
As software output is not always counted when judging the academic excellence
of applicants, which forces researchers who write software to spend extra
time on publicity~\cite{WieseEtAl2019}. In some domains, like biology%
\cite{HowisonAndBullard2016}, software is infrequently cited even when used
pervasively. In other words, there is a lack of a formal reward system for
research software~\cite{PintoEtAl2018}.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} Developers mentioned: hard to keep up with changes
in OS and libraries, difficult to transfer to new technologies, hard to support
multiple OSes, and hard to support lower-end
computers. Developers expressed difficulty balancing between four factors:
cross-platform compatibility, convenience of development and maintenance,
performance, and security.

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Correctness:} The most mentioned thread to correctness was complexity.
Example sources of complexity are: Interviewees identified multiple threats to
the large variety of data formats, complicated data standards, differing
outputs between medical imaging machines, and the addition of (non-viewing
related) functionality.  Other identified threats to correctness:

\begin{itemize}
\item Lack of real world image data for testing, in part because of patient
privacy concerns;
\item Tests are expensive and time-consuming because of the need for huge datasets;
\item Software releases are difficult to manage;
\item No systematic unit testing; and,
\item No dedicated quality assurance team.
\end{itemize}

\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:] \textbf{Usability:}
The discussion with the developers focused on usability issues for two classes
of users: the end users and other developers.  The threats to usability for end
users include an unintuitive user interface, inadequate feedback from the
interface (such as lack of a progress bar), users being unable to determine the
purpose of the software, not all users knowing if the software includes certain
features, not all users understanding how to use the command line tool, and not
all users understanding that the software is a web application. For developers
the threats to usability include not being able to find clear instructions on
how to deploy the software, and the architecture being difficult for new
developers to understand.

At least to some extent the problems for MI software users are due to holes in
their background knowledge.  For research software in general, users do not
always have the expertise required to install or use the software
\cite{WieseEtAl2019}. A similar pattern has been observed during interviews with
LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

\end{enumerate}

For each of the above, the MI developers suggested various potential
solutions, some of which have proven their worth in other domains.

\begin{enumerate}
\item[P\ref{P_LackDevTime}:] \textbf{Lack of Development Time:}
\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for CI/CD (Continuous Integration and Continuous
Delivery.)
\end{itemize}

\item[P\ref{P_LackFunding}:] \textbf{Lack of Funding:}
One interviewee proposed an interesting idea: Licensing the software
to commercial companies to integrate it into their products.
In general, solutions need to be more systemic, and we are seeing the rise
of ``research software engineers'' in various jurisdictions.

\item[P\ref{P_TechnologyHurdles}:] \textbf{Technology Hurdles:}
\begin{itemize}
\item Adopting a web-based approach with backend servers, to better support
lower-end computers;
\item Using memory-mapped files to consume less computer memory, to better
support lower-end computers; 
\item Using computing power from the computers GPU for web applications;
\item Maintaining better documentations to ease the development and maintenance
processes;
\item Improving performance via more powerful computers, which one interviewee
pointed out has already happened.
\end{itemize}

A number of developers perceive that some of the ``shifting technologies''
problems could be alleviated by moving to web-based applications.
Most of the teams (83\%) chose to develop
native applications. Three of the eight we interviewed were
building web applications, and another team was considering it too.

\item[P\ref{P_Correctness}:] \textbf{Correctness:}
Testing was the most often strategy for ensuring correctness, and teams
mentioned test related activities, including test-driven development, component
tests, integration tests, smoke tests, regression tests, self tests and
automated tests.  This puts MI software ahead of other domains where
insufficient testing is a problem~\cite{PintoEtAl2018} and
insufficiently well-understood~\cite{HannayEtAl2009}.

Another correctness strategy mentioned multiple times is a development process
that involves stable releases and nightly builds.  Other strategies that were
mentioned:
\begin{enumerate*}
\item using CI/CD, 
\item using de-identified copies of medical images for
debugging, 
\item sending beta versions to medical workers who can access the data to
do the tests, 
\item collecting/maintaining a dataset of problematic images,
\item using open datasets,
\item if (part of) the team belongs to a medical school or a hospital, using the
datasets they can access,
\item if the team has access to MRI scanners, self-building sample images for
testing, and
\item if the team has connections with MI equipment manufacturers, asking for
their help on data format problems.
\end{enumerate*}

Synthesizing the feedback from the interviewees, it is clear that increased
connections between the development team and medical professionals/institutions
could improve testing effectiveness and thus increase correctness.

\item[P\ref{P_Usability}:] \textbf{Usability:}

\begin{itemize}
    \item Use documentation (user manuals, mailing lists, forums) (mentioned by
    4 developers)
    \item Usability tests and interviews with end users; and, (mentioned by 3
    developers)
    \item Adjusting the software according to user feedback. (mentioned by 3
    developers)
\end{itemize}

Other suggested and practiced strategies include a graphical user interface,
testing every release with active users, making simple things simple and
complicated things possible, focusing on limited number of functions, icons with
clear visual expressions, designing the software to be intuitive, having a UX
(User eXperience) designer, dialog windows for important notifications,
providing an example for users to follow, downsampling images to consume less
memory, and providing an option to load only part of the data to boost
performance.  The last two points recognize that an important component of
usability is performance, since poor performance frustrates users.

\end{enumerate}

\section{Software Qualities}\label{sec:sq}
In addition to pain points, we summarize
MI developer strategies for improving maintainability and reproducibility.
Although the interviewees did not explicitly identify these two qualities as
pain points, we did discuss threats to these qualities and ways to improve them
as part of our interview process \cite{SmithEtAl2021}.  The interviewee's
practices for addressing pain points and improving quality can potentially be
emulated by other MI developers. Moreover, these practices may provide examples
%HERE
that can be followed by other research software domains.
Up to this point, we have covered the pain points that came up in interviews
with MI developers, along with a summary of the techniques that are currently
used to address these pain points.  Although the developers did not explicitly
identify the qualities of maintainability and reproducibility as pain points in
our interviews, as part of our interview questions
(Section~\ref{sec_interview_methods}) they did share their approaches for
improving these qualities, as discussed below.

\begin{enumerate} \item[Q\refstepcounter{qnum}\theqnum
\label{Q_Maintainability}:] \textbf{Maintainability:}  Maintainability has been
rated as the third most important software quality for research software
\cite{Nguyen-HoanEtAl2010}. The push for sustainable software
\cite{deSouzaEtAl2019} is motivated by the pain that past developers have had
with accumulating too much technical debt \cite{KruchtenEtAl2012}.  For LBM
software, technical debt has been identified as one of the developer pain points
[citation redacted for blind review]. %\cite{SmithEtAl2024}.

To improve maintainability, the most popular (with five out of nine interviewees
mentioning it) strategy is to use a modular approach, with often repeated
functions in a library.  Other strategies that were mentioned for improving
maintainability include supporting third-party extensions, an easy-to-understand
architecture, a dedicated architect, starting from simple solutions, and
documentation.  The \textit{3D Slicer} team used a well-defined structure for
the software, which they named as an ``event-driven MVC pattern''.  \textit{3D
Slicer} discovers and loads necessary modules at runtime. The \textit{BioImage
Suite Web} team had designed and re-designed their software multiple times in
the last 10+ years. They found that their modular approach effectively supports
maintainability \cite{Joshi2011}.

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  Although the MI developers did not mention
reproducibility explicitly as a pain point, they did mention the need to improve
documentation.  Good documentation does not just address the pain points of lack
of developer time (\ppref{P_LackDevTime}), technology hurdles
(\ppref{P_TechnologyHurdles}), usability \ppref{P_Usability}, and
maintainability.  Documentation is also necessary for reproducibility. The
challenges of inadequate documentation are a known problem for research software
\cite{PintoEtAl2018, WieseEtAl2019} and for non-research software
\cite{LethbridgeEtAl2003}.

In our interviews, we discussed threats to reproducibility and strategies for
improving it.  The threats that were mentioned include closed-source software,
no user interaction tests, no unit tests, the need to change versions of some
common libraries, variability between CPUs, and misinterpretation of how
manufacturers create medical images.

The most commonly cited (by 6 teams) strategy to improve reproducibility was
testing (regression tests, unit tests, having good tests). The second most
common strategy (mentioned by 5 teams) is making code, data, and documentation
available, possibly by creating open-source libraries.  Other ideas that were
mentioned include running the same tests on all platforms, a dockerized version
of the software to insulate it from the OS environment, using standard
libraries, monitoring the upgrades of the library dependencies, clearly
documenting the version information, bringing along the exact versions of all
the dependencies with the software, providing checksums of the data, and
benchmarking the software against other software that overlaps in functionality.
Specifically one interviewee suggested using \textit{3D Slicer} as the benchmark
to test their reproducibility.

\end{enumerate}

\section{Recommendations} \label{ch_recommendations}

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are suggestions for consideration in the future. We
expand on some of the ideas that came out of our interviews with developers
(Section~\ref{painpoints}), including continuous integration, moving to web
applications, and enriching the test data sets. We also bring in new like
employing linters, peer review, and design for change.  Our aim is to mention
ideas that are at least somewhat beyond conventional best practices. The ideas
listed here have the potential to become best practices in the medium to
long-term. We list the ideas roughly in the order of increasing implementation
effort.

\subsection{Use Continuous Integration} \label{Sec_ContinuousIntegration}

Continuous integration involves frequent pushes to a code repository.  With
every push the software is built and tested \cite[p.\ 13]
{HumbleAndFarley2010}, \cite{ShahinEtAl2017, Fowler2006}.  CI can take
significant time and effort to set up and integrate into a team's workflow, but
the benefits are impressive, as follows:

\begin{itemize}
    \item Elimination of headaches associated with a separate integration phase
    \cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
    \item Detection and removal of bugs \cite{Fowler2006} via
    automated testing.  
    \item Everyone is always working on a stable base, since the rejection of
    inadequate commits means that the main branch will always be working.  A
    stable base will always pass all tests.  If the CI system uses generators
    and linters, it will also have current documentation and standard compliant
    code.  A stable base improves developer productivity, allowing them to focus
    on coding, testing, and documentation.
\end{itemize}

Continuous integration and delivery helps with addressing several pain points.
For instance, CI/CD helps reduce development time (\ppref{P_LackDevTime}) by
removing the need for a time-consuming integration stage and by automating
regression testing.  Automated regression tests also help with ensuring
correctness (\ppref{P_Correctness}) and the quality of reproducibility
(\qref{Q_Reproducibility}). Details on CI for MI can be found in [citation
removed for blind review] %Smith et al. \cite{SmithEtAl2024_MI_SOP} 
, including an overview of the elements of CI, how to enable incorporation of CI
into a team's workflow, how to set-up a CI system, and challenges for CI.

\subsection{Move To Web Applications} \label{sec_webapps}

Section~\ref{painpoints} describes the pain point of technology hurdles
(\ppref{P_TechnologyHurdles}), which motivates considering the use of web
applications. Here we give further advice to help with deciding whether to adopt
a web application. The decision will be based on whether, on balance, the web
application improves the four factors identified by developers: compatibility,
maintainability, performance, and security. To enable decision-making, a team
will need to prioritize between these factors, based on their objectives and
experience. The suggestions are intended to provide ideas and avenues for
exploration; a web application will not be the right fit for all projects and
all teams.

\begin{itemize}

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React}, and \href{https://elm-lang.org/}{Elm}.

\item \textbf{Backend servers can potentially deliver high performance.} Web
applications with backend servers may perform even better than native
applications. If a team needs to support lower-end computers, it is good to use
back-end servers for heavy computing tasks.  For backend servers where traffic
and latency is not an issue, options include
\href{https://www.django-rest-framework.org/} {Django}, \href{https://laravel.com/} {Laravel} and
\href{https://nodejs.org/en/} {Node.js}.  The advantage of Django is that it provides access to Python
libraries.  For backend servers where traffic and latency is an issue,
\href{https://github.com/gin-gonic/gin} {Gin} is an option.

\item \textbf{Backend servers can have low costs.} Serverless solutions from
major cloud service providers (like Amazon Web Services (AWS) and Google Cloud
Platform) may be worth exploring. Serverless solutions still use a server, but
the server provider only charges the team when they use the server. The solution
is event-driven, and costs the team by the number of requests processed. Thus,
serverless can be very cost-effective for less intensively used functions.

\item \textbf{Web transmission may diminish security.} Transferring sensitive
data on-line can be a problem for projects requiring high security. Regulations
for some MI applications may forbid doing web transmissions. In this case, a web
application with a backend may not be an option.

\end{itemize}

\subsection{Enrich the Testing Datasets} 
\label{sec_recommendations_testing_dataset}

As described in Section~\ref{painpoints}, ensuring correctness
(\ppref{P_Correctness}) via testing can be problematic because of limited access
to real-world medical imaging datasets.  We build on the suggestions we heard
from our interviewees as follows:

\begin{itemize}
\item \textbf{Build and maintain good connections to datasets.} A team can build
connections with professionals working in the medical domain, who may have
access to private datasets and can perform tests for the team.

\item \textbf{Collect and maintain datasets over time.} A team may face problems
caused by various unique inputs over the years of software development. This
data should be collected and maintained over time to form a good, comprehensive,
dataset for testing.

\item \textbf{Search for open data sources.} In general, there are many open MI
datasets.  For instance, there are
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \cite{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\cite{PriorEtAl2017}, \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \cite{Smirniotopoulos2014}, and datasets for liver
\cite{BilicEtAl2019} and brain \cite{MenzeEtAl2015} tumor segmentation
benchmarks.

\item \textbf{Create sample data for testing.} If a team can access tools
creating sample data, they may also self-build datasets for testing. For
example, an MI software development team can use an MRI scanner to create images
of objects, animals, and volunteers. The team can build the images based on
specific testing requirements.

\item \textbf{Remove privacy from sensitive data.} For data with sensitive
information, a team can ask the data owner to remove such information or add
noise to protect privacy. One example is using de-identified copies of medical
images for testing.

\end{itemize}

\subsection{Employ Linters} \label{Sec_Linters}

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \cite{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a CI system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  Despite the lack of mention in the guidelines, we believe that
linters have the potential to improve code quality at a relatively low cost.

Linters have the following benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues \cite{SourceLevel2022_Lint}. Most popular programming languages have an
accompanying linter.  For example, Python has the options of PyLint, flake8 and
Black \cite{Zadka2018}.

Linters address several MI pain points.  For instance, linters address the lack
of development time (\ppref{P_LackDevTime}) by increasing the developer's
productive time via guarding against frustrating, time-consuming, and mundane
mistakes.  Moreover, since a linter can include rules that capture the wisdom of
senior programmers, it can help newer developers avoid common mistakes. With
respect to the technology hurdle pain point (\ppref{P_TechnologyHurdles}),
linters can assist with the move toward web applications
(Section~\ref{sec_webapps}).  For instance, ESLint in React is a pluggable
linter that lets the developer know if they have imported something and not used
it, if a function could be short-handed, if there are indentation
inconsistencies, etc. By insisting on code
standardization, linters can reduce technical debt and thus improve
maintainability (\qref{Q_Maintainability}). Although linters are tools for code
analysis, the idea of statically checking for adherence to basic rules can be
extended to check documentation. Tools to enforce documentation standards
partially explains the relatively higher quality of statistical tools that are
part of the Comprehensive R Archive Network (CRAN) [citation redacted for blind
review].  %\cite{SmithEtAl2018_StatSoft}.

\subsection{Conduct a Mix of Rigorous and Informal Peer Reviews} \label{Sec_PeerReview}

We advocate incorporating peer review into the development process, as
frequently recommended for research software \cite{HerouxEtAl2008,
OrvizEtAl2017, USGS2019}. In most cases a modern, lightweight review, should be
adequate.  Modern code review is informal, tool-based, asynchronous, and focused
on reviewing code changes \cite{SadowskiEtAl2018}. Managing a project via GitHub
pull requests is an example of a modern approach to reviewing code.  Software
development organizations have moved to this lightweight style of code review
because of the inefficiencies of rigorous inspections \cite{RigbyAndBird2013}.
However, for important parts of the code, developers may benefit from mixing in
a more rigorous approach.

Fagan \cite{Fagan1976} began work on rigorous review via code inspection.
Elements of a typical inspection include reviewing the code against a checklist
(checking the consistency of variable names, look for terminating loops, etc.),
performing specific review tasks (such as summarizing the code's purpose,
cross-referencing the code to the technical manual, creating a data dictionary
for a given module, etc.) Rigorous inspection finds 60-65\% of latent defects on
average, and often tops 85\% in defect removal efficiency \cite{Jones2008}. The
success rate of code inspection is generally higher than most forms of testing,
which average between 30 --- 35\% for defect removal efficiency
\cite{EbertAndJones2009, Jones2008}. For research software, Kelly and Shepard
\cite{KellyAndShepard2000} show a task based inspection approach can be
effective.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

\subsection{Design For Change} \label{Sec_DesForChange}

We recommend that MI developers use the principle of design for change for their
modularizations.  Although the advice to modularize research software to handle
complexity is common \cite{WilsonEtAl2014, StewartEtAl2017, Storer2017},
specific guidelines on how to divide the software into modules is less
prevalent.  Not every decomposition is a good design for supporting change, as
shown by \cite{Parnas1972a}.  For instance, a design with low cohesion and high
coupling \cite[p.\ 48]{GhezziEtAl2003} will make change difficult. Especially in
research software, where change is inevitable, designers need to produce a
modularization that supports change.  Ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code
\cite{JungEtAl2022}.

Developers can accomplish design for change by first identifying likely changes,
either implicitly or explicitly, and second by hiding each likely change behind
a well-defined module interface \cite{Parnas1972a}.

\section{Threats to Validity} \label{sec_threats_to_validity}

We follow Ampatzoglou et al.'s \cite{AmpatzoglouEtAl2019} analysis of
threats to validity in software engineering secondary studies.

\subsection{Reliability}

A study is reliable if repeating it by another researcher, using the same
methodology, would lead to the same results~\cite{RunesonAndHost2009}.
We identify the following threats:

\begin{itemize}
\item One individual does the measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{itemize}

The measurement process used has previously been shown to be reasonably
reproducible.  [citation redacted] 
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
reports grading five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

\subsection{Construct Validity}

Construct validity is when the adopted metrics represent what they are intended
to measure~\cite{RunesonAndHost2009}. We have identified the following
potential issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item We could not install or build \textit{dwv}, \textit{GATE}, and
\textit{DICOM Viewer}. We used a deployed on-line version for \textit{dwv}, a
virtual machine version for \textit{GATE}, but no alternative for \textit{DICOM
Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting.
\item We approximated popularity by stars and watches.
\item Table~\ref{Tbl_Guidelines} required judgement as
not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

\subsection{Internal Validity} \label{Sec_InternalValidity}

Internal validity means that discovered causal relations are trustworthy and
cannot be explained by other factors~\cite{RunesonAndHost2009}. We identify
the following:

\begin{itemize}
\item Our search (Section~\ref{SecWhatProjects}) could have missed a relevant
package.
\item Our methodology assumes that development activities
will leave a trace in the repositories, but this is not necessarily true. For
instance, we saw little evidence of requirements
(Section~\ref{Sec_CompareArtifacts}), but teams might keep this kind of
information outside their repos.
\item We interviewed a relatively small sample of 8 teams. Their pain points
(Section~\ref{painpoints}) may not be representative of the rest of their
community.
\end{itemize}

\subsection{External Validity}

If the results of a study can be generalized to other situations, then the
study is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if the development of MI software is fundamentally
different from other research software.  Although there are differences, like
the importance of data privacy for MI data, we found the approach to developing
[other software (citation redacted)]
%% DOUBLE BLIND
%LBM software \cite{SmithEtAl2024} 
and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

\section{Conclusions} \label{ch_conclusions}

Our analysis of the state of the practice for MI domain along nine software
qualities strongly indicates that ``higher quality'' is consistent with
community ranking proxies. Although our quality measures are rather shallow, we
see this as an advantage. The shallow measures are a proxy for the importance of
\emph{first impressions} for software adoption.

\begin{table*}[ht!]
    %\begin{center}
    \renewcommand{\arraystretch}{1.8}
    \begin{tabular}{ p{3cm}p{12.5cm} }
        \toprule

        Quality & Ranked 1st or 2nd\\

        \midrule

        Installability & 3D Slicer, BioImage Suite Web, Slice:Drop, INVESALIUS\\

        \pbox{3.0cm}{Correctness \\ and Verifiability} & OHIF Viewer, 3D
        Slicer, ImageJ\\

        Reliability & SMILI, ImageJ, Fiji, 3D Slicer, Slice:Drop, OHIF
        Viewer\\

        Robustness & XMedCon, Weasis, SMILI, ParaView, OsiriX Lite,
        MicroView, medInria, ITK-SNAP, INVESALIUS, ImageJ, Horos, Gwyddion,
        Fiji, dicompyler, DicomBrowser, BioImage Suite Web, AMIDE, 3DimViewer,
        3D Slicer, OHIF Viewer, DICOM Viewer\\

        Usability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, ParaView,
        INVESALIUS, Ginkgo CADx, SMILI, OsiriX Lite, BioImage Suite Web,
        ITK-SNAP, medInria, MicroView, Gwyddion\\

        Maintainability & 3D Slicer, Weasis, ImageJ, OHIF Viewer, ParaView\\

        Reusability & 3D Slicer, ImageJ, Fiji, OHIF Viewer, SMILI, dwv, BioImage
        Suite Web, GATE, ParaView\\

        \pbox{3.0cm}{Understandability} & 3D Slicer, ImageJ, Weasis,
        Fiji, Horos, OsiriX Lite, dwv, Drishti, OHIF Viewer, GATE, ITK-SNAP,
        ParaView, INVESALIUS\\

        \pbox{3.0cm}{Visibility and \\Transparency} & ImageJ, 3D Slicer, Fiji\\

        Overall Quality & 3D Slicer, ImageJ\\

        \bottomrule        
    \end{tabular}
    \caption{Top performers for each quality (sorted by order of quality
    measurement)} \label{topperformerstable}
    %\end{center}
\end{table*}

Our grading scores indicate that \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji} and \textit{OHIF Viewer} are the overall top four.  However, the
separation between the top performers and the others is not extreme.  Almost all
packages do well on at least a few qualities, as shown in
Table~\ref{topperformerstable}, which summarizes the packages ranked first and
second for each quality. Almost 70\% (20 of 29) of the software packages appear
in the top two for at least two qualities.  The only packages that do not appear
in Table~\ref{topperformerstable}, or only appear once, are \textit{Papaya},
\textit{MatrixUser}, \textit{MRIcroGL}, \textit{XMedCon}, \textit{dicompyler},
\textit{DicomBrowser}, \textit{AMIDE}, \textit{3DimViewer}, and
\textit{Drishti}.

While we did find a reasonable amount of documentation, especially when consider
all MI projects, there were definitely some holes. Some important documentation
(test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation)
(Section~\ref{Sec_CompareArtifacts}) were rare.

Our interviewees proposed strategies to improve the state of the practice, to
address the identified pain points, and to improve software quality. To their
list (Section~\ref{painpoints}) we added some of our own recommended strategies
(Section~\ref{ch_recommendations}). The strategies that emerged include
increasing documentation, increasing testing by enriching datasets, increasing
modularity, using continuous integration, moving to web applications, employing
linters, increasing peer review, and using design for change to guide
modularization.

A deeper understanding of the needs of the MI community will require data beyond
what is available in repositories.  Future work should involve interviewing MI
developers to better understand their ``pain points''.

\section*{Acknowledgements}

[redacted for blind review]
%% DOUBLE BLIND
% We would like to thank Peter Michalski and Oluwaseun Owojaiye for fruitful
% discussions on topics relevant to this paper.

\section*{Conflict of Interest}

On behalf of all authors, the corresponding author states that there is no
conflict of interest.

\bibliographystyle{Med_Bibliography_Style}
\bibliography{SOP_MI_CritRev}

\end{document}