### 2024-11-15 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/LetterToEditor.tex
Submitted version of paper, letter to editor and copyright transfer

\documentclass[12pt]{casletter}

\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

%\usepackage{fullpage}

% \oddsidemargin 0mm
% \evensidemargin 0mm
% \textwidth 160mm
% \textheight 180mm
%\renewcommand\baselinestretch{0.85}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 27929}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail: smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
~\\
~\\
~\\
Dr.\ Anthony McGoron\\
Editor-in-Chief\\
Critical Reviews in Biomedical Engineering\\
~\\
{\bf Re: Review Article Submission}}

\opening {Dear Dr.\ Anthony McGoron:}

We are submitting for your consideration a review article titled ``State of the
Practice for Medical Imaging Software Based on Open Source Repositories''.  This
article has not been published elsewhere and it has not been simultaneously
submitted for publication elsewhere. We have published a significantly extended
paper on the same topic on the arXiv.

We review the state of the practice for the development of Medical Imaging (MI)
software based on data available in open source repositories. We identify the
open source packages available, rank them by their application of software
engineering best practices, compare MI software to other research software and
provide recommendations to potentially improve MI software quality going
forward.

We suggest the following reviewers:

\begin{itemize}
    \item \href{https://bme.ubc.ca/?directory=alexander-weber} {Dr.\ Alexander Weber}: \texttt{aweber@bcchr.ca}
    \item \href{https://www.robarts.ca/scholl_group/people/bio/akbari_alireza.html} {Dr.\ Alireza Akbari}: \texttt{aakbari@uwo.ca}
    \item \href{https://www.se.informatik.uni-kiel.de/en/team/prof.-dr.-wilhelm-willi-hasselbring} {Dr.\ Wilhelm Hasselbring}: \texttt{hasselbring@email.uni-kiel.de}
\end{itemize}

Thank you for considering our work for publication in your journal.  If you have
any questions or concerns, please let me know.

\closing{Best regards,~\newline}
\vspace{-29mm}
\includegraphics[scale=0.7]{Signature.pdf}

\end {letter}

\end{document}