### 2024-11-26 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/AddressConcerns_LetterToEditor.tex
Letter to editor revised submission

\documentclass[12pt]{casletter}

\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

%\usepackage{fullpage}

% \oddsidemargin 0mm
% \evensidemargin 0mm
% \textwidth 160mm
% \textheight 180mm
%\renewcommand\baselinestretch{0.85}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 27929}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail: smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
~\\
~\\
~\\
Dr.\ Anthony McGoron\\
Editor-in-Chief\\
Critical Reviews in Biomedical Engineering\\
~\\
{\bf Re: Revisions to Review Article}}

\opening {Dear Dr.\ Anthony McGoron:}

Thank you for expressing your concerns. To address them, we are submitting a
revised version of our paper ``State of the Practice for Medical Imaging
Software Based on Open Source Repositories''.  The text has been modified to
eliminate the textual similarity to our paper: Smith, S., Michalski, P.,
Carette, J. et al.\ State of the Practice for Lattice Boltzmann Method Software.
Archives of Computational Methods in Engineering 31, 313--350 (2024).

The ``State of the Practice for Lattice Boltzmann Method Software'' paper is
frequently cited in the non-redacted version of our submission. Those citations
were redacted for the purpose of the double blind review.

Thank you for considering our work for publication in your journal.  If you have
any further questions or concerns, please let me know.

\closing{Best regards,~\newline}
\vspace{-29mm}
\includegraphics[scale=0.7]{Signature.pdf}

\end {letter}

\end{document}