### 2025-01-24 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/ResubmissionResponseToReviewer.tex
Preparing resubmission of SOP for MI for Crit Reviews in Bio Med Eng

\documentclass[12pt]{casletter}

\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 180mm
\textheight 200mm
%\renewcommand\baselinestretch{0.85}

\usepackage{xcolor}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 27929}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail:
                                  %  smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
    ~\\
    ~\\
    ~\\
    ~\\
    Dr.\ Anthony McGoron\\
    Editor-in-Chief\\
    Critical Reviews in Biomedical Engineering\\
    ~\\
    {\bf Re: Revisions to CRB-57015, State of the Practice for Medical Imaging
    Software Based on Open Source Repositories}}

  \opening {Dear Dr.\ McGoron:}

  Thank you and the reviewers for the feedback on our submission.  The reviewers
  provided thoughtful and constructive comments.  In response to your e-mail,
  dated January 3, 2025, we have revised the paper to incorporate the requested
  revisions.  We provide a summary below.  The revised submission also includes
  a ``diff'' version of the paper showing the additions and deletions.

  Some of the reviewer's questions are best answered by citing our related work.
  Therefore, the new version undoes the redaction originally added for the
  double blind review.

  \textbf{Reviewer \#1}

  \begin{enumerate}
  \item The relative priorities of each software quality were set equal.
  Couldn't some different weighting have been considered? Perhaps one that comes
  from interviews with developers or cited in papers? \medskip

  \emph{Papers don't agree and papers are from many domains of application for
  research software.  \cite{Nguyen-HoanEtAl2010}  Not aware of a study focused
  on medical imaging practitioners.  Interviewing a statistically significant
  number of developers is beyond the scope of this paper.  However, developers
  can see the impact of a different weighting themselves.  Point to spreadsheet.
  Say what text is added to the paper.  Put it in blue?  \smallskip}

  \item In addition to github, could any other reference have been used, such as
  the number of citations in papers or relevant websites in the field? \medskip

  \emph{Citations and websites are a good idea, but don't have this data.
  Papers unreliable because people don't tend to cite scientific softwarwe.
  cite Katz to support this.  Changes to paper.\smallskip}

  \item Where are the quality ratings for each software? \medskip

  \emph{Double blind review meant didn't show citations, but citations have the
  full quality ratings.  The overall quality rating is given in Figure 2. The
  actual numbers are not as relevant as the ranking. \smallskip}

  \item Caption tables above them, not below them. \medskip

  \emph{This has been fixed in the revised manuscript.  \smallskip}

  \item What are the possible reasons for the differences between the ranking
  found and that of GitHub, in addition to the impossibility of installing some
  of them? \medskip

  \emph{Add content from the arxiv version.  ``Further reason for discrepancies
  between our ranking and the community's ranking is that we weighted all
  qualities equally. This is not likely how users implicitly rank the different
  qualities. As a result, some projects with high community popularity may have
  scored lower with our method because of a relatively higher (compared to the
  scientific community's implicit ranking) weighting of the poor scores for some
  qualities. A further explanation for discrepancies between our measures and
  the star measures may also be due to inaccuracy with using stars to
  approximate popularity.  Stars are not an ideal measure because stars
  represent the community's feeling in the past more than they measure current
  preferences \citep{Szulik2017}.  The issue with stars is that they tend only
  to be added, not removed.  A final reason for inconsistencies between our
  ranking and the community's ranking is that, as for consumer products, more
  factors influence popularity than just quality.'')\smallskip}

  \item Some software could not be installed. What problems were found? Why not
  remove it from the list? \medskip

  \emph{Say the problems from installability section.  Didn't remove because
  only affected a few qualities.  Favoured including with a warning.
  \smallskip}

  \end{enumerate}

  \textbf{Reviewer \#2}

  \begin{enumerate}

  \item The introduction section should contain the scope, significance of the
  research by summarizing current understanding and background information,
  stating the purpose of the work, and highlighting the potential outcomes.
  Also, there's no need to divide various subsection in the introduction.
  \medskip

  \emph{Response.\smallskip}

  \item The title of the table should be placed above the table. \medskip

  \emph{This has been fixed in the revised manuscript.  \smallskip}

  \item It is not recommended to include tables in the conclusion. \medskip

  \emph{Move it.  \smallskip}

  \end{enumerate}

  \closing{Best regards,~\newline} \vspace{-29mm}
  \includegraphics[scale=0.75]{Signature.pdf}

\end {letter}

\end{document}